### 2023-06-14 app/melding_link_to_page.Rmd
initialize

---
title: "Bayesian Melding App"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: ../css/template.css
  html_document:
    df_print: paged
---

```{css, echo = FALSE}
#content{
max-width:1920px;
}

```

Shiny app exploring the implications of changing the parameters for $\alpha, \beta,$ and $P(S_1|\text{untested})$ is deployed [here](https://q-w-a.shinyapps.io/bayesian_melding_priors/).

<br><br><br><br><br><br><br><br><br>

### 2023-06-14 app/app.R

@@ -1,0 +1,917 @@
+
+# Packages ---------------------------------------------------------------
+library(shiny)
+library(tidyverse)
+library(latex2exp)
+library(future)
+library(truncdist)
+library(shinyscreenshot)
+library(shinyjs)
+library(patchwork)
+
+
+# Read Data -----------
+state_name <- "ma"
+covid_county <- "./example_county_data/ma_county_biweekly.RDS" %>%
+  readRDS()  %>%
+  select(-date) %>%
+  distinct() %>%
+  filter(fips == "25025")
+
+# covid_county <- covid_county[1:15,]
+
+corrected_original <- readRDS(
+               "./example_county_data/example_county.RDS")
+
+
+dates <- readRDS( "./example_county_data/date_to_biweek.RDS")
+
+
+
+background <- "<b></i><span style='font-size:120%;' >Background:</b></i> </span> In the implementation of Bayesian melding,
+we sample values of $\\theta = \\{ \\alpha, \\beta, P(S_1|\\text{untested})\\}$, where the sample size here is selected via the 
+<i>Sample Size</i> option.
+Then, we use the function $M: \\theta \\to \\phi$ defined by
+$M(\\theta) = \\dfrac{\\beta(1- P(S_1|\\text{untested}))}{\\beta(1-P(S_1|\\text{untested})) + \\alpha P(S_1|\\text{untested})}$,
+yielding the induced distribution $f_\\phi^{induced}$. Because we do not have an analytical form for $f_\\phi^{induced}$,
+we estimate it using kernel density estimation.
+
+Then, we compute the sampling weights $\\left(\\dfrac{f_\\phi^{direct}(\\phi_i)}{f_{\\phi}^{induced}(\\phi_i)}\\right)^{0.5}$.
+Here, $f_\\phi^{direct}(\\phi_i)$ is calculated using the density function $f_\\phi^{direct}$, which is defined based on
+information from meta-analyses on the asymptomatic rate, and $f_{\\phi}^{induced}(\\phi_i)$ is from the kernel density estimation.
+
+Sampling with these weights from our sample of $M(\\theta) = \\phi$ and our sample of $\\theta$ yields the post-melding distributions,
+where the post-melding density of $\\phi$ is the the logarithmic pool of $f_\\phi^{direct}$ and $f_\\phi^{induced}$, that is,
[…]

### 2023-06-14 R/data_cleaning.Rmd
covidestim data cleaning

---
title: "Data Cleaning"
author: "Quinn White"
date: '`r Sys.Date()`'
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: show
    css: ../../css/template.css
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval=FALSE, message=FALSE,warning=FALSE)

```

```{css, echo = FALSE}
#content{
max-width:1920px;
}

```

```{r,eval=TRUE}

library(tidyverse)
library(lubridate)

```

# File Descriptions

### Demographic 
* `population_2019.RDS` contains the population for each fips code
* `statecodes.csv` contains state names mapped to codes from [the census](https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/)
* `county_fips.tsv` contains fips code mapped to county name and state 
* `state_pop.RDS` contains state population estimates from [the census](https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/state/detail/)

### Covidestim
* `covidestim_original_all.RDS` is an RDS file of the covidestim csv from the link `https://covidestim.s3.us-east-2.amazonaws.com/latest/state/estimates.csv`, unmodified; includes dates through 2021 earlier than 2021-12-02
* `covidestim_last_weeks_all_states.RDS` contains estimates after 2021-12-02 from `https://api2.covidestim.org/latest_runs?geo_type=eq.state&select=*,timeseries(*)`
* `covidestim_biweekly_all_states.RDS` is the cases summed for each 2-week interval, including data from `https://covidestim.s3.us-east-2.amazonaws.com/latest/state/estimates.csv` to obtain dates before 2021-12-02 (latest date available is 2021-11-30) and data from `https://api2.covidestim.org/latest_runs?geo_type=eq.state&select=*,timeseries(*)` to obtain dates after  2021-12-02
* `covidestim_county_estimates.RDS` contains estimates of infections by date and fips from `https://covidestim.s3.us-east-2.amazonaws.com/latest/estimates.csv`, no modifications other than selecting variables of interest
* `covidestim_biweekly_all_counties.RDS` contains biweekly aggregates for all counties, data again from link `https://covidestim.s3.us-east-2.amazonaws.com/latest/estimates.csv`

### State-level
* `tests_daily_all_states.RDS` contains daily test results by date and state from the healthdata.gov endpoint [here](https://healthdata.gov/dataset/COVID-19-Diagnostic-Laboratory-Testing-PCR-Testing/j8mb-icvb) 
* `tests_biweekly_all_states.RDS` contains test results aggregated by two-week interval for each state

### County-level
* MA:
  * `ma/archived_webpage_mass_cases.html` is the webpage [here](https://www.mass.gov/info-details/archive-of-covid-19-cases-in-massachusetts) downloaded February 18, 2023. This page contains links to access the county-level data in Massachusetts.
  * `ma_county_original.RDS` contains the original county-level data in Massachusetts downloaded from a link on this page
  * `ma_county_biweekly` contains data aggregated by two-week interval by county
* MI:
  * `mi_county_original.RDS` contains the original county-level data in Michigan downloaded from the government website [here](https://www.michigan.gov/coronavirus/stats)
  * `mi_county_daily.RDS` contains daily data with small cleaning modifications from the original for clarity
  * `mi_county_biweekly.RDS.RDS` contains data aggregated by two-week interval by county

### Hospitalizations
* `hosp_biweekly_all_states.RDS` is biweekly hospitalizations by state, accessible [here](https://dev.socrata.com/foundry/healthdata.gov/g62h-syeh)
* `hosp_biweekly_all_counties.RDS` is biweekly hospitalizations by county, aggregated using functionality provided by [Covidestim](https://github.com/covidestim/covidestim-sources#user-content-fn-1-6778941a67ce95d7bd5fda0b62f3b0d7)
  * `hhs-hospitalizations-by-county.RDS` is the original file from Covidestim containing weekly counts of a collection of variables related to hospitalization

```{r demographic info, eval=TRUE}

# STRUCTURE: fips, state, week, negative, positive, total, posrate

data_path <- paste0(here::here(), "/data/")

########################################  
#CENSUS DATA POPULATION ESTIMATES 
########################################  
url_2019 <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv"

population_2019 <- read_csv(url_2019) %>%
  mutate(fips_code = paste0(STATE, COUNTY)) %>%
  select(fips_code, population =POPESTIMATE2019)

population_2019 %>%
  head()

saveRDS(population_2019,
        paste0(here::here(), 
        "/data/demographic/population_2019.RDS"))

```

# Covidestim

## Covidestim All States (State-level)

Due to changes in the model to handle the Omicron wave, we have to obtain dates before 2021-12-02 from a different source than the latest runs endpoint of the API.

```{r, eval = TRUE}

# set end date
end_date <- ymd("2022-02-25")

```

```{r covidestim all states last weeks, eval = TRUE}

########################################
# COVIDESTIM BIWEEKLY DATA ALL STATES
########################################

covidestim_api_link <- "https://api2.covidestim.org/latest_runs?geo_type=eq.state&select=*,timeseries(*)"
covidestim <- httr::GET(covidestim_api_link)

# only go to the 6th because these are already weekly counts
# get the last couple weeks of 2021
last_weeks <-jsonlite::fromJSON(
  httr::content(covidestim,
                as = "text",
                encoding = "UTF-8"),
  simplifyVector = TRUE,
  flatten = TRUE)  %>%
  unnest(timeseries, names_repair = "unique") %>%
  mutate( date = ymd(date),
          week = week( date)) %>%
  filter(date <= end_date & year(date) > 2020)

last_weeks <- last_weeks %>%
  select(date, 
         infections.lo = infections_p2_5,
         infections.hi =infections_p97_5,
         infections, created_at,
         state = geo_name) %>%
  mutate(week = week(date),
         created = substr(created_at, 1, 10),
         created = ymd(created))

# remove duplicates from multiple model runs by taking most recent
last_weeks <- last_weeks %>%
  group_by(week,date) %>%
  slice_max(order_by = created, n=1) %>%
  select(-c(created,created_at)) %>%
  mutate(year = year(date)) %>%
  mutate(week = case_when(
    year == 2022 ~ week + 52,
    year == 2021 ~ week
  )) %>%
  ungroup()

last_weeks %>%
  head()

saveRDS(last_weeks,
         file = paste0(
           here::here(), 
           "/data/covidestim/covidestim_last_weeks_all_states.RDS"))

```

```{r,eval=TRUE}

# access dates before 2021-12-02
covidestim_link <- "https://covidestim.s3.us-east-2.amazonaws.com/latest/state/estimates.csv"

covidestim_allstates <- read_csv(covidestim_link)

covidestim_allstates %>%
  head()

saveRDS(covidestim_allstates,
        file = paste0(here::here(), 
               "/data/covidestim/covidestim_original_all.RDS"))

```

```{r, eval = TRUE}

covidestim <- readRDS(
  paste0(
    here::here(),
    "/data/covidestim/covidestim_original_all.RDS"))

# join data from each source to include dates before and after 2021-12-02
covidestim <- covidestim %>%
  select(date, contains("infections"), state) %>%
  filter(date <= end_date & year(date) > 2020) %>%
  mutate(week = week(date), year = year(date)) %>%
  mutate(week = case_when(
    year == 2022 ~ week + 52,
    year == 2021 ~ week
  )) %>%
  bind_rows(last_weeks) %>%
  group_by(week, state, year) %>%
  summarize(across(contains("infections"), sum), date = min(date)) %>%
  ungroup()

# look at discontinuity for example
covidestim %>%
  filter(state == "Colorado") %>%
  mutate(source = ifelse(date <= ymd("2021-12-02"), "1", "2")) %>%
  ggplot(aes(x = week, 
             y = infections,
             color = source,
             ymin = infections.lo,
             ymax = infections.hi)) +
  geom_ribbon(alpha = .5, aes(fill = source)) +
  geom_line() +
  geom_point(alpha = .7) +
  theme_bw() +
  labs(title = "Change in Data Source, Colorado")

covidestim %>%
  filter(state == "Massachusetts") %>%
  mutate(source = ifelse(date <= ymd("2021-12-02"), "1", "2")) %>%
  ggplot(aes(x = week, 
             y = infections,
             color = source,
             ymin = infections.lo,
             ymax = infections.hi)) +
  geom_ribbon(alpha = .5, aes(fill = source)) +
  geom_line() +
  geom_point(alpha = .7) +
  theme_bw() +
  labs(title = "Change in Data Source, Massachusetts")

###############
# add biweek
###############

num_weeks <- covidestim %>% 
  pull(week) %>% 
  unique() %>% 
  length()

num_biweeks <- num_weeks/2

biweek <- tibble(biweek = c(rep(1:num_biweeks, 2))) %>%
  arrange(biweek)

biweek_to_week <- covidestim %>%
  select(week) %>%
  distinct() %>%
  arrange(week) %>%
  cbind(biweek =biweek)

# calculate biweekly sums
covidestim_biweekly <- covidestim %>%
  left_join(biweek_to_week) %>%
  group_by(biweek, state)  %>%
  mutate(across(
    contains("infections"), 
    sum)) %>%
  ungroup()

covidestim_biweekly %>%
  select(-week) %>%
  filter(state == "Massachusetts") %>%
  mutate(source = ifelse(date <= ymd("2021-12-02"), "1", "2")) %>%
  ggplot(aes(x = date, 
             y = infections,
             color = source,
             ymin = infections.lo,
             ymax = infections.hi)) +
  geom_ribbon(alpha = .5, aes(fill = source)) +
  geom_line() +
  geom_point(alpha = .7) +
  theme_bw() +
  labs(title = "Change in Data Source, Massachusetts")

# add state codes
statecodes <- read_csv(paste0(
  here::here(),
  "/data/demographic/statecodes.csv")) %>%
  rename(state_name = state,
         state = code) %>%
  select(-abbrev)

covidestim_biweekly %>%
  rename(state_name = state) %>%
  left_join(statecodes) %>%
  saveRDS(paste0(
    here::here(),
    "/data/covidestim/covidestim_biweekly_all_states.RDS"))

```

## Covidestim All States (County Level)

```{r county level estimates, eval=TRUE}

# if legacy file does not exist, acquire it from link
if(!file.exists(
  paste0(
      here::here(), 
      "/data/covidestim/covidestim_county_estimates.RDS"))) {

  legacy_link <- "https://covidestim.s3.us-east-2.amazonaws.com/latest/estimates.csv"

  estimates_legacy <- read_csv(legacy_link) %>%
    select(fips, date, infections)

  saveRDS(
    estimates_legacy,
    paste0(
      here::here(), 
      "/data/covidestim/covidestim_county_estimates.RDS"))
}

```

```{r,eval=TRUE}

county_estimates_path <- paste0(
  here::here(), 
  "/data/covidestim/covidestim_county_estimates.RDS")

covidestim_county <- readRDS(county_estimates_path) %>%
  filter(date <= end_date &  year(date) > 2020) %>%
  mutate(week = week(date),
         year = year(date)) %>%
  mutate(week = case_when(
    year == 2022 ~ week + 52,
    year == 2021 ~ week
  ))

num_weeks <- covidestim_county %>% 
  pull(week) %>% 
  unique() %>% 
  length()

num_biweeks <- num_weeks/2

biweek <- tibble(biweek = c(rep(1:num_biweeks, 2))) %>%
  arrange(biweek)

biweek_to_week <- covidestim_county %>%
  select(week) %>%
  distinct() %>%
  arrange(week) %>%
  cbind(biweek =biweek)

```

```{r, include = FALSE, eval = FALSE}
########################################
# HANDLE GROUPED COUNTIES IN MA
########################################
# these counties are grouped in the data reported by MA, 
# so grouping them here
county_fips <- read_tsv("./data/county_fips.tsv") %>%
  dplyr::rename_with(.cols =everything(), tolower)

data_ma <- readRDS("./data/ma/all.RDS")

together <- county_fips %>% 
  filter(name == "Dukes" | name == "Nantucket") %>%
  pull(fips) %>%
  paste0(collapse = ",")

#### fix population to add populations together
grouped <- county_fips %>% 
  filter(name == "Dukes" | name == "Nantucket") %>%
  pull(fips)

# NOTE = it turns out covidestim doesn't actually include these counties, so nothing to fix
covidestim_county <- covidestim_county %>%
  mutate(fips = ifelse(fips %in% grouped, paste0(fips, sep = ","), fips))

```

```{r,eval=TRUE}
############################################
# SUM FOR EACH BIWEEK IN EACH COUNTY
############################################

covidestim_biweekly_all_counties <- covidestim_county %>%
  left_join(biweek_to_week) %>%
  group_by(biweek, fips)  %>%
  mutate(across(contains("infections"), sum)) %>%
  ungroup()

covidestim_biweekly_all_counties %>%
  head()

covidestim_biweekly_all_counties %>% 
  saveRDS(file = paste0(
    here::here(), 
    "/data/covidestim/covidestim_biweekly_all_counties.RDS"))

```

# State-level Tests (CDC)

PCR test results at the state-level from healthdata.gov endpoint [here](https://healthdata.gov/dataset/COVID-19-Diagnostic-Laboratory-Testing-PCR-Testing/j8mb-icvb).

```{r, eval=TRUE}

dates <- readRDS(paste0(here::here(),
                 "/data/date_to_biweek.RDS"))

# PAGE: https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/state/detail/
state_population_link <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/state/detail/SCPRC-EST2019-18+POP-RES.csv"
state_pop <- read_csv(state_population_link)

saveRDS(state_pop, paste0(
  here::here(), 
  "/data/demographic/state_pop.RDS"))

statecodes <- read_csv(paste0(
  here::here(), 
  "/data/demographic/statecodes.csv"))

state_pop <- state_pop %>%
  left_join(statecodes, by = c("NAME" = "state")) %>%
  select(population = POPESTIMATE2019,
         state = code) %>%
  filter(!is.na(state))

```

```{r, eval=TRUE}

# split into two queries to ensure we obtain all data with the API limit

dat <- httr::GET(URLencode(
  paste0("https://healthdata.gov/resource/j8mb-icvb.json?",
  "$where=date between '2020-12-30T12:00:00' and '2021-10-15T12:00:00'&$limit=50000")))

cdc1 <-jsonlite::fromJSON(
      httr::content(dat,
                    as = "text", 
                    encoding = "UTF-8")) %>%
  as_tibble()

dat2 <- httr::GET(URLencode(
  paste0("https://healthdata.gov/resource/j8mb-icvb.json?",
  "$where=date between '2021-10-15T14:00:00' and '2022-02-25T14:00:00'&$limit=50000")))

cdc2 <-jsonlite::fromJSON(
      httr::content(dat2,
                    as = "text", 
                    encoding = "UTF-8")) %>%
  as_tibble()

cdc <- bind_rows(cdc1, cdc2) %>%
  mutate(date = ymd(substr(date,1,10))) %>%
  mutate(across(c(new_results_reported), as.numeric)) %>%
  filter(!state %in% c("MP", "AS", "GU", "PR", "VI", "MH"))

# overall_outcome is the outcome of the test (Inconclusive, Negative, or Positive)
# new_results_reported is the number with the given outcome
cdc_pos <- cdc %>%
  select(-c(fema_region, total_results_reported)) %>%
  pivot_wider(names_from = c("overall_outcome"),
              values_from = c("new_results_reported")) %>%
  mutate(total = Inconclusive + Negative + Positive) %>%
  rename_with(tolower) %>%
  select(state, positive, total, date)

saveRDS(cdc_pos, paste0(here::here(), 
                        "/data/state_level/tests_daily_all_states.RDS"))

cdc_pos %>%
  left_join(dates) %>%
  filter(!is.na(biweek)) %>%
  group_by(biweek, state) %>%
  mutate(positive = sum(positive, na.rm=TRUE),
         total = sum(total, na.rm = TRUE)) %>%
  ungroup() %>%
  mutate(posrate = positive/total) %>%
  left_join(state_pop) %>%
  saveRDS(paste0(here::here(), 
                 "/data/state_level/tests_biweekly_all_states.RDS"))

```

# State-level Tests (JHU)

```{r}

jhu <- read_csv('https://raw.githubusercontent.com/govex/COVID-19/master/data_tables/testing_data/time_series_covid19_US.csv')

jhu %>%
  mutate(date=as_date(date, format= "%m/%d/%Y")) %>%
  filter(date >= "2021-03-03" & date <= "2022-03-02") %>% 
  left_join(dates) %>%
  arrange(date) %>%
 # mutate(across(where(is.numeric), abs)) %>%
  mutate(people_viral_positive = people_viral_positive - lag(people_viral_positive, n = 1),
         people_viral_total = people_viral_total - lag(people_viral_total, n = 1)) %>%
#  filter(people_viral_positive < people_viral_total) %>%
  group_by(biweek,state) %>%
   # mutate(across(where(is.numeric), abs)) %>%
  mutate(across(c(people_viral_positive,people_viral_total), sum, na.rm=TRUE)) %>%
  select(date, biweek, people_viral_positive, people_viral_total, state) %>%
  group_by(state) %>%
  summarize(n=n_distinct(biweek))

dates <- readRDS(here('data/date_to_biweek.RDS'))

library(slider)

state_testing <- jhu %>%
  mutate(date=as_date(date, format= "%m/%d/%Y")) %>%
  filter(date >= "2021-02-26" & date <= "2022-02-25") %>% 
  group_by(state) %>%
  arrange(state, date) %>%
 # mutate(across(where(is.numeric), abs)) %>%
  mutate(people_viral_positive = people_viral_positive - lag(people_viral_positive, n = 1, order_by=date),
         people_viral_total = people_viral_total - lag(people_viral_total, n = 1,  order_by=date),
         ) %>%
  mutate(people_viral_positive = ifelse(people_viral_positive < 0, NA,people_viral_positive),
         people_viral_total = ifelse(people_viral_total < 0, NA,people_viral_total), 
  people_viral_positive_roll = slide_dbl(people_viral_positive, mean, .before=4, .after =4),
         people_viral_total_roll = slide_dbl(people_viral_total, mean, .before=4, .after =4),
         #  people_viral_positive = ifelse(is.na(people_viral_positive) | people_viral_positive==0,
         #                                 people_viral_positive_roll, 
         #                                 people_viral_positive),
         # people_viral_total = ifelse(is.na(people_viral_total) | people_viral_total==0,
         #                             people_viral_total_roll,
         #                             people_viral_total)
  ) %>%
  left_join(dates) %>%
#  filter(people_viral_positive < people_viral_total) %>%
  group_by(biweek,state) %>%
   # mutate(across(where(is.numeric), abs)) %>%
  mutate(across(c(people_viral_positive,people_viral_total),  function(x) sum(x, na.rm = FALSE))) %>%
  mutate(has_both = !is.na(people_viral_positive) & !is.na(people_viral_total)) %>%
  select(biweek, people_viral_positive, people_viral_total, state, has_both)  %>% 
  distinct() %>%
  ungroup()

state_testing %>%
  filter(has_both) %>%
  filter(biweek >=6)

states_keep <- state_testing %>%
  mutate(has_both = people_viral_positive > 0 & people_viral_total >0) %>%
  mutate(people_viral_total = ifelse(people_viral_total ==0, NA,  people_viral_total ),
         people_viral_total = ifelse(people_viral_positive ==0, NA,  people_viral_total )) %>%
  filter(has_both) %>%
  group_by(state)%>%
  summarize(n=n_distinct(biweek)) %>%
  filter(n >20)

state_testing %>%
  filter(biweek <28) %>%
  ggplot(aes(x=biweek, y = people_viral_positive/people_viral_total)) +
  geom_line() +
  facet_wrap(~state)

state_testing <- state_testing %>%
  group_by(state) %>%
  arrange(state,biweek) %>%
  mutate(pos_roll =slide_dbl(people_viral_positive, mean, .before=3, .after =2),
         total_roll =slide_dbl(people_viral_total, mean, .before=3, .after =2) ) %>%
  ungroup() %>%
  mutate(people_viral_total = ifelse(people_viral_total < people_viral_positive,
                                     total_roll,
                                     people_viral_total),
         people_viral_positive = ifelse(people_viral_total < people_viral_positive,
                                     pos_roll,
                                     people_viral_positive))  %>%
#  filter(people_viral_total < people_viral_positive) %>%
  mutate(people_viral_total = ifelse(people_viral_total < people_viral_positive,
                                     NA,
                                     people_viral_total),
         people_viral_positive = ifelse(people_viral_total < people_viral_positive,
                                     NA,
                                     people_viral_positive) ) %>%
  select(state, positive = people_viral_positive, total = people_viral_total, biweek) %>%
  mutate(posrate= positive/total) %>%
  filter(!is.na(positive) & !is.na(total) & !is.na(posrate)) %>%
  filter(state %in% states_keep$state) %>%
  left_join(state_pop)

state_all_biweeks <- state_testing %>% select(biweek, positive,state,population) %>%
  pivot_wider(names_from = biweek, values_from=positive) %>%
  pivot_longer(cols =3:ncol(.), values_to="positive", names_to="biweek") %>%
  mutate(biweek=as.integer(biweek)) %>%
  select(-positive)

state_testing <- state_testing %>%
  right_join(state_all_biweeks) %>%
  group_by(state) %>%
  arrange(biweek) %>%
  mutate(pos_roll =slide_dbl(positive, mean, na.rm=TRUE, .before=2, .after =2),
         total_roll =slide_dbl(total, mean, na.rm = TRUE, .before=2, .after =2)) %>%
  mutate(positive= ifelse(is.na(positive), pos_roll, positive),
         total = ifelse(is.na(total), total_roll, total)) %>%
  mutate(posrate=positive/total)

state_testing %>%
  filter(!state %in% c("MP", "AS", "GU", "PR", "VI", "MH")) %>%
   saveRDS(paste0(here::here(), 
                 "/data/state_level/tests_biweekly_all_states_JHU.RDS"))

```

# County-level Cases

## Michigan

```{r michigan county, eval=TRUE}

county_fips <- read_tsv(paste0(
  here::here(),
  "/data/demographic/county_fips.tsv")) %>%
  dplyr::rename_with(.cols =everything(), tolower)

###############################
# MICHIGAN BIWEEKLY DATA
###############################

if(!file.exists(paste0(here::here(),
                     "/data/county_level/mi/mi_county_original.RDS"))) {
  link <- paste0("https://www.michigan.gov/coronavirus/-/media/Project/Websites/coronavirus/",
                 "Michigan-Data/09-27-2022/Datasets/Diagnostic-Tests-by-Result-and-County-2022-09-27.xlsx?",
                 "rev=7ba61151dcff4b038e33ea9ead95137c&hash=6C1AA97E93A30A3C1F10171233FDF31A")
  httr::GET(link, httr::write_disk(tf <- tempfile(fileext = ".xlsx")))
  df <- readxl::read_excel(tf)
  saveRDS(df, file = paste0(
    here::here(),
    "/data/county_level/mi/mi_county_original.RDS"))
}

counts_raw <- readRDS(paste0(
  here::here(),
  "/data/county_level/mi/mi_county_original.RDS"))

# basic reformatting; join fips code and population
counts_raw <- counts_raw %>%
  mutate(week = week(MessageDate),
         date = ymd(MessageDate)) %>%
  dplyr::rename_with(.cols =everything(), tolower) %>%
  mutate(state = "MI") %>%
  left_join(county_fips,
            by = c("county"="name",
                   "state" = "state")) %>%
  select(-messagedate) %>%
  left_join(population_2019, by = c("fips" = "fips_code"))

counts_raw %>%
  saveRDS(., file =  paste0(
    here::here(),
  "/data/county_level/mi/mi_county_daily.RDS"))

county <- counts_raw %>% 
  mutate(
    week = week(date),
    year = year(date)) %>%
  filter(year > 2020 & date <= end_date) %>%
  mutate( week = case_when(
    year == 2021 ~ week,
    year == 2022 ~ week + 52)) %>%
  select(fips, county, positive, 
         date, total, week, 
         negative, population) %>%
  mutate(state = "MI")

######################
# ADD BIWEEK
######################
num_weeks <- unique(county$week) %>%
  length()

num_biweeks <- num_weeks/2

biweek <- tibble(biweek = c(rep(1:num_biweeks, 2))) %>%
  arrange(biweek)

biweek_to_week <- county %>%
  select(week) %>%
  distinct() %>%
  arrange(week) %>%
  cbind(biweek =biweek)

county_biweekly <- county %>%
  left_join(biweek_to_week) %>%
  group_by(biweek, fips)  %>%
  mutate(across(c(total,positive, negative),
                 sum)) %>%
  ungroup() %>%
  mutate(posrate = positive/ total) %>%
  select(-week) %>%
  filter(!is.na(fips))

county_biweekly %>%
  select(biweek,date) %>%
    distinct() %>%
  saveRDS(paste0(here::here(),
                 "/data/date_to_biweek.RDS"))

county_biweekly %>%
  select(-date) %>%
  distinct() %>%
  saveRDS(paste0(here::here(), 
               "/data/county_level/mi/mi_county_biweekly.RDS"))

```

## Massachusetts

Obtain data from the webpage [here](https://www.mass.gov/info-details/archive-of-covid-19-cases-in-massachusetts). A downloaded version of this webpage and data downloaded from a link it contains (downloaded Feb 18, 2023) is included in this repository.

```{r,eval=TRUE}

county_fips <- read_tsv(paste0(
  here::here(),
  "/data/demographic/county_fips.tsv")) %>%
  dplyr::rename_with(.cols =everything(), tolower)

url_2019 <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv"
population_2019 <- read_csv(url_2019) %>%
  mutate(fips_code = paste0(STATE, COUNTY)) %>%
  select(fips_code, 
         population = POPESTIMATE2019,
         county_name = CTYNAME)

```

```{r,eval=TRUE}

mass_html_page <- paste0(
  here::here(),
  "/data/county_level/ma/archived_webpage_mass_cases.html")

# SCRAPE LINKS FROM HTML
all_links <- tibble(lines = c(readLines(mass_html_page))) %>%
  filter(grepl("/doc/covid-19-raw-data-", lines)) %>%
  # extract the part in quotes
  mutate(lines = str_extract(lines, '".*."'))%>%
  # remove double quotes
  mutate(lines = gsub('"', "", lines)) %>%
  # remove any text before the www.mass.gov part of the link
  mutate(lines = gsub(".*www.mass.gov", "", lines)) %>%
  mutate(link = paste0("https://www.mass.gov", lines)) %>% 
  pull(link)

read_data <- function(link) {
  httr::GET(link, httr::write_disk(tf <- tempfile(fileext = ".xlsx")))
  df <- readxl::read_excel(tf, sheet = "County_Weekly")

  df %>% 
    select(contains("date"), 
           total = `Total Tests (Last 14 days)`,
           positive = `Total Positive Tests (Last 14 days)`,
           county_name =County) %>%
    mutate( negative = total - positive,
            state = "MA")

}

# take the most recent; will filter to only include needed dates
file_link <- all_links[1]

if(!file.exists(paste0(
    here::here(),
    "/data/county_level/ma/ma_county_original.RDS"))) {

  ma_data <- read_data(file_link)
  ma_data %>%
  saveRDS(paste0(
    here::here(),
    "/data/county_level/ma/ma_county_original.RDS"))

}

```

```{r,eval=TRUE}

data_ma <- readRDS(
  paste0(
    here::here(),
    "/data/county_level/ma/ma_county_original.RDS"))

# Dukes and Nantucket counties are combined
data_ma$county_name %>% unique()

# only NAs are from counties that were combined
data_ma %>%
  filter(is.na(total) | is.na(positive)) %>%
  pull(county_name) %>%
  unique()

data_ma <- data_ma %>%
  filter(!(is.na(total) | is.na(positive)))

together <- county_fips %>% 
  filter(name == "Dukes" | name == "Nantucket") %>%
  pull(fips) %>%
  paste0(collapse = ",")

##################################################################
# add Dukes and Nantucket county populations 
# since their test numbers are combined
##################################################################
grouped_counties_fips <- county_fips %>% 
  filter(name == "Dukes" | name == "Nantucket") %>%
  pull(fips)

county_fips <- county_fips %>%
  filter(!(fips %in% grouped_counties_fips)) %>%
  bind_rows(tibble(name = "Dukes and Nantucket", 
                   fips = paste0(grouped_counties_fips, 
                                 collapse=","),
                   state = "MA"))

# sum populations for grouped counties
population_2019 <- population_2019 %>%
  mutate(fips_code = ifelse(fips_code %in% grouped_counties_fips,
                            paste0(grouped_counties_fips, 
                                 collapse=","), 
                            fips_code)) %>%
  group_by(fips_code) %>%
  summarize(population = sum(population))

```

```{r, eval =TRUE}

data_ma <- data_ma %>%
  rename(report_date = `Report Date`,
         start_period_date = `Start Date`,
         end_period_date = `End Date`)%>%
  filter(!county_name %in% c("All of Massachusetts", 
                             "Unknown County")) %>%
  mutate(year = year(start_period_date),
         week = week(start_period_date)) %>%
  filter(year> 2020 & start_period_date <= end_date) %>%
  mutate( week = case_when(
    year == 2021 ~ week,
    year == 2022 ~ week + 52)) %>%
  mutate(across(contains("date"),ymd))

# add fips 
data_ma <- data_ma %>%
  mutate(county_name = gsub(" County", "", county_name),
         county_name = gsub(" Counties", "", county_name),
         county_name = trimws(county_name)) %>%
  left_join(county_fips, 
            by = c("state"="state", "county_name"="name"))

# note that dates are overlapping, so remove the duplicates
# so that we have non-overlapping two-week intervals
data_ma %>% 
  select(start_period_date, end_period_date, week) %>% 
  distinct() %>% 
  head(n = 30)

# note that dates are overlapping, so remove the duplicates
startdate_to_biweek <- data_ma %>%
  filter(!week %% 2 ==0) %>%
  select(-week) %>%
  select(start_period_date) %>%
  distinct() %>%
  arrange(start_period_date) %>%
  mutate(biweek = row_number())

# note that summing by biweek is not needed because structure is already in 
# 2 week interval format
data_ma  <- data_ma %>%
  filter(!week %% 2 ==0) %>%
  left_join(startdate_to_biweek) %>%
  select(-c(week, end_period_date,report_date)) %>%
  left_join(population_2019, by = c("fips" = "fips_code")) %>%
  rename(date = start_period_date) %>%
  mutate(posrate = positive/total)

data_ma %>%
  saveRDS(
    paste0(here::here(),
           "/data/county_level/ma/ma_county_biweekly.RDS"))

```

# Hospitalizations

## State-level Hospitalizations

Hospitalization data from the healthdata.gov endpoint [here](https://dev.socrata.com/foundry/healthdata.gov/g62h-syeh).

```{r,eval=TRUE}

##### WEEKLY
# https://dev.socrata.com/foundry/healthdata.gov/g62h-syeh
hosp <- httr::GET(URLencode(
  paste0("https://healthdata.gov/resource/g62h-syeh.json?",
  "$where=date between '2020-12-31T12:00:00' and '2022-02-25T14:00:00'&$",
  "select=deaths_covid,date,percent_of_inpatients_with_covid,",
  "inpatient_beds_used_covid,inpatient_beds_used,state&$limit=50000")))

hosp <-jsonlite::fromJSON(
  httr::content(hosp,
                as="text",
                encoding = "UTF-8")) %>%
  as_tibble()

hosp <- hosp %>%
  # reformat dates from 2021-01-01T00:00:00.000 format
  # to date foramt
  mutate(date = gsub("T.*", "",date),
         date = as.Date(date, format = "%Y-%m-%d"),
         week = week(date)) %>%
  mutate(across(-c(date,state), as.numeric))

hosp <- hosp %>%
  mutate(
    year = year(date)) %>%
  mutate(across(-c(date,state), as.numeric)) %>%
  mutate(week = case_when(
    year == 2021 ~ week,
    year == 2022 ~ week + 52
  ))

num_weeks <- unique(hosp$week) %>%
  length()

num_biweeks <- num_weeks/2

biweek <- tibble(biweek = c(rep(1:num_biweeks, 2))) %>%
  arrange(biweek)

biweek_to_week <- hosp %>%
  select(week) %>%
  distinct() %>%
  arrange(week) %>%
  cbind(biweek =biweek)

####################################
#save biweekly hospitalizations
####################################
hosp_biweekly <- hosp %>%
  left_join(biweek_to_week) %>%
  group_by(biweek, state) %>%
  mutate(across(c(deaths_covid, 
                  inpatient_beds_used_covid,
                  inpatient_beds_used),
                sum))

hosp_biweekly %>%
  head(n= 30)

hosp_biweekly %>%
  saveRDS(paste0(here::here(),
                 "/data/state_level/hosp_biweekly_all_states.RDS"))

```

## County-level Hospitalizations

Hospitalization data is from the Covidestim data sources repository [here](https://github.com/covidestim/covidestim-sources); details available under the *Hospitalization data pipeline section* of their Readme. Briefly, they use HHS facility-level hospitalization data and using geographic information to aggregate by county.

Values between 1 and 3 are censored, so the Covidestim developers included two estimates to bound the possible values:

* `admissionsAdultsConfirmed_min`: The smallest the outcome could be - all censored values, which each represent a possible range of 1-3, will be resolved to 1.
* `admissionsAdultsConfirmed_max:`  censored values taken to be 3

```{r, eval = TRUE}

dates <- readRDS(paste0(here::here(),
                 "/data/date_to_biweek.RDS"))

county_hosp <- readRDS(paste0(
    here::here(),
    "/data/county_level/hosp/hhs-hospitalizations-by-county.RDS")) %>%
  filter(year(weekstart) >= 2021 & weekstart <= end_date)

county_hosp %>% head()

# sum by biweek
county_hosp <- county_hosp %>%
  select(fips,
         weekstart,
         admissionsAdultsConfirmed_max,
         admissionsAdultsConfirmed_min) %>%
  left_join(dates, by = c("weekstart"="date")) %>%
  rename(date=weekstart)%>%
  group_by(fips,biweek) %>%
  summarize(across(contains("admissions"), sum),
            date = min(date)) %>%
  ungroup()

county_hosp %>% head()

county_hosp %>%
  saveRDS(paste0(
    here::here(),
    "/data/county_level/hosp/hosp_biweekly_all_counties.RDS"))

```

# COVID-19 Trends and Impact Survey Data

Accessed through the [Covidcast API](https://cmu-delphi.github.io/delphi-epidata/api/covidcast-signals/fb-survey.html).

From the documentation:

* `smoothed_wtested_positive_14d`: positivity rate among those tested for COVID-19 in previous 14 days
* `smoothed_wscreening_tested_positive_14d`: positivity rate among those screened for COVID-19 without symptoms or known exposure

```{r, eval=TRUE}

screening_data_link <- paste0(
  "https://api.covidcast.cmu.edu/epidata/covidcast/?data_source=fb-survey",
  "&signal=smoothed_wscreening_tested_positive_14d,smoothed_wtested_positive_14d,smoothed_wcli",
  "&geo_type=state&time_type=day&time_values=20210320-20221212&geo_value=*")

fb_screening <- httr::GET(screening_data_link)

fb_symptoms <-jsonlite::fromJSON(
      httr::content(fb_screening,
                    as = "text", 
                    encoding = "UTF-8"))$epidata %>%
   mutate(date = lubridate::ymd(time_value),
          week = lubridate::week(date),
          state = geo_value,
          value = value/100,
          stderr = stderr/100) %>% 
  filter(date <= lubridate::ymd("2022-03-01")) %>%
  as_tibble()

fb_symptoms %>%
  saveRDS(paste0(
    here::here(), 
    "/data/state_level/screeningpos_all_states.RDS"))

```

# Biobot Wastewater Data

```{r}

biobot_link <- "https://raw.githubusercontent.com/biobotanalytics/covid19-wastewater-data/master/wastewater_by_county.csv"

dates <- readRDS(here("data", "date_to_biweek.RDS"))

w_data <- read_csv(biobot_link)%>% 
  filter(sampling_week >= ymd("2021-03-01") &  sampling_week <= ymd("2022-03-01")) %>%
  mutate(fips = as.character(fipscode)) %>%
  select(-fipscode)  %>%
  left_join(dates, by = c("sampling_week" = "date")) %>%
  group_by(fips, biweek, state,name) %>%
  summarize(mean_conc = mean(effective_concentration_rolling_average, na.rm= TRUE)) %>%
  ungroup()

saveRDS(w_data,
        here("data/county_level", "biobot_wastewater_county.RDS"))

```

### 2023-06-14 R/base_functions.R

@@ -1,0 +1,319 @@
+
+
+
+
+###############################################################
+# BETA PARAMETERS FROM DESIRED MEAN AND VARIANCE
+###############################################################
+get_beta_params <- function(mu, sd){
+  var = sd ^ 2
+  a = ((1 - mu) / var - 1 / mu) * mu ^ 2
+  b = a * (1 / mu - 1)
+  return (list(a = a, b = b))
+}
+
+
+# vectorized version to get each parameter separately
+get_shape1 <- function(mu, sd) {
+  var = sd^2
+  alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
+  beta <- alpha * (1 / mu - 1)
+  return(alpha)
+}
+
+get_shape2 <- function(mu, sd) {
+  var = sd^2
+  alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
+  beta <- alpha * (1 / mu - 1)
+  return(beta)
+}
+
+
+
+
+###############################################################
+# BETA DENSITY WITH DESIRED MEAN AND VARIANCE
+###############################################################
+beta_density <- function(x, mean, sd, bounds=NA) {
+  shape_params <-  get_beta_params(
+    mu = mean,
+    sd = sd)
+  
+  if(!length(bounds) == 1){
+    # message("here")
+    dtrunc(x,
+           spec = "beta",
+           a = bounds[1],
+           b = bounds[2],
+           shape1 = shape_params$a,
+           shape2 = shape_params$b) %>%
+      return()
+  }else{
+    dbeta(x,
+          shape1 = shape_params$a,
+          shape2 = shape_params$b)  %>%
+      return()
+  }
+}
+
+
+
+
+###############################################################
+# SAMPLE FROM BETA DENSITY WITH DESIRED MEAN AND VARIANCE
+###############################################################
+
+sample_beta_density <- function(n, mean, sd, bounds = NA) {
+  
+  shape_params <-  get_beta_params(
+    mu = mean,
+    sd = sd)
+  
+  rbeta(n,
+        shape1 = shape_params$a,
+        shape2 = shape_params$b)
+  
+  if(!length(bounds) == 1){
+    # message("here")
+    rtrunc(n,
+           spec = "beta",
+           a = bounds[1],
+           b = bounds[2],
+           shape1 = shape_params$a,
+           shape2 = shape_params$b) %>%
+      return()
+  }else{
+    rbeta(n,
+          shape1 = shape_params$a,
[…]

### 2023-06-14 R/data_cleaning.R

@@ -1,0 +1,147 @@
+
+library(here)
+library(tidyverse)
+library(httr)
+
+#------------------ State Level Cleaning -----------------
+
+data_path <- here("data_clean")
+
+#------Census Population Estimates -----
+
+url_2019 <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv"
+
+population_2019 <- read_csv(url_2019) %>%
+  mutate(fips_code = paste0(STATE, COUNTY)) %>%
+  select(fips_code, population =POPESTIMATE2019)
+
+
+
+
+#------- Covidestim (State Level) -----
+
+# Due to changes in the model to handle the Omicron wave, 
+# we have to obtain dates before 2021-12-02 from a different
+# source than the latest runs endpoint of the API.
+
+
+get_covidestim_biweekly <- function() {
+  
+  # set end date
+  end_date <- ymd("2022-02-25")
+  
+  
+  covidestim_api_link <- "https://api2.covidestim.org/latest_runs?geo_type=eq.state&select=*,timeseries(*)"
+  covidestim <- GET(covidestim_api_link)
+  
+  
+  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+  # access dates BEFORE 2021-12-02
+  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+  last_weeks <-jsonlite::fromJSON(
+    httr::content(covidestim,
+                  as = "text",
+                  encoding = "UTF-8"),
+    simplifyVector = TRUE,
+    flatten = TRUE)  %>%
+    unnest(cols=timeseries, 
+           names_repair = "unique") %>%
+    mutate( date = ymd(date),
+            week = week( date)) %>%
+    filter(date <= end_date & year(date) > 2020)
+  
+  last_weeks <- last_weeks %>%
+    select(date, 
+           infections.lo = infections_p2_5,
+           infections.hi =infections_p97_5,
+           infections, created_at,
+           state = geo_name) %>%
+    mutate(week = week(date),
+           created = substr(created_at, 1, 10),
+           created = ymd(created))
+  
+  # remove duplicates from multiple model runs by taking most recent
+  last_weeks <- last_weeks %>%
[…]

### 2023-06-14 R/get_melded.R

@@ -1,0 +1,285 @@
+
+
+get_melded <- function(alpha_mean = 0.9,
+                       alpha_sd = 0.04,
+                       alpha_bounds = NA,
+                       beta_mean = .15,
+                       beta_sd =.09,
+                       beta_bounds = NA,
+                       s_untested_mean = .025,
+                       s_untested_sd = .0225,
+                       s_untested_bounds = NA,
+                       p_s0_pos_mean = .4,
+                       p_s0_pos_sd = .1225,
+                       p_s0_pos_bounds = NA,
+                       pre_nsamp = 1e4,
+                       post_nsamp = 1e3,
+                       include_corrected = TRUE,
+                       bde = FALSE) {
+  
+
+  given_args <- as.list(environment())
+  cat("Arguments to get_melded:\n")
+  print(given_args)
+  
+  
+  theta <- tibble(alpha = sample_gamma_density(pre_nsamp,
+                                               mean = alpha_mean,
+                                               sd = alpha_sd,
+                                               bounds = alpha_bounds),
+                  beta= sample_beta_density(pre_nsamp,
+                                            mean = beta_mean,
+                                            sd = beta_sd,
+                                            bounds = beta_bounds),
+                  P_S_untested = sample_beta_density(pre_nsamp,
+                                                     mean = s_untested_mean,
+                                                     sd = s_untested_sd,
+                                                     bounds = s_untested_bounds)) %>%
+    mutate(phi_induced = est_P_A_testpos(P_S_untested = P_S_untested,
+                                         alpha = alpha,
+                                         beta=beta))
+  
+  # theta contains values sampled from alpha, beta, P_S_untested, and M(theta) = phi_induced
+  # induced phi
+  phi <- theta$phi_induced
+
+  # approximate induced distribution via a density approximation
+  if(bde) {
+    message("Beta Kernel Density Estimation")
+    phi_induced_density <- kdensity::kdensity(x = phi,  bw="SJ", adjust = 1, start = "beta", kernel = "beta", support = c(0,1))
[…]

### 2023-06-14 _targets.R
start targets workflow

@@ -9,7 +9,7 @@ library(targets)
 
 # Set target options:
 tar_option_set(
-  packages = c("tibble"), # packages that your targets need to run
+  packages = c("tidyverse", "lubridate", "here", "httr"), # packages that your targets need to run
   format = "rds" # default storage format
   # Set other options as needed.
 )
@@ -21,18 +21,16 @@ options(clustermq.scheduler = "multicore")
 # Install packages {{future}}, {{future.callr}}, and {{future.batchtools}} to allow use_targets() to configure tar_make_future() options.
 
 # Run the R scripts in the R/ folder with your custom functions:
-tar_source()
+#tar_source("R/priors.R")
+tar_source("R/data_cleaning.R")
+
 # source("other_functions.R") # Source other scripts as needed. # nolint
 
 # Replace the target list below with your own:
 list(
   tar_target(
-    name = data,
-    command = tibble(x = rnorm(100), y = rnorm(100))
-#   format = "feather" # efficient storage of large data frames # nolint
-  ),
-  tar_target(
-    name = model,
-    command = coefficients(lm(y ~ x, data = data))
-  )
+    name = covidestim_biweekly_state,
+    command = get_covidestim_biweekly(),
+    format = "rds")
+  
 )

### 2023-06-15 R/data_cleaning.R
archive linked data

@@ -25,19 +25,16 @@ population_2019 <- read_csv(url_2019) %>%
 # source than the latest runs endpoint of the API.
 
 
-get_covidestim_biweekly <- function() {
+get_covidestim_state_biweekly <- function(end_date = "2022-02-25") {
   
-  # set end date
-  end_date <- ymd("2022-02-25")
-  
+  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+  # access dates AFTER 2021-12-02
+  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
   
   covidestim_api_link <- "https://api2.covidestim.org/latest_runs?geo_type=eq.state&select=*,timeseries(*)"
   covidestim <- GET(covidestim_api_link)
   
   
-  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-  # access dates BEFORE 2021-12-02
-  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
   last_weeks <-jsonlite::fromJSON(
     httr::content(covidestim,
                   as = "text",
@@ -73,7 +70,7 @@ get_covidestim_biweekly <- function() {
     ungroup()
   
   #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-  # access dates AFTER 2021-12-02
+  # access dates BEFORE 2021-12-02
   #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
   covidestim_link <- "https://covidestim.s3.us-east-2.amazonaws.com/latest/state/estimates.csv"
   
@@ -137,11 +134,101 @@ get_covidestim_biweekly <- function() {
   
 }
 
-# covidestim <- get_covidestim_biweekly()
-# saveRDS(here("data/data_clean/covidestim_biweekly_all_states.RDS"))
 
 
 
 
 
 
+
+
+
+#------- Covidestim (County Level) -----
+
+get_covidestim_county_biweekly <- function(end_date = "2022-02-25") {
+  legacy_link <- "https://covidestim.s3.us-east-2.amazonaws.com/latest/estimates.csv"
+  
+  covidestim_county <- read_csv(legacy_link) %>%
[…]

### 2023-06-15 _targets.R
targets

@@ -28,9 +28,15 @@ tar_source("R/data_cleaning.R")
 
 # Replace the target list below with your own:
 list(
+  #---------------- data -------------------
+  # state-level covidestim
   tar_target(
     name = covidestim_biweekly_state,
-    command = get_covidestim_biweekly(),
+    command = get_covidestim_state_biweekly(),
+    format = "rds"),
+  # county-level covidestim
+  tar_target(
+    name = covidestim_biweekly_county,
+    command = get_covidestim_county_biweekly(),
     format = "rds")
-  
 )

### 2023-06-15 R/data_cleaning.R
clean state level testing

@@ -232,3 +232,83 @@ archive_data <- function() {
 }
 
 # archive_data()
+
+
+#----------- State-Level Testing -------
+
+get_state_testing <- function() {
+
+    
+  dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
+  
+  
+  # PAGE: https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/state/detail/
+  state_population_link <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/state/detail/SCPRC-EST2019-18+POP-RES.csv"
+  state_pop <- read_csv(state_population_link)
+  
+  
+  
+  statecodes <- read_csv(paste0(
+    here::here(), 
+    "/data/demographic/statecodes.csv"))
+  
+  state_pop <- state_pop %>%
+    left_join(statecodes, by = c("NAME" = "state")) %>%
+    select(population = POPESTIMATE2019,
+           state = code) %>%
+    filter(!is.na(state))
+  
+  
+  
+  # split into two queries to ensure we obtain all data with the API limit
+  
+  dat <- httr::GET(URLencode(
+    paste0("https://healthdata.gov/resource/j8mb-icvb.json?",
+           "$where=date between '2020-12-30T12:00:00' and '2021-10-15T12:00:00'&$limit=50000")))
+  
+  cdc1 <-jsonlite::fromJSON(
+    httr::content(dat,
+                  as = "text", 
+                  encoding = "UTF-8")) %>%
+    as_tibble()
+  
+  
+  dat2 <- httr::GET(URLencode(
+    paste0("https://healthdata.gov/resource/j8mb-icvb.json?",
+           "$where=date between '2021-10-15T14:00:00' and '2022-02-25T14:00:00'&$limit=50000")))
+  
+  
+  cdc2 <-jsonlite::fromJSON(
+    httr::content(dat2,
+                  as = "text", 
+                  encoding = "UTF-8")) %>%
+    as_tibble()
+  
+  
+  cdc <- bind_rows(cdc1, cdc2) %>%
+    mutate(date = ymd(substr(date,1,10))) %>%
+    mutate(across(c(new_results_reported), as.numeric)) %>%
+    filter(!state %in% c("MP", "AS", "GU", "PR", "VI", "MH"))
+  
+  # overall_outcome is the outcome of the test (Inconclusive, Negative, or Positive)
+  # new_results_reported is the number with the given outcome
[…]

### 2023-06-15 _targets.R

@@ -38,5 +38,9 @@ list(
   tar_target(
     name = covidestim_biweekly_county,
     command = get_covidestim_county_biweekly(),
+    format = "rds"),
+  tar_target(
+    name = tests_biweekly_state,
+    command = get_state_testing(),
     format = "rds")
 )

### 2023-06-15 R/data_cleaning.R
update cleaning, don't keep week column

@@ -128,7 +128,8 @@ get_covidestim_state_biweekly <- function(end_date = "2022-02-25") {
   
   covidestim_biweekly <- covidestim_biweekly %>%
     rename(state_name = state) %>%
-    left_join(statecodes)
+    left_join(statecodes) %>%
+    select(-week)
   
   return(covidestim_biweekly)
   
@@ -183,7 +184,11 @@ get_covidestim_county_biweekly <- function(end_date = "2022-02-25") {
     left_join(biweek_to_week) %>%
     group_by(biweek, fips)  %>%
     mutate(across(contains("infections"), sum)) %>%
-    ungroup() 
+    ungroup() %>%
+    select(-week)
+  
+  return(covidestim_biweekly_all_counties)
+  
 }
 
 
@@ -300,7 +305,7 @@ get_state_testing <- function() {
     rename_with(tolower) %>%
     select(state, positive, total, date)
   
-  cdc_pos %>%
+  cdc_pos <- cdc_pos %>%
     left_join(dates) %>%
     filter(!is.na(biweek)) %>%
     group_by(biweek, state) %>%
@@ -309,6 +314,9 @@ get_state_testing <- function() {
     ungroup() %>%
     mutate(posrate = positive/total) %>%
     left_join(state_pop) 
+  
+  
+  return(cdc_pos)
 
 
 } 

### 2023-06-15 _targets.R

@@ -39,6 +39,7 @@ list(
     name = covidestim_biweekly_county,
     command = get_covidestim_county_biweekly(),
     format = "rds"),
+  # state-level positive tests
   tar_target(
     name = tests_biweekly_state,
     command = get_state_testing(),

### 2023-06-15 R/data_cleaning.R
archive michigan data

@@ -5,15 +5,15 @@ library(httr)
 
 #------------------ State Level Cleaning -----------------
 
-data_path <- here("data_clean")
+# data_path <- here("data/data_clean")
 
 #------Census Population Estimates -----
 
-url_2019 <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv"
-
-population_2019 <- read_csv(url_2019) %>%
-  mutate(fips_code = paste0(STATE, COUNTY)) %>%
-  select(fips_code, population =POPESTIMATE2019)
+# url_2019 <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv"
+# 
+# population_2019 <- read_csv(url_2019) %>%
+#   mutate(fips_code = paste0(STATE, COUNTY)) %>%
+#   select(fips_code, population =POPESTIMATE2019)
 
 
 
@@ -320,3 +320,6 @@ get_state_testing <- function() {
 
 
 } 
+
+
+

### 2023-06-15 R/data_cleaning.R
michigan county target

@@ -9,11 +9,16 @@ library(httr)
 
 #------Census Population Estimates -----
 
-# url_2019 <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv"
-# 
-# population_2019 <- read_csv(url_2019) %>%
-#   mutate(fips_code = paste0(STATE, COUNTY)) %>%
-#   select(fips_code, population =POPESTIMATE2019)
+
+get_county_pop <- function() {
+  url_2019 <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv"
+  
+  population_2019 <- read_csv(url_2019) %>%
+    mutate(fips_code = paste0(STATE, COUNTY)) %>%
+    select(fips_code, population =POPESTIMATE2019)
+  
+  return(population_2019)
+}
 
 
 
@@ -195,9 +200,9 @@ get_covidestim_county_biweekly <- function(end_date = "2022-02-25") {
 
 
 
-#--------- Archive Data -----------
+#--------- Archive Covidestim Data -----------
 
-archive_data <- function() {
+archive_covidestim_data <- function() {
   
   # save county estimates
   legacy_link <- "https://covidestim.s3.us-east-2.amazonaws.com/latest/estimates.csv"
@@ -322,4 +327,92 @@ get_state_testing <- function() {
 } 
 
 
+#--------- Archive County Data -----------
+
+
+archive_county_data <- function() {
+  
+  link <- paste0("https://www.michigan.gov/coronavirus/-/media/Project/Websites/coronavirus/",
+                 "Michigan-Data/09-27-2022/Datasets/Diagnostic-Tests-by-Result-and-County-2022-09-27.xlsx?",
+                 "rev=7ba61151dcff4b038e33ea9ead95137c&hash=6C1AA97E93A30A3C1F10171233FDF31A")
+  httr::GET(link, httr::write_disk(tf <- tempfile(fileext = ".xlsx")))
+  df <- readxl::read_excel(tf)
+  saveRDS(df, file =here("data/data_raw/mi_county_original.RDS"))
+  
+}
+
+
+# archive_county_data()
+
+
+#--------- Michigan County Data -----------
+
+
+get_michigan_county <- function(end_date = "2022-02-25") {
+  
+  counts_raw <- readRDS(here(
+    "data/data_raw/mi_county_original.RDS"))
[…]

### 2023-06-15 _targets.R

@@ -43,5 +43,9 @@ list(
   tar_target(
     name = tests_biweekly_state,
     command = get_state_testing(),
+    format = "rds"),
+  tar_target(
+    name = mi_biweekly_county,
+    command = get_michigan_county(),
     format = "rds")
 )

### 2023-06-15 R/data_cleaning.R
ma county data cleaning

@@ -332,12 +332,28 @@ get_state_testing <- function() {
 
 archive_county_data <- function() {
   
+  # Michigan
   link <- paste0("https://www.michigan.gov/coronavirus/-/media/Project/Websites/coronavirus/",
                  "Michigan-Data/09-27-2022/Datasets/Diagnostic-Tests-by-Result-and-County-2022-09-27.xlsx?",
                  "rev=7ba61151dcff4b038e33ea9ead95137c&hash=6C1AA97E93A30A3C1F10171233FDF31A")
   httr::GET(link, httr::write_disk(tf <- tempfile(fileext = ".xlsx")))
   df <- readxl::read_excel(tf)
   saveRDS(df, file =here("data/data_raw/mi_county_original.RDS"))
+  
+  
+  # Massachusetts
+  mass_link <- "https://www.mass.gov/doc/covid-19-raw-data-april-12-2022/download"
+  httr::GET(mass_link, httr::write_disk(tf <- tempfile(fileext = ".xlsx")))
+  
+  df <- readxl::read_excel(tf, sheet = "County_Weekly")
+  df <- df %>% 
+    select(contains("date"), 
+           total = `Total Tests (Last 14 days)`,
+           positive = `Total Positive Tests (Last 14 days)`,
+           county_name =County) %>%
+    mutate( negative = total - positive,
+            state = "MA") 
+  saveRDS(df, file =here("data/data_raw/ma_county_original.RDS"))
   
 }
 
@@ -415,4 +431,105 @@ get_michigan_county <- function(end_date = "2022-02-25") {
   return(county_biweekly) 
   
 }
+
+#---------- Massachusetts County Data -----------
+
+
+get_mass_county <- function(end_date = "2022-02-25") {
+  
+  county <- readRDS(here("data/data_raw/ma_county_original.RDS"))
+  
+  county_fips <- read_tsv(here("data/demographic/county_fips.tsv")) %>%
+    dplyr::rename_with(.cols =everything(), tolower)
+  
+  population_2019 <- get_county_pop()
+  
+  # Dukes and Nantucket are combined
+  together <- county_fips %>% 
+    filter(name == "Dukes" | name == "Nantucket") %>%
+    pull(fips)
+  
+  county_fips <- county_fips %>%
+    filter(!(fips %in% together)) %>%
+    bind_rows(tibble(name = "Dukes and Nantucket", 
+                     fips = paste0(together, 
+                                   collapse=","),
[…]

### 2023-06-15 _targets.R

@@ -39,13 +39,19 @@ list(
     name = covidestim_biweekly_county,
     command = get_covidestim_county_biweekly(),
     format = "rds"),
-  # state-level positive tests
+  # state-level testing data
   tar_target(
     name = tests_biweekly_state,
     command = get_state_testing(),
     format = "rds"),
+  # county-level testing data
   tar_target(
     name = mi_biweekly_county,
     command = get_michigan_county(),
+    format = "rds"),
+  tar_target(
+    name = ma_biweekly_county,
+    command = get_mass_county(),
     format = "rds")
+
 )

### 2023-06-16 R/03-base-functions.R
rename

@@ -1,7 +1,4 @@
 
-
-
-
 ###############################################################
 # BETA PARAMETERS FROM DESIRED MEAN AND VARIANCE
 ###############################################################

### 2023-06-16 R/01-data_cleaning.R
rename

@@ -1,0 +1,535 @@
+
+library(here)
+library(tidyverse)
+library(httr)
+
+#------------------ State Level Cleaning -----------------
+
+# data_path <- here("data/data_clean")
+
+#------Census Population Estimates -----
+
+
+get_county_pop <- function() {
+  url_2019 <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv"
+  
+  population_2019 <- read_csv(url_2019) %>%
+    mutate(fips_code = paste0(STATE, COUNTY)) %>%
+    select(fips_code, population =POPESTIMATE2019)
+  
+  return(population_2019)
+}
+
+
+
+
+#------- Covidestim (State Level) -----
+
+# Due to changes in the model to handle the Omicron wave, 
+# we have to obtain dates before 2021-12-02 from a different
+# source than the latest runs endpoint of the API.
+
+
+get_covidestim_state_biweekly <- function(end_date = "2022-02-25") {
+  
+  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+  # access dates AFTER 2021-12-02
+  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
+  
+  covidestim_api_link <- "https://api2.covidestim.org/latest_runs?geo_type=eq.state&select=*,timeseries(*)"
+  covidestim <- GET(covidestim_api_link)
+  
+  
+  last_weeks <-jsonlite::fromJSON(
+    httr::content(covidestim,
+                  as = "text",
+                  encoding = "UTF-8"),
+    simplifyVector = TRUE,
+    flatten = TRUE)  %>%
+    unnest(cols=timeseries, 
+           names_repair = "unique") %>%
+    mutate( date = ymd(date),
+            week = week( date)) %>%
+    filter(date <= end_date & year(date) > 2020)
+  
+  last_weeks <- last_weeks %>%
+    select(date, 
+           infections.lo = infections_p2_5,
+           infections.hi =infections_p97_5,
+           infections, created_at,
+           state = geo_name) %>%
+    mutate(week = week(date),
+           created = substr(created_at, 1, 10),
+           created = ymd(created))
+  
+  # remove duplicates from multiple model runs by taking most recent
[…]

### 2023-06-16 R/02-priors.R

@@ -1,0 +1,25 @@
+
+get_priors <- function() { 
+  
+  list(
+    alpha_mean = .95,
+    alpha_sd = 0.08,
+    alpha_bounds = NA,
+    # alpha_bounds = c(.8,1),
+    beta_mean = .15,
+    beta_sd =.09,
+    beta_bounds = NA,
+    #  beta_bounds = c(0.002, 0.4),
+    s_untested_mean = .03,
+    s_untested_sd = .0225,
+    #  s_untested_bounds = c(0.0018, Inf),
+    s_untested_bounds = NA,
+    p_s0_pos_mean = .4,
+    p_s0_pos_sd = .1225,
+    p_s0_pos_bounds = NA,
+    #  p_s0_pos_bounds = c(.25, .7),
+    pre_nsamp = 1e6,
+    post_nsamp = 1e5)
+  
+}
+

### 2023-06-16 R/03-state-analysis.R

@@ -1,0 +1,24 @@
+
+
+get_v1_state <- function(data, params) {
+  
+  state_testing <- data
+  
+  
+  state_testing <- state_testing %>%
+    rename(fips = state) %>%
+    select(-date) %>% 
+    distinct()
+  
+  
+  # only use a few rows if testing
+  state_testing <- if(testing) state_testing %>% filter(biweek >=6) %>% slice_sample(n=15) else state_testing
+  
+  results_path <- here("analysis/results/adj_biweekly_state/")
+  
+  
+  melded <- do.call(get_melded, prior_params)
+  
+  
+  
+}

### 2023-06-16 R/04-melding.R

@@ -1,0 +1,107 @@
+
+get_melded <- function(alpha_mean = 0.9,
+                       alpha_sd = 0.04,
+                       alpha_bounds = NA,
+                       beta_mean = .15,
+                       beta_sd =.09,
+                       beta_bounds = NA,
+                       s_untested_mean = .025,
+                       s_untested_sd = .0225,
+                       s_untested_bounds = NA,
+                       p_s0_pos_mean = .4,
+                       p_s0_pos_sd = .1225,
+                       p_s0_pos_bounds = NA,
+                       pre_nsamp = 1e4,
+                       post_nsamp = 1e3,
+                       include_corrected = TRUE,
+                       bde = FALSE) {
+  
+  
+  given_args <- as.list(environment())
+  cat("Arguments to get_melded:\n")
+  print(given_args)
+  
+  
+  theta <- tibble(alpha = sample_gamma_density(pre_nsamp,
+                                               mean = alpha_mean,
+                                               sd = alpha_sd,
+                                               bounds = alpha_bounds),
+                  beta= sample_beta_density(pre_nsamp,
+                                            mean = beta_mean,
+                                            sd = beta_sd,
+                                            bounds = beta_bounds),
+                  P_S_untested = sample_beta_density(pre_nsamp,
+                                                     mean = s_untested_mean,
+                                                     sd = s_untested_sd,
+                                                     bounds = s_untested_bounds)) %>%
+    mutate(phi_induced = est_P_A_testpos(P_S_untested = P_S_untested,
+                                         alpha = alpha,
+                                         beta=beta))
+  
+  # theta contains values sampled from alpha, beta, P_S_untested, and M(theta) = phi_induced
+  # induced phi
+  phi <- theta$phi_induced
+  
+  # approximate induced distribution via a density approximation
+  if(bde) {
+    message("Beta Kernel Density Estimation")
+    phi_induced_density <- kdensity::kdensity(x = phi,  bw="SJ", adjust = 1, start = "beta", kernel = "beta", support = c(0,1))
+    phi_sampled_density <- phi_induced_density(phi) 
+  }
+  else {
+    phi_induced_density <- density(x = phi, n = pre_nsamp, adjust = 2, kernel = "gaussian")
[…]

### 2023-06-16 R/data_cleaning.R
rename

@@ -1,535 +1,0 @@
-
-library(here)
-library(tidyverse)
-library(httr)
-
-#------------------ State Level Cleaning -----------------
-
-# data_path <- here("data/data_clean")
-
-#------Census Population Estimates -----
-
-
-get_county_pop <- function() {
-  url_2019 <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/counties/totals/co-est2019-alldata.csv"
-  
-  population_2019 <- read_csv(url_2019) %>%
-    mutate(fips_code = paste0(STATE, COUNTY)) %>%
-    select(fips_code, population =POPESTIMATE2019)
-  
-  return(population_2019)
-}
-
-
-
-
-#------- Covidestim (State Level) -----
-
-# Due to changes in the model to handle the Omicron wave, 
-# we have to obtain dates before 2021-12-02 from a different
-# source than the latest runs endpoint of the API.
-
-
-get_covidestim_state_biweekly <- function(end_date = "2022-02-25") {
-  
-  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-  # access dates AFTER 2021-12-02
-  #~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~~
-  
-  covidestim_api_link <- "https://api2.covidestim.org/latest_runs?geo_type=eq.state&select=*,timeseries(*)"
-  covidestim <- GET(covidestim_api_link)
-  
-  
-  last_weeks <-jsonlite::fromJSON(
-    httr::content(covidestim,
-                  as = "text",
-                  encoding = "UTF-8"),
-    simplifyVector = TRUE,
-    flatten = TRUE)  %>%
-    unnest(cols=timeseries, 
-           names_repair = "unique") %>%
-    mutate( date = ymd(date),
-            week = week( date)) %>%
-    filter(date <= end_date & year(date) > 2020)
-  
-  last_weeks <- last_weeks %>%
-    select(date, 
-           infections.lo = infections_p2_5,
-           infections.hi =infections_p97_5,
-           infections, created_at,
-           state = geo_name) %>%
-    mutate(week = week(date),
-           created = substr(created_at, 1, 10),
-           created = ymd(created))
-  
-  # remove duplicates from multiple model runs by taking most recent
[…]

### 2023-06-16 _targets.R
priors target

@@ -17,12 +17,11 @@ tar_option_set(
 # tar_make_clustermq() configuration (okay to leave alone):
 options(clustermq.scheduler = "multicore")
 
-# tar_make_future() configuration (okay to leave alone):
-# Install packages {{future}}, {{future.callr}}, and {{future.batchtools}} to allow use_targets() to configure tar_make_future() options.
-
 # Run the R scripts in the R/ folder with your custom functions:
 #tar_source("R/priors.R")
-tar_source("R/data_cleaning.R")
+tar_source("R/01-data_cleaning.R")
+tar_source("R/02-priors.R")
+
 
 # source("other_functions.R") # Source other scripts as needed. # nolint
 
@@ -32,26 +31,32 @@ list(
   # state-level covidestim
   tar_target(
     name = covidestim_biweekly_state,
-    command = get_covidestim_state_biweekly(),
-    format = "rds"),
+    command = get_covidestim_state_biweekly()),
   # county-level covidestim
   tar_target(
     name = covidestim_biweekly_county,
-    command = get_covidestim_county_biweekly(),
-    format = "rds"),
+    command = get_covidestim_county_biweekly()),
   # state-level testing data
   tar_target(
     name = tests_biweekly_state,
-    command = get_state_testing(),
-    format = "rds"),
+    command = get_state_testing()),
   # county-level testing data
   tar_target(
     name = mi_biweekly_county,
-    command = get_michigan_county(),
-    format = "rds"),
+    command = get_michigan_county()),
   tar_target(
     name = ma_biweekly_county,
-    command = get_mass_county(),
-    format = "rds")
+    command = get_mass_county()),
+  #---------------- analysis -------------------
+  # set priors
+  tar_target(
+    name = prior_params,
+    command= get_priors()
+  )
+  
+  
+  
+  
+  
 
 )

### 2023-06-16 R/01-data_cleaning.R
state level analysis, version 1

@@ -308,16 +308,18 @@ get_state_testing <- function() {
                 values_from = c("new_results_reported")) %>%
     mutate(total = Inconclusive + Negative + Positive) %>%
     rename_with(tolower) %>%
-    select(state, positive, total, date)
+    select(state, positive, total, negative, date)
   
   cdc_pos <- cdc_pos %>%
     left_join(dates) %>%
     filter(!is.na(biweek)) %>%
     group_by(biweek, state) %>%
     mutate(positive = sum(positive, na.rm=TRUE),
-           total = sum(total, na.rm = TRUE)) %>%
+           total = sum(total, na.rm = TRUE),
+           negative=sum(negative)) %>%
     ungroup() %>%
-    mutate(posrate = positive/total) %>%
+    mutate(posrate = positive/total,
+           negative = total) %>%
     left_join(state_pop) 
   
   

### 2023-06-16 R/03-base-functions.R

@@ -193,8 +193,7 @@ process_priors_per_county <-  function(priors, county_df, nsamp){
       empirical_testpos = county_df$posrate,
       population = county_df$population,
       total = county_df$total,
-      positive = county_df$positive,
-      negative = county_df$negative) %>%
+      positive = county_df$positive) %>%
     mutate(Se = dist_Se,
            Sp = dist_Sp,
            biweek = county_df$biweek,
@@ -306,7 +305,6 @@ summarize_corrected_sample <- function(priors_by_county_df_exp_cases) {
     fips = unique(priors_by_county_df_exp_cases$fips),
     empirical_testpos = unique(priors_by_county_df_exp_cases$empirical_testpos),
     population = unique(priors_by_county_df_exp_cases$population),
-    negative = unique(priors_by_county_df_exp_cases$negative),
     positive = unique(priors_by_county_df_exp_cases$positive),
     total = unique(priors_by_county_df_exp_cases$total))
   

### 2023-06-16 R/03-state-analysis.R

@@ -1,24 +1,0 @@
-
-
-get_v1_state <- function(data, params) {
-  
-  state_testing <- data
-  
-  
-  state_testing <- state_testing %>%
-    rename(fips = state) %>%
-    select(-date) %>% 
-    distinct()
-  
-  
-  # only use a few rows if testing
-  state_testing <- if(testing) state_testing %>% filter(biweek >=6) %>% slice_sample(n=15) else state_testing
-  
-  results_path <- here("analysis/results/adj_biweekly_state/")
-  
-  
-  melded <- do.call(get_melded, prior_params)
-  
-  
-  
-}

### 2023-06-16 R/04-melding.R

@@ -1,4 +1,6 @@
 
+#--------- Melding -------------
+
 get_melded <- function(alpha_mean = 0.9,
                        alpha_sd = 0.04,
                        alpha_bounds = NA,
@@ -94,14 +96,204 @@ get_corrected_counts <- function(county_df,
   #   rename(Z_S = alpha,
   #          Z_A = beta)
   
-  corrected <- pmap_df(county_df, ~ {
+  corrected <- pmap_df(county_df, function(posrate,
+                                           population,
+                                           total, 
+                                           positive,
+                                           biweek,
+                                           fips,
+                                           ...) {
     process_priors_per_county(
       priors = melded_df,
-      county_df = list(...),
+      county_df = tibble(posrate, 
+                         population,
+                         total, 
+                         positive,
+                         biweek,
+                         fips),
       nsamp = post_nsamp) %>%
-      generate_corrected_sample(., num_reps = num_reps_counts) %>%
+      generate_corrected_sample(., num_reps = 1e3) %>%
       summarize_corrected_sample() })
   
   corrected %>%
     left_join(dates)
+}
+
+
+#--------- Plotting -------------
+
+
+
+#' reformat for plot generation
+reformat_melded <- function(melded_df,
+                            theta_df,
+                            pre_nsamp,
+                            p_s0_pos_mean,
+                            p_s0_pos_sd,
+                            p_s0_pos_bounds) {
+  
+  melded_df_long <- melded_df %>%
+    pivot_longer(cols=everything()) %>%
+    mutate(type = "After Melding")
+  
+  
+  melded <- theta_df %>%
+    mutate(P_A_testpos = sample_beta_density(pre_nsamp,
+                                             mean = p_s0_pos_mean,
+                                             sd = p_s0_pos_sd,
+                                             bounds = p_s0_pos_bounds)) %>%
+    pivot_longer(cols=everything()) %>%
+    mutate(type = ifelse(
+      name == "phi_induced",
+      "Induced", "Before Melding")) %>%
+    mutate(name = ifelse(name == "phi_induced",
+                         "P_A_testpos",
+                         name)) %>%
[…]

### 2023-06-16 R/05-state-analysis.R

@@ -1,0 +1,43 @@
+
+
+get_v1_state <- function(data, params, testing = TRUE) {
+  
+  state_testing <- data
+  
+  state_testing <- state_testing %>%
+    rename(fips = state) %>%
+    select(-date) %>% 
+    distinct()
+  
+  
+  # only use a few rows if testing
+  state_testing <- if(testing) state_testing %>% 
+    filter(biweek >=6) %>% slice_sample(n=15) else state_testing
+  
+  results_path <- here("results/adj_biweekly_state/")
+  
+  
+  melded <- do.call(get_melded, params)
+  
+  corrected <- pmap_df(state_testing, 
+                       function(posrate,
+                                population,
+                                total, 
+                                positive,
+                                biweek,
+                                fips,
+                                ...) {
+    process_priors_per_county(
+      priors = melded$post_melding,
+      county_df = tibble(posrate, 
+                         population,
+                         total, 
+                         positive,
+                         biweek,
+                         fips),
+      nsamp = params$post_nsamp) %>%
+      generate_corrected_sample(., num_reps = 1e3) %>%
+      summarize_corrected_sample() })
+  
+  return(corrected)
+}

### 2023-06-16 R/get_melded.R

@@ -1,285 +1,0 @@
-
-
-get_melded <- function(alpha_mean = 0.9,
-                       alpha_sd = 0.04,
-                       alpha_bounds = NA,
-                       beta_mean = .15,
-                       beta_sd =.09,
-                       beta_bounds = NA,
-                       s_untested_mean = .025,
-                       s_untested_sd = .0225,
-                       s_untested_bounds = NA,
-                       p_s0_pos_mean = .4,
-                       p_s0_pos_sd = .1225,
-                       p_s0_pos_bounds = NA,
-                       pre_nsamp = 1e4,
-                       post_nsamp = 1e3,
-                       include_corrected = TRUE,
-                       bde = FALSE) {
-  
-
-  given_args <- as.list(environment())
-  cat("Arguments to get_melded:\n")
-  print(given_args)
-  
-  
-  theta <- tibble(alpha = sample_gamma_density(pre_nsamp,
-                                               mean = alpha_mean,
-                                               sd = alpha_sd,
-                                               bounds = alpha_bounds),
-                  beta= sample_beta_density(pre_nsamp,
-                                            mean = beta_mean,
-                                            sd = beta_sd,
-                                            bounds = beta_bounds),
-                  P_S_untested = sample_beta_density(pre_nsamp,
-                                                     mean = s_untested_mean,
-                                                     sd = s_untested_sd,
-                                                     bounds = s_untested_bounds)) %>%
-    mutate(phi_induced = est_P_A_testpos(P_S_untested = P_S_untested,
-                                         alpha = alpha,
-                                         beta=beta))
-  
-  # theta contains values sampled from alpha, beta, P_S_untested, and M(theta) = phi_induced
-  # induced phi
-  phi <- theta$phi_induced
-
-  # approximate induced distribution via a density approximation
-  if(bde) {
-    message("Beta Kernel Density Estimation")
-    phi_induced_density <- kdensity::kdensity(x = phi,  bw="SJ", adjust = 1, start = "beta", kernel = "beta", support = c(0,1))
[…]

### 2023-06-16 _targets.R

@@ -21,6 +21,11 @@ options(clustermq.scheduler = "multicore")
 #tar_source("R/priors.R")
 tar_source("R/01-data_cleaning.R")
 tar_source("R/02-priors.R")
+tar_source("R/03-base-functions.R")
+tar_source("R/04-melding.R")
+tar_source("R/05-state-analysis.R")
+
+
 
 
 # source("other_functions.R") # Source other scripts as needed. # nolint
@@ -47,13 +52,19 @@ list(
   tar_target(
     name = ma_biweekly_county,
     command = get_mass_county()),
+  
   #---------------- analysis -------------------
   # set priors
   tar_target(
     name = prior_params,
     command= get_priors()
+  ),
+  tar_target(
+    name = state_v1,
+    command = get_v1_state(
+        data = tests_biweekly_state,
+        params = prior_params)
   )
-  
   
   
   

### 2023-06-16 R/01-data_cleaning.R
save CTIS survey data

@@ -526,8 +526,37 @@ get_mass_county <- function(end_date = "2022-02-25") {
 
 
 
+#---------- COVID-19 Trends and Impact Survey Data --------------
 
+save_survey_data <- function() {
+  
+  screening_data_link <- paste0(
+    "https://api.covidcast.cmu.edu/epidata/covidcast/?data_source=fb-survey",
+    "&signal=smoothed_wscreening_tested_positive_14d,smoothed_wtested_positive_14d,smoothed_wcli",
+    "&geo_type=state&time_type=day&time_values=20210320-20221212&geo_value=*&api_key=TEMP-API-KEY-EXPIRES-2023-06-28")
+  
+  fb_screening <- httr::GET(screening_data_link)
+  
+  fb_symptoms <-jsonlite::fromJSON(
+    httr::content(fb_screening,
+                  as = "text", 
+                  encoding = "UTF-8"))$epidata %>%
+    mutate(date = lubridate::ymd(time_value),
+           week = lubridate::week(date),
+           state = geo_value,
+           value = value/100,
+           stderr = stderr/100) %>% 
+    filter(date <= lubridate::ymd("2022-03-01")) %>%
+    as_tibble()
+  
+  
+  fb_symptoms %>%
+    saveRDS(here(
+      "data/data_raw/ctis_all_states.RDS"))
+  
+}
 
+save_survey_data()
 
 
 

### 2023-06-16 R/04-melding.R
add ctis smoothing

@@ -16,7 +16,8 @@ get_melded <- function(alpha_mean = 0.9,
                        pre_nsamp = 1e4,
                        post_nsamp = 1e3,
                        include_corrected = TRUE,
-                       bde = FALSE) {
+                       bde = FALSE,
+                       quiet=FALSE) {
   
   
   given_args <- as.list(environment())

### 2023-06-16 R/05-state-analysis.R

@@ -1,6 +1,8 @@
 
 
-get_v1_state <- function(data, params, testing = TRUE) {
+
+#---------- Version 1 -------------
+get_v1_state <- function(data, params, testing = FALSE) {
   
   state_testing <- data
   
@@ -13,9 +15,6 @@ get_v1_state <- function(data, params, testing = TRUE) {
   # only use a few rows if testing
   state_testing <- if(testing) state_testing %>% 
     filter(biweek >=6) %>% slice_sample(n=15) else state_testing
-  
-  results_path <- here("results/adj_biweekly_state/")
-  
   
   melded <- do.call(get_melded, params)
   
@@ -40,4 +39,81 @@ get_v1_state <- function(data, params, testing = TRUE) {
       summarize_corrected_sample() })
   
   return(corrected)
+  
 }
+
+
+#------------- Using CTIS Data ----------------
+
+
+get_smoothed_ctis <- function(ctis,
+                              smooth_beta_span = .33,
+                              smooth_s_untested_span =.2) {
+  
+  dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
+  
+  symp <- ctis %>% 
+    select(signal, date, value, state = geo_value) %>% 
+    pivot_wider(names_from = signal,
+                values_from = c(value)) %>%
+    mutate(beta_est = smoothed_wscreening_tested_positive_14d/smoothed_wtested_positive_14d)
+  
+  symp <- symp %>%
+    group_by(state) %>%
+    mutate(m = sum(!is.na(beta_est)),
+           n = n()) %>%
+    mutate(prop_with_data = m/n) %>%
+    # keep states with less than 40% of observations missing
+    mutate(keep = ifelse(prop_with_data > .6, TRUE, FALSE)) %>%
+    ungroup() 
+  
+  # impute missing values 
+  symp <- symp %>%
+    group_by(state) %>%
+    arrange(date) %>%
+    mutate(imputed_beta = ifelse(
+      keep,
+      imputeTS::na_ma(beta_est, k = 10, weighting = "simple"),
+      beta_est),
+      imputed_s_untested  =ifelse(
+        keep,
+        imputeTS::na_ma(smoothed_wcli, k = 10, weighting = "simple"),
+        smoothed_wcli ))
+  
+  
+  # loess smoothing
+  
+  
+  symp_smooth <- symp %>%
+    group_by(state) %>%
+    group_split() %>%
+    map_df(~{
+      if(unique(.x$keep)) {
[…]

### 2023-06-16 _targets.R

@@ -9,7 +9,9 @@ library(targets)
 
 # Set target options:
 tar_option_set(
-  packages = c("tidyverse", "lubridate", "here", "httr"), # packages that your targets need to run
+  # packages needed for targets
+  packages = c("tidyverse", "lubridate", 
+               "here", "httr", "imputeTS"), 
   format = "rds" # default storage format
   # Set other options as needed.
 )
@@ -59,15 +61,19 @@ list(
     name = prior_params,
     command= get_priors()
   ),
+  # tar_target(
+  #   name = state_v1,
+  #   command = get_v1_state(
+  #       data = tests_biweekly_state,
+  #       params = prior_params)
+  # ),
+  # get smoothed CTIS survey data
   tar_target(
-    name = state_v1,
-    command = get_v1_state(
-        data = tests_biweekly_state,
-        params = prior_params)
+    name = ctis_smoothed,
+    command = get_smoothed_ctis(
+      ctis = readRDS(here('data/data_raw/ctis_all_states.RDS')),
+      smooth_beta_span = .33,
+      smooth_s_untested_span =.2)
   )
-  
-  
-  
-  
 
 )

### 2023-06-16 R/01-data_cleaning.R
state-level analysis version 2

@@ -556,7 +556,7 @@ save_survey_data <- function() {
   
 }
 
-save_survey_data()
+# save_survey_data()
 
 
 

### 2023-06-16 R/05-state-analysis.R

@@ -2,7 +2,7 @@
 
 
 #---------- Version 1 -------------
-get_v1_state <- function(data, params, testing = FALSE) {
+get_v1_state <- function(data, params, testing = TRUE) {
   
   state_testing <- data
   
@@ -46,13 +46,13 @@ get_v1_state <- function(data, params, testing = FALSE) {
 #------------- Using CTIS Data ----------------
 
 
-get_smoothed_ctis <- function(ctis,
+get_smoothed_ctis <- function(ctis_raw,
                               smooth_beta_span = .33,
                               smooth_s_untested_span =.2) {
   
   dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
   
-  symp <- ctis %>% 
+  symp <- ctis_raw %>% 
     select(signal, date, value, state = geo_value) %>% 
     pivot_wider(names_from = signal,
                 values_from = c(value)) %>%
@@ -67,47 +67,64 @@ get_smoothed_ctis <- function(ctis,
     mutate(keep = ifelse(prop_with_data > .6, TRUE, FALSE)) %>%
     ungroup() 
   
-  # impute missing values 
-  symp <- symp %>%
+  # impute missing values
+  
+  
+  # make sure each state has all dates
+  symp_all <- symp %>% 
+    select(date,state,beta_est) %>% 
+    pivot_wider(names_from = date, 
+                values_from =beta_est) %>% 
+    pivot_longer(cols =2:ncol(.), 
+                 values_to = "beta_est",
+                 names_to = "date") %>%
+    mutate(date=as_date(date))
+  
+  symp <- symp_all %>%
+    left_join(symp)
+  
+  symp_imputed <- symp %>%
     group_by(state) %>%
     arrange(date) %>%
-    mutate(imputed_beta = ifelse(
-      keep,
-      imputeTS::na_ma(beta_est, k = 10, weighting = "simple"),
-      beta_est),
-      imputed_s_untested  =ifelse(
-        keep,
-        imputeTS::na_ma(smoothed_wcli, k = 10, weighting = "simple"),
-        smoothed_wcli ))
+    mutate(imputed_beta = ifelse(keep,
+                                 imputeTS::na_ma(beta_est,
+                                          k = 20, 
+                                          weighting = "simple"),
+                                 beta_est),
+           imputed_s_untested  = imputeTS::na_ma(smoothed_wcli,
+                                                 k = 20, 
[…]

### 2023-06-16 _targets.R

@@ -55,25 +55,35 @@ list(
     name = ma_biweekly_county,
     command = get_mass_county()),
   
-  #---------------- analysis -------------------
+  #---------------- state-level analysis -------------------
   # set priors
   tar_target(
     name = prior_params,
     command= get_priors()
   ),
-  # tar_target(
-  #   name = state_v1,
-  #   command = get_v1_state(
-  #       data = tests_biweekly_state,
-  #       params = prior_params)
-  # ),
+  # version 1
+  tar_target(
+    name = state_v1,
+    command = get_v1_state(
+        data = tests_biweekly_state,
+        params = prior_params)
+  ),
   # get smoothed CTIS survey data
   tar_target(
     name = ctis_smoothed,
     command = get_smoothed_ctis(
-      ctis = readRDS(here('data/data_raw/ctis_all_states.RDS')),
+      ctis_raw = readRDS(here('data/data_raw/ctis_all_states.RDS')),
       smooth_beta_span = .33,
       smooth_s_untested_span =.2)
+  ),
+  # version 2
+  tar_target(
+    name = state_v2,
+    command = get_v2_state(
+      data = tests_biweekly_state,
+      params = prior_params,
+      ctis=ctis_smoothed
+    )
   )
 
 )

### 2023-06-16 R/04-melding.R
update v2 state level

@@ -17,12 +17,15 @@ get_melded <- function(alpha_mean = 0.9,
                        post_nsamp = 1e3,
                        include_corrected = TRUE,
                        bde = FALSE,
-                       quiet=FALSE) {
+                       quiet=TRUE) {
   
   
   given_args <- as.list(environment())
-  cat("Arguments to get_melded:\n")
-  print(given_args)
+  
+  if (!quiet) {
+    cat("Arguments to get_melded:\n")
+    print(given_args)
+  } 
   
   
   theta <- tibble(alpha = sample_gamma_density(pre_nsamp,

### 2023-06-16 R/05-state-analysis.R

@@ -104,6 +104,7 @@ get_smoothed_ctis <- function(ctis_raw,
     mutate(index = row_number()) %>%
     group_split() %>%
     map_df(~ {
+      # message(paste0(unique(.x$state), unique(.x$keep)))
       smoothed_s_untested <- loess(imputed_s_untested~index, 
                                    data = .x,
                                    span = smooth_s_untested_span)
@@ -134,7 +135,9 @@ get_smoothed_ctis <- function(ctis_raw,
 
 #---------- Version 2 -------------
 
-get_v2_state <- function(data, params, ctis, testing = TRUE) {
+get_v2_state <- function(data, params, ctis, vary, testing = TRUE) {
+  
+  dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
   
   state_testing <- data
   
@@ -148,9 +151,6 @@ get_v2_state <- function(data, params, ctis, testing = TRUE) {
   state_testing <- if(testing) state_testing %>% 
     filter(biweek >=6) %>% slice_sample(n=15) else state_testing
   
-  # only keep states that have enough observations of beta
-  ctis <- ctis %>% 
-    filter(keep == TRUE) 
   
   ctis_biweekly <- ctis %>%
     left_join(dates) %>%
@@ -160,34 +160,61 @@ get_v2_state <- function(data, params, ctis, testing = TRUE) {
     mutate(state =toupper(state)) %>%
     select(fips = state,
            biweek, 
-           smoothed_s_untested, 
-           smoothed_beta) 
+           s_untested_smoothed, 
+           beta_estimate_smoothed,
+           keep) 
     
     
     
     
   state_testing <- state_testing %>% 
   #  mutate(fips = tolower(fips)) %>%
-    inner_join(ctis_biweekly, by =c('fips'='fips', 'biweek'='biweek')) %>%
+    inner_join(ctis_biweekly, by =c('fips'='fips', 'biweek'='biweek')) 
+  
+  glimpse(state_testing)
+  
+  corrected <- case_when(vary ==  "beta" ~ 
+                           get_v2_corrected(state_testing, params) %>%
+                           mutate(version="v2"))
+    
+  
+  
+  return(corrected)
+  
+}
+
+
+get_v2_corrected <- function(state_testing, params) {
+  
+  
+  state_testing <- state_testing %>% 
     # only have CTIS data starting at week 6
[…]

### 2023-06-16 _targets.R

@@ -82,7 +82,8 @@ list(
     command = get_v2_state(
       data = tests_biweekly_state,
       params = prior_params,
-      ctis=ctis_smoothed
+      ctis=ctis_smoothed,
+      vary = "beta"
     )
   )
 

### 2023-06-17 R/05-state-analysis.R
function names

@@ -135,7 +135,7 @@ get_smoothed_ctis <- function(ctis_raw,
 
 #---------- Version 2 -------------
 
-get_v2_state <- function(data, params, ctis, vary, testing = TRUE) {
+get_corrected_state <- function(data, params, ctis, vary, testing = TRUE) {
   
   dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
   

### 2023-06-17 _targets.R

@@ -79,7 +79,7 @@ list(
   # version 2
   tar_target(
     name = state_v2,
-    command = get_v2_state(
+    command = get_corrected_state(
       data = tests_biweekly_state,
       params = prior_params,
       ctis=ctis_smoothed,

### 2023-06-17 R/05-state-analysis.R
full state-level targets

@@ -2,7 +2,7 @@
 
 
 #---------- Version 1 -------------
-get_v1_state <- function(data, params, testing = TRUE) {
+get_v1_state <- function(data, params, testing = FALSE) {
   
   state_testing <- data
   
@@ -14,7 +14,7 @@ get_v1_state <- function(data, params, testing = TRUE) {
   
   # only use a few rows if testing
   state_testing <- if(testing) state_testing %>% 
-    filter(biweek >=6) %>% slice_sample(n=15) else state_testing
+    filter(biweek >=6) %>% slice_sample(n=5) else state_testing
   
   melded <- do.call(get_melded, params)
   
@@ -133,9 +133,7 @@ get_smoothed_ctis <- function(ctis_raw,
 
 
 
-#---------- Version 2 -------------
-
-get_corrected_state <- function(data, params, ctis, vary, testing = TRUE) {
+get_corrected_state <- function(data, params, ctis, vary, testing = FALSE) {
   
   dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
   
@@ -149,7 +147,7 @@ get_corrected_state <- function(data, params, ctis, vary, testing = TRUE) {
   
   # only use a few rows if testing
   state_testing <- if(testing) state_testing %>% 
-    filter(biweek >=6) %>% slice_sample(n=15) else state_testing
+    filter(biweek >=6) %>% slice_sample(n=5) else state_testing
   
   
   ctis_biweekly <- ctis %>%
@@ -171,12 +169,36 @@ get_corrected_state <- function(data, params, ctis, vary, testing = TRUE) {
   #  mutate(fips = tolower(fips)) %>%
     inner_join(ctis_biweekly, by =c('fips'='fips', 'biweek'='biweek')) 
   
-  glimpse(state_testing)
+  # glimpse(state_testing)
   
-  corrected <- case_when(vary ==  "beta" ~ 
-                           get_v2_corrected(state_testing, params) %>%
-                           mutate(version="v2"))
-    
+  if(vary=="beta") {
+    message("v2")
+    corrected <- get_v2_corrected(state_testing, params) %>%
+    mutate(version="v2") }
+  
+  if(vary=="s_untested") {
+    message("v3")
+    corrected <- get_v3_corrected(state_testing, params) %>%
+      mutate(version="v3") }
+  
+  if(vary=="s_untested_and_beta") {
+    message("v4")
+    corrected <- get_v4_corrected(state_testing, params) %>%
+      mutate(version="v4") }
+  
[…]

### 2023-06-17 _targets.R

@@ -85,6 +85,54 @@ list(
       ctis=ctis_smoothed,
       vary = "beta"
     )
+  ),
+  # version 3
+  tar_target(
+    name = state_v3,
+    command = get_corrected_state(
+      data = tests_biweekly_state,
+      params = prior_params,
+      ctis=ctis_smoothed,
+      vary = "s_untested"
+    )
+  ),
+  # version 4
+  tar_target(
+    name = state_v4,
+    command = get_corrected_state(
+      data = tests_biweekly_state,
+      params = prior_params,
+      ctis=ctis_smoothed,
+      vary = "s_untested_and_beta"
+    )
   )
+  
+  
 
 )
+
+
+
+# for testing
+# ctis <- tar_read(ctis_smoothed)
+# params <-  list(
+# alpha_mean = .95,
+# alpha_sd = 0.08,
+# alpha_bounds = NA,
+# # alpha_bounds = c(.8,1),
+# beta_mean = .15,
+# beta_sd =.09,
+# beta_bounds = NA,
+# #  beta_bounds = c(0.002, 0.4),
+# s_untested_mean = .03,
+# s_untested_sd = .0225,
+# #  s_untested_bounds = c(0.0018, Inf),
+# s_untested_bounds = NA,
+# p_s0_pos_mean = .4,
+# p_s0_pos_sd = .1225,
+# p_s0_pos_bounds = NA,
+# #  p_s0_pos_bounds = c(.25, .7),
+# pre_nsamp = 1e6,
+# post_nsamp = 1e5)
+# data <- tar_read(tests_biweekly_state)
+

### 2023-06-17 R/05-state-analysis.R
update county-level targets

@@ -185,21 +185,7 @@ get_corrected_state <- function(data, params, ctis, vary, testing = FALSE) {
     message("v4")
     corrected <- get_v4_corrected(state_testing, params) %>%
       mutate(version="v4") }
-  
-  
-  
-  
-  # corrected <- case_when(vary ==  "beta" ~ 
-  #                          get_v2_corrected(state_testing, params) %>%
-  #                          mutate(version="v2"),
-  #                        vary == "s_untested" ~ 
-  #                          get_v3_corrected(state_testing, params) %>%
-  #                          mutate(version="v3"),
-  #                        vary == "s_untested_and_beta" ~ 
-  #                          get_v4_corrected(state_testing, params) %>%
-  #                          mutate(version="v4"))
-  #   
-  
+
   
   return(corrected)
   
@@ -266,13 +252,7 @@ get_v2_corrected <- function(state_testing, params) {
 get_v3_corrected <- function(state_testing, params) {
   
   message("Running version 3")
-  
-  
-  state_testing <- state_testing %>% 
-    # only have CTIS data starting at week 6
-    # filter out the beginning dates where beta_estimate_smoothed is NA
-    filter(!is.na(beta_estimate_smoothed))
-  
+
   
   corrected <- state_testing %>% 
     arrange(biweek) %>%

### 2023-06-17 R/06-county-analysis.R

@@ -1,0 +1,279 @@
+
+#---------- Version 1 -------------
+get_county_v1 <- function(data, params, testing = FALSE) {
+
+    county_testing <- data
+    
+    county_testing <- county_testing %>%
+      select(-date) %>% 
+      distinct()
+    
+    
+    # only use a few rows if testing
+    county_testing <- if(testing) county_testing %>% 
+      filter(biweek >=6) %>% slice_sample(n=5) else county_testing
+    
+    melded <- do.call(get_melded, params)
+    
+    corrected <- pmap_df(county_testing, 
+                         function(posrate,
+                                  population,
+                                  total, 
+                                  positive,
+                                  biweek,
+                                  fips,
+                                  ...) {
+                           process_priors_per_county(
+                             priors = melded$post_melding,
+                             county_df = tibble(posrate, 
+                                                population,
+                                                total, 
+                                                positive,
+                                                biweek,
+                                                fips),
+                             nsamp = params$post_nsamp) %>%
+                             generate_corrected_sample(., num_reps = 1e3) %>%
+                             summarize_corrected_sample() })
+    
+    return(corrected)
+    
+    
+}
+
+
+#---------------- Version 2 --------------
+
+get_ctis_biweekly <- function(ctis) {
+  
+  dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
+  
+  ctis_biweekly <- ctis %>%
+    left_join(dates) %>%
+    group_by(biweek, state) %>%
+    slice_max(n=1, order_by=date) %>%
+    select(-date) %>%
+    mutate(state =toupper(state)) %>%
+    select(state,
+           biweek, 
+           s_untested_smoothed, 
+           beta_estimate_smoothed,
+           keep) 
+  
+  return(ctis_biweekly)
+  
+}
+
+
+
+get_corrected_county <- function(data, params, ctis,
+                                 vary, testing = FALSE) {
+  
+  dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
+  
+  county_testing <- data
+  
+  county_testing <- county_testing %>%
+    select(-date) %>% 
+    distinct()
+  
+  
+  # only use a few rows if testing
+  county_testing <- if(testing) county_testing %>% 
+    filter(biweek >=6) %>% slice_sample(n=5) else county_testing
+  
+  ctis_biweekly <- get_ctis_biweekly(ctis)
+  
[…]

### 2023-06-17 _targets.R

@@ -26,6 +26,7 @@ tar_source("R/02-priors.R")
 tar_source("R/03-base-functions.R")
 tar_source("R/04-melding.R")
 tar_source("R/05-state-analysis.R")
+tar_source("R/06-county-analysis.R")
 
 
 
@@ -105,12 +106,52 @@ list(
       ctis=ctis_smoothed,
       vary = "s_untested_and_beta"
     )
+  ),
+  
+  
+  #---------------- county-level analysis -------------------
+  
+  #------ MA ---------
+  # version 1
+  tar_target(
+    name = ma_v1,
+    command = get_county_v1(
+      data = ma_biweekly_county,
+      params = prior_params
+    )
+  ),
+  tar_target(
+    name = ma_v2,
+    command = get_corrected_county(
+      data = ma_biweekly_county,
+      params = prior_params,
+      ctis=ctis_smoothed,
+      vary = "beta"
+    )
+  ),
+  tar_target(
+    name = ma_v3,
+    command = get_corrected_county(
+      data = ma_biweekly_county,
+      params = prior_params,
+      ctis=ctis_smoothed,
+      vary = "s_untested"
+    )
+  ),
+  tar_target(
+    name = ma_v4,
+    command = get_corrected_county(
+      data = ma_biweekly_county,
+      params = prior_params,
+      ctis=ctis_smoothed,
+      vary = "s_untested_and_beta"
+    )
   )
-  
-  
-
 )
 
+# tar_read(ma_v2)
+# tar_read(ma_v3)
+# tar_read(ma_v4)
 
 
 # for testing

### 2023-06-18 reports/priors.Rmd
visualize priors

---
title: "Priors"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

library(tidyverse)
library(targets)
library(here)

# read priors from target
prior_params <- tar_read(prior_params,
                         store = here("_targets"))
# load beta_density, gamma_density functions
source(here("R/03-base-functions.R"))

```

```{r}

theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
    theme_bw() %+replace%    #replace elements we want to change

    theme(

      #text elements
      plot.title = element_text(             #title
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),

      plot.subtitle = element_text(          #subtitle
                   size = 12,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size

      axis.title = element_text(             #axis titles
                   size = 12),               #font size

      axis.text.x = element_text(              #axis text
                   size = 12),
      legend.text = element_text(size = 18),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 16, face = "bold", color="white"),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

}

```

# $\alpha$

```{r}

tibble(x = seq(0, 1.3, length = 10^(5)),
       y = with(prior_params, gamma_density(x,
                                     mean = alpha_mean,
                                     sd = alpha_sd,
                                     bounds = alpha_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $\\alpha$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$alpha_mean,
                         ", SD: ", prior_params$alpha_sd))

```

# $\beta$

```{r}

tibble(x = seq(0, 1, length = 10^(5)),
       y = with(prior_params, beta_density(x,
                                     mean = beta_mean,
                                     sd = beta_sd,
                                     bounds = beta_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $\\beta$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$beta_mean,
                         ", SD: ", prior_params$beta_sd))

```

# $P(S_1|\text{untested})$

```{r}

tibble(x = seq(0, 1, length = 10^(5)),
       y = with(prior_params, beta_density(x,
                                     mean = s_untested_mean,
                                     sd = s_untested_sd,
                                     bounds = s_untested_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $P(S_1|untested)$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$s_untested_mean,
                         ", SD: ", prior_params$s_untested_sd))

```

# $P(S_0|  \text{test}_+, \text{untested})$

```{r}

tibble(x = seq(0, 1, length = 10^(5)),
       y = with(prior_params, beta_density(x,
                                     mean = p_s0_pos_mean,
                                     sd = p_s0_pos_sd,
                                     bounds = p_s0_pos_bounds))) %>%
   ggplot(aes(x=x, y = y)) +
  geom_line(alpha = .8) + 
  geom_ribbon(aes(x=x,ymin=0,ymax=y),
              fill="black",
              alpha=.6) +
  theme_c(legend.text=element_text(size = 10)) +
  viridis::scale_fill_viridis(discrete=TRUE,
                             option = "rocket", begin=.3,end=.8) +
  labs(x = "Value",
       y = "Probability Density", 
       title = latex2exp::TeX("Definition of Prior for $P(S_0|test_+,untested)$"),
       fill ='',
       subtitle = paste0("Mean: ", prior_params$p_s0_pos_mean,
                         ", SD: ", prior_params$p_s0_pos_sd))

```

### 2023-06-18 _targets.R
update targets

@@ -120,6 +120,7 @@ list(
       params = prior_params
     )
   ),
+  # version 2
   tar_target(
     name = ma_v2,
     command = get_corrected_county(
@@ -129,6 +130,7 @@ list(
       vary = "beta"
     )
   ),
+  # version 3
   tar_target(
     name = ma_v3,
     command = get_corrected_county(
@@ -138,6 +140,7 @@ list(
       vary = "s_untested"
     )
   ),
+  # version 4
   tar_target(
     name = ma_v4,
     command = get_corrected_county(
@@ -146,7 +149,14 @@ list(
       ctis=ctis_smoothed,
       vary = "s_untested_and_beta"
     )
-  )
+  ),
+  
+  tar_target(
+    name = waste,
+    command = get_biweekly_wastewater())
+  
+  
+  
 )
 
 # tar_read(ma_v2)

### 2023-06-18 reports/county_correlations.Rmd
county correlations

---
title: "County Correlations"
author: "Quinn White"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### 2023-06-18 R/01-data_cleaning.R
wastewater data

@@ -560,6 +560,52 @@ save_survey_data <- function() {
 
 
 
+#----------- Wastewater Data ----------------
+
+get_biweekly_wastewater <- function() {
+  
+  biobot_link <- "https://raw.githubusercontent.com/biobotanalytics/covid19-wastewater-data/master/wastewater_by_county.csv"
+  
+  dates <- readRDS(here("data", "date_to_biweek.RDS"))
+  
+  w_data <- read_csv(biobot_link)%>% 
+    filter(sampling_week >= ymd("2021-03-01") &  sampling_week <= ymd("2022-03-01")) %>%
+    mutate(fips = as.character(fipscode)) %>%
+    select(-fipscode)  %>%
+    left_join(dates, by = c("sampling_week" = "date")) %>%
+    group_by(fips, biweek, state,name) %>%
+    summarize(mean_conc = mean(effective_concentration_rolling_average, na.rm= TRUE)) %>%
+    ungroup()
+  
+  
+  
+  # fill in missing values with rolling mean
+  w_data <- w_data %>% 
+    group_by(fips) %>%
+    # keep track of min biweek make sure to not fill in values
+    # before first data point when using rolled mean
+    mutate(min_biweek = min(biweek)) %>%
+    ungroup() %>%
+    pivot_wider(names_from = biweek,
+                values_from = mean_conc) %>% 
+    pivot_longer(cols = -c(fips,state, name, min_biweek), 
+                 names_to = "biweek", 
+                 values_to = "mean_conc") %>%
+    mutate(biweek = as.numeric(biweek)) %>%
+    group_by(fips) %>%
+    arrange(biweek) %>%
+    mutate(rolled_mean = RcppRoll::roll_mean(mean_conc,
+                                             n = 4,
+                                             na.rm = TRUE,
+                                             fill = NA),
+           mean_conc = ifelse(is.na(mean_conc) & biweek >= min_biweek, 
+                              rolled_mean, mean_conc)) %>%
+    ungroup()
+  
+  return(w_data)
+  
+}
+
 
 
 

### 2023-06-18 R/05-state-analysis.R
add v1 label

@@ -38,6 +38,9 @@ get_v1_state <- function(data, params, testing = FALSE) {
       generate_corrected_sample(., num_reps = 1e3) %>%
       summarize_corrected_sample() })
   
+  corrected <- corrected %>%
+    mutate(version="v1")
+  
   return(corrected)
   
 }

### 2023-06-18 R/06-county-analysis.R

@@ -35,6 +35,10 @@ get_county_v1 <- function(data, params, testing = FALSE) {
                              generate_corrected_sample(., num_reps = 1e3) %>%
                              summarize_corrected_sample() })
     
+    
+    corrected <- corrected %>%
+      mutate(version="v1")
+    
     return(corrected)
     
     

### 2023-06-18 R/01-data_cleaning.R
waste target

@@ -566,10 +566,11 @@ get_biweekly_wastewater <- function() {
   
   biobot_link <- "https://raw.githubusercontent.com/biobotanalytics/covid19-wastewater-data/master/wastewater_by_county.csv"
   
-  dates <- readRDS(here("data", "date_to_biweek.RDS"))
+  dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
   
   w_data <- read_csv(biobot_link)%>% 
-    filter(sampling_week >= ymd("2021-03-01") &  sampling_week <= ymd("2022-03-01")) %>%
+    filter(sampling_week >= ymd("2021-03-01") &  
+             sampling_week <= ymd("2022-03-01")) %>%
     mutate(fips = as.character(fipscode)) %>%
     select(-fipscode)  %>%
     left_join(dates, by = c("sampling_week" = "date")) %>%

### 2023-06-19 _targets.R
add targets for spline smoothing

@@ -107,6 +107,28 @@ list(
       vary = "s_untested_and_beta"
     )
   ),
+  # version 5
+  tar_target(
+    name = state_v5,
+    command = get_corrected_state(
+      data = tests_biweekly_state,
+      params = prior_params,
+      ctis=ctis_smoothed,
+      spline=TRUE,
+      vary = "beta"
+    )
+  ),
+  # version 6
+  tar_target(
+    name = state_v6,
+    command = get_corrected_state(
+      data = tests_biweekly_state,
+      params = prior_params,
+      ctis=ctis_smoothed,
+      spline=TRUE,
+      vary = "s_untested_and_beta"
+    )
+  ),
   
   
   #---------------- county-level analysis -------------------
@@ -151,10 +173,33 @@ list(
     )
   ),
   
+  # version 5
+  tar_target(
+    name = ma_v5,
+    command = get_corrected_county(
+      data = ma_biweekly_county,
+      params = prior_params,
+      ctis=ctis_smoothed,
+      spline=TRUE,
+      vary = "beta"
+    )
+  ),
+  
+  # version 6
+  tar_target(
+    name = ma_v6,
+    command = get_corrected_county(
+      data = ma_biweekly_county,
+      params = prior_params,
+      ctis=ctis_smoothed,
+      spline=TRUE,
+      vary = "s_untested_and_beta"
+    )
+  ),
+  
   tar_target(
     name = waste,
     command = get_biweekly_wastewater())
-  
   
   
 )

### 2023-06-19 R/05-state-analysis.R
spline smoothing functions county + state-level

@@ -47,8 +47,6 @@ get_v1_state <- function(data, params, testing = FALSE) {
 
 
 #------------- Using CTIS Data ----------------
-
-
 get_smoothed_ctis <- function(ctis_raw,
                               smooth_beta_span = .33,
                               smooth_s_untested_span =.2) {
@@ -117,26 +115,37 @@ get_smoothed_ctis <- function(ctis_raw,
         smoothed_beta <- loess(imputed_beta~index, 
                                data = .x, 
                                span = smooth_beta_span)
+        # 257 corresponds to 2021-12-01
+        spline_smoothed_beta <- lm(imputed_beta ~ splines::bs(index,knots = c(118, 257)), 
+                                   data = .x )
+        
+        # include both spline smoothing and loess smoothing for beta
         .x %>%
-          mutate(beta_estimate_smoothed = predict(smoothed_beta),
-                 s_untested_smoothed = predict(smoothed_s_untested)) 
+          mutate(beta_estimate_smoothed = predict(smoothed_beta, newdata = index),
+                 beta_estimate_spline_smoothed = predict(spline_smoothed_beta, newx= index),
+                 s_untested_smoothed = predict(smoothed_s_untested, newdata = index)) 
       }
       
       else {
         .x %>%
-          mutate(s_untested_smoothed = predict(smoothed_s_untested)) 
+          mutate(s_untested_smoothed = predict(smoothed_s_untested,newdata = index)) 
   
       }
      
     } ) 
   
-  return(symp_smooth)
+  symp_smooth %>% 
+    select(-c(m,n,prop_with_data, index)) %>%
+    return()
   
 }
 
 
 
-get_corrected_state <- function(data, params, ctis, vary, testing = FALSE) {
+get_corrected_state <- function(data, params,
+                                ctis, vary, 
+                                spline=FALSE,
+                                testing = FALSE) {
   
   dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
   
@@ -165,29 +174,45 @@ get_corrected_state <- function(data, params, ctis, vary, testing = FALSE) {
            beta_estimate_smoothed,
            keep) 
     
-    
-    
-    
   state_testing <- state_testing %>% 
   #  mutate(fips = tolower(fips)) %>%
     inner_join(ctis_biweekly, by =c('fips'='fips', 'biweek'='biweek')) 
   
   # glimpse(state_testing)
   
-  if(vary=="beta") {
+  
[…]

### 2023-06-19 R/06-county-analysis.R

@@ -70,7 +70,8 @@ get_ctis_biweekly <- function(ctis) {
 
 
 get_corrected_county <- function(data, params, ctis,
-                                 vary, testing = FALSE) {
+                                 vary, spline = FALSE,
+                                 testing = FALSE) {
   
   dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
   
@@ -94,22 +95,36 @@ get_corrected_county <- function(data, params, ctis,
   
   # glimpse(state_testing)
   
-  if(vary=="beta") {
+  if(vary=="beta" & spline == FALSE) {
     message("v2")
     corrected <- get_v2_corrected(county_testing, params) %>%
       mutate(version="v2") }
   
-  if(vary=="s_untested") {
+  if(vary=="s_untested" & spline == FALSE) {
     message("v3")
     corrected <- get_v3_corrected(county_testing, params) %>%
       mutate(version="v3") }
   
-  if(vary=="s_untested_and_beta") {
+  if(vary=="s_untested_and_beta" & spline == FALSE) {
     message("v4")
     corrected <- get_v4_corrected(county_testing, params) %>%
       mutate(version="v4") }
   
+  ###########################
+  # spline-smoothed beta
+  ###########################
   
+  if(vary=="beta" & spline == TRUE) {
+    message("v5")
+    corrected <- get_v5_corrected(county_testing, params) %>%
+      mutate(version="v5") }
+  
+  if(vary=="s_untested_and_beta" & spline == TRUE) {
+    message("v6")
+    corrected <- get_v6_corrected(county_testing, params) %>%
+      mutate(version="v6") }
+  
+
   return(corrected)
   
 }
@@ -170,7 +185,7 @@ get_v2_corrected <- function(county_testing, params) {
 
 
 #---------- Version 3 -------------
-# center prior for beta at survey estimates 
+# center prior for P(S_1|untested) at survey estimates 
 get_v3_corrected <- function(county_testing, params) {
   
   message("Running version 3")
@@ -227,7 +242,7 @@ get_v3_corrected <- function(county_testing, params) {
 
 
 #---------- Version 4 -------------
-# center prior for beta at survey estimates 
+# center prior for BOTH beta and P(S_1|untested) at survey estimates 
 get_v4_corrected <- function(county_testing, params) {
   
   message("Running version 4")
@@ -261,6 +276,119 @@ get_v4_corrected <- function(county_testing, params) {
[…]

### 2023-06-19 reports/county_correlations.Rmd
update report files

@@
 date: "`r Sys.Date()`"
-output: html_document
+output:
+  rmdformats::readthedown:
+    df_print: paged
+    code_folding: hide
+    css: !expr here::here('css/template.css')
+  html_document:
+    df_print: paged
+knit: (function(input, ...) {
+    rmarkdown::render(
+      input,
+      output_dir = "../output_html")})
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE)
+
+library(tidyverse)
+library(targets)
+library(here)
+library(latex2exp)
+
 ```
+
+```{r theme_c}
+
+theme_c <- function(...){ 
+   # font <- "Helvetica"   #assign font family up front
+    theme_bw() %+replace%    #replace elements we want to change
+
+    theme(
+
+      #text elements
+      plot.title = element_text(             #title
+                   size = 16,                #set font size
+                   face = 'bold',            #bold typeface
+                   hjust = .5,
+                   vjust = 3),
+
+      plot.subtitle = element_text(          #subtitle
+                   size = 12,
+                   hjust = .5,
+                   face = 'italic',
+                   vjust = 3),               #font size
+
+      axis.title = element_text(             #axis titles
+                   size = 12),               #font size
+
+      axis.text = element_text(              #axis text
+                   size = 9),
+      legend.text = element_text(size = 12),
+      # t, r, b, l
+      plot.margin = unit(c(1,.5,.5,.5), "cm"),
+      legend.position = "right",
+      strip.text.x = element_text(size = 11, face = "bold", color="white"),
+      strip.background = element_rect(fill = "#3E3D3D")
+      ) %+replace%
+      theme(...)
+
+}
+
+```
+
+```{r load-data}
+
+county <- tar_read(ma_v1, 
+                   store = here("_targets")) %>% 
+  mutate(version = "v1") %>%
+  bind_rows(
+    tar_read(ma_v2,
+             store = here("_targets"))
+  ) %>%
+   bind_rows(
+    tar_read(ma_v3, 
+             store = here("_targets"))
+  ) %>%
+   bind_rows(
+    tar_read(ma_v4, 
+             store = here("_targets"))
+  )
+
+waste <- tar_read(waste, store = here("_targets"))
+covidestim <- tar_read(covidestim_biweekly_county,
+                        store = here("_targets"))
+
+```
+
+```{r}
+
+wjoined <- county %>%
+  # set Nantucket, Duke fips to Nantucket since we have wastewater data
+  # for Nantucket
+  mutate(fips = ifelse(grepl("25019", fips), "25019", fips)) %>%
+  inner_join(waste) %>%
+  left_join(covidestim, relationship = "many-to-many") %>%
+  group_by(biweek) %>%
+  mutate(mindate = min(date),
+         maxdate = max(date)) %>%
+  ungroup()
+
+counties_with_waste <- wjoined %>%
+  group_by(fips) %>%
+  mutate(obs_w_notna = sum(!is.na(mean_conc))) %>%
+  filter(obs_w_notna != 0) %>%
+  pull(fips) %>%
+  unique()
+
+versions <- tibble(
+  version = c("v1", "v2", "v3", "v4"),
+  vlabel = factor(
+    c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
+    "$\\beta$ Centered at Emp. Value",
+    "$P(S_1|untested)$ Centered at Emp. Value",
+    "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values"
+    ),
+    levels = c(
+        "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
+        "$\\beta$ Centered at Emp. Value",
+         "$P(S_1|untested)$ Centered at Emp. Value",
+        "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values"
+
+    ) )
+)
+
+wjoined <- wjoined %>%
+  left_join(versions)
+
+w_data <- wjoined %>% 
+  select(-date) %>%
+  distinct() %>%
+  filter(fips %in% counties_with_waste) %>%
+  filter(!is.na(mean_conc)) %>%
+  group_by(fips) %>%
+  mutate(mean_conc_std = mean_conc/ max(mean_conc),
+         exp_cases_median_std = exp_cases_median / max(exp_cases_median),
+         exp_cases_lb_std = exp_cases_lb /max(exp_cases_median),
+         exp_cases_ub_std = exp_cases_ub / max(exp_cases_median),
+         infections_std  = infections/max(infections),
+         positive_std = positive/max(positive)) %>%
+  ungroup()
+
+```
+
+# Correlations
+
+For each county, normalize the wastewater concentration by the maximum concentration for that county and the probabilistic bias counts by the maximum *median* estimated counts for that county (2.5th percentile and 97.5th percentile also standardized by the maximum *median* estimated counts for that county).
+
+## Wastewater
+
+```{r}
+
+# correlations between PB counts and wastewater concentrations
+
+correlations <- w_data %>%
+  group_by(version) %>%
+  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
+            vlabel =unique(vlabel)) %>%
+  mutate(x=1.3,
+         y=.1)
+
+w_data %>%
+  group_by(version) %>%
+  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = mean_conc_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= correlations ) +
+  facet_wrap(~vlabel,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Mean Wastewater Concentration\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")
+
+```
+
+## Covidestim
+
+```{r}
+
+# correlations between PB counts and Covidestim estimates
+
+covidestim_correlations <- w_data %>%
+  filter(!is.na(infections_std)) %>%
+  group_by(vlabel) %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(exp_cases_median_std,infections_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1.3,
+         y=.1)
+
+w_data %>%
+  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = infections_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  facet_wrap(~vlabel,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Covidestim Estimates\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+
+```
+
+## Covidestim and Wastewater
+
+```{r}
+
+covidestim_correlations <- w_data %>%
+  filter(version=="v1") %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(mean_conc_std,infections_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1.3,
+         y=.1)
+
+w_data %>%
+  filter(version=="v1") %>%
+  ggplot(aes(x=infections_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  facet_wrap(~vlabel,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Covidestim Estimates\n(Standardized)",
+       y = "Wastewater Concentrations\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+
+pb_correlations <- w_data %>%
+  filter(version=="v1") %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(positive_std,mean_conc_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1.3,
+         y=.1)
+
+```

### 2023-06-19 reports/priors.Rmd
update report files

@@
 source(here("R/03-base-functions.R"))
-
-```
-
-```{r}
 
@@
 ```
+
+# COVID-19 Trends and Impact Survey
+
+```{r}
+
+ctis_smoothed <- tar_read(ctis_smoothed,store = here("_targets"))
+
+```
+
+```{r}
+
+ctis_smoothed  %>%
+  filter(keep) %>%
+  mutate(state=toupper(state)) %>%
+  select(date,
+         state, 
+         imputed_beta,
+         beta_estimate_smoothed, 
+         beta_estimate_spline_smoothed) %>%
+  pivot_longer(contains("beta")) %>%
+  mutate(name = case_when(
+    name == "beta_estimate_smoothed" ~ "LOESS smoothed",
+    name == "beta_estimate_spline_smoothed" ~ "Spline smoothed",
+    name == "imputed_beta" ~ "Survey Value"
+  )) %>%
+  ggplot(aes(x=date, y=value, color = name, 
+             alpha = name, linewidth=name)) +
+  geom_line() +
+  facet_wrap(~state, ncol=4) +
+  scale_alpha_manual(values=c("LOESS smoothed" = .9, 
+                            "Spline smoothed" = .9, 
+                              "Survey Value"=.3),
+                     name='') +
+  scale_linewidth_manual(values = c("LOESS smoothed" = 1.05, 
+                            "Spline smoothed" = 1.05, 
+                              "Survey Value"=.5)) +
+  scale_color_manual(values=c("#3381FF", "#B58746", "#26900F"),
+                     name ='') +
+  guides(alpha="none",
+         linewidth="none",
+         color = guide_legend(override.aes = list(linewidth = 3,
+                                                  alpha = c("LOESS smoothed" = .9, 
+                            "Spline smoothed" = .9, 
+                              "Survey Value"=.3)))) +
+  theme_c() +
+  ylim(0,1) +
+  scale_x_date(date_breaks="3 months",
+               date_labels = "%b %Y")
+
+```

### 2023-06-20 reports/county_correlations.Rmd
update rmarkdown files

--- 2023-06-18 reports/county_correlations.Rmd
@@
 ---
-title: "County Correlations"
+title: "County Results"
 author: "Quinn White"
 date: "`r Sys.Date()`"
-output: html_document
+output:
+  rmdformats::readthedown:
+    df_print: paged
+    code_folding: hide
+    css: !expr here::here('css/template.css')
+  html_document:
+    df_print: paged
+    toc-depth: 4
+knit: (function(input, ...) {
+    rmarkdown::render(
+      input,
+      output_dir = "../output_html")})
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE)
+
+library(tidyverse)
+library(targets)
+library(here)
+library(latex2exp)
+
 ```
+
+```{r theme_c}
+
+theme_c <- function(...){ 
+   # font <- "Helvetica"   #assign font family up front
+    theme_bw() %+replace%    #replace elements we want to change
+
+    theme(
+
+      #text elements
+      plot.title = element_text(             #title
+                   size = 16,                #set font size
+                   face = 'bold',            #bold typeface
+                   hjust = .5,
+                   vjust = 3),
+
+      plot.subtitle = element_text(          #subtitle
+                   size = 12,
+                   hjust = .5,
+                   face = 'italic',
+                   vjust = 3),               #font size
+
+      axis.title = element_text(             #axis titles
+                   size = 12),               #font size
+
+      axis.text = element_text(              #axis text
+                   size = 9),
+      legend.text = element_text(size = 12),
+      # t, r, b, l
+      plot.margin = unit(c(1,.5,.5,.5), "cm"),
+      legend.position = "right",
+      strip.text.x = element_text(size = 11, face = "bold", color="white"),
+      strip.text.y = element_text(size = 11, face = "bold", color="white",angle=270),
+      strip.background = element_rect(fill = "#3E3D3D")
+      ) %+replace%
+      theme(...)
+
+}
+
+```
+
+```{r load-data}
+
+targets_dir <- here("_targets_6_19")
+
+county <- tar_read(ma_v1, 
+                   store =targets_dir) %>% 
+  mutate(version = "v1") %>%
+  bind_rows(
+    tar_read(ma_v2,
+             store =targets_dir)
+  ) %>%
+   bind_rows(
+    tar_read(ma_v3, 
+             store =targets_dir)
+  ) %>%
+   bind_rows(
+    tar_read(ma_v4, 
+             store =targets_dir)
+  # )%>%
+  #  bind_rows(
+  #   tar_read(ma_v5, 
+  #            store =targets_dir)
+  # )%>%
+  #  bind_rows(
+  #   tar_read(ma_v6, 
+  #            store =targets_dir)
+   )
+
+waste <- tar_read(waste, store =targets_dir)
+covidestim <- tar_read(covidestim_biweekly_county,
+                        store =targets_dir)
+
+dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
+
+```
+
+```{r}
+
+wjoined <- county %>%
+  # set Nantucket, Duke fips to Nantucket since we have wastewater data
+  # for Nantucket
+  mutate(fips = ifelse(grepl("25019", fips), "25019", fips)) %>%
+  inner_join(waste) %>%
+  left_join(covidestim, relationship = "many-to-many") %>%
+  group_by(biweek) %>%
+  mutate(mindate = min(date),
+         maxdate = max(date)) %>%
+  ungroup()
+
+counties_with_waste <- wjoined %>%
+  group_by(fips) %>%
+  mutate(obs_w_notna = sum(!is.na(mean_conc))) %>%
+  filter(obs_w_notna != 0) %>%
+  pull(fips) %>%
+  unique()
+
+versions <- tibble(
+  version = c("v1", "v2", "v3", "v4", "v5", "v6"),
+  vlabel = factor(
+    c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
+    "$\\beta$ Centered at Emp. Value",
+    "$P(S_1|untested)$ Centered at Emp. Value",
+    "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
+        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
+         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"
+    ),
+    levels = c(
+        "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
+        "$\\beta$ Centered at Emp. Value",
+         "$P(S_1|untested)$ Centered at Emp. Value",
+        "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
+        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
+         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"
+
+    ) )
+)
+
+wjoined <- wjoined %>%
+  left_join(versions)
+
+w_data <- wjoined %>% 
+  select(-date) %>%
+  distinct() %>%
+  filter(fips %in% counties_with_waste) %>%
+  filter(!is.na(mean_conc)) %>%
+  group_by(fips, version) %>%
+  mutate(mean_conc_std = mean_conc/ max(mean_conc),
+         exp_cases_median_std = exp_cases_median / max(exp_cases_median),
+         exp_cases_lb_std = exp_cases_lb /max(exp_cases_median),
+         exp_cases_ub_std = exp_cases_ub / max(exp_cases_median),
+         infections_std  = infections/max(infections),
+         positive_std = positive/max(positive)) %>%
+  ungroup()
+
+```
+
+# Correlations
+
+## Normalization Strategy 1: By Values at Biweek Where Wastewater Concentration is Maximized
+
+```{r}
+
+w_data <- wjoined %>% 
+  select(-date) %>%
+  distinct() %>%
+  filter(fips %in% counties_with_waste) %>%
+  filter(!is.na(mean_conc)) %>%
+  group_by(fips, version) %>%
+  mutate(mean_conc_std = mean_conc/ max(mean_conc),
+         max_conc = max(mean_conc),
+         exp_cases_median_std = exp_cases_median / exp_cases_median[which.max(mean_conc)],
+         exp_cases_lb_std = exp_cases_lb /exp_cases_median[which.max(mean_conc)],
+         exp_cases_ub_std = exp_cases_ub / exp_cases_median[which.max(mean_conc)],
+         infections_std  = infections/ infections[which.max(mean_conc)],
+         positive_std = positive/max(positive)) %>%
+  ungroup()
+
+```
+
+### PB Counts and Wastewater
+
+```{r}
+
+# correlations between PB counts and wastewater concentrations
+
+correlations <- w_data %>%
+  group_by(version) %>%
+  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
+            vlabel =unique(vlabel)) %>%
+  mutate(x=1.8,
+         y=.1)
+
+w_data %>%
+  group_by(version) %>%
+  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = mean_conc_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= correlations ) +
+  facet_wrap(~vlabel,
+             ncol=2,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Mean Wastewater Concentration\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")
+
+plt_wastewater_pb <- w_data %>%
+  group_by(version) %>%
+  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = mean_conc_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= correlations ) +
+  facet_wrap(~vlabel,
+             ncol=1,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Mean Wastewater Concentration\n(Standardized)",
+       title = "")
+
+```
+
+### PB Counts and Covidestim
+
+```{r}
+
+# correlations between PB counts and Covidestim estimates
+
+covidestim_correlations <- w_data %>%
+  filter(!is.na(infections_std)) %>%
+  group_by(vlabel) %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(exp_cases_median_std,infections_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1.8,
+         y=.1)
+
+w_data %>%
+  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = infections_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  facet_wrap(~vlabel,
+              ncol=2,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Covidestim Estimates\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+
+plt_covidestim_pb <- w_data %>%
+  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = infections_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  facet_wrap(~vlabel,
+              ncol=1,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Covidestim Estimates\n(Standardized)",
+       title = "")
+
+```
+
+### Combined Figure
+
+```{r, fig.height=10, fig.width=10}
+
+cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)
+
+```
+
+### Covidestim and Wastewater
+
+```{r}
+
+covidestim_correlations <- w_data %>%
+  filter(version=="v1") %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(mean_conc_std,infections_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1,
+         y=.1)
+
+w_data %>%
+  filter(version=="v1") %>%
+  ggplot(aes(x=infections_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  theme_c() +
+  labs(x = "Covidestim Estimates\n(Standardized)",
+       y = "Wastewater Concentrations\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+
+pb_correlations <- w_data %>%
+  filter(version=="v1") %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(positive_std,mean_conc_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1.8,
+         y=.1)
+
+```
+
+### Compare to Wastewater Over Time
+
+```{r,fig.width=12, fig.height=15}
+
+library(scales)
+
+pal <- c("#10BAC5", "#1B10C5", "#EFB719", "#900C3F")
+
+names(pal) <- c(#"$observed$", 
+                '$P(S_1|untested)$ Centered at Emp. Value',
+                '$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values',
+               "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$",
+               "$\\beta$ Centered at Emp. Value"
+               )
+
+waste_df <- w_data %>%
+  select(biweek, mean_conc_std, name) %>%
+  distinct() %>%
+  left_join(dates,relationship = "many-to-many" ) %>%
+  group_by(name, biweek) %>%
+  # center date within 2-week interval
+  summarize(date = min(date) + days(7),
+            mean_conc_std=unique(mean_conc_std)) %>%
+  mutate(name = gsub("County, MA", "", name))
+
+w_data%>%
+  mutate(name = gsub("County, MA", "", name)) %>%
+  left_join(dates,relationship = "many-to-many" ) %>%
+  filter(version %in% c("v1","v3")) %>%
+  filter(biweek>=6) %>%
+  group_by(biweek) %>%
+  mutate(mindate = min(date),
+         maxdate=max(date)) %>%
+  ungroup() %>% 
+ #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
+    ggplot() +
+      geom_ribbon(aes(x = date, 
+                 ymin = exp_cases_lb_std,
+                 ymax = exp_cases_ub_std,
+                 fill = vlabel),
+                 alpha = .5) +
+      geom_line(data = waste_df,
+                aes(x = date, y =mean_conc_std, 
+                    color = 'Wastewater Effective Concentrations'),
+               # color = "#DB4048",
+                size = 1.1) +
+      geom_linerange(aes(xmin = mindate, 
+                         xmax = maxdate, 
+                         y = infections_std,
+                         color = 'Covidestim Infection Estimates')) + 
+      geom_point(data = waste_df,
+                aes(x = date, y =mean_conc_std ),
+                color = "#DB4048",
+                alpha = .5,
+                size = 1.2) +
+      facet_grid(name~vlabel, 
+                 labeller=labeller(vlabel =as_labeller(
+                   TeX, default=label_parsed))) +
+      scale_fill_manual(values = pal, labels = c('','','',''), 
+                        name = "Probabilistic Bias Intervals") +
+      theme_bw() +
+     theme_c() +
+      labs(y = "Value (Standardized)",
+           title = "",
+           x= "") +
+      scale_x_date(date_labels = "%b %Y") +
+    scale_color_manual(
+      name = '', 
+      values = c(
+        'Covidestim Infection Estimates' = 'darkblue',
+        'Wastewater Effective Concentrations' = '#DB4048')) +
+    guides(color = guide_legend(override.aes = list(size = 12,
+                                                    linewidth=3)),
+           fill = guide_legend(override.aes =list(size = 6)))
+
+```
+
+## Normalization Strategy 2: By Maximum (Separately)
+
+For each county, normalize the wastewater concentration by the maximum concentration for that county and the probabilistic bias counts by the maximum *median* estimated counts for that county (2.5th percentile and 97.5th percentile also standardized by the maximum *median* estimated counts for that county).
+
+### PB Counts and Wastewater
+
+```{r}
+
+# correlations between PB counts and wastewater concentrations
+
+correlations <- w_data %>%
+  group_by(version) %>%
+  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
+            vlabel =unique(vlabel)) %>%
+  mutate(x=1.8,
+         y=.1)
+
+w_data %>%
+  group_by(version) %>%
+  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = mean_conc_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= correlations ) +
+  facet_wrap(~vlabel,
+             ncol=2,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Mean Wastewater Concentration\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")
+
+plt_wastewater_pb <- w_data %>%
+  group_by(version) %>%
+  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = mean_conc_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= correlations ) +
+  facet_wrap(~vlabel,
+             ncol=1,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Mean Wastewater Concentration\n(Standardized)",
+       title = "")
+
+```
+
+### PB Counts and Covidestim
+
+```{r}
+
+# correlations between PB counts and Covidestim estimates
+
+covidestim_correlations <- w_data %>%
+  filter(!is.na(infections_std)) %>%
+  group_by(vlabel) %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(exp_cases_median_std,infections_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1.8,
+         y=.1)
+
+w_data %>%
+  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = infections_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  facet_wrap(~vlabel,
+             ncol=2,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Covidestim Estimates\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+
+plt_covidestim_pb <- w_data %>%
+  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = infections_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  facet_wrap(~vlabel,
+              ncol=1,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Covidestim Estimates\n(Standardized)",
+       title = "")
+```
+
+### Combined Figure
+
+```{r, fig.height=10, fig.width=10}
+
+cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)
+
+```
+
+### Covidestim and Wastewater
+
+```{r}
+
+covidestim_correlations <- w_data %>%
+  filter(version=="v1") %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(mean_conc_std,infections_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1,
+         y=.1)
+
+w_data %>%
+  filter(version=="v1") %>%
+  ggplot(aes(x=infections_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  theme_c() +
+  labs(x = "Covidestim Estimates\n(Standardized)",
+       y = "Wastewater Concentrations\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+
+pb_correlations <- w_data %>%
+  filter(version=="v1") %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(positive_std,mean_conc_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1,
+         y=.1)
+
+```

### 2023-06-20 reports/priors.Rmd
update rmarkdown files

--- 2023-06-18 reports/priors.Rmd
@@
 library(here)
+library(latex2exp)
 
@@
                    size = 12),
-      legend.text = element_text(size = 18),
+      legend.text = element_text(size = 14),
       # t, r, b, l
@@
 ```
+
+# COVID-19 Trends and Impact Survey
+
+```{r}
+
+ctis_smoothed <- tar_read(ctis_smoothed,store = here("_targets"))
+
+```
+
+## $\beta$
+
+```{r beta-smoothing, fig.cap="The ratio of the screening test positivity over the overall test positivity from the COVID-19 Trends and Impact Survey is taken to be the estimate of $\\beta$.
+We compare two approaches to smoothing: cubic spline smoothing with 2 knots (July 15th, 2021 and December 1st, 2021) to LOESS smoothing with a span of 0.33."}
+
+ctis_smoothed  %>%
+  filter(keep) %>%
+  mutate(state=toupper(state)) %>%
+  select(date,
+         state, 
+         imputed_beta,
+         beta_estimate_smoothed, 
+         beta_estimate_spline_smoothed) %>%
+  pivot_longer(contains("beta")) %>%
+  mutate(name = case_when(
+    name == "beta_estimate_smoothed" ~ "LOESS smoothed",
+    name == "beta_estimate_spline_smoothed" ~ "Spline smoothed",
+    name == "imputed_beta" ~ "Survey Value"
+  )) %>%
+  ggplot(aes(x=date, y=value, color = name, 
+             alpha = name, linewidth=name)) +
+  geom_line() +
+  facet_wrap(~state, ncol=4) +
+  scale_alpha_manual(values=c("LOESS smoothed" = .9, 
+                            "Spline smoothed" = .9, 
+                              "Survey Value"=.3),
+                     name='') +
+  scale_linewidth_manual(values = c("LOESS smoothed" = 1.05, 
+                            "Spline smoothed" = 1.05, 
+                              "Survey Value"=.5)) +
+  scale_color_manual(values=c("#3381FF", "#B58746", "#26900F"),
+                     name ='') +
+  guides(alpha="none",
+         linewidth="none",
+         color = guide_legend(override.aes = list(linewidth = 3,
+                                                  alpha = c("LOESS smoothed" = .9, 
+                            "Spline smoothed" = .9, 
+                              "Survey Value"=.3)),
+                            nrow=3)) +
+  theme_c(legend.position="top") +
+  ylim(0,1) +
+  scale_x_date(date_breaks="3 months",
+               date_labels = "%b %Y") + 
+  labs(y = TeX("Survey Estimate of $\\beta$"),
+       x= "",
+       title = TeX("Comparing Approaches for Smoothing Survey Estimates of $\\beta$"))
+
+```
+
+## $\Pr(S_1 | \text{untested})$
+
+```{r s-untested-smoothing, fig.cap="The percentage of the population experiencing COVID-19-like illness is taken to be the estimate of $\\Pr(S_1|\\text{untested}).
+The LOESS smoothed estimate with a span of 0.2 is shown in red."}
+
+ctis_smoothed  %>%
+  mutate(state=toupper(state)) %>%
+  select(date,
+         state, 
+         contains("s_untested")) %>%
+  pivot_longer(contains("s_untested")) %>%
+  mutate(name = case_when(
+    name == "s_untested_smoothed" ~ "LOESS smoothed",
+    name == "imputed_s_untested" ~ "Survey Value"
+  )) %>%
+  ggplot(aes(x=date, y=value, color = name, 
+             alpha = name, linewidth=name)) +
+  geom_line() +
+  facet_wrap(~state, ncol=5) +
+  scale_alpha_manual(values=c("LOESS smoothed" = .9, 
+                              "Survey Value"=.6),
+                     name='') +
+  scale_linewidth_manual(values = c("LOESS smoothed" = 1.02, 
+                              "Survey Value"=.9)) +
+  scale_color_manual(values=c("Survey Value" ="black", "LOESS smoothed"="darkred"),
+                     name ='') +
+  guides(alpha="none",
+         linewidth="none",
+         color = guide_legend(override.aes = list(
+           linewidth = 3,
+           alpha = c("LOESS smoothed" = .9, 
+                     "Survey Value"=.3)),
+                            nrow=3)) +
+  theme_c(legend.position="top") +
+  scale_x_date(date_breaks="3 months",
+               date_labels = "%b %Y") + 
+  labs(y = TeX("Survey Estimate of $Pr(S_1|untested)$"),
+       x= "",
+       title = TeX("Comparing Approaches for Smoothing Survey Estimates of $Pr(S_1|untested)$"))
+
+```

### 2023-06-20 R/05-state-analysis.R
add additional analysis versions (v5, v6)

@@ -2,7 +2,7 @@
 
 
 #---------- Version 1 -------------
-get_v1_state <- function(data, params, testing = FALSE) {
+get_v1_state <- function(data, params, testing = TRUE) {
   
   state_testing <- data
   
@@ -145,7 +145,7 @@ get_smoothed_ctis <- function(ctis_raw,
 get_corrected_state <- function(data, params,
                                 ctis, vary, 
                                 spline=FALSE,
-                                testing = FALSE) {
+                                testing = TRUE) {
   
   dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
   
@@ -172,7 +172,11 @@ get_corrected_state <- function(data, params,
            biweek, 
            s_untested_smoothed, 
            beta_estimate_smoothed,
-           keep) 
+           beta_estimate_spline_smoothed,
+           keep)  %>%
+    ungroup()
+  
+  glimpse(ctis_biweekly)
     
   state_testing <- state_testing %>% 
   #  mutate(fips = tolower(fips)) %>%
@@ -209,7 +213,7 @@ get_corrected_state <- function(data, params,
   
   if(vary=="s_untested_and_beta" & spline == TRUE) {
     message("v6")
-    corrected <- get_v5_corrected(state_testing, params) %>%
+    corrected <- get_v6_corrected(state_testing, params) %>%
       mutate(version="v6") }
   
   
@@ -387,8 +391,11 @@ get_v4_corrected <- function(state_testing, params) {
 # at survey estimate (loess smoothed) 
 get_v5_corrected <- function(state_testing, params) {
   
-  message("Running version 5")
+  glimpse(state_testing)
+  nrow(state_testing)
   
+  message("Running version 5")
+
   state_testing <- state_testing %>% 
     # only have CTIS data starting at week 6
     # filter out the beginning dates where beta_estimate_smoothed is NA
@@ -407,6 +414,7 @@ get_v5_corrected <- function(state_testing, params) {
                       posrate,
                       population,
                       beta_estimate_smoothed,
+                      beta_estimate_spline_smoothed,
                       s_untested_smoothed,
                       ...) {
       
@@ -414,9 +422,14 @@ get_v5_corrected <- function(state_testing, params) {
                            total,  biweek,
                            posrate, population,
                            beta_estimate_smoothed,
-                           s_untested_smoothed)
+                           s_untested_smoothed,
+                           beta_estimate_spline_smoothed)
       # message(paste0("before: ",prior_params$beta_mean))
[…]

### 2023-06-20 R/06-county-analysis.R

@@ -61,6 +61,7 @@ get_ctis_biweekly <- function(ctis) {
            biweek, 
            s_untested_smoothed, 
            beta_estimate_smoothed,
+           beta_estimate_spline_smoothed,
            keep) 
   
   return(ctis_biweekly)
@@ -97,17 +98,17 @@ get_corrected_county <- function(data, params, ctis,
   
   if(vary=="beta" & spline == FALSE) {
     message("v2")
-    corrected <- get_v2_corrected(county_testing, params) %>%
+    corrected <- get_v2_county_corrected(county_testing, params) %>%
       mutate(version="v2") }
   
   if(vary=="s_untested" & spline == FALSE) {
     message("v3")
-    corrected <- get_v3_corrected(county_testing, params) %>%
+    corrected <- get_v3_county_corrected(county_testing, params) %>%
       mutate(version="v3") }
   
   if(vary=="s_untested_and_beta" & spline == FALSE) {
     message("v4")
-    corrected <- get_v4_corrected(county_testing, params) %>%
+    corrected <- get_v4_county_corrected(county_testing, params) %>%
       mutate(version="v4") }
   
   ###########################
@@ -116,12 +117,12 @@ get_corrected_county <- function(data, params, ctis,
   
   if(vary=="beta" & spline == TRUE) {
     message("v5")
-    corrected <- get_v5_corrected(county_testing, params) %>%
+    corrected <- get_v5_county_corrected(county_testing, params) %>%
       mutate(version="v5") }
   
   if(vary=="s_untested_and_beta" & spline == TRUE) {
     message("v6")
-    corrected <- get_v6_corrected(county_testing, params) %>%
+    corrected <- get_v6_county_corrected(county_testing, params) %>%
       mutate(version="v6") }
   
 
@@ -133,7 +134,7 @@ get_corrected_county <- function(data, params, ctis,
 
 #---------- Version 2 -------------
 # center prior for beta at survey estimates 
-get_v2_corrected <- function(county_testing, params) {
+get_v2_county_corrected <- function(county_testing, params) {
   
   message("Running version 2")
   
@@ -186,7 +187,7 @@ get_v2_corrected <- function(county_testing, params) {
 
 #---------- Version 3 -------------
 # center prior for P(S_1|untested) at survey estimates 
-get_v3_corrected <- function(county_testing, params) {
+get_v3_county_corrected <- function(county_testing, params) {
   
   message("Running version 3")
   
[…]

### 2023-06-20 app/app.R
delete unnecessary files

@@ -1,917 +1,0 @@
-
-# Packages ---------------------------------------------------------------
-library(shiny)
-library(tidyverse)
-library(latex2exp)
-library(future)
-library(truncdist)
-library(shinyscreenshot)
-library(shinyjs)
-library(patchwork)
-
-
-# Read Data -----------
-state_name <- "ma"
-covid_county <- "./example_county_data/ma_county_biweekly.RDS" %>%
-  readRDS()  %>%
-  select(-date) %>%
-  distinct() %>%
-  filter(fips == "25025")
-
-# covid_county <- covid_county[1:15,]
-
-corrected_original <- readRDS(
-               "./example_county_data/example_county.RDS")
-
-
-dates <- readRDS( "./example_county_data/date_to_biweek.RDS")
-
-
-
-background <- "<b></i><span style='font-size:120%;' >Background:</b></i> </span> In the implementation of Bayesian melding,
-we sample values of $\\theta = \\{ \\alpha, \\beta, P(S_1|\\text{untested})\\}$, where the sample size here is selected via the 
-<i>Sample Size</i> option.
-Then, we use the function $M: \\theta \\to \\phi$ defined by
-$M(\\theta) = \\dfrac{\\beta(1- P(S_1|\\text{untested}))}{\\beta(1-P(S_1|\\text{untested})) + \\alpha P(S_1|\\text{untested})}$,
-yielding the induced distribution $f_\\phi^{induced}$. Because we do not have an analytical form for $f_\\phi^{induced}$,
-we estimate it using kernel density estimation.
-
-Then, we compute the sampling weights $\\left(\\dfrac{f_\\phi^{direct}(\\phi_i)}{f_{\\phi}^{induced}(\\phi_i)}\\right)^{0.5}$.
-Here, $f_\\phi^{direct}(\\phi_i)$ is calculated using the density function $f_\\phi^{direct}$, which is defined based on
-information from meta-analyses on the asymptomatic rate, and $f_{\\phi}^{induced}(\\phi_i)$ is from the kernel density estimation.
-
-Sampling with these weights from our sample of $M(\\theta) = \\phi$ and our sample of $\\theta$ yields the post-melding distributions,
-where the post-melding density of $\\phi$ is the the logarithmic pool of $f_\\phi^{direct}$ and $f_\\phi^{induced}$, that is,
[…]

### 2023-06-20 app/base_functions.R

@@ -1,319 +1,0 @@
-
-
-
-
-###############################################################
-# BETA PARAMETERS FROM DESIRED MEAN AND VARIANCE
-###############################################################
-get_beta_params <- function(mu, sd){
-  var = sd ^ 2
-  a = ((1 - mu) / var - 1 / mu) * mu ^ 2
-  b = a * (1 / mu - 1)
-  return (list(a = a, b = b))
-}
-
-
-# vectorized version to get each parameter separately
-get_shape1 <- function(mu, sd) {
-  var = sd^2
-  alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
-  beta <- alpha * (1 / mu - 1)
-  return(alpha)
-}
-
-get_shape2 <- function(mu, sd) {
-  var = sd^2
-  alpha <- ((1 - mu) / var - 1 / mu) * mu ^ 2
-  beta <- alpha * (1 / mu - 1)
-  return(beta)
-}
-
-
-
-
-###############################################################
-# BETA DENSITY WITH DESIRED MEAN AND VARIANCE
-###############################################################
-beta_density <- function(x, mean, sd, bounds=NA) {
-  shape_params <-  get_beta_params(
-    mu = mean,
-    sd = sd)
-  
-  if(!length(bounds) == 1){
-    # message("here")
-    dtrunc(x,
-           spec = "beta",
-           a = bounds[1],
-           b = bounds[2],
-           shape1 = shape_params$a,
-           shape2 = shape_params$b) %>%
-      return()
-  }else{
-    dbeta(x,
-          shape1 = shape_params$a,
-          shape2 = shape_params$b)  %>%
-      return()
-  }
-}
-
-
-
-
-###############################################################
-# SAMPLE FROM BETA DENSITY WITH DESIRED MEAN AND VARIANCE
-###############################################################
-
-sample_beta_density <- function(n, mean, sd, bounds = NA) {
-  
-  shape_params <-  get_beta_params(
-    mu = mean,
-    sd = sd)
-  
-  rbeta(n,
-        shape1 = shape_params$a,
-        shape2 = shape_params$b)
-  
-  if(!length(bounds) == 1){
-    # message("here")
-    rtrunc(n,
-           spec = "beta",
-           a = bounds[1],
-           b = bounds[2],
-           shape1 = shape_params$a,
-           shape2 = shape_params$b) %>%
-      return()
-  }else{
-    rbeta(n,
-          shape1 = shape_params$a,
[…]

### 2023-06-20 reports/state_results.Rmd
update state results

---
title: "State Results"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
    toc-depth: 4
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE)

library(tidyverse)
library(targets)
library(here)
library(latex2exp)

```

```{r load-data}

# clear labels for versions
versions <- tibble(
  version = c("v1", "v2", "v3", "v4", "v5", "v6"),
  vlabel = factor(
    c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
    "$\\beta$ Centered at Emp. Value",
    "$P(S_1|untested)$ Centered at Emp. Value",
    "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"
    ),
    levels = c(
        "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
        "$\\beta$ Centered at Emp. Value",
         "$P(S_1|untested)$ Centered at Emp. Value",
        "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"

    ) )
)

# read state-level results
targets_dir <- here("_targets_6_19")

state_res <- tar_read(state_v1, 
                   store =targets_dir) %>% 
  mutate(version = "v1") %>%
  bind_rows(
    tar_read(state_v2,
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(state_v3, 
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(state_v4, 
             store =targets_dir)
   )
  # )%>%
  #  bind_rows(
  #   tar_read(state_v5, 
  #            store =targets_dir)
  # )%>%
  #  bind_rows(
  #   tar_read(state_v6, 
  #            store =targets_dir)
  # )

covidestim <- tar_read(covidestim_biweekly_state,
                        store =targets_dir)

dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
codes <- read_csv(here('data/demographic/statecodes.csv'))

state_res <- state_res %>%
  rename(state=fips) %>%
  left_join(versions) %>%
  mutate(state=toupper(state)) %>%
  left_join(covidestim, relationship= "many-to-many") %>%
  left_join(dates, relationship = "many-to-many")

```

```{r theme_c}

theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
    theme_bw() %+replace%    #replace elements we want to change

    theme(

      #text elements
      plot.title = element_text(             #title
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),

      plot.subtitle = element_text(          #subtitle
                   size = 12,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size

      axis.title = element_text(             #axis titles
                   size = 12),               #font size

      axis.text = element_text(              #axis text
                   size = 9),
      legend.text = element_text(size = 12),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 11, face = "bold", color="white"),
      strip.text.y = element_text(size = 11, face = "bold", color="white",angle=270),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

}

```

# Summarizing Concordance with Covidestim

```{r,fig.width=11}

# corrected %>%
#   mutate(in_interval = case_when(
#     exp_cases_lb <= infections & infections <= exp_cases_ub ~ "In Interval",
#     exp_cases_lb  > infections  ~ "Below Interval",
#     exp_cases_ub < infections ~ "Above Interval"
#   )) %>%
#   filter(!is.na(in_interval)) %>%
#   group_by(in_interval, state, vlabel) %>%
#   summarize(n_per_county=n()) %>%
#   group_by(vlabel, in_interval) %>%
#   summarize(n_per_version = sum(n_per_county)) %>%
#   group_by(vlabel) %>%
#   mutate(total = sum(n_per_version)) %>%
#   ungroup() %>%
#   mutate(prop=n_per_version/total)

by_version <- state_res %>%
  select(-date) %>%
  distinct() %>% 
  mutate(in_interval = case_when(
    exp_cases_lb <= infections & infections <= exp_cases_ub ~ "In Interval",
    exp_cases_lb  > infections  ~ "Below Interval",
    exp_cases_ub < infections ~ "Above Interval"
  )) %>%
  filter(!is.na(in_interval)) %>%
  group_by(in_interval, vlabel) %>%
  summarize(n=n()) %>%
  group_by(vlabel) %>%
  mutate(total = sum(n)) %>%
  mutate(prop=n/total)

labels <- by_version %>%
  arrange(prop) %>%
  pull(vlabel) %>%
  as.character() %>%
  unique()

by_version %>%
  mutate(in_interval = factor(in_interval, 
                             levels = c("Above Interval",
                                        "In Interval",
                                        "Below Interval"
                                        ))) %>%
   ggplot(aes(x= fct_reorder(vlabel,prop,max), 
             fill = in_interval, y =prop)) + 
  geom_bar(stat="identity",position="stack") +
  theme_c() +
  coord_flip()+ scale_fill_manual(values=c("In Interval" = "#79D2AF",
                             "Above Interval" = "#D2AF79",
                      "Below Interval"="#7997D2"),
                       breaks = c("Below Interval",
                               "In Interval",
                               "Above Interval")) +
  labs(x="", 
       y= "Proportion",
       fill = "Covidestim Median:",
       title = "Proportion Where Covidestim Median\nWas Within, Above, or Below the Median",
       subtitle = "Including All States") +
  theme_c() +
#  viridis::scale_fill_viridis(option="mako", begin=.2, discrete=TRUE) +
  theme_c(legend.position="right",
          legend.title = element_text(face="bold", size = 15),
          legend.text= element_text( size = 15),
          legend.spacing.y = unit(3, 'pt'),
          axis.text.y = element_text(size = 17, hjust=1)) +
  theme(
          plot.subtitle = element_text(size=18, 
                                       face='italic', 
                                       margin=margin(4,0,4,0))) +
  scale_x_discrete(labels = (TeX(labels)))

```

# Ratio of Estimated to Observed During Omicron

```{r,results='hide', fig.height=11,fig.width=8}

subset <- state_res %>%
  filter(vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$")

subset %>% 
  left_join(codes) %>%
  filter(biweek == 27) %>%
  pull(date) %>% 
  range()

subset %>% 
  left_join(codes) %>%
  filter(biweek == 27) %>%
  ggplot(aes(y =fct_reorder(state_name,exp_cases_median/positive),
             xmin = exp_cases_lb/positive,
             xmax = exp_cases_ub/positive)) +
  geom_errorbar() +
  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
  #            color = "darkred", size=.8, alpha=.5) +
  theme_c() + 
  theme(axis.text.y=element_text(size=15),
        axis.text.x = element_text(size=13),
        axis.title = element_text(size = 18),
        plot.subtitle = element_text(size=14)) +
  labs(y="", x="Ratio of Estimated Infections to Observed",
       title ="Ratio of Estimated Infections to Observed Infections\nDuring Omicron",
       subtitle = "December 31, 2021 through January 14, 2022")

```

### 2023-06-20 reports/county_correlations.Rmd
update rmd files

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE,cache=TRUE)
 
@@
 library(latex2exp)
+library(kableExtra)
 
@@
 
-targets_dir <- here("_targets_6_19")
+targets_dir <- here("_targets")
 
@@
              store =targets_dir)
-  # )%>%
-  #  bind_rows(
-  #   tar_read(ma_v5, 
-  #            store =targets_dir)
-  # )%>%
-  #  bind_rows(
-  #   tar_read(ma_v6, 
-  #            store =targets_dir)
+  )%>%
+   bind_rows(
+    tar_read(ma_v5,
+             store =targets_dir)
+  )%>%
+   bind_rows(
+    tar_read(ma_v6,
+             store =targets_dir)
    )
@@
 
-```{r, fig.height=10, fig.width=10}
+```{r, fig.height=10, fig.width=12}
 
 cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)
+
+cor_covid <- covidestim_correlations %>% 
+  select(vlabel, 
+         `Correlation with Covidestim Estimates`=Correlation)
+
+cor_waste <- correlations %>% 
+  select(vlabel, 
+         `Correlation with Wastewater Concentrations`=Correlation)
+
+cor_waste %>%
+  left_join(cor_covid) %>%
+  mutate(vlabel = gsub(
+    "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
+    "Priors Do Not Vary by County and Date",
+    vlabel, fixed=TRUE),vlabel=gsub(" at","\nat", vlabel)) %>%
+  arrange(desc( `Correlation with Wastewater Concentrations`)) %>%
+  rename(Implementation = vlabel) %>%
+  kbl() %>%
+  kable_styling(full_width = TRUE)
 

### 2023-06-20 reports/priors.Rmd
update rmd files

@@
 ```{r beta-smoothing, fig.cap="The ratio of the screening test positivity over the overall test positivity from the COVID-19 Trends and Impact Survey is taken to be the estimate of $\\beta$.
-We compare two approaches to smoothing: cubic spline smoothing with 2 knots (July 15th, 2021 and December 1st, 2021) to LOESS smoothing with a span of 0.33."}
+We compare two approaches to smoothing: cubic spline smoothing with 2 knots (July 15th, 2021 and December 1st, 2021) to LOESS smoothing with a span of 0.33.", fig.width=12,fig.height=12}
 
@@
 ```{r s-untested-smoothing, fig.cap="The percentage of the population experiencing COVID-19-like illness is taken to be the estimate of $\\Pr(S_1|\\text{untested}).
-The LOESS smoothed estimate with a span of 0.2 is shown in red."}
+The LOESS smoothed estimate with a span of 0.2 is shown in red.",fig.width=12,fig.height=16}
 
@@
   geom_line() +
-  facet_wrap(~state, ncol=5) +
+  facet_wrap(~state, ncol=4) +
   scale_alpha_manual(values=c("LOESS smoothed" = .9, 

### 2023-06-24 R/05-state-analysis.R
filter out NAs in spline smoothed

@@ -2,7 +2,7 @@
 
 
 #---------- Version 1 -------------
-get_v1_state <- function(data, params, testing = TRUE) {
+get_v1_state <- function(data, params, testing = FALSE) {
   
   state_testing <- data
   
@@ -116,7 +116,7 @@ get_smoothed_ctis <- function(ctis_raw,
                                data = .x, 
                                span = smooth_beta_span)
         # 257 corresponds to 2021-12-01
-        spline_smoothed_beta <- lm(imputed_beta ~ splines::bs(index,knots = c(118, 257)), 
+        spline_smoothed_beta <- lm(imputed_beta ~ splines::ns(index,knots = c(118, 257)), 
                                    data = .x )
         
         # include both spline smoothing and loess smoothing for beta
@@ -145,7 +145,7 @@ get_smoothed_ctis <- function(ctis_raw,
 get_corrected_state <- function(data, params,
                                 ctis, vary, 
                                 spline=FALSE,
-                                testing = TRUE) {
+                                testing = FALSE) {
   
   dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
   
@@ -285,6 +285,11 @@ get_v3_corrected <- function(state_testing, params) {
   
   message("Running version 3")
 
+  
+  state_testing <- state_testing %>% 
+    # only have CTIS data starting at week 6
+    # filter out the beginning dates where beta_estimate_smoothed is NA
+    filter(!is.na(beta_estimate_smoothed))
   
   corrected <- state_testing %>% 
     arrange(biweek) %>%

### 2023-06-24 reports/county_correlations.Rmd
update output files

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE,cache=TRUE)
+knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,cache=TRUE, fig.width=10, fig.height=7)
 
@@
 
-## Normalization Strategy 1: By Values at Biweek Where Wastewater Concentration is Maximized
+## Normalization Strategy 1: Standardization
+
+```{r}
+
+w_data <- wjoined %>% 
+  select(-date) %>%
+  distinct() %>%
+  filter(fips %in% counties_with_waste) %>%
+  filter(!is.na(mean_conc)) %>%
+  group_by(fips, version) %>%
+  mutate(mean_conc_std = (mean_conc - mean(mean_conc))/ 
+           sd(mean_conc),
+         exp_cases_median_std = (exp_cases_median -
+                                   mean(exp_cases_median)/sd(exp_cases_median)),
+         infections_std = (infections - mean(infections))/sd(infections),
+         exp_cases_lb_std = (exp_cases_lb - mean(exp_cases_lb)/ sd(exp_cases_lb)),
+        exp_cases_ub_std = (exp_cases_ub - mean(exp_cases_ub)/ sd(exp_cases_ub)),
+         positive_std = (positive - mean(positive))/sd(positive))  %>%
+  ungroup()
+
+```
+
+### PB Counts and Wastewater
+
+```{r}
+
+# correlations between PB counts and wastewater concentrations
+
+correlations <- w_data %>%
+  group_by(version) %>%
+  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
+            vlabel =unique(vlabel)) %>%
+  mutate(x=1.8,
+         y=.1)
+
+w_data %>%
+  group_by(version) %>%
+  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = mean_conc_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= correlations ) +
+  facet_wrap(~vlabel,
+             ncol=2,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Mean Wastewater Concentration\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")
+
+plt_wastewater_pb <- w_data %>%
+  group_by(version) %>%
+  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = mean_conc_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= correlations ) +
+  facet_wrap(~vlabel,
+             ncol=1,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Mean Wastewater Concentration\n(Standardized)",
+       title = "")
+
+```
+
+### PB Counts and Covidestim
+
+```{r}
+
+# correlations between PB counts and Covidestim estimates
+
+covidestim_correlations <- w_data %>%
+  filter(!is.na(infections_std)) %>%
+  group_by(vlabel) %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(exp_cases_median_std,infections_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1.8,
+         y=.1)
+
+w_data %>%
+  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = infections_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  facet_wrap(~vlabel,
+              ncol=2,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Covidestim Estimates\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+
+plt_covidestim_pb <- w_data %>%
+  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = infections_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  facet_wrap(~vlabel,
+              ncol=1,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized)",
+       y = "Covidestim Estimates\n(Standardized)",
+       title = "")
+
+```
+
+### Combined Figure
+
+```{r, fig.height=10, fig.width=12}
+
+cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)
+
+cor_covid <- covidestim_correlations %>% 
+  select(vlabel, 
+         `Correlation with Covidestim Estimates`=Correlation)
+
+cor_waste <- correlations %>% 
+  select(vlabel, 
+         `Correlation with Wastewater Concentrations`=Correlation)
+
+cor_waste %>%
+  left_join(cor_covid) %>%
+  mutate(vlabel = gsub(
+    "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
+    "Priors Do Not Vary by County and Date",
+    vlabel, fixed=TRUE),vlabel=gsub(" at","\nat", vlabel)) %>%
+  arrange(desc( `Correlation with Wastewater Concentrations`)) %>%
+  rename(Implementation = vlabel) %>%
+  kbl() %>%
+  kable_styling(full_width = TRUE)
+
+```
+
+### Covidestim and Wastewater
+
+```{r}
+
+covidestim_correlations <- w_data %>%
+  filter(version=="v1") %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(mean_conc_std,infections_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1,
+         y=.1)
+
+w_data %>%
+  filter(version=="v1") %>%
+  ggplot(aes(x=infections_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  theme_c() +
+  labs(x = "Covidestim Estimates\n(Standardized)",
+       y = "Wastewater Concentrations\n(Standardized)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+
+pb_correlations <- w_data %>%
+  filter(version=="v1") %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(positive_std,mean_conc_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1.8,
+         y=.1)
+
+```
+
+### Compare to Wastewater Over Time
+
+```{r,fig.width=12, fig.height=15}
+
+library(scales)
+
+pal <- c("#10BAC5", "#1B10C5", "#EFB719", "#900C3F")
+
+names(pal) <- c(#"$observed$", 
+                '$P(S_1|untested)$ Centered at Emp. Value',
+                '$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values',
+               "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$",
+               "$\\beta$ Centered at Emp. Value"
+               )
+
+waste_df <- w_data %>%
+  select(biweek, mean_conc_std, name) %>%
+  distinct() %>%
+  left_join(dates,relationship = "many-to-many" ) %>%
+  group_by(name, biweek) %>%
+  # center date within 2-week interval
+  summarize(date = min(date) + days(7),
+            mean_conc_std=unique(mean_conc_std)) %>%
+  mutate(name = gsub("County, MA", "", name))
+
+w_data%>%
+  mutate(name = gsub("County, MA", "", name)) %>%
+  left_join(dates,relationship = "many-to-many" ) %>%
+  filter(version %in% c("v1","v3")) %>%
+  filter(biweek>=6) %>%
+  group_by(biweek) %>%
+  mutate(mindate = min(date),
+         maxdate=max(date)) %>%
+  ungroup() %>% 
+ #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
+    ggplot() +
+      geom_ribbon(aes(x = date, 
+                 ymin = exp_cases_lb_std,
+                 ymax = exp_cases_ub_std,
+                 fill = vlabel),
+                 alpha = .5) +
+      geom_line(data = waste_df,
+                aes(x = date, y =mean_conc_std, 
+                    color = 'Wastewater Effective Concentrations'),
+               # color = "#DB4048",
+                size = 1.1) +
+      geom_linerange(aes(xmin = mindate, 
+                         xmax = maxdate, 
+                         y = infections_std,
+                         color = 'Covidestim Infection Estimates')) + 
+      geom_point(data = waste_df,
+                aes(x = date, y =mean_conc_std ),
+                color = "#DB4048",
+                alpha = .5,
+                size = 1.2) +
+      facet_grid(name~vlabel, 
+                 labeller=labeller(vlabel =as_labeller(
+                   TeX, default=label_parsed))) +
+      scale_fill_manual(values = pal, labels = c('','','',''), 
+                        name = "Probabilistic Bias Intervals") +
+      theme_bw() +
+     theme_c() +
+      labs(y = "Value (Standardized)",
+           title = "",
+           x= "") +
+      scale_x_date(date_labels = "%b %Y") +
+    scale_color_manual(
+      name = '', 
+      values = c(
+        'Covidestim Infection Estimates' = 'darkblue',
+        'Wastewater Effective Concentrations' = '#DB4048')) +
+    guides(color = guide_legend(override.aes = list(size = 12,
+                                                    linewidth=3)),
+           fill = guide_legend(override.aes =list(size = 6)))
+
+```
+
+## Normalization Strategy 2: By Values at Biweek Where Wastewater Concentration is Maximized
 
@@
 
-## Normalization Strategy 2: By Maximum (Separately)
+## Normalization Strategy 3: By Maximum (Separately)
 

### 2023-06-24 writing/manuscript.Rmd
initialize writing rmds

---
title: "Approximating the Number of COVID-19 Infections in the United States Across 2021"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: false
    number_sections: false

---

\newpage

# Abstract

# Introduction

# Results

## County Level in Massachusetts

## State-level (1 version)

# Methods

## Data

### Massachusetts County Level
### State Level
### Survey Data

## Statistical Methods

### Probabilistic bias analysis

### Bayesian Melding

### Specification of Priors

### 2023-06-24 writing/supplement.Rmd
initialize writing rmds

---
title: "Supplement"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  bookdown::pdf_document2:
    toc: true
    number_sections: false

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
library(knitr)
library(here)

```

# Specification of the Priors

```{r ctis-by-state, fig.cap ="Since $\\beta$ represents the ratio of $\\Pr(\\text{test}_+|\\text{untested}, S_0)$ to $\\Pr(\\text{test}_+|\\text{tested})$, we estimate $\\beta$ from the COVID-19 Trends and Impact Survey as the ratio of the screening test positivity over the overall test positivity. Here, we consider the 2.5\\% and 97.5\\% percentiles and mean for each state. Most have a mean near $20%$."}

include_graphics(here("figures/emp-estimate-beta-dist-by-state.pdf"))

```

# Sensitivity Analyses

### 2023-06-24 reports/county_correlations.Rmd
update county results

@@
        y = "Wastewater Concentrations\n(Standardized)",
-       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+       title = "Correlations Between Wastewater Concentrations\nand Covidestim Estimates")
 

### 2023-06-24 reports/priors.Rmd
add priors figures on beta distributions by state

@@
 library(latex2exp)
+library(viridis)
 
@@
 ```
+
+## Distributions of $\beta$ by State
+
+```{r}
+
+# raw ctis data
+
+ctis_raw <- readRDS(here('data/data_raw/ctis_all_states.RDS'))
+
+states_keep <- ctis_smoothed  %>%
+  filter(keep) %>% pull(state) %>% unique()
+
+ctis <- ctis_raw %>%
+  select(signal, state=geo_value, date, value) %>%
+  pivot_wider(names_from=signal, values_from=value) %>%
+  filter(state %in% states_keep) %>%
+  mutate(beta=smoothed_wscreening_tested_positive_14d/
+           smoothed_wtested_positive_14d,
+         state=toupper(state))
+
+```
+
+```{r ctis-by-state, fig.cap ="Since $\\beta$ represents the ratio of $\\Pr(\\text{test}_+|\\text{untested}, S_0)$ to $\\Pr(\\text{test}_+|\\text{tested})$, we estimate $\\beta$ from the COVID-19 Trends and Impact Survey as the ratio of the screening test positivity over the overall test positivity.
+Here, we consider the 2.5\\% and 97.5\\% percentiles and mean for each state.
+Most have a mean near $20%$.", fig.width=9}
+
+ctis %>%
+  group_by(state) %>%
+  summarize(median = median(beta, na.rm=TRUE),
+            mean = mean(beta, na.rm=TRUE),
+            q1 = quantile(beta, .025,  na.rm=TRUE),
+            q2 = quantile(beta, .975,  na.rm=TRUE)) %>%
+  ggplot(aes(x= fct_reorder(state, mean, .desc=TRUE),
+             y = mean)) +
+  geom_errorbar(aes(ymin=q1,ymax=q2),width=.2) +
+  geom_point(color="darkred", size=2) +
+  scale_y_continuous(n.breaks=10) +
+  labs(y= TeX("Empirical Estimates of $\\beta$"),
+       x="State",
+       title = TeX("Empirical Estimates of $\\beta$ from the COVID-19 Trends and Impact Survey"),
+       subtitle = "2.7% Percentile, 97.5% Percentile, and Mean By State" ) +
+  theme_c(axis.title=element_text(size=12))
+
+ggsave(here("figures/emp-estimate-beta-dist-by-state.pdf"))
+
+```
+
+```{r}
+
+ctis %>%
+  ggplot(aes(x = beta, fill=state)) +
+  geom_density() +
+  scale_fill_viridis(option="mako", discrete=TRUE) +
+  theme_c()
+
+ctis %>%
+  ggplot(aes(x = beta, fill=state)) +
+  geom_density() +
+  scale_fill_viridis(option="mako", discrete=TRUE) +
+  theme_c() +
+  facet_wrap(~state)
+
+```

### 2023-06-25 reports/county_figures.Rmd
rename

---
title: "County Figures"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: hide
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
    toc-depth: 4
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,cache=TRUE, fig.width=10, fig.height=7)

library(tidyverse)
library(targets)
library(here)
library(latex2exp)
library(kableExtra)

```

```{r theme_c}

theme_c <- function(...){ 
   # font <- "Helvetica"   #assign font family up front
    theme_bw() %+replace%    #replace elements we want to change

    theme(

      #text elements
      plot.title = element_text(             #title
                   size = 16,                #set font size
                   face = 'bold',            #bold typeface
                   hjust = .5,
                   vjust = 3),

      plot.subtitle = element_text(          #subtitle
                   size = 12,
                   hjust = .5,
                   face = 'italic',
                   vjust = 3),               #font size

      axis.title = element_text(             #axis titles
                   size = 12),               #font size

      axis.text = element_text(              #axis text
                   size = 9),
      legend.text = element_text(size = 12),
      # t, r, b, l
      plot.margin = unit(c(1,.5,.5,.5), "cm"),
      legend.position = "right",
      strip.text.x = element_text(size = 11, face = "bold", color="white"),
      strip.text.y = element_text(size = 11, face = "bold", color="white",angle=270),
      strip.background = element_rect(fill = "#3E3D3D")
      ) %+replace%
      theme(...)

}

```

```{r load-data}

targets_dir <- here("_targets")

county <- tar_read(ma_v1, 
                   store =targets_dir) %>% 
  mutate(version = "v1") %>%
  bind_rows(
    tar_read(ma_v2,
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(ma_v3, 
             store =targets_dir)
  ) %>%
   bind_rows(
    tar_read(ma_v4, 
             store =targets_dir)
  )%>%
   bind_rows(
    tar_read(ma_v5,
             store =targets_dir)
  )%>%
   bind_rows(
    tar_read(ma_v6,
             store =targets_dir)
   )

waste <- tar_read(waste, store =targets_dir)
covidestim <- tar_read(covidestim_biweekly_county,
                        store =targets_dir)

dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))

```

```{r}

wjoined <- county %>%
  # set Nantucket, Duke fips to Nantucket since we have wastewater data
  # for Nantucket
  mutate(fips = ifelse(grepl("25019", fips), "25019", fips)) %>%
  inner_join(waste) %>%
  left_join(covidestim, relationship = "many-to-many") %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate = max(date)) %>%
  ungroup()

counties_with_waste <- wjoined %>%
  group_by(fips) %>%
  mutate(obs_w_notna = sum(!is.na(mean_conc))) %>%
  filter(obs_w_notna != 0) %>%
  pull(fips) %>%
  unique()

versions <- tibble(
  version = c("v1", "v2", "v3", "v4", "v5", "v6"),
  vlabel = factor(
    c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
    "$\\beta$ Centered at Emp. Value",
    "$P(S_1|untested)$ Centered at Emp. Value",
    "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"
    ),
    levels = c(
        "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
        "$\\beta$ Centered at Emp. Value",
         "$P(S_1|untested)$ Centered at Emp. Value",
        "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"

    ) )
)

wjoined <- wjoined %>%
  left_join(versions)

w_data <- wjoined %>% 
  select(-date) %>%
  distinct() %>%
  filter(fips %in% counties_with_waste) %>%
  filter(!is.na(mean_conc)) %>%
  group_by(fips, version) %>%
  mutate(mean_conc_std = mean_conc/ max(mean_conc),
         exp_cases_median_std = exp_cases_median / max(exp_cases_median),
         exp_cases_lb_std = exp_cases_lb /max(exp_cases_median),
         exp_cases_ub_std = exp_cases_ub / max(exp_cases_median),
         infections_std  = infections/max(infections),
         positive_std = positive/max(positive)) %>%
  ungroup()

```

# Correlations

## Normalization Strategy 1: Standardization

```{r}

w_data <- wjoined %>% 
  select(-date) %>%
  distinct() %>%
  filter(fips %in% counties_with_waste) %>%
  filter(!is.na(mean_conc)) %>%
  group_by(fips, version) %>%
  mutate(mean_conc_std = (mean_conc - mean(mean_conc))/ 
           sd(mean_conc),
         exp_cases_median_std = (exp_cases_median -
                                   mean(exp_cases_median)/sd(exp_cases_median)),
         infections_std = (infections - mean(infections))/sd(infections),
         exp_cases_lb_std = (exp_cases_lb - mean(exp_cases_lb)/ sd(exp_cases_lb)),
        exp_cases_ub_std = (exp_cases_ub - mean(exp_cases_ub)/ sd(exp_cases_ub)),
         positive_std = (positive - mean(positive))/sd(positive))  %>%
  ungroup()

```

### PB Counts and Wastewater

```{r}

# correlations between PB counts and wastewater concentrations

correlations <- w_data %>%
  group_by(version) %>%
  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
            vlabel =unique(vlabel)) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")

plt_wastewater_pb <- w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County)",
       title = "")

```

### PB Counts and Covidestim

```{r}

# correlations between PB counts and Covidestim estimates

covidestim_correlations <- w_data %>%
  filter(!is.na(infections_std)) %>%
  group_by(vlabel) %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(exp_cases_median_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Covidestim Estimates\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

plt_covidestim_pb <- w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Covidestim Estimates\n(Standardized by County)",
       title = "")

```

### Combined Figure

```{r, fig.height=10, fig.width=12}

cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)

cor_covid <- covidestim_correlations %>% 
  select(vlabel, 
         `Correlation with Covidestim Estimates`=Correlation)

cor_waste <- correlations %>% 
  select(vlabel, 
         `Correlation with Wastewater Concentrations`=Correlation)

cor_waste %>%
  left_join(cor_covid) %>%
  mutate(vlabel = gsub(
    "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
    "Priors Do Not Vary by County and Date",
    vlabel, fixed=TRUE),vlabel=gsub(" at","\nat", vlabel)) %>%
  arrange(desc( `Correlation with Wastewater Concentrations`)) %>%
  rename(Implementation = vlabel) %>%
  kbl() %>%
  kable_styling(full_width = TRUE)

```

### Covidestim and Wastewater

```{r}

covidestim_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(mean_conc_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1,
         y=.1)

w_data %>%
  filter(version=="v1") %>%
  ggplot(aes(x=infections_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  theme_c() +
  labs(x = "Covidestim Estimates\n(Standardized by County)",
       y = "Wastewater Concentrations\n(Standardized by County)",
       title = "Correlations Between Wastewater Concentrations\nand Covidestim Estimates")

pb_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(positive_std,mean_conc_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

```

### Compare to Wastewater Over Time

```{r,fig.width=12, fig.height=15}

library(scales)

pal <- c("#10BAC5", "#1B10C5", "#EFB719", "#900C3F")

names(pal) <- c(#"$observed$", 
                '$P(S_1|untested)$ Centered at Emp. Value',
                '$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values',
               "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$",
               "$\\beta$ Centered at Emp. Value"
               )

waste_df <- w_data %>%
  select(biweek, mean_conc_std, name) %>%
  distinct() %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  group_by(name, biweek) %>%
  # center date within 2-week interval
  summarize(date = min(date) + days(7),
            mean_conc_std=unique(mean_conc_std)) %>%
  mutate(name = gsub("County, MA", "", name))

w_data%>%
  mutate(name = gsub("County, MA", "", name)) %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  filter(version %in% c("v1","v3")) %>%
  filter(biweek>=6) %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate=max(date)) %>%
  ungroup() %>% 
 #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
    ggplot() +
      geom_ribbon(aes(x = date, 
                 ymin = exp_cases_lb_std,
                 ymax = exp_cases_ub_std,
                 fill = vlabel),
                 alpha = .5) +
      geom_line(data = waste_df,
                aes(x = date, y =mean_conc_std, 
                    color = 'Wastewater Effective Concentrations'),
               # color = "#DB4048",
                size = 1.1) +
      geom_linerange(aes(xmin = mindate, 
                         xmax = maxdate, 
                         y = infections_std,
                         color = 'Covidestim Infection Estimates')) + 
      geom_point(data = waste_df,
                aes(x = date, y =mean_conc_std ),
                color = "#DB4048",
                alpha = .5,
                size = 1.2) +
      facet_grid(name~vlabel, 
                 labeller=labeller(vlabel =as_labeller(
                   TeX, default=label_parsed))) +
      scale_fill_manual(values = pal, labels = c('','','',''), 
                        name = "Probabilistic Bias Intervals") +
      theme_bw() +
     theme_c() +
      labs(y = "Value (Standardized by County)",
           title = "",
           x= "") +
      scale_x_date(date_labels = "%b %Y") +
    scale_color_manual(
      name = '', 
      values = c(
        'Covidestim Infection Estimates' = 'darkblue',
        'Wastewater Effective Concentrations' = '#DB4048')) +
    guides(color = guide_legend(override.aes = list(size = 12,
                                                    linewidth=3)),
           fill = guide_legend(override.aes =list(size = 6)))

```

## Normalization Strategy 2: By Values at Biweek Where Wastewater Concentration is Maximized

```{r}

w_data <- wjoined %>% 
  select(-date) %>%
  distinct() %>%
  filter(fips %in% counties_with_waste) %>%
  filter(!is.na(mean_conc)) %>%
  group_by(fips, version) %>%
  mutate(mean_conc_std = mean_conc/ max(mean_conc),
         max_conc = max(mean_conc),
         exp_cases_median_std = exp_cases_median / exp_cases_median[which.max(mean_conc)],
         exp_cases_lb_std = exp_cases_lb /exp_cases_median[which.max(mean_conc)],
         exp_cases_ub_std = exp_cases_ub / exp_cases_median[which.max(mean_conc)],
         infections_std  = infections/ infections[which.max(mean_conc)],
         positive_std = positive/max(positive)) %>%
  ungroup()

```

### PB Counts and Wastewater

```{r}

# correlations between PB counts and wastewater concentrations

correlations <- w_data %>%
  group_by(version) %>%
  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
            vlabel =unique(vlabel)) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")

plt_wastewater_pb <- w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County by County)",
       title = "")

```

### PB Counts and Covidestim

```{r}

# correlations between PB counts and Covidestim estimates

covidestim_correlations <- w_data %>%
  filter(!is.na(infections_std)) %>%
  group_by(vlabel) %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(exp_cases_median_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
       y = "Covidestim Estimates\n(Standardized by County by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

plt_covidestim_pb <- w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
       y = "Covidestim Estimates\n(Standardized by County by County)",
       title = "")

```

### Combined Figure

```{r, fig.height=10, fig.width=12}

cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)

cor_covid <- covidestim_correlations %>% 
  select(vlabel, 
         `Correlation with Covidestim Estimates`=Correlation)

cor_waste <- correlations %>% 
  select(vlabel, 
         `Correlation with Wastewater Concentrations`=Correlation)

cor_waste %>%
  left_join(cor_covid) %>%
  mutate(vlabel = gsub(
    "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
    "Priors Do Not Vary by County and Date",
    vlabel, fixed=TRUE),vlabel=gsub(" at","\nat", vlabel)) %>%
  arrange(desc( `Correlation with Wastewater Concentrations`)) %>%
  rename(Implementation = vlabel) %>%
  kbl() %>%
  kable_styling(full_width = TRUE)

```

### Covidestim and Wastewater

```{r}

covidestim_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(mean_conc_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1,
         y=.1)

w_data %>%
  filter(version=="v1") %>%
  ggplot(aes(x=infections_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  theme_c() +
  labs(x = "Covidestim Estimates\n(Standardized by County by County)",
       y = "Wastewater Concentrations\n(Standardized by County by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

pb_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(positive_std,mean_conc_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

```

### Compare to Wastewater Over Time

```{r,fig.width=12, fig.height=15}

library(scales)

pal <- c("#10BAC5", "#1B10C5", "#EFB719", "#900C3F")

names(pal) <- c(#"$observed$", 
                '$P(S_1|untested)$ Centered at Emp. Value',
                '$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values',
               "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$",
               "$\\beta$ Centered at Emp. Value"
               )

waste_df <- w_data %>%
  select(biweek, mean_conc_std, name) %>%
  distinct() %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  group_by(name, biweek) %>%
  # center date within 2-week interval
  summarize(date = min(date) + days(7),
            mean_conc_std=unique(mean_conc_std)) %>%
  mutate(name = gsub("County, MA", "", name))

w_data%>%
  mutate(name = gsub("County, MA", "", name)) %>%
  left_join(dates,relationship = "many-to-many" ) %>%
  filter(version %in% c("v1","v3")) %>%
  filter(biweek>=6) %>%
  group_by(biweek) %>%
  mutate(mindate = min(date),
         maxdate=max(date)) %>%
  ungroup() %>% 
 #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
    ggplot() +
      geom_ribbon(aes(x = date, 
                 ymin = exp_cases_lb_std,
                 ymax = exp_cases_ub_std,
                 fill = vlabel),
                 alpha = .5) +
      geom_line(data = waste_df,
                aes(x = date, y =mean_conc_std, 
                    color = 'Wastewater Effective Concentrations'),
               # color = "#DB4048",
                size = 1.1) +
      geom_linerange(aes(xmin = mindate, 
                         xmax = maxdate, 
                         y = infections_std,
                         color = 'Covidestim Infection Estimates')) + 
      geom_point(data = waste_df,
                aes(x = date, y =mean_conc_std ),
                color = "#DB4048",
                alpha = .5,
                size = 1.2) +
      facet_grid(name~vlabel, 
                 labeller=labeller(vlabel =as_labeller(
                   TeX, default=label_parsed))) +
      scale_fill_manual(values = pal, labels = c('','','',''), 
                        name = "Probabilistic Bias Intervals") +
      theme_bw() +
     theme_c() +
      labs(y = "Value (Standardized by County by County)",
           title = "",
           x= "") +
      scale_x_date(date_labels = "%b %Y") +
    scale_color_manual(
      name = '', 
      values = c(
        'Covidestim Infection Estimates' = 'darkblue',
        'Wastewater Effective Concentrations' = '#DB4048')) +
    guides(color = guide_legend(override.aes = list(size = 12,
                                                    linewidth=3)),
           fill = guide_legend(override.aes =list(size = 6)))

```

## Normalization Strategy 3: By Maximum (Separately)

For each county, normalize the wastewater concentration by the maximum concentration for that county and the probabilistic bias counts by the maximum *median* estimated counts for that county (2.5th percentile and 97.5th percentile also standardized by the maximum *median* estimated counts for that county).

### PB Counts and Wastewater

```{r}

# correlations between PB counts and wastewater concentrations

correlations <- w_data %>%
  group_by(version) %>%
  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
            vlabel =unique(vlabel)) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")

plt_wastewater_pb <- w_data %>%
  group_by(version) %>%
  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = mean_conc_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= correlations ) +
  facet_wrap(~vlabel,
             ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Mean Wastewater Concentration\n(Standardized by County)",
       title = "")

```

### PB Counts and Covidestim

```{r}

# correlations between PB counts and Covidestim estimates

covidestim_correlations <- w_data %>%
  filter(!is.na(infections_std)) %>%
  group_by(vlabel) %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(exp_cases_median_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1.8,
         y=.1)

w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
             ncol=2,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Covidestim Estimates\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

plt_covidestim_pb <- w_data %>%
  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
  geom_point(alpha=.5) +
  geom_linerange(aes(xmin = exp_cases_lb_std,
                     xmax=exp_cases_ub_std, 
                     y = infections_std),
                 alpha = .3) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  facet_wrap(~vlabel,
              ncol=1,
             labeller= as_labeller(TeX, 
                                   default = label_parsed)) +
  theme_c() +
  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
       y = "Covidestim Estimates\n(Standardized by County)",
       title = "")
```

### Combined Figure

```{r, fig.height=10, fig.width=10}

cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)

```

### Covidestim and Wastewater

```{r}

covidestim_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(mean_conc_std,infections_std,
                              use = "complete.obs")) %>%
  mutate(x=1,
         y=.1)

w_data %>%
  filter(version=="v1") %>%
  ggplot(aes(x=infections_std, y = mean_conc_std)) +
  geom_point(alpha=.5) +
  geom_text(
           aes(x=x,y=y,  label=paste0("Correlation: ", 
                                      round(Correlation,3))), 
                data= covidestim_correlations ) +
  theme_c() +
  labs(x = "Covidestim Estimates\n(Standardized by County)",
       y = "Wastewater Concentrations\n(Standardized by County)",
       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")

pb_correlations <- w_data %>%
  filter(version=="v1") %>%
  # missing values are handled by casewise deletion
  summarize(Correlation = cor(positive_std,mean_conc_std,
                              use = "complete.obs")) %>%
  mutate(x=1,
         y=.1)

```

# Concordance with Covidestim

```{r}

ma_concordance <- wjoined %>%
  filter(biweek >=6) %>%
  # covidestim does not report estimates for Nantucket 
  filter(!grepl("Nantucket", name )) %>%
  select(-date) %>%
  distinct() %>%
  select(exp_cases_lb, exp_cases_ub, name, biweek, vlabel, infections) %>%
  mutate(in_interval = case_when(
    exp_cases_lb <= infections & infections <= exp_cases_ub ~ "In Interval",
    exp_cases_lb  > infections  ~ "Below Interval",
    exp_cases_ub < infections ~ "Above Interval"
  ) ) %>%
  group_by(in_interval, name, vlabel) %>%
  summarize(n_per_county=n()) %>%
  group_by(vlabel, in_interval) %>%
  summarize(n_per_version = sum(n_per_county)) %>%
  group_by(vlabel) %>%
  mutate(total = sum(n_per_version)) %>%
  ungroup() %>%
  mutate(prop=n_per_version/total)

```

## Concordance by County

```{r}

ma_concordance_by_county %>%
  filter(vlabel=="$P(S_1|untested)$ Centered at Emp. Value")

##############################################
# stacked bar chart of concordance by county
##############################################
ma_concordance_by_county %>%
  filter(vlabel=="$P(S_1|untested)$ Centered at Emp. Value") %>%
  group_by(county_name) %>%
  mutate(m = prop[which(in_interval=="In Interval")]) %>%
  ungroup() %>%
  mutate(in_interval = factor(in_interval, levels = c("Above Interval",
                            "In Interval", "Below Interval"))) %>%
  ggplot(aes(x= fct_reorder(county_name,m), 
             fill = (in_interval), y =prop)) + 
  geom_bar(stat="identity",position="stack") +
  theme_c() +
  coord_flip()+
  scale_fill_manual(values=c("In Interval" = "#79D2AF",
                             "Above Interval" = "#D2AF79",
                      "Below Interval"="#7997D2")) +
  labs(x="", 
       y= "Proportion",
       fill = "Covidestim Median:",
       title = "Massachusetts: Proportion Where Covidestim Median\nWas Within, Above, or Below the Median",
       subtitle = TeX("For Implementation with $P(S_1|untested)$ Centered at Survey Value")) +
  theme_c() +
  theme(legend.position="right",
          legend.title = element_text(face="bold", size = 15),
          legend.text= element_text( size = 15),
          legend.spacing.y = unit(3, 'pt'),
          axis.text.y = element_text(size = 17, hjust=1))

```

## Concordance by Version

```{r}

in_interval <- ma_concordance %>%
  filter(in_interval == "In Interval") %>%
  mutate(n_interval = n_per_version) %>%
  select(vlabel, n_interval)

labels <- ma_concordance %>%
  filter(in_interval=="In Interval") %>%
  arrange((prop)) %>%
  pull(vlabel) %>%
  unique() %>%
  as.character()

#########################################################
# CONCORDANCE BY VERSION 
#########################################################
ma_concordance %>%
  left_join(in_interval) %>%
  mutate(in_interval=factor(in_interval, 
                            levels=c("Above Interval",
                                      "In Interval",
                                     "Below Interval"))) %>%
  ggplot(aes(y =fct_reorder(vlabel,n_interval),
             x = prop, 
             fill=in_interval)) +
  #facet_wrap(~fct_reorder((vlabel),m), ncol=1) +
  geom_bar(stat="identity",position="stack", color="darkgray", linewidth=.2) +
#  viridis::scale_fill_viridis(option="mako", begin=.2, discrete=TRUE) +
  theme_c(legend.position="right",
          legend.title = element_text(face="bold", size = 15),
          legend.text= element_text( size = 15),
          legend.spacing.y = unit(3, 'pt'),
          axis.text.y = element_text(size = 17, hjust=1)) +
  theme(
          plot.subtitle = element_text(size=18, face='italic', 
                                       margin=margin(4,0,4,0))) +
  scale_y_discrete(labels=unname(TeX(labels))) +
  labs(y ="",
       x="Proportion",
       fill = "Covidestim Median:",
       title = "Where Covidestim Medians Fall\nRelative to Probabilistic Bias Intervals",
       subtitle = "By Version, in Massachusetts") +
  scale_fill_manual(values=c("Below Interval"="#7997D2",
                             "In Interval" = "#79D2AF",
                             "Above Interval" = "#D2AF79"),
                     breaks = c("Below Interval",
                               "In Interval",
                               "Above Interval")) +
  guides(fill=guide_legend(byrow=TRUE)) +
  scale_x_continuous(n.breaks = 7,
                     expand=c(0,0),
                     limits=c(-.05,1))

```

### 2023-06-25 reports/priors.Rmd
update figures

@@
 
+```{r, fig.height=7}
+
+ctis %>%
+  group_by(state) %>%
+  summarize(median = median(beta, na.rm=TRUE),
+            mean = mean(beta, na.rm=TRUE),
+            q1 = quantile(beta, .025,  na.rm=TRUE),
+            q2 = quantile(beta, .975,  na.rm=TRUE)) %>%
+  ggplot(aes(y=fct_reorder(state, mean, .desc=TRUE))) +
+  ggridges::geom_density_ridges(
+    aes(x=beta,
+        y=fct_reorder(state, beta, .fun =mean)),
+    data=ctis,
+    fill="#596281",
+    quantile_lines=TRUE,
+    quantiles = c(.025,.975),
+    quantile_fun=function(beta,...) c(
+                                    #  quantile(beta,.025, na.rm=TRUE),
+                                      mean(beta,na.rm=TRUE)
+                                    #  quantile(beta,.975, na.rm=TRUE)
+                                      )) +
+  # geom_errorbar(aes( xmin=q1,xmax=q2),width=.2) +
+  scale_x_continuous(n.breaks=10, limits=c(0,1)) +
+  labs(x= TeX("Empirical Estimates of $\\beta$"),
+       y="State",
+       title = TeX("Empirical Estimates of $\\beta$ from the COVID-19 Trends and Impact Survey"),
+       subtitle = "Density Estimate of Distribution and Mean by State" ) +
+  theme_c(axis.title=element_text(size=12),
+          axis.text.x=element_text(size=11))
+
+```
+
 ```{r}
@@
   ggplot(aes(x = beta, fill=state)) +
-  geom_density() +
+  geom_density(alpha=.6) +
   scale_fill_viridis(option="mako", discrete=TRUE) +
@@
 
+```
+
+```{r,include=FALSE}
 ctis %>%

### 2023-06-25 reports/county_figures.Rmd
concordance w/ covidestim figure

@@
   group_by(fips, version) %>%
-  mutate(mean_conc_std = (mean_conc - mean(mean_conc))/ 
+  mutate(
+    mean_conc_std = (mean_conc - mean(mean_conc))/ 
            sd(mean_conc),
-         exp_cases_median_std = (exp_cases_median -
-                                   mean(exp_cases_median)/sd(exp_cases_median)),
-         infections_std = (infections - mean(infections))/sd(infections),
-         exp_cases_lb_std = (exp_cases_lb - mean(exp_cases_lb)/ sd(exp_cases_lb)),
-        exp_cases_ub_std = (exp_cases_ub - mean(exp_cases_ub)/ sd(exp_cases_ub)),
-         positive_std = (positive - mean(positive))/sd(positive))  %>%
+    exp_cases_median_std = (exp_cases_median -  mean(exp_cases_median))/
+                                   sd(exp_cases_median),
+         infections_std = (infections - mean(infections))/
+      sd(infections),
+         exp_cases_lb_std = (exp_cases_lb - mean(exp_cases_lb))/
+      sd(exp_cases_lb),
+        exp_cases_ub_std = (exp_cases_ub - mean(exp_cases_ub))/
+      sd(exp_cases_ub),
+         positive_std = (positive - mean(positive))
+    /sd(positive))  %>%
   ungroup()
@@
   geom_text(
-           aes(x=x,y=y,  label=paste0("Correlation: ", 
+           aes(x=x,y=y,  
+               label=paste0("Correlation: ", 
                                       round(Correlation,3))), 
@@
                               use = "complete.obs")) %>%
-  mutate(x=1.8,
+  mutate(x=2.5,
          y=.1)
@@
 
+wjoined <- wjoined %>%
+  mutate(name = gsub("County, MA","", name))
+
 ma_concordance <- wjoined %>%
@@
   distinct() %>%
-  select(exp_cases_lb, exp_cases_ub, name, biweek, vlabel, infections) %>%
+  select(exp_cases_lb, exp_cases_ub,
+         name, biweek, vlabel, infections) %>%
   mutate(in_interval = case_when(
-    exp_cases_lb <= infections & infections <= exp_cases_ub ~ "In Interval",
+    exp_cases_lb <= infections & 
+      infections <= exp_cases_ub ~ "In Interval",
     exp_cases_lb  > infections  ~ "Below Interval",
@@
 
+ma_concordance_by_county <- wjoined %>%
+  filter(biweek >=6) %>%
+  # covidestim does not report estimates for Nantucket 
+  filter(!grepl("Nantucket", name )) %>%
+  select(-date) %>%
+  mutate(name = gsub("$", "", 
+                            name, fixed=TRUE)) %>%
+  select(exp_cases_lb, exp_cases_ub, 
+         name, biweek, vlabel, infections) %>%
+   distinct() %>%
+  mutate(in_interval = case_when(
+    exp_cases_lb <= infections & 
+      infections <= exp_cases_ub ~ "In Interval",
+    exp_cases_lb  > infections  ~ "Below Interval",
+    exp_cases_ub < infections ~ "Above Interval"
+  ) ) %>%
+  group_by(in_interval, name, vlabel) %>%
+  summarize(n_per_county=n()) %>%
+  group_by(name,vlabel) %>%
+  mutate(total = sum(n_per_county),
+         prop = n_per_county/total) %>%
+  mutate(state="MA") %>%
+  ungroup()
+
 ```
@@
 ```{r}
-
-ma_concordance_by_county %>%
-  filter(vlabel=="$P(S_1|untested)$ Centered at Emp. Value")
 
@@
   filter(vlabel=="$P(S_1|untested)$ Centered at Emp. Value") %>%
-  group_by(county_name) %>%
+  group_by(name) %>%
   mutate(m = prop[which(in_interval=="In Interval")]) %>%
@@
                             "In Interval", "Below Interval"))) %>%
-  ggplot(aes(x= fct_reorder(county_name,m), 
+  ggplot(aes(x= fct_reorder(name,m), 
              fill = (in_interval), y =prop)) + 
@@
 
-```{r}
-
+```{r, fig.width=15}
+
+# 
 in_interval <- ma_concordance %>%
@@
 
-```
+ma_concordance %>%
+  group_by(vlabel) %>%
+  mutate(total=sum(n_per_version)) %>%
+  filter(in_interval == "In Interval") %>%
+  mutate(n_interval = n_per_version) %>%
+  ggplot(aes(y=fct_reorder(vlabel,prop), x=prop)) +
+  geom_text(aes(label=round(prop,3), x= prop+.03), 
+             angle=0) +
+  geom_bar(stat="identity", fill="#596281") +
+  scale_y_discrete(breaks= unique(ma_concordance$vlabel),
+                   labels=function(x)TeX(x)) +
+  scale_x_continuous(expand=c(0,0), 
+                     limits=c(0,.9), n.breaks=7)  +
+  theme_c() +
+  labs(y="",
+       x="Proportion Containing the Covidestim Median",
+       title="Proportion Containing the Covidestim Median\nBy Implementation")
+
+```

### 2023-06-25 R/05-state-analysis.R
state-specific priors target

@@ -519,3 +519,71 @@ get_v6_corrected <- function(state_testing, params) {
   
 }
 
+
+
+#---------------- Version 7 --------------
+
+
+#---------- Version 1 -------------
+get_v7_state <- function(data, testing = FALSE) {
+  
+  params <-  list(
+    alpha_mean = .95,
+    alpha_sd = 0.08,
+    alpha_bounds = NA,
+    # alpha_bounds = c(.8,1),
+    beta_mean = .216,
+    beta_sd =.09,
+    beta_bounds = NA,
+    #  beta_bounds = c(0.002, 0.4),
+    s_untested_mean = .016,
+    s_untested_sd = .0225,
+    #  s_untested_bounds = c(0.0018, Inf),
+    s_untested_bounds = NA,
+    p_s0_pos_mean = .4,
+    p_s0_pos_sd = .1225,
+    p_s0_pos_bounds = NA,
+    #  p_s0_pos_bounds = c(.25, .7),
+    pre_nsamp = 1e6,
+    post_nsamp = 1e5)
+  
+  state_testing <- data
+  
+  state_testing <- state_testing %>%
+    rename(fips = state) %>%
+    select(-date) %>% 
+    distinct()
+  
+  
+  # only use a few rows if testing
+  state_testing <- if(testing) state_testing %>% 
+    filter(biweek >=6) %>% slice_sample(n=5) else state_testing
+  
+  melded <- do.call(get_melded, params)
+  
+  corrected <- pmap_df(state_testing, 
+                       function(posrate,
+                                population,
+                                total, 
+                                positive,
+                                biweek,
+                                fips,
+                                ...) {
+                         process_priors_per_county(
+                           priors = melded$post_melding,
+                           county_df = tibble(posrate, 
+                                              population,
+                                              total, 
+                                              positive,
+                                              biweek,
+                                              fips),
+                           nsamp = params$post_nsamp) %>%
+                           generate_corrected_sample(., num_reps = 1e3) %>%
+                           summarize_corrected_sample() })
+  
+  corrected <- corrected %>%
+    mutate(version="v7")
+  
+  return(corrected)
+  
+}

### 2023-06-25 R/06-county-analysis.R

@@ -420,3 +420,72 @@ get_v6_county_corrected <- function(county_testing, params) {
 }
 
 
+#----------------- Version 7 ------------------------
+#---------- Version 1 -------------
+get_county_v7 <- function(data, testing = FALSE) {
+  
+  params <-  list(
+    alpha_mean = .95,
+    alpha_sd = 0.08,
+    alpha_bounds = NA,
+    # alpha_bounds = c(.8,1),
+    beta_mean = .216,
+    beta_sd =.09,
+    beta_bounds = NA,
+    #  beta_bounds = c(0.002, 0.4),
+    s_untested_mean = .016,
+    s_untested_sd = .0225,
+    #  s_untested_bounds = c(0.0018, Inf),
+    s_untested_bounds = NA,
+    p_s0_pos_mean = .4,
+    p_s0_pos_sd = .1225,
+    p_s0_pos_bounds = NA,
+    #  p_s0_pos_bounds = c(.25, .7),
+    pre_nsamp = 1e6,
+    post_nsamp = 1e5)
+  
+  
+  county_testing <- data
+  
+  county_testing <- county_testing %>%
+    select(-date) %>% 
+    distinct()
+  
+  
+  # only use a few rows if testing
+  county_testing <- if(testing) county_testing %>% 
+    filter(biweek >=6) %>% slice_sample(n=5) else county_testing
+  
+  melded <- do.call(get_melded, params)
+  
+  corrected <- pmap_df(county_testing, 
+                       function(posrate,
+                                population,
+                                total, 
+                                positive,
+                                biweek,
+                                fips,
+                                ...) {
+                         process_priors_per_county(
+                           priors = melded$post_melding,
+                           county_df = tibble(posrate, 
+                                              population,
+                                              total, 
+                                              positive,
+                                              biweek,
+                                              fips),
+                           nsamp = params$post_nsamp) %>%
+                           generate_corrected_sample(., num_reps = 1e3) %>%
+                           summarize_corrected_sample() })
+  
+  
+  corrected <- corrected %>%
+    mutate(version="v7")
+  
+  return(corrected)
+  
+  
+}
+
+
+

### 2023-06-25 _targets.R

@@ -129,6 +129,13 @@ list(
       vary = "s_untested_and_beta"
     )
   ),
+  # version 7
+  tar_target(
+    name = state_v7,
+    command = get_v7_state(
+      data = tests_biweekly_state
+    )
+  ),
   
   
   #---------------- county-level analysis -------------------
@@ -194,6 +201,14 @@ list(
       ctis=ctis_smoothed,
       spline=TRUE,
       vary = "s_untested_and_beta"
+    )
+  ),
+  
+  # version 7
+  tar_target(
+    name = ma_v7,
+    command = get_county_v7(
+      data = ma_biweekly_county
     )
   ),
   

### 2023-06-25 reports/county_figures.Rmd
add version 7

@@
              store =targets_dir)
+   )%>%
+   bind_rows(
+    tar_read(ma_v7,
+             store =targets_dir)
    )
@@
 versions <- tibble(
-  version = c("v1", "v2", "v3", "v4", "v5", "v6"),
+  version = c("v1", "v2", "v3", "v4", "v5", "v6", "v7"),
   vlabel = factor(
-    c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
+    c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
     "$\\beta$ Centered at Emp. Value",
@@
         "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
-         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"
+         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
+        "$\\,$CTIS Priors that Do Not Vary by Date$\\,$"
     ),
     levels = c(
-        "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
+        "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
         "$\\beta$ Centered at Emp. Value",
@@
         "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
-         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"
+         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
+        "$\\,$CTIS Priors that Do Not Vary by Date$\\,$"
 
@@
 
+```
+
+```{r}
 plt_wastewater_pb <- w_data %>%
@@
                               use = "complete.obs")) %>%
-  mutate(x=2.5,
+  mutate(x=2.8,
          y=.1)
@@
   mutate(vlabel = gsub(
-    "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
-    "Priors Do Not Vary by County and Date",
+    "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
+    "Priors Do Not Vary by State and Date",
     vlabel, fixed=TRUE),vlabel=gsub(" at","\nat", vlabel)) %>%
@@
 
-```{r, fig.width=15}
+```{r, fig.width=13}
 
@@
   geom_bar(stat="identity",position="stack", color="darkgray", linewidth=.2) +
-#  viridis::scale_fill_viridis(option="mako", begin=.2, discrete=TRUE) +
-  theme_c(legend.position="right",
+   theme_c(axis.text.y = element_text(size = 14, hjust=1),
+          axis.title.x = element_text(size=14),
+          axis.text.x=element_text(size=10),
+          legend.position="right",
           legend.title = element_text(face="bold", size = 15),
           legend.text= element_text( size = 15),
-          legend.spacing.y = unit(3, 'pt'),
-          axis.text.y = element_text(size = 17, hjust=1)) +
-  theme(
-          plot.subtitle = element_text(size=18, face='italic', 
-                                       margin=margin(4,0,4,0))) +
+          legend.spacing.y = unit(3, 'pt')) +
   scale_y_discrete(labels=unname(TeX(labels))) +
   labs(y ="",
-       x="Proportion",
+       x="Proportion Containing the Covidestim Median",
        fill = "Covidestim Median:",
        title = "Where Covidestim Medians Fall\nRelative to Probabilistic Bias Intervals",
-       subtitle = "By Version, in Massachusetts") +
+       subtitle = "By Implementation, in Massachusetts") +
   scale_fill_manual(values=c("Below Interval"="#7997D2",
@@
                      limits=c(-.05,1))
+
+```
+
+```{r, fig.width=12.3}
 
@@
                      limits=c(0,.9), n.breaks=7)  +
-  theme_c() +
+  theme_c(axis.text.y = element_text(size = 13, hjust=1),
+          axis.title.x = element_text(size=14),
+          axis.text.x=element_text(size=10)) +
   labs(y="",
        x="Proportion Containing the Covidestim Median",
-       title="Proportion Containing the Covidestim Median\nBy Implementation")
-
-```
+       title="Proportion Containing the Covidestim Median",
+       subtitle = "By Implementation")
+
+```
+
+# Summary of Comparison Between Implementations
+
+Although the spline-smoothing improved the correlations, it did not improve the proportion containing the Covidestim median.
+
+The implementation centering $\Pr(S_1|\text{untested})$ at the survey value along with the implementation that does not vary the priors by state or date are the most correlated with Covidestim estimates, wastewater concentrations, and  have the highest proportion of county-biweeks containing the Covidestim median.

### 2023-06-25 reports/priors.Rmd
add version 7

@@
 ```{r,include=FALSE}
+
 ctis %>%
@@
 ```
+
+```{r}
+
+ctis %>%
+  group_by(state) %>%
+  summarize(mean=mean(beta,na.rm=TRUE)) %>%
+  pull(mean) %>%
+  mean()
+
+ctis %>%
+  group_by(state) %>%
+  summarize(mean=mean(smoothed_wcli,na.rm=TRUE)) %>%
+  pull(mean) %>%
+  mean()
+
+```

### 2023-06-25 R/05-state-analysis.R

@@ -523,8 +523,6 @@ get_v6_corrected <- function(state_testing, params) {
 
 #---------------- Version 7 --------------
 
-
-#---------- Version 1 -------------
 get_v7_state <- function(data, testing = FALSE) {
   
   params <-  list(
@@ -587,3 +585,5 @@ get_v7_state <- function(data, testing = FALSE) {
   return(corrected)
   
 }
+
+

### 2023-06-26 reports/county_figures.Rmd
improve and add county-level figures

@@
 library(kableExtra)
-
+library(scales)
+library(cowplot)
 ```
@@
 waste <- tar_read(waste, store =targets_dir)
+county_names <- read_tsv(here("data/demographic/county_fips.tsv")) %>%
+  rename_with(.cols = everything(), .fn =tolower) %>%
+  bind_rows()
+
 covidestim <- tar_read(covidestim_biweekly_county,
-                        store =targets_dir)
+                        store =targets_dir) %>%
+  select(-c(date)) %>% 
+  distinct()
 
@@
   left_join(covidestim, relationship = "many-to-many") %>%
+  left_join(dates,  relationship = "many-to-many") %>%
   group_by(biweek) %>%
@@
        title = "Correlations Between Wastewater Concentrations\nand Covidestim Estimates")
+
+pb_correlations <- w_data %>%
+  filter(version=="v1") %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(positive_std,mean_conc_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1.8,
+         y=.1)
+
+```
+
+### Compare to Wastewater Over Time
+
+```{r,fig.width=12, fig.height=15}
+
+library(scales)
+
+pal <- c("#10BAC5", "#1B10C5", "#EFB719", "#900C3F")
+
+names(pal) <- c(#"$observed$", 
+                '$P(S_1|untested)$ Centered at Emp. Value',
+                '$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values',
+               "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$",
+               "$\\beta$ Centered at Emp. Value"
+               )
+
+w_data <- w_data %>%
+  group_by(name) %>%
+  mutate(adjust = 1/(max(mean_conc))*(exp_cases_median[which.max(mean_conc)])) %>%
+  ungroup()
+
+waste_df <- w_data %>%
+  select(biweek, mean_conc, name, adjust, fips) %>%
+  distinct() %>%
+  left_join(dates,relationship = "many-to-many" ) %>%
+  group_by(name, biweek) %>%
+  # center date within 2-week interval
+  summarize(date = min(date) + days(7),
+            mean_conc =unique(mean_conc),
+            adjust=unique(adjust),
+            fips = unique(fips)) %>%
+  mutate(name = gsub("County, MA", "", name))
+
+pltlist <- w_data %>%
+  mutate(name = gsub("County, MA", "", name)) %>%
+  left_join(dates,relationship = "many-to-many" ) %>%
+  # filter(version %in% c("v1","v3")) %>%
+  filter(version=="v1") %>%
+  filter(biweek>=6) %>%
+  group_by(biweek) %>%
+  mutate(mindate = min(date),
+         maxdate=max(date)) %>%
+  ungroup() %>% 
+  group_by(name) %>%
+  group_split() %>%
+  map(~ {
+
+    county_fips <- .x$fips %>% unique()
+    adj <- .x$adjust %>% unique()
+
+    .x %>% 
+ #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
+    ggplot() +
+      geom_ribbon(aes(x = date, 
+                 ymin = exp_cases_lb,
+                 ymax = exp_cases_ub,
+                 fill = vlabel),
+                 alpha = .5) +
+      geom_line(data = waste_df[which(waste_df$fips == county_fips),],
+                aes(x = date, y =mean_conc*adj, 
+                    color = 'Wastewater Effective Concentrations'),
+               # color = "#DB4048",
+                size = 1.1) +
+      geom_linerange(aes(xmin = mindate, 
+                         xmax = maxdate, 
+                         y = infections,
+                         color = 'Covidestim Infection Estimates')) + 
+      geom_point(data =  waste_df[which(waste_df$fips == county_fips),],
+                aes(x = date, y =mean_conc*adj ),
+                color = "#DB4048",
+                alpha = .5,
+                size = 1.2) +
+      scale_y_continuous(sec.axis=sec_axis(~.x/adj,
+                                           name = 'Mean Wastewater Effective Concentration',
+                                           labels=comma),
+                         labels=comma
+                    ) +
+      # facet_grid(name~vlabel, 
+      #            labeller=labeller(vlabel =as_labeller(
+      #              TeX, default=label_parsed))) +
+      facet_wrap(~name, ncol=3) + 
+      scale_fill_manual(values = pal, labels = c('','','',''), 
+                        name = "Probabilistic Bias Intervals") +
+     theme_c(legend.position="none") +
+      labs(y = "Value (Standardized by County)",
+           title = "",
+           x= "") +
+      scale_x_date(date_labels = "%b %Y") +
+    scale_color_manual(
+      name = '', 
+      values = c(
+        'Covidestim Infection Estimates' = 'darkblue',
+        'Wastewater Effective Concentrations' = '#DB4048')) +
+    guides(color = guide_legend(override.aes = list(size = 12,
+                                                    linewidth=3)),
+           fill = guide_legend(override.aes =list(size = 6))) 
+    })
+
+legend <- cowplot::get_legend(
+  pltlist[[1]] + theme(legend.position="top")
+)
+
+grid <- cowplot::plot_grid(plotlist=pltlist, ncol =3)
+
+cowplot::plot_grid(legend, grid, nrow=2, rel_heights=c(.05,.95))
+
+```
+
+```{r, eval=FALSE,include=FALSE}
+
+ w_data_std <- w_data %>%
+  group_by(name, version) %>%
+  mutate(exp_cases_lb = exp_cases_lb /max(exp_cases_median),
+         exp_cases_ub = exp_cases_ub /max(exp_cases_median),
+         exp_cases_median = exp_cases_median / max(exp_cases_median),
+         mean_conc=mean_conc/max(mean_conc),
+         infections=infections/max(infections)) %>%
+  mutate(name = gsub("County, MA", "", name)) %>%
+  left_join(dates,relationship = "many-to-many" ) %>%
+  filter(version %in% c("v1","v3")) %>%
+  filter(biweek>=6) %>%
+  group_by(biweek) %>%
+  mutate(mindate = min(date),
+         maxdate=max(date)) %>%
+  ungroup()
+
+waste_df <- w_data_std %>%
+  select(biweek, mean_conc, name, fips) %>%
+  distinct() %>%
+  left_join(dates,relationship = "many-to-many" ) %>%
+  group_by(name, biweek) %>%
+  # center date within 2-week interval
+  summarize(date = min(date) + days(7),
+            mean_conc =unique(mean_conc),
+            fips = unique(fips)) %>%
+  mutate(name = gsub("County, MA", "", name))
+
+w_data_std %>% 
+ #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
+    ggplot() +
+      geom_ribbon(aes(x = date, 
+                 ymin = exp_cases_lb,
+                 ymax = exp_cases_ub,
+                 fill = vlabel),
+                 alpha = .5) +
+      geom_line(data = waste_df,
+                aes(x = date, y =mean_conc, 
+                    color = 'Wastewater Effective Concentrations'),
+               # color = "#DB4048",
+                size = 1.1) +
+      geom_linerange(aes(xmin = mindate, 
+                         xmax = maxdate, 
+                         y = infections,
+                         color = 'Covidestim Infection Estimates')) + 
+      geom_point(data =  waste_df,
+                aes(x = date, y =mean_conc),
+                color = "#DB4048",
+                alpha = .5,
+                size = 1.2) +
+   #   scale_y_continuous(sec.axis=sec_axis(~.x/adj)) +
+      facet_grid(name~vlabel, 
+                 labeller=labeller(vlabel =as_labeller(
+                   TeX, default=label_parsed))) +
+      scale_fill_manual(values = pal, labels = c('','','',''), 
+                        name = "Probabilistic Bias Intervals") +
+      theme_bw() +
+     theme_c() +
+      labs(y = "Value (Standardized by County)",
+           title = "",
+           x= "") +
+      scale_x_date(date_labels = "%b %Y") +
+    scale_color_manual(
+      name = '', 
+      values = c(
+        'Covidestim Infection Estimates' = 'darkblue',
+        'Wastewater Effective Concentrations' = '#DB4048')) +
+    guides(color = guide_legend(override.aes = list(size = 12,
+                                                    linewidth=3)),
+           fill = guide_legend(override.aes =list(size = 6)))
+
+```
+
+## Normalization Strategy 2: By Values at Biweek Where Wastewater Concentration is Maximized
+
+```{r}
+
+w_data <- wjoined %>% 
+  select(-date) %>%
+  distinct() %>%
+  filter(fips %in% counties_with_waste) %>%
+  filter(!is.na(mean_conc)) %>%
+  group_by(fips, version) %>%
+  mutate(mean_conc_std = mean_conc/ max(mean_conc),
+         max_conc = max(mean_conc),
+         exp_cases_median_std = exp_cases_median / exp_cases_median[which.max(mean_conc)],
+         exp_cases_lb_std = exp_cases_lb /exp_cases_median[which.max(mean_conc)],
+         exp_cases_ub_std = exp_cases_ub / exp_cases_median[which.max(mean_conc)],
+         infections_std  = infections/ infections[which.max(mean_conc)],
+         positive_std = positive/max(positive)) %>%
+  ungroup()
+
+```
+
+### PB Counts and Wastewater
+
+```{r}
+
+# correlations between PB counts and wastewater concentrations
+
+correlations <- w_data %>%
+  group_by(version) %>%
+  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
+            vlabel =unique(vlabel)) %>%
+  mutate(x=1.8,
+         y=.1)
+
+w_data %>%
+  group_by(version) %>%
+  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = mean_conc_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= correlations ) +
+  facet_wrap(~vlabel,
+             ncol=2,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
+       y = "Mean Wastewater Concentration\n(Standardized by County)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")
+
+plt_wastewater_pb <- w_data %>%
+  group_by(version) %>%
+  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = mean_conc_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= correlations ) +
+  facet_wrap(~vlabel,
+             ncol=1,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
+       y = "Mean Wastewater Concentration\n(Standardized by County by County)",
+       title = "")
+
+```
+
+### PB Counts and Covidestim
+
+```{r}
+
+# correlations between PB counts and Covidestim estimates
+
+covidestim_correlations <- w_data %>%
+  filter(!is.na(infections_std)) %>%
+  group_by(vlabel) %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(exp_cases_median_std,infections_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1.8,
+         y=.1)
+
+w_data %>%
+  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = infections_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  facet_wrap(~vlabel,
+              ncol=2,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
+       y = "Covidestim Estimates\n(Standardized by County by County)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+
+plt_covidestim_pb <- w_data %>%
+  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
+  geom_point(alpha=.5) +
+  geom_linerange(aes(xmin = exp_cases_lb_std,
+                     xmax=exp_cases_ub_std, 
+                     y = infections_std),
+                 alpha = .3) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  facet_wrap(~vlabel,
+              ncol=1,
+             labeller= as_labeller(TeX, 
+                                   default = label_parsed)) +
+  theme_c() +
+  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
+       y = "Covidestim Estimates\n(Standardized by County by County)",
+       title = "")
+
+```
+
+### Combined Figure
+
+```{r, fig.height=10, fig.width=12}
+
+cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)
+
+cor_covid <- covidestim_correlations %>% 
+  select(vlabel, 
+         `Correlation with Covidestim Estimates`=Correlation)
+
+cor_waste <- correlations %>% 
+  select(vlabel, 
+         `Correlation with Wastewater Concentrations`=Correlation)
+
+cor_waste %>%
+  left_join(cor_covid) %>%
+  mutate(vlabel = gsub(
+    "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
+    "Priors Do Not Vary by County and Date",
+    vlabel, fixed=TRUE),vlabel=gsub(" at","\nat", vlabel)) %>%
+  arrange(desc( `Correlation with Wastewater Concentrations`)) %>%
+  rename(Implementation = vlabel) %>%
+  kbl() %>%
+  kable_styling(full_width = TRUE)
+
+```
+
+### Covidestim and Wastewater
+
+```{r}
+
+covidestim_correlations <- w_data %>%
+  filter(version=="v1") %>%
+  # missing values are handled by casewise deletion
+  summarize(Correlation = cor(mean_conc_std,infections_std,
+                              use = "complete.obs")) %>%
+  mutate(x=1,
+         y=.1)
+
+w_data %>%
+  filter(version=="v1") %>%
+  ggplot(aes(x=infections_std, y = mean_conc_std)) +
+  geom_point(alpha=.5) +
+  geom_text(
+           aes(x=x,y=y,  label=paste0("Correlation: ", 
+                                      round(Correlation,3))), 
+                data= covidestim_correlations ) +
+  theme_c() +
+  labs(x = "Covidestim Estimates\n(Standardized by County by County)",
+       y = "Wastewater Concentrations\n(Standardized by County by County)",
+       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
 
@@
      theme_c() +
-      labs(y = "Value (Standardized by County)",
+      labs(y = "Value (Standardized by County by County)",
            title = "",
@@
 
-## Normalization Strategy 2: By Values at Biweek Where Wastewater Concentration is Maximized
-
-```{r}
-
-w_data <- wjoined %>% 
-  select(-date) %>%
-  distinct() %>%
-  filter(fips %in% counties_with_waste) %>%
-  filter(!is.na(mean_conc)) %>%
-  group_by(fips, version) %>%
-  mutate(mean_conc_std = mean_conc/ max(mean_conc),
-         max_conc = max(mean_conc),
-         exp_cases_median_std = exp_cases_median / exp_cases_median[which.max(mean_conc)],
-         exp_cases_lb_std = exp_cases_lb /exp_cases_median[which.max(mean_conc)],
-         exp_cases_ub_std = exp_cases_ub / exp_cases_median[which.max(mean_conc)],
-         infections_std  = infections/ infections[which.max(mean_conc)],
-         positive_std = positive/max(positive)) %>%
-  ungroup()
-
-```
+## Normalization Strategy 3: By Maximum (Separately)
+
+For each county, normalize the wastewater concentration by the maximum concentration for that county and the probabilistic bias counts by the maximum *median* estimated counts for that county (2.5th percentile and 97.5th percentile also standardized by the maximum *median* estimated counts for that county).
 
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
-       y = "Mean Wastewater Concentration\n(Standardized by County by County)",
+  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
+       y = "Mean Wastewater Concentration\n(Standardized by County)",
        title = "")
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
-       y = "Covidestim Estimates\n(Standardized by County by County)",
+  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
+       y = "Covidestim Estimates\n(Standardized by County)",
        title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
-       y = "Covidestim Estimates\n(Standardized by County by County)",
+  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
+       y = "Covidestim Estimates\n(Standardized by County)",
        title = "")
-
 ```
@@
 
-```{r, fig.height=10, fig.width=12}
+```{r, fig.height=10, fig.width=10}
 
 cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)
-
-cor_covid <- covidestim_correlations %>% 
-  select(vlabel, 
-         `Correlation with Covidestim Estimates`=Correlation)
-
-cor_waste <- correlations %>% 
-  select(vlabel, 
-         `Correlation with Wastewater Concentrations`=Correlation)
-
-cor_waste %>%
-  left_join(cor_covid) %>%
-  mutate(vlabel = gsub(
-    "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
-    "Priors Do Not Vary by County and Date",
-    vlabel, fixed=TRUE),vlabel=gsub(" at","\nat", vlabel)) %>%
-  arrange(desc( `Correlation with Wastewater Concentrations`)) %>%
-  rename(Implementation = vlabel) %>%
-  kbl() %>%
-  kable_styling(full_width = TRUE)
 
@@
   theme_c() +
-  labs(x = "Covidestim Estimates\n(Standardized by County by County)",
-       y = "Wastewater Concentrations\n(Standardized by County by County)",
-       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
-
-pb_correlations <- w_data %>%
-  filter(version=="v1") %>%
-  # missing values are handled by casewise deletion
-  summarize(Correlation = cor(positive_std,mean_conc_std,
-                              use = "complete.obs")) %>%
-  mutate(x=1.8,
-         y=.1)
-
-```
-
-### Compare to Wastewater Over Time
-
-```{r,fig.width=12, fig.height=15}
-
-library(scales)
-
-pal <- c("#10BAC5", "#1B10C5", "#EFB719", "#900C3F")
-
-names(pal) <- c(#"$observed$", 
-                '$P(S_1|untested)$ Centered at Emp. Value',
-                '$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values',
-               "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$",
-               "$\\beta$ Centered at Emp. Value"
-               )
-
-waste_df <- w_data %>%
-  select(biweek, mean_conc_std, name) %>%
-  distinct() %>%
-  left_join(dates,relationship = "many-to-many" ) %>%
-  group_by(name, biweek) %>%
-  # center date within 2-week interval
-  summarize(date = min(date) + days(7),
-            mean_conc_std=unique(mean_conc_std)) %>%
-  mutate(name = gsub("County, MA", "", name))
-
-w_data%>%
-  mutate(name = gsub("County, MA", "", name)) %>%
-  left_join(dates,relationship = "many-to-many" ) %>%
-  filter(version %in% c("v1","v3")) %>%
-  filter(biweek>=6) %>%
-  group_by(biweek) %>%
-  mutate(mindate = min(date),
-         maxdate=max(date)) %>%
-  ungroup() %>% 
- #   filter(fips == county_fips  & date >= begin_date & date <= end_date) %>%
-    ggplot() +
-      geom_ribbon(aes(x = date, 
-                 ymin = exp_cases_lb_std,
-                 ymax = exp_cases_ub_std,
-                 fill = vlabel),
-                 alpha = .5) +
-      geom_line(data = waste_df,
-                aes(x = date, y =mean_conc_std, 
-                    color = 'Wastewater Effective Concentrations'),
-               # color = "#DB4048",
-                size = 1.1) +
-      geom_linerange(aes(xmin = mindate, 
-                         xmax = maxdate, 
-                         y = infections_std,
-                         color = 'Covidestim Infection Estimates')) + 
-      geom_point(data = waste_df,
-                aes(x = date, y =mean_conc_std ),
-                color = "#DB4048",
-                alpha = .5,
-                size = 1.2) +
-      facet_grid(name~vlabel, 
-                 labeller=labeller(vlabel =as_labeller(
-                   TeX, default=label_parsed))) +
-      scale_fill_manual(values = pal, labels = c('','','',''), 
-                        name = "Probabilistic Bias Intervals") +
-      theme_bw() +
-     theme_c() +
-      labs(y = "Value (Standardized by County by County)",
-           title = "",
-           x= "") +
-      scale_x_date(date_labels = "%b %Y") +
-    scale_color_manual(
-      name = '', 
-      values = c(
-        'Covidestim Infection Estimates' = 'darkblue',
-        'Wastewater Effective Concentrations' = '#DB4048')) +
-    guides(color = guide_legend(override.aes = list(size = 12,
-                                                    linewidth=3)),
-           fill = guide_legend(override.aes =list(size = 6)))
-
-```
-
-## Normalization Strategy 3: By Maximum (Separately)
-
-For each county, normalize the wastewater concentration by the maximum concentration for that county and the probabilistic bias counts by the maximum *median* estimated counts for that county (2.5th percentile and 97.5th percentile also standardized by the maximum *median* estimated counts for that county).
-
-### PB Counts and Wastewater
-
-```{r}
-
-# correlations between PB counts and wastewater concentrations
-
-correlations <- w_data %>%
-  group_by(version) %>%
-  summarize(Correlation = cor(exp_cases_median_std,mean_conc_std),
-            vlabel =unique(vlabel)) %>%
-  mutate(x=1.8,
-         y=.1)
-
-w_data %>%
-  group_by(version) %>%
-  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
-  geom_point(alpha=.5) +
-  geom_linerange(aes(xmin = exp_cases_lb_std,
-                     xmax=exp_cases_ub_std, 
-                     y = mean_conc_std),
-                 alpha = .3) +
-  geom_text(
-           aes(x=x,y=y,  label=paste0("Correlation: ", 
-                                      round(Correlation,3))), 
-                data= correlations ) +
-  facet_wrap(~vlabel,
-             ncol=2,
-             labeller= as_labeller(TeX, 
-                                   default = label_parsed)) +
-  theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
-       y = "Mean Wastewater Concentration\n(Standardized by County)",
-       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")
-
-plt_wastewater_pb <- w_data %>%
-  group_by(version) %>%
-  ggplot(aes(x=exp_cases_median_std, y = mean_conc_std)) +
-  geom_point(alpha=.5) +
-  geom_linerange(aes(xmin = exp_cases_lb_std,
-                     xmax=exp_cases_ub_std, 
-                     y = mean_conc_std),
-                 alpha = .3) +
-  geom_text(
-           aes(x=x,y=y,  label=paste0("Correlation: ", 
-                                      round(Correlation,3))), 
-                data= correlations ) +
-  facet_wrap(~vlabel,
-             ncol=1,
-             labeller= as_labeller(TeX, 
-                                   default = label_parsed)) +
-  theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
-       y = "Mean Wastewater Concentration\n(Standardized by County)",
-       title = "")
-
-```
-
-### PB Counts and Covidestim
-
-```{r}
-
-# correlations between PB counts and Covidestim estimates
-
-covidestim_correlations <- w_data %>%
-  filter(!is.na(infections_std)) %>%
-  group_by(vlabel) %>%
-  # missing values are handled by casewise deletion
-  summarize(Correlation = cor(exp_cases_median_std,infections_std,
-                              use = "complete.obs")) %>%
-  mutate(x=1.8,
-         y=.1)
-
-w_data %>%
-  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
-  geom_point(alpha=.5) +
-  geom_linerange(aes(xmin = exp_cases_lb_std,
-                     xmax=exp_cases_ub_std, 
-                     y = infections_std),
-                 alpha = .3) +
-  geom_text(
-           aes(x=x,y=y,  label=paste0("Correlation: ", 
-                                      round(Correlation,3))), 
-                data= covidestim_correlations ) +
-  facet_wrap(~vlabel,
-             ncol=2,
-             labeller= as_labeller(TeX, 
-                                   default = label_parsed)) +
-  theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
-       y = "Covidestim Estimates\n(Standardized by County)",
-       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
-
-plt_covidestim_pb <- w_data %>%
-  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
-  geom_point(alpha=.5) +
-  geom_linerange(aes(xmin = exp_cases_lb_std,
-                     xmax=exp_cases_ub_std, 
-                     y = infections_std),
-                 alpha = .3) +
-  geom_text(
-           aes(x=x,y=y,  label=paste0("Correlation: ", 
-                                      round(Correlation,3))), 
-                data= covidestim_correlations ) +
-  facet_wrap(~vlabel,
-              ncol=1,
-             labeller= as_labeller(TeX, 
-                                   default = label_parsed)) +
-  theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
-       y = "Covidestim Estimates\n(Standardized by County)",
-       title = "")
-```
-
-### Combined Figure
-
-```{r, fig.height=10, fig.width=10}
-
-cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)
-
-```
-
-### Covidestim and Wastewater
-
-```{r}
-
-covidestim_correlations <- w_data %>%
-  filter(version=="v1") %>%
-  # missing values are handled by casewise deletion
-  summarize(Correlation = cor(mean_conc_std,infections_std,
-                              use = "complete.obs")) %>%
-  mutate(x=1,
-         y=.1)
-
-w_data %>%
-  filter(version=="v1") %>%
-  ggplot(aes(x=infections_std, y = mean_conc_std)) +
-  geom_point(alpha=.5) +
-  geom_text(
-           aes(x=x,y=y,  label=paste0("Correlation: ", 
-                                      round(Correlation,3))), 
-                data= covidestim_correlations ) +
-  theme_c() +
   labs(x = "Covidestim Estimates\n(Standardized by County)",
@@
          y=.1)
+
+```
+
+# Simulation Intervals Over Time
+
+```{r, fig.width=11, fig.height=8}
+
+county_dat <- county %>%
+  # set Nantucket, Duke fips to Nantucket since we have wastewater data
+  # for Nantucket
+  left_join(dates,  relationship = "many-to-many") %>%
+  group_by(biweek) %>%
+  mutate(mindate = min(date),
+         maxdate = max(date)) %>%
+  ungroup() %>%
+  left_join(county_names, by=c('fips'='fips')) %>%
+  mutate(name = ifelse(fips =="25007,25019", "Nantucket and Dukes", name))
+
+county_dat %>%
+  filter(version=="v1")  %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub),
+             alpha = .5,
+             show.legend=FALSE,
+             fill="#596281") +
+   geom_linerange(aes(xmin = mindate,
+                      xmax=maxdate,
+                      y = positive,
+                      color = 'obs'),
+                  size=.5) +
+  facet_wrap(~name, scales="free_y", ncol=3) +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(
+          legend.position = "top",
+          ) +
+  scale_fill_manual(values = pal)  +
+  labs(x="",
+       y = "Estimated Infections",
+       title = "Simulation Intervals Over Time",
+       subtitle = "Counties in Massachusetts") +
+  scale_color_manual(values = c('obs' = 'red'), labels = 'Positive Tests',
+                     name = '') +
+  guides(color = guide_legend(override.aes = list(linewidth =2)))
+
+```
+
+## Ratio of Estimated Infections to Observed
+
+```{r, fig.width=12, fig.height=9}
+
+county_dat %>%
+  filter(version=="v1")  %>%
+  #filter(grepl("Nan",name)) %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb/positive,
+             ymax = exp_cases_ub/positive),
+             alpha = .5,
+             show.legend=FALSE,
+             fill="#596281") +
+   # geom_linerange(aes(xmin = mindate,
+   #                    xmax=maxdate,
+   #                    y = positive,
+   #                    color = 'obs')) +
+  facet_wrap(~name,ncol=3) +
+ # facet_grid(name~vlabel, 
+ #                 labeller=labeller(vlabel =as_labeller(
+ #                   TeX, default=label_parsed))) +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(
+          legend.position = "top") +
+  scale_fill_manual(values = pal)  +
+  labs(x="",
+       y = "Estimated Infections",
+       title = "Ratio of Estimated to Observed Over Time",
+       subtitle = "Counties in Massachusetts") +
+  scale_color_manual(values = c('obs' = 'red'), labels = 'Positive Tests',
+                     name = '') +
+  guides(color = guide_legend(override.aes = list(linewidth =2)))
+
+```
+
+# Trends in Total Tests and Positive Tests
+
+Informs our understanding of the trends in the ratio of estimated to observed infections.
+
+## Positive Tests and Test Positivity
+
+```{r, fig.height=12, fig.width=12}
+
+############################################
+# POSITIVE TESTS AND TEST POSITIVITY
+############################################
+pltlist <- county_dat %>%filter(version=="v1") %>%
+    mutate(testpos = positive/total) %>%
+  group_by(across(-date)) %>%
+  summarize(date=min(date) + days(7)) %>%
+  group_by(fips) %>%
+  mutate(adj = (1/max(testpos)) * max(positive)) %>%
+  group_split() %>%
+  map(~ {
+    adj <-unique(.x$adj)
+    .x %>%
+      ggplot(aes(x=date))+
+    geom_line(aes(y=positive, color = 'Positive Tests')) +
+      geom_point(aes(y=positive, color = 'Positive Tests'),
+              alpha=.5,
+              size=.8) +
+    geom_line(aes(y=testpos*adj, color='Test Positivity'),
+              linetype=2) +
+      geom_point(aes(y=testpos*adj, color='Test Positivity'),
+              alpha=.5,
+              size=.8) +
+    #scale_linetype_discrete(breaks = c(1,1,2)) +
+    scale_color_manual(values=c(
+                                "Positive Tests"= "#279143",
+                                "Test Positivity"= '#E38E4F')) +
+    guides(color = guide_legend(override.aes = list(linetype = c(1,2),
+                                                    linewidth=c(2,2)))) +
+    theme_c() +
+    theme(axis.title = element_text(size=10),
+          axis.text =element_text(size=9),
+          legend.position="none") +
+    scale_y_continuous(labels=comma, sec.axis = sec_axis(~./adj , name = 'Test Positivity')) +
+    labs(color ='',
+         y="Positive Tests",
+         x= "Date") +
+      facet_wrap(~name)
+
+  })
+
+legend <- get_legend(
+  pltlist[[1]] +
+    theme(legend.position = "top",
+          legend.text=element_text(size=16))
+)
+
+title_gg <- ggplot() + 
+  labs(title = "Comparing Trends in Positive Tests and Test Positivity") + 
+  theme(plot.title=element_text(face="bold",
+                                hjust = .5, 
+                                size = 18,
+                                margin =margin(5,0,2,0)))
+
+plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)
+
+plot_grid(title_gg, legend, plt, rel_heights=c(.03, .03,.94), ncol=1)
+
+```
+
+## Total Tests and Positive Tests
+
+```{r, fig.height=12, fig.width=12}
+
+#####################################
+# TOTAL TESTS AND POSITIVE TESTS
+#####################################
+pltlist <- county_dat %>%filter(version=="v1") %>%
+    mutate(testpos = positive/total) %>%
+  group_by(across(-date)) %>%
+  summarize(date=min(date) + days(7)) %>%
+  group_by(fips) %>%
+  mutate(adj = (1/max(total)) * max(positive)) %>%
+  group_split() %>%
+  map(~ {
+    adj <-unique(.x$adj)
+    .x %>%
+      ggplot(aes(x=date))+
+    geom_line(aes(y=positive, color = 'Positive Tests')) +
+      geom_point(aes(y=positive, color = 'Positive Tests'),
+              alpha=.5,
+              size=.8) +
+    geom_line(aes(y=total*adj, color='Total Tests'),
+              linetype=2) +
+      geom_point(aes(y=total*adj, color='Total Tests'),
+              alpha=.5,
+              size=.8) +
+    #scale_linetype_discrete(breaks = c(1,1,2)) +
+    scale_color_manual(values=c(
+                                "Positive Tests"= "#279143",
+                                "Total Tests"= '#E38E4F')) +
+    guides(color = guide_legend(override.aes = list(linetype = c(1,2)))) +
+    theme_c() +
+    theme(axis.title = element_text(size=11),
+          axis.text =element_text(size=8),
+          legend.position="none") +
+    scale_y_continuous(labels=label_scientific(), 
+                       n.breaks=4,
+                       sec.axis = sec_axis(~./adj ,
+                                           name = 'Total Tests',
+                                           labels=label_scientific())) +
+    labs(color ='',
+         y="Positive Tests",
+         x= "Date") +
+      facet_wrap(~name)
+})
+
+legend <- cowplot::get_legend(
+  pltlist[[1]] +
+  theme(legend.position="top") +
+    guides(color = guide_legend(keywidth=2,
+                                override.aes = list(linewidth=c(2,2),
+                                                    linetype =c(1,2))))
+)
+
+title_gg <- ggplot() + 
+  labs(title = "Comparing Trends in Total Tests and Positive Tests") + 
+  theme(plot.title=element_text(face="bold",
+                                hjust = .5, 
+                                size = 18,
+                                margin =margin(5,0,2,0)))
+
+plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)
+
+cowplot::plot_grid(title_gg, legend, plt, rel_heights =c(.03, .03,.94),ncol=1)
+
+```
+
+## Percent Change in Total Tests and Positive Tests
+
+```{r, fig.height=10, fig.width=12}
+
+###################################################
+# PERCENT CHANGE: TOTAL TESTS AND POSITIVE TESTS
+###################################################
+
+county_dat %>%
+  filter(version=="v1") %>%
+  filter(!grepl("Nantucket",name)) %>%
+    mutate(testpos = positive/total) %>%
+  group_by(across(-date)) %>%
+  summarize(date=min(date) + days(7)) %>%
+  group_by(fips) %>%
+  arrange(biweek) %>%
+  mutate(pct_change_positive = (positive - lag(positive, 1))/lag(positive, 1),
+         pct_change_total = (total - lag(total, 1))/lag(total, 1)) %>%
+
+   ggplot(aes(x=date))+
+    geom_line(aes(y=pct_change_positive, color = 'Positive Tests')) +
+      geom_point(aes(y=pct_change_positive, color = 'Positive Tests'),
+              alpha=.5,
+              size=.8) +
+    geom_line(aes(y=pct_change_total, color='Total Tests'),
+              linetype=2) +
+      geom_point(aes(y=pct_change_total, color='Total Tests'),
+              alpha=.5,
+              size=.8) +
+    #scale_linetype_discrete(breaks = c(1,1,2)) +
+    scale_color_manual(values=c(
+                                "Positive Tests"= "#279143",
+                                "Total Tests"= '#E38E4F')) +
+    guides(color = guide_legend(override.aes = list(linetype = c(1,2)))) +
+    theme_c() +
+    theme(axis.title = element_text(size=15),
+          axis.text.x=element_text(size=9),
+          legend.position="top") +
+   # scale_y_continuous(labels=comma, sec.axis = sec_axis(~./adj , name = 'Test Positivity')) +
+    labs(color ='',
+         y="Percent Change",
+         x= "Date",
+         title = "Percent Change in Positive Tests and Total Tests") +
+      facet_wrap(~name, ncol=3, scales="free_x") +
+  scale_y_continuous(limits = c(-1,5.5),
+                     labels=percent) +
+  scale_x_date(date_breaks="3 months", date_labels="%b %Y")
 
@@
 
-Although the spline-smoothing improved the correlations, it did not improve the proportion containing the Covidestim median.
+Although the spline-smoothing improved the correlations with wastewater concentrations and the correlations with Covidestim estimates, it did not improve the proportion containing the Covidestim median.
 

### 2023-06-26 reports/state_figures.Rmd
rename

--- 2023-06-20 reports/state_results.Rmd
@@
 
-```{r load-data}
+```{r load-data, fig.width=10}
 
@@
         "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
-         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"
+         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
+        "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$"
     ),
     levels = c(
-        "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
+        "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
         "$\\beta$ Centered at Emp. Value",
@@
         "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
-         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)"
-
+         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
+        "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$"
     ) )
@@
 # read state-level results
-targets_dir <- here("_targets_6_19")
+targets_dir <- here("_targets")
 
@@
              store =targets_dir)
+   )%>%
+   bind_rows(
+    tar_read(state_v5,
+             store =targets_dir)
+  )%>%
+   bind_rows(
+    tar_read(state_v6,
+             store =targets_dir)
+  ) %>%
+   bind_rows(
+    tar_read(state_v7,
+             store =targets_dir)
   )
-  # )%>%
-  #  bind_rows(
-  #   tar_read(state_v5, 
-  #            store =targets_dir)
-  # )%>%
-  #  bind_rows(
-  #   tar_read(state_v6, 
-  #            store =targets_dir)
-  # )
 
@@
 
-```{r,fig.width=11}
+```{r,fig.width=14}
 
@@
        fill = "Covidestim Median:",
-       title = "Proportion Where Covidestim Median\nWas Within, Above, or Below the Median",
+       title = "Proportion of Simulation Intervals Where Covidestim Median\nWas Within, Above, or Below the Median",
        subtitle = "Including All States") +
   theme_c() +
-#  viridis::scale_fill_viridis(option="mako", begin=.2, discrete=TRUE) +
   theme_c(legend.position="right",
@@
           axis.text.y = element_text(size = 17, hjust=1)) +
-  theme(
+  theme(plot.title = element_text(size = 20),
           plot.subtitle = element_text(size=18, 
                                        face='italic', 
-                                       margin=margin(4,0,4,0))) +
+                                       margin=margin(4,0,4,0)),
+        axis.text.x=element_text(size=12),
+        axis.title = element_text(size=14)) +
   scale_x_discrete(labels = (TeX(labels)))
+
+```
+
+# Simulation Intervals
+
+```{r}
+
+subset <- state_res %>%
+  filter(vlabel ==  "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$")
 

### 2023-06-26 R/05-state-analysis.R
update CTIS-informed priors

@@ -530,7 +530,7 @@ get_v7_state <- function(data, testing = FALSE) {
     alpha_sd = 0.08,
     alpha_bounds = NA,
     # alpha_bounds = c(.8,1),
-    beta_mean = .216,
+    beta_mean = .18,
     beta_sd =.09,
     beta_bounds = NA,
     #  beta_bounds = c(0.002, 0.4),

### 2023-06-26 R/06-county-analysis.R

@@ -429,7 +429,7 @@ get_county_v7 <- function(data, testing = FALSE) {
     alpha_sd = 0.08,
     alpha_bounds = NA,
     # alpha_bounds = c(.8,1),
-    beta_mean = .216,
+    beta_mean = .18,
     beta_sd =.09,
     beta_bounds = NA,
     #  beta_bounds = c(0.002, 0.4),

### 2023-06-27 _targets.R
update targets based on CTIS priors

@@ -133,7 +133,11 @@ list(
   tar_target(
     name = state_v7,
     command = get_v7_state(
-      data = tests_biweekly_state
+      data = tests_biweekly_state,
+      beta_shape = get_shape(ctis_smoothed,
+                             option="beta"),
+      s_untested_shape = get_shape(ctis_smoothed,
+                                   option="s_untested")
     )
   ),
   
@@ -208,7 +212,11 @@ list(
   tar_target(
     name = ma_v7,
     command = get_county_v7(
-      data = ma_biweekly_county
+      data = ma_biweekly_county,
+      beta_shape = get_shape(ctis_smoothed,
+                             option="beta"),
+      s_untested_shape = get_shape(ctis_smoothed,
+                                   option="s_untested")
     )
   ),
   

### 2023-06-27 R/01-data_cleaning.R
fix biweekly covidestim cleaning

@@ -119,7 +119,7 @@ get_covidestim_state_biweekly <- function(end_date = "2022-02-25") {
   covidestim_biweekly <- covidestim %>%
     left_join(biweek_to_week) %>%
     group_by(biweek, state)  %>%
-    mutate(across(
+    summarize(across(
       contains("infections"), 
       sum)) %>%
     ungroup()
@@ -133,8 +133,7 @@ get_covidestim_state_biweekly <- function(end_date = "2022-02-25") {
   
   covidestim_biweekly <- covidestim_biweekly %>%
     rename(state_name = state) %>%
-    left_join(statecodes) %>%
-    select(-week)
+    left_join(statecodes) 
   
   return(covidestim_biweekly)
   

### 2023-06-27 R/04-melding.R
option to handle alpha,beta params rather than alpha and beta

@@ -15,34 +15,58 @@ get_melded <- function(alpha_mean = 0.9,
                        p_s0_pos_bounds = NA,
                        pre_nsamp = 1e4,
                        post_nsamp = 1e3,
-                       include_corrected = TRUE,
                        bde = FALSE,
-                       quiet=TRUE) {
+                       quiet=TRUE,
+                       direct_params=FALSE,
+                       beta_shape1=NULL,
+                       beta_shape2=NULL,
+                       s_untested_shape1=NULL,
+                       s_untested_shape2) {
   
   
   given_args <- as.list(environment())
   
-  if (!quiet) {
+  if(!quiet) {
     cat("Arguments to get_melded:\n")
     print(given_args)
   } 
   
-  
-  theta <- tibble(alpha = sample_gamma_density(pre_nsamp,
-                                               mean = alpha_mean,
-                                               sd = alpha_sd,
-                                               bounds = alpha_bounds),
-                  beta= sample_beta_density(pre_nsamp,
-                                            mean = beta_mean,
-                                            sd = beta_sd,
-                                            bounds = beta_bounds),
-                  P_S_untested = sample_beta_density(pre_nsamp,
-                                                     mean = s_untested_mean,
-                                                     sd = s_untested_sd,
-                                                     bounds = s_untested_bounds)) %>%
-    mutate(phi_induced = est_P_A_testpos(P_S_untested = P_S_untested,
-                                         alpha = alpha,
-                                         beta=beta))
+  if(!direct_params) {
+    
+    theta <- tibble(alpha = sample_gamma_density(pre_nsamp,
+                                                 mean = alpha_mean,
+                                                 sd = alpha_sd,
+                                                 bounds = alpha_bounds),
+                    beta= sample_beta_density(pre_nsamp,
+                                              mean = beta_mean,
+                                              sd = beta_sd,
+                                              bounds = beta_bounds),
+                    P_S_untested = sample_beta_density(pre_nsamp,
+                                                       mean = s_untested_mean,
+                                                       sd = s_untested_sd,
+                                                       bounds = s_untested_bounds)) %>%
+      mutate(phi_induced = est_P_A_testpos(P_S_untested = P_S_untested,
+                                           alpha = alpha,
+                                           beta=beta))
+    
+  }
+  if(direct_params) {
[…]

### 2023-06-27 R/05-state-analysis.R
get alpha and beta parameters based on given quantiles

@@ -519,23 +519,82 @@ get_v6_corrected <- function(state_testing, params) {
   
 }
 
+objective <- function(shapes, x, prob, ...) {
+  fit <- pbeta(x, shapes[1], shapes[2])
+  # return sum of squares
+  return(sum((fit-prob)^2))
+}
+
+
+
+get_shape <- function(ctis_smoothed, option, quiet=FALSE) {
+  
+  if (option=="beta") {
+    ctis_quantiles <- ctis_smoothed %>%
+      summarize(q1 = quantile(beta_estimate_spline_smoothed,.025, na.rm=TRUE),
+                median = median(beta_estimate_spline_smoothed, na.rm=TRUE),
+                q2 = quantile(beta_estimate_spline_smoothed,.975, na.rm=TRUE),
+                sd=sd(beta_estimate_spline_smoothed,na.rm=TRUE))
+    
+
+  }
+  if (option=="s_untested") {
+    ctis_quantiles <- ctis_smoothed %>%
+      summarize(q1 = quantile(s_untested_smoothed,.025, na.rm=TRUE),
+                median = median(s_untested_smoothed, na.rm=TRUE),
+                q2 = quantile(s_untested_smoothed,.975, na.rm=TRUE),
+                sd=sd(s_untested_smoothed, na.rm=TRUE))
+    
+  }
+  
+  start <- get_beta_params(ctis_quantiles$q1, ctis_quantiles$sd) %>%
+    unlist()
+  
+  optim_results <- nlm(objective, start, 
+             x=c(ctis_quantiles$q1, ctis_quantiles$median, ctis_quantiles$q2),
+             prob=c(.025,.5,.975), lower=0, upper=1,
+             typsize=c(1,1), fscale=1e-14, gradtol=1e-14)
+  
+  params <- (optim_results$estimate) 
+  
+  if(!quiet) {
+    message(paste0("Input quantiles (0.025, .5, .975):\n",
+                   round(ctis_quantiles$q1,4), ", ",
+                   round(ctis_quantiles$median,4), ", ",
+                   round(ctis_quantiles$q2,4)))
+    sim <- rbeta(1e4, params[1], params[2])
+    output_quantiles <- list(q1=quantile(sim,.025),
+                          median=median(sim),
+                          q2=quantile(sim, .975))
+    message(paste0("Output quantiles (0.025, .5, .975):\n",
[…]

### 2023-06-27 R/06-county-analysis.R

@@ -421,20 +421,19 @@ get_v6_county_corrected <- function(county_testing, params) {
 
 
 #----------------- Version 7 ------------------------
-#---------- Version 1 -------------
-get_county_v7 <- function(data, testing = FALSE) {
+get_county_v7 <- function(data, beta_shape, s_untested_shape, testing = FALSE) {
   
   params <-  list(
     alpha_mean = .95,
     alpha_sd = 0.08,
     alpha_bounds = NA,
     # alpha_bounds = c(.8,1),
-    beta_mean = .18,
-    beta_sd =.09,
+    beta_shape1 = beta_shape[1],
+    beta_shape2 = beta_shape[2],
     beta_bounds = NA,
     #  beta_bounds = c(0.002, 0.4),
-    s_untested_mean = .016,
-    s_untested_sd = .0225,
+    s_untested_shape1 = s_untested_shape[1],
+    s_untested_shape2 = s_untested_shape[2],
     #  s_untested_bounds = c(0.0018, Inf),
     s_untested_bounds = NA,
     p_s0_pos_mean = .4,

### 2023-06-27 reports/county_figures.Rmd
update reports

@@
          "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
-        "$\\,$CTIS Priors that Do Not Vary by Date$\\,$"
+        "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$"
     ),
@@
          "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
-        "$\\,$CTIS Priors that Do Not Vary by Date$\\,$"
+        "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$"
 
@@
   left_join(county_names, by=c('fips'='fips')) %>%
+  left_join(versions) %>%
   mutate(name = ifelse(fips =="25007,25019", "Nantucket and Dukes", name))
@@
   facet_wrap(~name, scales="free_y", ncol=3) +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(
+          legend.position = "top",
+          ) +
+  scale_fill_manual(values = pal)  +
+  labs(x="",
+       y = "Estimated Infections",
+       title = "Simulation Intervals Over Time",
+       subtitle = "Counties in Massachusetts") +
+  scale_color_manual(values = c('obs' = 'red'), labels = 'Positive Tests',
+                     name = '') +
+  guides(color = guide_legend(override.aes = list(linewidth =2)))
+
+```
+
+```{r}
+
+county_dat %>%
+  filter(fips %in% sample(unique(.data$fips),2))  %>% 
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub),
+             alpha = .5,
+             show.legend=FALSE,
+             fill="#596281") +
+   geom_linerange(aes(xmin = mindate,
+                      xmax=maxdate,
+                      y = positive,
+                      color = 'obs'),
+                  size=.5) +
+  facet_grid(name~vlabel, 
+             labeller=labeller(vlabel =as_labeller(
+               TeX, default=label_parsed)),
+             scales="free_y") +
   scale_y_continuous(labels = comma) +

### 2023-06-27 reports/priors.Rmd
update reports

--- 2023-06-24 reports/priors.Rmd
@@
        title = TeX("Empirical Estimates of $\\beta$ from the COVID-19 Trends and Impact Survey"),
-       subtitle = "2.7% Percentile, 97.5% Percentile, and Mean By State" ) +
+       subtitle = "2.7% Percentile, 97.5% Percentile, and Median By State" ) +
   theme_c(axis.title=element_text(size=12))
@@
 ```{r}
+#
+# Logistic transformation of the Beta CDF.
+#
+f.beta <- function(alpha, beta, x, lower=0, upper=1) {
+  p <- pbeta(x, alpha, beta)
+ log(p/(1-p))
+}
+#
+# Sums of squares.
+#
+delta <- function(fit, actual) sum((fit-actual)^2)
+#
+# The objective function handles the transformed parameters `theta` and
+# uses `f.beta` and `delta` to fit the values and measure their discrepancies.
+#
+objective <- function(theta, x, prob, ...) {
+  ab <- exp(theta) # Parameters are the *logs* of alpha and beta
+  fit <- f.beta(ab[1], ab[2], x, ...)
+  return (delta(fit, prob))
+}
+#
+# Solve two problems.
+#
+par(mfrow=c(1,2))
+alpha <- 15; beta <- 22 # The true parameters
+for (x in list(c(1e-3, 2e-3), c(1/3, 2/3))) {
+  x.p <- f.beta(alpha, beta, x)        # The correct values of the p_i
+  start <- log(c(1e1, 1e1))            # A good guess is useful here
+  sol <- nlm(objective, start, x=x, prob=x.p, lower=0, upper=1,
+             typsize=c(1,1), fscale=1e-12, gradtol=1e-12)
+  parms <- exp(sol$estimate)           # Estimates of alpha and beta
+  #
+  # Display the actual and estimated values.
+  #
+  print(rbind(Actual=c(alpha=alpha, beta=beta), Fit=parms))
+  #
+  # Plot the true and estimated CDFs.
+  #      
+  curve(pbeta(x, alpha, beta), 0, 1, n=1001, lwd=2)
+  curve(pbeta(x, parms[1], parms[2]), n=1001, add=TRUE, col="Red")
+  points(x, pbeta(x, alpha, beta))
+}
+```
+
+```{r}
+#
+# Logistic transformation of the Beta CDF.
+#
+f.beta <- function(alpha, beta, x, lower=0, upper=1) {
+  p <- pbeta(x, alpha, beta)
+ return(p)
+}
+#
+# Sums of squares.
+#
+delta <- function(fit, actual) sum((fit-actual)^2)
+#
+# The objective function handles the transformed parameters `theta` and
+# uses `f.beta` and `delta` to fit the values and measure their discrepancies.
+#
+objective <- function(shapes, x, prob, ...) {
+  fit <- pbeta(x, shapes[1], shapes[2], ...)
+  # return sum of squares
+  return(sum((fit, prob)^2))
+}
+
+# Solve two problems.
+
+par(mfrow=c(1,2))
+alpha <- 15; beta <- 22 # The true parameters
+for (x in list(c(1e-3, 2e-3), c(1/3, 2/3))) {
+  x.p <- f.beta(alpha, beta, x)        # The correct values of the p_i
+  start <- (c(1e1, 1e1))            # A good guess is useful here
+  sol <- nlm(objective, start, x=x, prob=x.p, lower=0, upper=1,
+             typsize=c(1,1), fscale=1e-12, gradtol=1e-12)
+  parms <- (sol$estimate)           # Estimates of alpha and beta
+  #
+  # Display the actual and estimated values.
+  #
+  print(rbind(Actual=c(alpha=alpha, beta=beta), Fit=parms))
+  #
+  # Plot the true and estimated CDFs.
+  #      
+  curve(pbeta(x, alpha, beta), 0, 1, n=1001, lwd=2)
+  curve(pbeta(x, parms[1], parms[2]), n=1001, add=TRUE, col="Red")
+  points(x, pbeta(x, alpha, beta))
+}
+
+for (x in list(c(.1200,.2188,.3683))) {
+  # x.p <- f.beta(alpha, beta, x)        # The correct values of the p_i
+  start <- (c(1e1, 1e1))            # A good guess is useful here
+  sol <- nlm(objective, start, x=x, prob=c(.025,.5,.975), lower=0, upper=1,
+             typsize=c(1,1), fscale=1e-12, gradtol=1e-13)
+  parms <- (sol$estimate)           # Estimates of alpha and beta
+  #
+  # Display the actual and estimated values.
+  #
+  # print(rbind(Actual=c(alpha=alpha, beta=beta), Fit=parms))
+  #
+  # Plot the true and estimated CDFs.
+  #      
+  # curve(pbeta(x, alpha, beta), 0, 1, n=1001, lwd=2)
+  # curve(pbeta(x, parms[1], parms[2]), n=1001, add=TRUE, col="Red")
+  # points(x, pbeta(x, alpha, beta))
+}
+parms
+
+rbeta(1e4, parms[1], parms[2]) %>% median()
+rbeta(1e4, parms[1], parms[2]) %>% quantile(.025)
+rbeta(1e4, parms[1], parms[2]) %>% quantile(.975)
+
+rbeta(1e3, parms[1], parms[2]) %>% mean()
+
+```
+
+```{r, fig.height=7}
+
+ctis %>%
+  group_by(state) %>%
+  summarize(median = median(beta, na.rm=TRUE),
+            mean = mean(beta, na.rm=TRUE),
+            q1 = quantile(beta, .025,  na.rm=TRUE),
+            q2 = quantile(beta, .975,  na.rm=TRUE)) %>%
+  ggplot(aes(y=fct_reorder(state, mean, .desc=TRUE))) +
+  ggridges::geom_density_ridges(
+    aes(x=beta,
+        y=fct_reorder(state, beta, .fun =median)),
+    data=ctis,
+    fill="#596281",
+    quantile_lines=TRUE,
+    quantiles = c(.025,.975),
+    quantile_fun=function(beta,...) c(
+                                    #  quantile(beta,.025, na.rm=TRUE),
+                                      median(beta,na.rm=TRUE)
+                                    #  quantile(beta,.975, na.rm=TRUE)
+                                      )) +
+  # geom_errorbar(aes( xmin=q1,xmax=q2),width=.2) +
+  scale_x_continuous(n.breaks=10, limits=c(0,1)) +
+  labs(x= TeX("Empirical Estimates of $\\beta$"),
+       y="State",
+       title = TeX("Empirical Estimates of $\\beta$ from the COVID-19 Trends and Impact Survey"),
+       subtitle = "Density Estimate of Distribution and Median by State" ) +
+  theme_c(axis.title=element_text(size=12),
+          axis.text.x=element_text(size=11))
+
+```
+
+```{r}
 
@@
   ggplot(aes(x = beta, fill=state)) +
-  geom_density() +
+  geom_density(alpha=.6) +
   scale_fill_viridis(option="mako", discrete=TRUE) +
   theme_c()
+
+```
+
+```{r,include=FALSE}
 

### 2023-06-27 reports/state_figures.Rmd
update reports

@@
 library(here)
+library(scales)
 library(latex2exp)
+library(cowplot)
 
@@
 # read state-level results
-targets_dir <- here("_targets")
+targets_dir <- here("_targets_6_26")
 
@@
                    store =targets_dir) %>% 
-  mutate(version = "v1") %>%
   bind_rows(
@@
   left_join(covidestim, relationship= "many-to-many") %>%
-  left_join(dates, relationship = "many-to-many")
+  left_join(dates, relationship = "many-to-many")  %>%
+  rename(name=state_name) %>%
+  group_by(biweek) %>%
+  mutate(mindate=min(date),
+         maxdate=max(date)) %>%
+  ungroup()
+
+state_res <- state_res %>%
+  filter(version %in% c("v2", "v5", "v6", "v7"))
 
@@
 
-# Simulation Intervals
+# Simulation Intervals Over Time
+
+```{r}
+
+all_versions <- state_res %>% 
+  group_by(state) %>%
+  summarize(state_n=n_distinct(version)) %>%
+  filter(state_n == max(state_n)) %>%
+  pull(state)
+```
+
+```{r, eval=FALSE}
+
+state_res %>%
+   filter(state %in% sample(unique(all_versions),5))  %>% 
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub),
+             alpha = .8,
+             show.legend=FALSE,
+             fill="#596281") +
+   geom_linerange(aes(xmin = mindate,
+                      xmax=maxdate,
+                      y = positive,
+                      color = 'obs'),
+                  size=.5) +
+  facet_grid(name~vlabel, 
+             labeller=labeller(vlabel =as_labeller(
+               TeX, default=label_parsed)),
+             scales="free_y") +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(
+          legend.position = "top",
+          ) +
+  scale_fill_manual(values = pal)  +
+  labs(x="",
+       y = "Estimated Infections",
+       title = "Simulation Intervals Over Time",
+       subtitle = "By State") +
+  scale_color_manual(values = c('obs' = 'red'), labels = 'Positive Tests',
+                     name = '') +
+  guides(color = guide_legend(override.aes = list(linewidth =2)))
+
+```
+
+## Compare Versions
 
@@
 
+pal <- c("#247C90", "red")
+
+names(pal) <- c(unique(subset$vlabel), "Covidestim")
+
+state_res %>%
+   filter(state %in% sample(unique(all_versions),5))  %>% 
+  filter(version %in% c("v2", "v5", "v6", "v7")) %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub,
+             fill=vlabel),
+             alpha = .8,
+             show.legend=FALSE #,
+           #  fill="#596281"
+             ) +
+   geom_linerange(aes(xmin = mindate,
+                      xmax=maxdate,
+                      y = positive,
+                      color = 'obs'),
+                  size=.5) +
+  facet_grid(name~vlabel, 
+             labeller=labeller(vlabel =as_labeller(
+               TeX, default=label_parsed)),
+             scales="free_y") +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(
+          legend.position = "top",
+          ) +
+  scale_fill_manual(values = pal)  +
+  labs(x="",
+       y = "Estimated Infections",
+       title = "Simulation Intervals Over Time",
+       subtitle = "By State") +
+  scale_color_manual(values = c('obs' = 'red'), labels = 'Positive Tests',
+                     name = '') +
+  guides(color = guide_legend(override.aes = list(linewidth =2)))
+
 ```
 
-# Ratio of Estimated to Observed During Omicron
+## Compare States
 
-```{r,results='hide', fig.height=11,fig.width=8}
+```{r, fig.height=11, fig.height=9}
 
-subset <- state_res %>%
-  filter(vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$")
+subset %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub,
+             fill = vlabel),
+             alpha = .8,
+             show.legend=FALSE) +
+  geom_ribbon(aes(x = date,
+                  fill = 'Covidestim',
+                  ymin = infections.lo,
+                  ymax = infections.hi),
+              alpha = .6) +
+  facet_wrap(~state, ncol=5, scales = "free_y",
+              labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(axis.text.x=element_text(size=6),
+          axis.text.y = element_text(size=8),
+          legend.position = "top") +
+  scale_fill_manual(values =pal,
+                    labels = TeX(names(pal)),
+                    name='',
+                    breaks="Covidestim")  +
+  labs(y = "Estimated Infections",
+       title = paste0("Estimated Infections by State")) +
+  guides(color = guide_legend(override.aes = list(linewidth =2),
+                              nrow=2,
+                              byrow=TRUE))
+
+ggsave(here('figures/intervals-and-covidestim-all-states.pdf'))
+
+```
+
+## Ratio of Estimated to Observed
+
+```{r, fig.height=10,fig.width=12}
+
+subset %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb/positive,
+             ymax = exp_cases_ub/positive,
+             fill = vlabel),
+             alpha = .9,
+             show.legend=FALSE,
+             fill= "#247C90") +
+  facet_wrap(~state, ncol=5, 
+              labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(axis.text.x=element_text(size=6),
+          axis.text.y = element_text(size=8),
+          legend.position = "top") +
+  scale_fill_manual(values =pal,
+                    labels = TeX(names(pal)),
+                    name='',
+                    breaks="Covidestim")  +
+  labs(y = "Ratio of Estimated Infections to Observed",
+       x="",
+       title = paste0("Ratio of Estimated Infections to Observed by State")) +
+  guides(color = guide_legend(override.aes = list(linewidth =2),
+                              nrow=2,
+                              byrow=TRUE))
+
+ggsave(here('figures/ratio-estimated-to-observed-all-states.pdf'))
+
+```
+
+# Comparing States at Specific Two-week Intervals
+
+## During Alpha Wave
+
+```{r, fig.height=9,fig.width=8}
+
+plt1 <- subset %>% 
+  # left_join(codes) %>%
+  filter(biweek == 7) %>%
+  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
+             xmin = exp_cases_lb/positive,
+             xmax = exp_cases_ub/positive)) +
+  geom_errorbar() +
+  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
+  #            color = "darkred", size=.8, alpha=.5) +
+  theme_c() + 
+  theme(axis.text.y=element_text(size=10),
+        axis.text.x = element_text(size=11),
+        axis.title = element_text(size = 17),
+        plot.subtitle = element_text(size=14),
+        plot.title = element_text(size=17, face="plain")) +
+  labs(y="", x="Ratio of Estimated Infections\nto Observed",
+       title ="During Two-week Period\nin the Alpha Wave",
+       subtitle = "March 26, 2021 through April 8, 2021")
+
+plt1
+
+```
+
+## During Delta Wave
+
+```{r, fig.height=9,fig.width=8, results='hide'}
+
+subset %>%
+  mutate(ratio = exp_cases_median/positive) %>%
+  group_by(biweek) %>%
+  summarize(mean = mean(ratio),
+            mindate=min(date),
+            maxdate=max(date)) %>%
+  arrange(desc(mean))
+
+```
+
+```{r, fig.height=9,fig.width=8}
+
+plt2 <- subset %>% 
+  # left_join(codes) %>%
+  filter(biweek == 13) %>%
+  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
+             xmin = exp_cases_lb/positive,
+             xmax = exp_cases_ub/positive)) +
+  geom_errorbar() +
+  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
+  #            color = "darkred", size=.8, alpha=.5) +
+  theme_c() + 
+ theme(axis.text.y=element_text(size=10),
+        axis.text.x = element_text(size=11),
+        axis.title = element_text(size = 17),
+        plot.subtitle = element_text(size=14),
+        plot.title = element_text(size=17, face="plain")) +
+  labs(y="", x="Ratio of Estimated Infections\nto Observed",
+       title ="During Two-week Period\nin the Delta Wave",
+       subtitle = "June 18, 2021 through July 1, 2021")
+
+plt2
+
+```
+
+## During Omicron Wave
+
+```{r,results='hide', fig.height=9,fig.width=8}
 
@@
   left_join(codes) %>%
+  filter(biweek == 13) %>%
+  pull(date) %>% 
+  range()
+
+plt3 <- subset %>% 
+  # left_join(codes) %>%
   filter(biweek == 27) %>%
-  ggplot(aes(y =fct_reorder(state_name,exp_cases_median/positive),
+  ggplot(aes(y =fct_reorder(name,exp_cases_median/positive),
              xmin = exp_cases_lb/positive,
@@
   theme_c() + 
-  theme(axis.text.y=element_text(size=15),
-        axis.text.x = element_text(size=13),
-        axis.title = element_text(size = 18),
-        plot.subtitle = element_text(size=14)) +
-  labs(y="", x="Ratio of Estimated Infections to Observed",
-       title ="Ratio of Estimated Infections to Observed Infections\nDuring Omicron",
+  theme(axis.text.y=element_text(size=10),
+        axis.text.x = element_text(size=11),
+        axis.title = element_text(size = 17),
+        plot.subtitle = element_text(size=14),
+        plot.title = element_text(size=17, face="plain")) +
+  labs(y="", x="Ratio of Estimated Infections\nto Observed",
+       title ="During Two-week Period\nin the Omicron Wave",
        subtitle = "December 31, 2021 through January 14, 2022")
 
+plt3
+
 ```
+
+## All Waves Together
+
+```{r, fig.height=10, fig.width=17}
+
+plt <- plot_grid(plt1,plt2, plt3, nrow=1, align="none")
+
+title <- ggplot() +
+  labs(title = "Ratio of Estimated Infections to Observed Infections") +
+  theme_c(plot.title=element_text(size=25))
+
+plot_grid(title, plt, nrow=2, rel_heights= c(.05,.95))
+
+```

### 2023-06-30 writing/supplement.Rmd
update reports
update supplement

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
 library(knitr)
@@
 
-# Sensitivity Analyses
+# Details on Each Implementation
+
+```{r covidestim-concordance-state, fig.cap="For each implementation, the proportions of simulation intervals where the Covidestim median was below, within, or above the interval are shown, across all states used for the implementation."}
+
+include_graphics(here('figures/concordance-covidestim-state.pdf'))
+
+```
+
+```{r, fig.cap="The proportions of simulation intervals where the Covidestim median was below, within, or above the intervals are shown, considering all counties except Dukes/Nantucket, where Covidestim estimates were not produced."}
+
+include_graphics(here('figures/concordance-covidestim-county.pdf'))
+
+```
+
+\newpage
+
+# Full Set of Simulation Intervals by State
+
+```{r sim-intervals-state}
+
+include_graphics(here('figures/intervals-and-covidestim-all-states.pdf'))
+
+```
+
+## COVID-19 Trends and Impact Survey
+
+# County Results by Implementation

### 2023-06-30 reports/priors.Rmd
comparison to CTIS priors

@@
 Here, we consider the 2.5\\% and 97.5\\% percentiles and mean for each state.
-Most have a mean near $20%$.", fig.width=9}
+Most have a mean near 0.2.", fig.width=9}
 
@@
 }
-```
-
-```{r}
-#
-# Logistic transformation of the Beta CDF.
-#
-f.beta <- function(alpha, beta, x, lower=0, upper=1) {
-  p <- pbeta(x, alpha, beta)
- return(p)
-}
-#
-# Sums of squares.
-#
-delta <- function(fit, actual) sum((fit-actual)^2)
-#
-# The objective function handles the transformed parameters `theta` and
-# uses `f.beta` and `delta` to fit the values and measure their discrepancies.
-#
-objective <- function(shapes, x, prob, ...) {
-  fit <- pbeta(x, shapes[1], shapes[2], ...)
-  # return sum of squares
-  return(sum((fit, prob)^2))
-}
-
-# Solve two problems.
-
-par(mfrow=c(1,2))
-alpha <- 15; beta <- 22 # The true parameters
-for (x in list(c(1e-3, 2e-3), c(1/3, 2/3))) {
-  x.p <- f.beta(alpha, beta, x)        # The correct values of the p_i
-  start <- (c(1e1, 1e1))            # A good guess is useful here
-  sol <- nlm(objective, start, x=x, prob=x.p, lower=0, upper=1,
-             typsize=c(1,1), fscale=1e-12, gradtol=1e-12)
-  parms <- (sol$estimate)           # Estimates of alpha and beta
-  #
-  # Display the actual and estimated values.
-  #
-  print(rbind(Actual=c(alpha=alpha, beta=beta), Fit=parms))
-  #
-  # Plot the true and estimated CDFs.
-  #      
-  curve(pbeta(x, alpha, beta), 0, 1, n=1001, lwd=2)
-  curve(pbeta(x, parms[1], parms[2]), n=1001, add=TRUE, col="Red")
-  points(x, pbeta(x, alpha, beta))
-}
-
-for (x in list(c(.1200,.2188,.3683))) {
-  # x.p <- f.beta(alpha, beta, x)        # The correct values of the p_i
-  start <- (c(1e1, 1e1))            # A good guess is useful here
-  sol <- nlm(objective, start, x=x, prob=c(.025,.5,.975), lower=0, upper=1,
-             typsize=c(1,1), fscale=1e-12, gradtol=1e-13)
-  parms <- (sol$estimate)           # Estimates of alpha and beta
-  #
-  # Display the actual and estimated values.
-  #
-  # print(rbind(Actual=c(alpha=alpha, beta=beta), Fit=parms))
-  #
-  # Plot the true and estimated CDFs.
-  #      
-  # curve(pbeta(x, alpha, beta), 0, 1, n=1001, lwd=2)
-  # curve(pbeta(x, parms[1], parms[2]), n=1001, add=TRUE, col="Red")
-  # points(x, pbeta(x, alpha, beta))
-}
-parms
-
-rbeta(1e4, parms[1], parms[2]) %>% median()
-rbeta(1e4, parms[1], parms[2]) %>% quantile(.025)
-rbeta(1e4, parms[1], parms[2]) %>% quantile(.975)
-
-rbeta(1e3, parms[1], parms[2]) %>% mean()
-
 ```
@@
 ```
+
+## Compare Priors
+
+```{r}
+
+source(here("R/05-state-analysis.R"))
+
+beta_shape = get_shape(ctis_smoothed,
+                             option="beta",
+                       quiet=TRUE)
+
+tibble(x = seq(0, 1, length = 10^(4)),
+       original = with(prior_params, beta_density(x,
+                                     mean = beta_mean,
+                                     sd = beta_sd,
+                                     bounds = beta_bounds)),
+       ctis= dbeta(x, shape1=beta_shape[1], 
+                   shape2=beta_shape[2])) %>%
+  pivot_longer(c(original,ctis))%>%
+   ggplot(aes(x=x, y = value, fill=name)) +
+  geom_line(alpha = .8) + 
+  geom_ribbon(aes(x=x,ymin=0, ymax=value,fill=name),
+              alpha=.6) +
+  theme_c(legend.text=element_text(size = 10)) +
+  viridis::scale_fill_viridis(discrete=TRUE,
+                             option = "rocket", begin=.3,end=.8) +
+  labs(x = "Value",
+       y = "Probability Density", 
+       title = latex2exp::TeX("Definition of Prior for $\\beta$"),
+       fill ='')
+
+```
+
+```{r}
+
+s_untested_shape = get_shape(ctis_smoothed,
+                             option="s_untested",
+                       quiet=TRUE)
+
+tibble(x = seq(0, 1, length = 10^(4)),
+       original = with(prior_params, beta_density(x,
+                                     mean = s_untested_mean,
+                                     sd = s_untested_sd,
+                                     bounds = s_untested_bounds)),
+       ctis= dbeta(x, shape1=s_untested_shape[1], 
+                   shape2=s_untested_shape[2])) %>%
+  pivot_longer(c(original,ctis))%>%
+   ggplot(aes(x=x, y = value, fill=name)) +
+  geom_line(alpha = .8) + 
+  geom_ribbon(aes(x=x,ymin=0, ymax=value,fill=name),
+              alpha=.6) +
+  theme_c(legend.text=element_text(size = 10)) +
+  viridis::scale_fill_viridis(discrete=TRUE,
+                             option = "rocket", begin=.3,end=.8) +
+  labs(x = "Value",
+       y = "Probability Density", 
+       title = latex2exp::TeX("Definition of Prior for $Pr(S_1|untested)$"),
+       fill ='')
+
+```

### 2023-06-30 writing/manuscript.Rmd
update writing/supplement

@@
     number_sections: false
+header-includes:
+ \usepackage{float}
 
 ---
+
+```{r setup, include=FALSE}
+knitr::opts_chunk$set(echo = FALSE,  message=FALSE,
+                      warning=FALSE, fig.pos="H",
+                      out.extra = "")
+library(knitr)
+library(here)
+
+```
 
@@
 
-## County Level in Massachusetts
+## County Level Estimates in Massachusetts
 
-## State-level (1 version)
+### Comparison to Wastewater Data and Covidestim Estimates
+
+Because there is no established ground truth to compare to regarding the true number of infections for any time-interval, at the county level we compare our results to two distincts sources of information: wastewater data aggregated at the county-level, and results from a previously published Bayesian evidence synthesis model, Covidestim.
+This also allows us to compare how different implementations of probabilistic bias analysis compare to Covidestim estimates and wastewater concentrations.
+
+Our first comparison considers the correlations between wastewater concentrations and the probabilistic bias estimates as well as between the Covidestim estimates and the probabilistic bias estimates (Figure \@ref(fig:correlations)).
+
+```{r correlations, fig.cap ="Considering the correlations between probablistic bias estimates and wastewater concentrations (a) and between probablistic bias estimates and Covidestim estimates (b).
+We see that all implementations considered are highly correlated with both wastewater concentrations and Covidestim estimates.
+The implementation that does not allow priors to vary by state or date has the highest correlations."}
+
+include_graphics(
+  here('figures/correlations-wastewater-by-version.pdf')
+  )
+
+```
+
+NOTE add (a), (b) to this figure
+
+```{r}
+
+include_graphics(here('figures/simulation-intervals-county.pdf'))
+
+```
+
+## State-level Estimates
+
+For simplicity, at the state-level we focus on the implementation that does not vary by state or date.
+This also allows us to consider the entirety of 2021, since survey data is only available for dates after March 20, 2021.
+A full comparison of implementations is included in Supplementary Figure covidestim-concordance-state.
+
+In Figure \@ref(fig:ratio-estimated-to-observed-by-wave), we consider three distinct two-week intervals during waves of the pandemic in 2021.
+Simulation intervals for all two-week intervals for all states are shown in Supplementary Figure sim-intervals-state.
+
+Although prevalence of COVID-19 was highest during the time interval during the Omicron wave, we see that the ratio of estimated infections to observed infections is higher during the time intervals in the alpha and delta waves.
+This distinction is explained by the differences in testing rates during these period: the testing rate during this two-week interval during the omicron wave was 2.4 times that of the alpha wave and 4.9 times that of the delta wave.
+
+Several states consistently have among the highest or lowest ratios of estimated to observed infections.
+In particular, there are 6 states with among the lowest 10 ratios of estimated infections to observed infections, and as such the highest case ascertainment, for more than 80% of time intervals considered.
+These states were Rhode Island, Massachusetts, District of Columbia, Alaska, New York, and Vermont.
+Meanwhile, states that had the highest ratios, and equivalently the lowest case ascertainment, include Mississippi, South Dakota, Oklahoma, Nebraska, and Tennessee.
+
+```{r ratio-estimated-to-observed-by-wave, fig.cap="The ratio of estimated infections to observed infections for three time intervals of interest: one during the alpha wave, one during the delta wave, and one during the omicron wave.
+Although the prevalence of COVID-19 was highest during the omicron wave, the ratios of estimated to observed infections are higher for the time intervals during the alpha and delta waves, a difference that was driven by lower testing rates during these times.
+The trend we see in these three time intervals where Mississippi, South Dakota, Oklahoma, Nebraska, and Tennessee have among the highest ratios of estimated infections to observed infections, and as such the lowest case ascertainment rates, is consistent across the full set of time intervals considered from January of 2021 to March of 2022."}
+
+include_graphics(here('figures/ratio-estimated-to-observed-multiple-waves.pdf'))
+
+```
 
@@
 ### Massachusetts County Level
+
 ### State Level
+
 ### Survey Data
+
+The COVID-19 Trends and Impact Survey was run in collaboration by ...
+
+### Wastewater Data
+
+Biobot analytics ...
 

### 2023-06-30 R/05-state-analysis.R
don't smooth for v7

@@ -527,7 +527,7 @@ objective <- function(shapes, x, prob, ...) {
 
 
 
-get_shape <- function(ctis_smoothed, option, quiet=FALSE) {
+get_shape_smoothed <- function(ctis_smoothed, option, quiet=FALSE) {
   
   if (option=="beta") {
     ctis_quantiles <- ctis_smoothed %>%
@@ -566,6 +566,71 @@ get_shape <- function(ctis_smoothed, option, quiet=FALSE) {
     output_quantiles <- list(q1=quantile(sim,.025),
                           median=median(sim),
                           q2=quantile(sim, .975))
+    message(paste0("Output quantiles (0.025, .5, .975):\n",
+                   round(output_quantiles$q1,4), ", ",
+                   round(output_quantiles$median,4), ", ",
+                   round(output_quantiles$q2,4)))
+    
+    
+  }
+  
+  return(params)
+  
+  
+}
+
+
+
+
+
+get_shape <- function(ctis_smoothed, option, quiet=FALSE) {
+  
+  dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
+  
+  if (option=="beta") {
+    ctis_quantiles <- ctis_smoothed %>% 
+      left_join(dates) %>%
+      group_by(biweek,state) %>%
+      summarize(beta_est=mean(beta_est,na.rm=TRUE)) %>%
+      ungroup() %>%
+      summarize(q1 = quantile(beta_est,.025, na.rm=TRUE),
+                median = median(beta_est, na.rm=TRUE),
+                q2 = quantile(beta_est,.975, na.rm=TRUE),
+                sd=sd(beta_est,na.rm=TRUE))
+    
+  }
+  if (option=="s_untested") {
+    ctis_quantiles <- ctis_smoothed %>% 
+      left_join(dates) %>%
+      group_by(biweek,state) %>%
+      summarize(smoothed_wcli=mean(smoothed_wcli,na.rm=TRUE))  %>%
+      ungroup() %>%
+      summarize(q1 = quantile(smoothed_wcli,.025, na.rm=TRUE),
+                median = median(smoothed_wcli, na.rm=TRUE),
+                q2 = quantile(smoothed_wcli,.975, na.rm=TRUE),
+                sd=sd(smoothed_wcli, na.rm=TRUE))
+    
+  }
+  
+  start <- get_beta_params(ctis_quantiles$median, ctis_quantiles$sd) %>%
+    unlist()
+  
+  optim_results <- nlm(objective, start, 
[…]

### 2023-06-30 writing/manuscript.Rmd
update manuscript

@@
 
+## State-level Estimates
+
+For simplicity, at the state-level we focus on the implementation that does not vary by state or date.
+This also allows us to consider the entirety of 2021, since survey data is only available for dates after March 20, 2021.
+A full comparison of implementations is included in Supplementary Figure covidestim-concordance-state.
+
+In Figure \@ref(fig:ratio-estimated-to-observed-by-wave), we consider three distinct two-week intervals during waves of the pandemic in 2021.
+
+Although prevalence of COVID-19 was highest during the time interval during the Omicron wave, we see that the ratio of estimated infections to observed infections is higher during the time intervals in the alpha and delta waves.
+This distinction is explained by the differences in testing rates during these period: the testing rate during this two-week interval during the omicron wave was 2.4 times that of the alpha wave and 4.9 times that of the delta wave.
+
+Several states consistently have among the highest or lowest ratios of estimated to observed infections.
+In particular, there are 6 states with among the lowest 10 ratios of estimated infections to observed infections, and as such the highest case ascertainment rates, for more than 80% of time intervals considered.
+These states were Rhode Island, Massachusetts, District of Columbia, Alaska, New York, and Vermont.
+Meanwhile, states that had the highest ratios, and equivalently the lowest case ascertainment rates, include Mississippi, South Dakota, Oklahoma, Nebraska, and Tennessee.
+
+```{r ratio-estimated-to-observed-by-wave, fig.cap="The ratio of estimated infections to observed infections for three time intervals of interest: one during the alpha wave, one during the delta wave, and one during the omicron wave.
+Although the prevalence of COVID-19 was highest during the omicron wave, the ratios of estimated to observed infections are higher for the time intervals during the alpha and delta waves, a difference that was driven by lower testing rates during these times.
+The trend we see in these three time intervals where Mississippi, South Dakota, Oklahoma, Nebraska, and Tennessee have among the highest ratios of estimated infections to observed infections, and as such the lowest case ascertainment rates, is consistent across the full set of time intervals considered from January of 2021 to March of 2022."}
+
+include_graphics(here('figures/ratio-estimated-to-observed-multiple-waves.pdf'))
+
+```
+
+In Figure \@ref(fig:simulation-intervals-state), we see the simulation intervals for all two-week intervals and all states.
+
+```{r simulation-intervals-state, fig.cap="Simulation intervals for each 2-week interval considered, for all states.
+For any given state, each vertical bar shows the 2.5\\% percentile and 97.5\\% percentile for the total number of infections in that two-week interval.
+Covidestim intervals summed over the same two-week time-scale are shown in red.
+The scale on the $y$-axis is distinct across states to highlight differences across time within each state."}
+
+include_graphics(here('figures/intervals-and-covidestim-all-states.pdf'))
+
+```
+
 ## County Level Estimates in Massachusetts
@@
 
-## State-level Estimates
-
-For simplicity, at the state-level we focus on the implementation that does not vary by state or date.
-This also allows us to consider the entirety of 2021, since survey data is only available for dates after March 20, 2021.
-A full comparison of implementations is included in Supplementary Figure covidestim-concordance-state.
-
-In Figure \@ref(fig:ratio-estimated-to-observed-by-wave), we consider three distinct two-week intervals during waves of the pandemic in 2021.
-Simulation intervals for all two-week intervals for all states are shown in Supplementary Figure sim-intervals-state.
-
-Although prevalence of COVID-19 was highest during the time interval during the Omicron wave, we see that the ratio of estimated infections to observed infections is higher during the time intervals in the alpha and delta waves.
-This distinction is explained by the differences in testing rates during these period: the testing rate during this two-week interval during the omicron wave was 2.4 times that of the alpha wave and 4.9 times that of the delta wave.
-
-Several states consistently have among the highest or lowest ratios of estimated to observed infections.
-In particular, there are 6 states with among the lowest 10 ratios of estimated infections to observed infections, and as such the highest case ascertainment, for more than 80% of time intervals considered.
-These states were Rhode Island, Massachusetts, District of Columbia, Alaska, New York, and Vermont.
-Meanwhile, states that had the highest ratios, and equivalently the lowest case ascertainment, include Mississippi, South Dakota, Oklahoma, Nebraska, and Tennessee.
-
-```{r ratio-estimated-to-observed-by-wave, fig.cap="The ratio of estimated infections to observed infections for three time intervals of interest: one during the alpha wave, one during the delta wave, and one during the omicron wave.
-Although the prevalence of COVID-19 was highest during the omicron wave, the ratios of estimated to observed infections are higher for the time intervals during the alpha and delta waves, a difference that was driven by lower testing rates during these times.
-The trend we see in these three time intervals where Mississippi, South Dakota, Oklahoma, Nebraska, and Tennessee have among the highest ratios of estimated infections to observed infections, and as such the lowest case ascertainment rates, is consistent across the full set of time intervals considered from January of 2021 to March of 2022."}
-
-include_graphics(here('figures/ratio-estimated-to-observed-multiple-waves.pdf'))
-
-```
-
 # Methods

### 2023-06-30 R/05-state-analysis.R
update CTIS priors

@@ -520,70 +520,23 @@ get_v6_corrected <- function(state_testing, params) {
 }
 
 objective <- function(shapes, x, prob, ...) {
+  # parameters are log(shape1),log(shape2)
+  shapes <- exp(shapes)
   fit <- pbeta(x, shapes[1], shapes[2])
+  # logistic transformation
+  fit <- log(fit/(1-fit))
   # return sum of squares
   return(sum((fit-prob)^2))
 }
 
 
-
-get_shape_smoothed <- function(ctis_smoothed, option, quiet=FALSE) {
-  
-  if (option=="beta") {
-    ctis_quantiles <- ctis_smoothed %>%
-      summarize(q1 = quantile(beta_estimate_spline_smoothed,.025, na.rm=TRUE),
-                median = median(beta_estimate_spline_smoothed, na.rm=TRUE),
-                q2 = quantile(beta_estimate_spline_smoothed,.975, na.rm=TRUE),
-                sd=sd(beta_estimate_spline_smoothed,na.rm=TRUE))
-    
-
-  }
-  if (option=="s_untested") {
-    ctis_quantiles <- ctis_smoothed %>%
-      summarize(q1 = quantile(s_untested_smoothed,.025, na.rm=TRUE),
-                median = median(s_untested_smoothed, na.rm=TRUE),
-                q2 = quantile(s_untested_smoothed,.975, na.rm=TRUE),
-                sd=sd(s_untested_smoothed, na.rm=TRUE))
-    
-  }
-  
-  start <- get_beta_params(ctis_quantiles$q1, ctis_quantiles$sd) %>%
-    unlist()
-  
-  optim_results <- nlm(objective, start, 
-             x=c(ctis_quantiles$q1, ctis_quantiles$median, ctis_quantiles$q2),
-             prob=c(.025,.5,.975), lower=0, upper=1,
-             typsize=c(1,1), fscale=1e-14, gradtol=1e-14)
-  
-  params <- (optim_results$estimate) 
-  
-  if(!quiet) {
-    message(paste0("Input quantiles (0.025, .5, .975):\n",
-                   round(ctis_quantiles$q1,4), ", ",
-                   round(ctis_quantiles$median,4), ", ",
-                   round(ctis_quantiles$q2,4)))
-    sim <- rbeta(1e4, params[1], params[2])
-    output_quantiles <- list(q1=quantile(sim,.025),
-                          median=median(sim),
-                          q2=quantile(sim, .975))
-    message(paste0("Output quantiles (0.025, .5, .975):\n",
[…]

### 2023-06-30 _targets.R

@@ -135,6 +135,7 @@ list(
     command = get_v7_state(
       data = tests_biweekly_state,
       beta_shape = get_shape(ctis_smoothed,
+                             probs=c(.1,.5,.9),
                              option="beta"),
       s_untested_shape = get_shape(ctis_smoothed,
                                    option="s_untested")
@@ -214,6 +215,7 @@ list(
     command = get_county_v7(
       data = ma_biweekly_county,
       beta_shape = get_shape(ctis_smoothed,
+                             probs=c(.1,.5,.9),
                              option="beta"),
       s_untested_shape = get_shape(ctis_smoothed,
                                    option="s_untested")

### 2023-06-30 reports/county_figures.Rmd
figures

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,cache=TRUE, fig.width=10, fig.height=7)
+knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,cache=FALSE, fig.width=10, fig.height=7)
 
@@
 library(cowplot)
+
+filter_versions <- FALSE 
+save_figures <- FALSE
+
+include_other_norm <- FALSE
+
 ```
@@
 }
+
+pal <- c("red", "#10BAC5", "#1B10C5", "#EFB719", "#900C3F")
+
+names(pal) <- c("Covidestim",
+                '$P(S_1|untested)$ Centered at Survey Value',
+                '$P(S_1|untested)$ and $\\beta$ Centered at Survey Values',
+               "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
+               "$\\beta$ Centered at Survey Value"
+               )
 
@@
 
+if(filter_versions) {
+  county <- county %>%
+    filter(version %in% c("v3", "v5", "v6", "v7")) 
+}
+
 ```
@@
     c( "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
-    "$\\beta$ Centered at Emp. Value",
-    "$P(S_1|untested)$ Centered at Emp. Value",
-    "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
-        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
-         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
+    "$\\beta$ Centered at Survey Value",
+    "$P(S_1|untested)$ Centered at Survey Value",
+    "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+        "$\\beta$ Centered at Survey Value (Spline Smoothed)",
+         "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values ($\\beta$ Spline Smoothed)",
         "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$"
@@
         "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
-        "$\\beta$ Centered at Emp. Value",
-         "$P(S_1|untested)$ Centered at Emp. Value",
-        "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values",
-        "$\\beta$ Centered at Emp. Value (Spline Smoothed)",
-         "$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values ($\\beta$ Spline Smoothed)",
+        "$\\beta$ Centered at Survey Value",
+         "$P(S_1|untested)$ Centered at Survey Value",
+        "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+        "$\\beta$ Centered at Survey Value (Spline Smoothed)",
+         "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values ($\\beta$ Spline Smoothed)",
         "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$"
@@
   left_join(versions)
-
-w_data <- wjoined %>% 
-  select(-date) %>%
-  distinct() %>%
-  filter(fips %in% counties_with_waste) %>%
-  filter(!is.na(mean_conc)) %>%
-  group_by(fips, version) %>%
-  mutate(mean_conc_std = mean_conc/ max(mean_conc),
-         exp_cases_median_std = exp_cases_median / max(exp_cases_median),
-         exp_cases_lb_std = exp_cases_lb /max(exp_cases_median),
-         exp_cases_ub_std = exp_cases_ub / max(exp_cases_median),
-         infections_std  = infections/max(infections),
-         positive_std = positive/max(positive)) %>%
-  ungroup()
 
@@
 
-```{r, fig.height=10, fig.width=12}
-
-cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2)
+```{r, fig.height=10, fig.width=14}
+
+cowplot::plot_grid(plt_wastewater_pb, plt_covidestim_pb, ncol=2,  labels="auto", label_size=19)
+
+```
+
+```{r,eval=save_figures}
+
+ggsave(here("figures/correlations-wastewater-by-version.pdf"))
+
+```
+
+```{r}
 
@@
 covidestim_correlations <- w_data %>%
-  filter(version=="v1") %>%
+  filter(version=="v7") %>%
   # missing values are handled by casewise deletion
@@
 w_data %>%
-  filter(version=="v1") %>%
+  filter(version=="v7") %>%
   ggplot(aes(x=infections_std, y = mean_conc_std)) +
@@
 pb_correlations <- w_data %>%
-  filter(version=="v1") %>%
+  filter(version=="v7") %>%
   # missing values are handled by casewise deletion
@@
 names(pal) <- c(#"$observed$", 
-                '$P(S_1|untested)$ Centered at Emp. Value',
-                '$P(S_1|untested)$ and $\\beta$ Centered at Emp. Values',
+                '$P(S_1|untested)$ Centered at Survey Value',
+                '$P(S_1|untested)$ and $\\beta$ Centered at Survey Values',
                "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$",
-               "$\\beta$ Centered at Emp. Value"
+               "$\\beta$ Centered at Survey Value"
                )
@@
   left_join(dates,relationship = "many-to-many" ) %>%
-  # filter(version %in% c("v1","v3")) %>%
-  filter(version=="v1") %>%
+  # filter(version %in% c("v7","v3")) %>%
+  filter(version=="v7") %>%
   filter(biweek>=6) %>%
@@
 cowplot::plot_grid(legend, grid, nrow=2, rel_heights=c(.05,.95))
+```
+
+```{r, eval=save_figures}
+
+ggsave(here('figures/simulation-intervals-wastewater.pdf'))
 
@@
   left_join(dates,relationship = "many-to-many" ) %>%
-  filter(version %in% c("v1","v3")) %>%
+  filter(version %in% c("v7","v3")) %>%
   filter(biweek>=6) %>%
@@
 
-```{r}
+```{r, eval=include_other_norm}
 
@@
 
-```{r}
+```{r,eval=include_other_norm}
 
@@
 
-```{r, fig.height=10, fig.width=12}
+```{r, fig.height=10, fig.width=12,eval=include_other_norm}
 
@@
 
-```{r}
+```{r,eval=include_other_norm}
 
 covidestim_correlations <- w_data %>%
-  filter(version=="v1") %>%
+  filter(version=="v7") %>%
   # missing values are handled by casewise deletion
@@
 
-```{r,fig.width=12, fig.height=15}
+```{r,fig.width=12, fig.height=15,eval=include_other_norm}
 
@@
 
-```{r, fig.height=10, fig.width=10}
+```{r, fig.height=10, fig.width=10,eval=include_other_norm}
 
@@
 
-```{r, fig.width=11, fig.height=8}
+```{r, fig.width=13, fig.height=8}
 
@@
 county_dat %>%
-  filter(version=="v1")  %>%
+  filter(version=="v7")  %>%
    ggplot() +
@@
   theme_c(
+    axis.text.x=element_text(size=7),
           legend.position = "top",
@@
 
-```{r}
+```{r, eval=save_figures}
+
+ggsave(here('figures/simulation-intervals-county.pdf'))
+
+```
+
+```{r,fig.width=14, fig.height=8}
 
@@
 county_dat %>%
-  filter(version=="v1")  %>%
+  filter(version=="v7")  %>%
   #filter(grepl("Nan",name)) %>%
@@
 ############################################
-pltlist <- county_dat %>%filter(version=="v1") %>%
+pltlist <- county_dat %>%filter(version=="v7") %>%
     mutate(testpos = positive/total) %>%
@@
 #####################################
-pltlist <- county_dat %>%filter(version=="v1") %>%
+pltlist <- county_dat %>%filter(version=="v7") %>%
     mutate(testpos = positive/total) %>%
@@
 county_dat %>%
-  filter(version=="v1") %>%
+  filter(version=="v7") %>%
   filter(!grepl("Nantucket",name)) %>%
@@
 ma_concordance_by_county %>%
-  filter(vlabel=="$P(S_1|untested)$ Centered at Emp. Value") %>%
+  filter(vlabel=="$P(S_1|untested)$ Centered at Survey Value") %>%
   group_by(name) %>%
@@
   labs(y ="",
-       x="Proportion Containing the Covidestim Median",
+       x="Proportion Containing\nthe Covidestim Median",
        fill = "Covidestim Median:",
@@
                      limits=c(-.05,1))
+
+ggsave(here('figures/concordance-covidestim-county.pdf'))
 

### 2023-06-30 reports/priors.Rmd
figures

@@
                              option="beta",
-                       quiet=TRUE)
+                       quiet=FALSE)
 
@@
                              option="s_untested",
-                       quiet=TRUE)
+                       quiet=FALSE)
 
@@
 ```
+
+```{r,eval=FALSE}
+
+ctis_smoothed <- tar_read(ctis_smoothed,
+                          store=here("_targets"))
+
+ctis_smoothed %>% left_join(dates) %>%
+  group_by(biweek,state) %>%
+  summarize(beta_est=mean(beta_est,na.rm=TRUE)) %>%
+  ungroup() %>%
+  summarize(q1 = quantile(beta_est,.025, na.rm=TRUE),
+                median = median(beta_est, na.rm=TRUE),
+                q2 = quantile(beta_est,.975, na.rm=TRUE),
+                sd=sd(beta_est,na.rm=TRUE))
+
+ctis_smoothed %>%
+      summarize(q1 = quantile(
+        beta_estimate_spline_smoothed,.025, na.rm=TRUE),
+        median = median(beta_estimate_spline_smoothed, na.rm=TRUE),
+        q2 = quantile(beta_estimate_spline_smoothed,.975, na.rm=TRUE),
+        sd=sd(beta_estimate_spline_smoothed,na.rm=TRUE))
+
+```

### 2023-06-30 reports/state_figures.Rmd
figures

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE,cache=FALSE)
 
@@
 library(cowplot)
+
+filter_versions <- FALSE
+save_figures <- FALSE
 
@@
 # read state-level results
-targets_dir <- here("_targets_6_26")
+targets_dir <- here("_targets")
 
@@
 
+if(filter_versions){
   state_res <- state_res %>%
-  filter(version %in% c("v2", "v5", "v6", "v7"))
+  filter(version %in% c("v3", "v5", "v6", "v7"))
+
+}
 
@@
 by_version <- state_res %>%
+ # filter(biweek >=6) %>% 
   select(-date) %>%
@@
        fill = "Covidestim Median:",
-       title = "Proportion of Simulation Intervals Where Covidestim Median\nWas Within, Above, or Below the Median",
-       subtitle = "Including All States") +
+       title = "Proportion of Simulation Intervals Where Covidestim Median\nWas Within, Above, or Below the Median") +
   theme_c() +
@@
   scale_x_discrete(labels = (TeX(labels)))
+
+```
+
+```{r,eval=save_figures, fig.width=14}
+
+ggsave(here('figures/concordance-covidestim-state.pdf'))
+
+```
+
+```{r, fig.width=17}
+
+by_version %>%
+  group_by(vlabel) %>%
+  filter(in_interval == "In Interval") %>%
+  mutate(n_interval = n) %>%
+  ggplot(aes(y=fct_reorder(vlabel,prop), x=prop)) +
+  geom_text(aes(label=round(prop,3), x= prop+.03), 
+             angle=0) +
+  geom_bar(stat="identity", fill="#596281") +
+  scale_y_discrete(breaks= unique(by_version$vlabel),
+                   labels=function(x)TeX(x)) +
+  scale_x_continuous(expand=c(0,0), 
+                     limits=c(0,.9), n.breaks=7)  +
+  theme_c(axis.text.y = element_text(size = 13, hjust=1),
+          axis.title.x = element_text(size=14),
+          axis.text.x=element_text(size=10)) +
+  labs(y="",
+       x="Proportion Containing the Covidestim Median",
+       title="Proportion Containing the Covidestim Median",
+       subtitle = "By Implementation")
 
@@
 
-```{r}
+### All Versions
+
+```{r, fig.width=18,fig.height=8}
 
@@
 
-names(pal) <- c(unique(subset$vlabel), "Covidestim")
+names(pal) <- c(as.character(unique(subset$vlabel)), "Covidestim")
 
@@
 
+### Original Priors versus CTIS Priors (that do not vary)
+
+```{r}
+
+state_res %>%
+  filter(version %in% c("v1","v7") & state %in% sample(unique(.$state),15)) %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub,
+             fill=vlabel),
+             alpha = .6,
+             show.legend=FALSE,
+             ) +
+  facet_wrap(~name, 
+             labeller=labeller(vlabel =as_labeller(
+               TeX, default=label_parsed),
+               nrow=3),
+             scales="free_y") +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(
+          legend.position = "top",
+          ) +
+  viridis::scale_fill_viridis(option="magma", begin=.1,end=.9, discrete=TRUE)  +
+  labs(x="",
+       y = "Estimated Infections",
+       title = "Simulation Intervals Over Time",
+       subtitle = "By State")
+
+```
+
 ## Compare States
 
-```{r, fig.height=11, fig.height=9}
+```{r, fig.height=14, fig.width=11}
 
@@
               alpha = .6) +
-  facet_wrap(~state, ncol=5, scales = "free_y",
+  facet_wrap(~name, ncol=4, scales = "free",
               labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
   scale_y_continuous(labels = comma) +
-  scale_x_date(date_breaks = "3 months", 
+  scale_x_date(date_breaks = "3.5 months", 
                date_labels = "%b %Y") +
-  theme_c(axis.text.x=element_text(size=6),
-          axis.text.y = element_text(size=8),
-          legend.position = "top") +
+  theme_c(axis.text.x=element_text(size=5),
+          axis.text.y = element_text(size=4.5),
+          legend.position = "top",
+          legend.text=element_text(size=12),
+          strip.text.x=element_text(size=8,
+                                    face='plain',
+                                    color='white')) +
   scale_fill_manual(values =pal,
                     labels = TeX(names(pal)),
-                    name='',
-                    breaks="Covidestim")  +
+                    name=''
+                    #,
+                  # , breaks="Covidestim"
+                    )  +
   labs(y = "Estimated Infections",
@@
                               byrow=TRUE))
+
+```
+
+```{r,eval=save_figures, fig.height=14, fig.width=9}
 
@@
 
+```
+
+```{r,eval=save_figures, fig.height=10,fig.width=12}
+
 ggsave(here('figures/ratio-estimated-to-observed-all-states.pdf'))
@@
        title ="During Two-week Period\nin the Alpha Wave",
-       subtitle = "March 26, 2021 through April 8, 2021")
+       subtitle = "March 26, 2021 through April 8, 2021") +
+  xlim(0,65)
 
@@
        title ="During Two-week Period\nin the Delta Wave",
-       subtitle = "June 18, 2021 through July 1, 2021")
+       subtitle = "June 18, 2021 through July 1, 2021") +
+  xlim(0,65)
 
@@
   geom_errorbar() +
-  # geom_point(aes(y=state_name,x=exp_cases_median/positive),
-  #            color = "darkred", size=.8, alpha=.5) +
   theme_c() + 
@@
        title ="During Two-week Period\nin the Omicron Wave",
-       subtitle = "December 31, 2021 through January 14, 2022")
+       subtitle = "December 31, 2021 through January 14, 2022") +
+  xlim(0,65)
 
@@
 
-plt <- plot_grid(plt1,plt2, plt3, nrow=1, align="none")
+plt <- plot_grid(plt1,plt2, plt3, nrow=1, align="none", labels="auto", label_size=19)
 
@@
 
-plot_grid(title, plt, nrow=2, rel_heights= c(.05,.95))
+plot_grid(title, plt, nrow=2, 
+          rel_heights= c(.05,.95))
 
 ```
+
+```{r,eval=save_figures, fig.height=10, fig.width=17}
+
+ggsave(here('figures/ratio-estimated-to-observed-multiple-waves.pdf'))
+
+```
+
+```{r}
+
+subset %>% 
+  # left_join(codes) %>%
+  filter(biweek %in% c(7, 13, 27))  %>%
+  select(-date) %>%
+  distinct() %>%
+  group_by(state) %>%
+  mutate(ord = empirical_testpos[which(biweek==7)]) %>%
+  mutate(biweek=factor(biweek)) %>%
+  ggplot(aes(y=fct_reorder(state,ord),
+             x=empirical_testpos,fill=biweek)) +
+  geom_bar(stat="identity",position="dodge")+
+  theme_c() +
+  theme_c() +
+  labs(title = "Testing Positivity by State",
+       y="",
+       x="Biweekly Testing Positivity",
+       fill="") +
+  scale_x_continuous(n.breaks=10)
+
+subset %>% 
+  # left_join(codes) %>%
+  filter(biweek %in% c(7, 13, 27))  %>%
+  group_by(biweek) %>%
+  mutate(daterange= paste(format(min(date), "%B %d %Y"),
+                          "to", 
+                          format(max(date), "%B %d %Y")),
+         daterange=factor(
+           daterange,
+           levels=c(
+             "March 26 2021 to April 08 2021",
+             "June 18 2021 to July 01 2021",
+             "December 31 2021 to January 14 2022"))) %>%
+  select(-date) %>%
+  distinct() %>%
+  group_by(state) %>%
+  mutate(testrate=total/population,
+         ord = testrate[which(biweek==7)]) %>%
+  mutate(biweek=factor(biweek)) %>%
+  ggplot(aes(y=fct_reorder(name,ord,.desc=TRUE),
+             x=testrate,fill=daterange)) +
+  geom_bar(stat="identity",position="dodge") +
+  theme_c() +
+  labs(title = "Testing Rate by State",
+       y="",
+       x="Biweekly Testing Rate",
+       fill="") +
+  scale_x_continuous(n.breaks=10)
+
+compare_testrates <- subset %>%
+  mutate(testrate=total/population) %>%
+  select(-date) %>%
+  distinct() %>%
+  filter(biweek %in% c(7, 13, 27))  %>%
+  select(biweek,state,testrate) %>%
+  pivot_wider(names_from=biweek, values_from=testrate, names_prefix="biweek_") %>%
+  mutate(omicron_over_alpha = biweek_27/biweek_7,
+         omicron_over_delta = biweek_27/biweek_13) %>%
+  select(state,contains("omicron")) %>%
+  pivot_longer(c(omicron_over_alpha,omicron_over_delta)) %>%
+  group_by(name) %>%
+  summarize(mean_val = mean(value))
+
+cat("The testing rate during this two-week interval during the omicron wave was", compare_testrates %>% filter(name=="omicron_over_alpha") %>% pull(mean_val), "times the testing rate during this time interval in the alpha wave.")
+
+cat("The testing rate during this two-week interval during the omicron wave was", compare_testrates %>% filter(name=="omicron_over_delta") %>% pull(mean_val), "times the testing rate during this time interval in the delta wave.")
+
+```
+
+# Rank Ratio of Estimated to Observed Over Time
+
+```{r,fig.height=8}
+
+ranks <- subset %>%
+  mutate(ratio = exp_cases_median/positive,
+         tested = total/population) %>%
+  # one observation per biweek per state
+  select(biweek, date, ratio, name,tested) %>%
+  group_by(biweek,ratio,name,tested) %>%
+  summarize(date=min(date)) %>%
+  # rank for each time interval
+  group_by(biweek) %>%
+  mutate(rank = rank(ratio),
+         rank_tested=rank(-tested))
+
+r <- ranks %>%
+  mutate(rank_group = case_when(
+    rank <= 10 ~  "Top 10",
+    rank > 41 ~ "Bottom 10",
+    TRUE ~ "Middle" )) %>%
+  group_by(name, rank_group) %>%
+  mutate(n=n()) %>%
+  group_by(name) %>%
+  mutate(total =n(),
+         max_group = rank_group[which.max(n)]) %>%
+  filter( max(n) >24) %>%
+  filter(max_group != "Middle")  %>%
+  ungroup()
+
+top_10 <- r %>% filter(rank_group=="Top 10") %>%
+  arrange(desc(n)) %>% pull(name) %>%
+  unique() %>%paste(collapse=", ")
+
+cat("States that consistently have the lowest ratios:", top_10)
+
+bottom_10 <- r %>% filter(rank_group=="Bottom 10") %>%
+  arrange(desc(n)) %>% pull(name) %>%
+  unique() %>%paste(collapse=", ")
+
+cat("States that consistently have the highest ratios:", bottom_10)
+
+end_labs <- r %>%
+  ungroup() %>%
+  filter(date==max(date)) %>%
+  mutate(date = date + days(10)) %>%
+  select(name, rank, date) %>%
+  ungroup()
+
+```
+
+```{r}
+
+##############################
+# RANK PLOT OVER TIME
+##############################
+# 
+# set.seed(999)
+# 
+# pal1 <-viridis_pal(option="viridis", end = .9, begin=.3)(3)
+# pal2 <- viridis_pal(option="rocket", end = .9, begin=.3)(2)
+# pal3 <- viridis_pal(option="magma", end = .9, begin=.3)(2)
+# pal4 <- viridis_pal(option="mako", end = .8, begin=.3)(2)
+# pal5 <- viridis_pal(option="cividis", end = .9, begin=.1)(3)
+# 
+# 
+# rankpal <- c(pal1, pal2, pal3,pal4,pal5)
+# indexes <- sample.int(length(rankpal), replace=FALSE)
+# rankpal <- rankpal[indexes]
+
+set.seed(123)
+
+rankpal <- c("#b4ddd4", "#7b2c31", "#65d04b",
+             "#da239b", "#b7d165", "#453fbc", 
+             "#658114", "#af3fe8", "#ffb947",
+             "#0a60a8", "#f87945", "#16d0ae")
+
+rankpal <- c("#851657", "#be3acd", "#19a71f",
+             "#824598", "#ed820a", "#BB8E37",
+             "#974810", "#1f84ec", "#d02023",
+             "#059dc5", "#f23387", "#16d0ae")
+
+indexes <- sample.int(length(rankpal), replace=FALSE)
+rankpal <- rankpal[indexes]
+
+r %>%
+ # filter(rank_group!="Middle") %>%
+  ggplot() +
+  geom_point(aes(x=date,y=rank, 
+             group=name,
+             color= name),
+             size=.5) +
+  geom_line(aes(x=date,y=rank, 
+             group=name,
+             color= name)) +
+ # facet_wrap(~max_group) +
+  ggrepel::geom_text_repel( aes(x= date-days(4),
+                y = rank,
+                color=name,
+                label = name),
+           end_labs,
+           min.segment.length=2,
+           nudge_y=0,
+           hjust=0,
+           size=3,
+           direction="y",
+           force_pull=9,
+           box.padding=.15) +
+  theme_c(legend.position="none",
+          axis.text.x = element_text(angle =0,size=14)) +
+  # theme(plot.subtitle =element_text(size=18),
+  #       axis.title.x=element_text(size=16),
+  #       axis.title.y=element_text(size=16)) +
+ # scale_color_brewer(palette="Dark2")  +
+  scale_color_manual(values=rankpal) +
+  scale_x_date(date_breaks="2 months", date_labels = "%b %Y",
+               limits = c(ymd("2021-01-01"),
+                          ymd("2022-04-15"))) +
+  labs(y = "Rank of Ratio",
+       title = "Rank of Estimated Infections to Observed Infections Over Time",
+       subtitle = "For States Ranked in the Top or Bottom 10\nfor More Than 80% of Time Intervals",
+       x="Date") +
+  scale_y_reverse(breaks=seq(1,51, by=9))
+
+```

### 2023-07-03 reports/county_figures.Rmd
update figures

@@
 
-filter_versions <- FALSE 
-save_figures <- FALSE
+filter_versions <- TRUE 
+save_figures <- TRUE
 
@@
 
+if(filter_versions) {
+
+  county <- county %>%
+    left_join(versions) %>%
+    mutate(vlabel=case_when(
+      version == "v7" ~  "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
+      version == "v5" ~ "$\\beta$ Centered at Survey Value",
+      version =="v6" ~  "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+      TRUE ~ vlabel
+    )) %>%
+    mutate(vlabel =factor(vlabel,levels= c(
+        "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
+         "$P(S_1|untested)$ Centered at Survey Value",
+         "$\\beta$ Centered at Survey Value",
+        "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values"))) %>%
+    left_join(county_names) %>%
+    mutate(name = ifelse(fips =="25007,25019", "Nantucket and Dukes", name)) %>%
+    left_join(dates,  relationship = "many-to-many") %>%
+    group_by(biweek) %>%
+    mutate(mindate = min(date),
+           maxdate = max(date)) %>%
+    ungroup()
+
+}
+
 ```
@@
             vlabel =unique(vlabel)) %>%
-  mutate(x=1.8,
+  mutate(x=2.8,
          y=.1)
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
+  labs(x = "Probabilistic Bias Infections\n(Standardized by County)",
        y = "Mean Wastewater Concentration\n(Standardized by County)",
-       title = "Correlations Between Probabilistic Bias Counts\nand Mean Wastewater Concentration")
+       title = "Correlations Between Probabilistic Bias Infections\nand Mean Wastewater Concentration")
 
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
+  labs(x = "Probabilistic Bias Infections\n(Standardized by County)",
        y = "Mean Wastewater Concentration\n(Standardized by County)",
-       title = "")
+       title = "Correlations with Wastewater Concentrations")
 
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
+  labs(x = "Probabilistic Bias Infections\n(Standardized by County)",
        y = "Covidestim Estimates\n(Standardized by County)",
-       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+       title = "Correlations Between Probabilistic Bias Infections\nand Covidestim Estimates")
 
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
+  labs(x = "Probabilistic Bias Infections\n(Standardized by County)",
        y = "Covidestim Estimates\n(Standardized by County)",
-       title = "")
+       title = "Correlations with Covidestim Infections")
 
@@
                 '$P(S_1|untested)$ and $\\beta$ Centered at Survey Values',
-               "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$",
+               "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
                "$\\beta$ Centered at Survey Value"
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
+  labs(x = "Probabilistic Bias Infections\n(Standardized by County by County)",
        y = "Mean Wastewater Concentration\n(Standardized by County by County)",
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
+  labs(x = "Probabilistic Bias Infections\n(Standardized by County by County)",
        y = "Covidestim Estimates\n(Standardized by County by County)",
-       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+       title = "Correlations Between Probabilistic Bias Infections\nand Covidestim Estimates")
 
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County by County)",
+  labs(x = "Probabilistic Bias Infections\n(Standardized by County by County)",
        y = "Covidestim Estimates\n(Standardized by County by County)",
@@
   mutate(vlabel = gsub(
-    "$Priors\\,Do\\,Not\\,Vary\\,by\\,County\\,and\\,Date$", 
+    "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
     "Priors Do Not Vary by County and Date",
@@
        y = "Wastewater Concentrations\n(Standardized by County by County)",
-       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+       title = "Correlations Between Probabilistic Bias Infections\nand Covidestim Estimates")
 
@@
 
-For each county, normalize the wastewater concentration by the maximum concentration for that county and the probabilistic bias counts by the maximum *median* estimated counts for that county (2.5th percentile and 97.5th percentile also standardized by the maximum *median* estimated counts for that county).
+For each county, normalize the wastewater concentration by the maximum concentration for that county and the Probabilistic Bias Infections by the maximum *median* estimated counts for that county (2.5th percentile and 97.5th percentile also standardized by the maximum *median* estimated counts for that county).
 
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
+  labs(x = "Probabilistic Bias Infections\n(Standardized by County)",
        y = "Mean Wastewater Concentration\n(Standardized by County)",
@@
   theme_c() +
-  labs(x = "Probabilistic Bias Counts\n(Standardized by County)",
+  labs(x = "Probabilistic Bias Infections\n(Standardized by County)",
        y = "Covidestim Estimates\n(Standardized by County)",
@@
        y = "Wastewater Concentrations\n(Standardized by County)",
-       title = "Correlations Between Probabilistic Bias Counts\nand Covidestim Estimates")
+       title = "Correlations Between Probabilistic Bias Infections\nand Covidestim Estimates")
 
@@
 
-county_dat <- county %>%
-  # set Nantucket, Duke fips to Nantucket since we have wastewater data
-  # for Nantucket
-  left_join(dates,  relationship = "many-to-many") %>%
-  group_by(biweek) %>%
-  mutate(mindate = min(date),
-         maxdate = max(date)) %>%
-  ungroup() %>%
-  left_join(county_names, by=c('fips'='fips')) %>%
-  left_join(versions) %>%
-  mutate(name = ifelse(fips =="25007,25019", "Nantucket and Dukes", name))
-
-county_dat %>%
+county %>%
   filter(version=="v7")  %>%
@@
 
-county_dat %>%
+county %>%
   filter(fips %in% sample(unique(.data$fips),2))  %>% 
@@
 
-county_dat %>%
+county %>%
   filter(version=="v7")  %>%
@@
 ############################################
-pltlist <- county_dat %>%filter(version=="v7") %>%
+pltlist <- county %>%filter(version=="v7") %>%
     mutate(testpos = positive/total) %>%
@@
 #####################################
-pltlist <- county_dat %>%filter(version=="v7") %>%
+pltlist <- county %>%filter(version=="v7") %>%
     mutate(testpos = positive/total) %>%
@@
 
-county_dat %>%
+county %>%
   filter(version=="v7") %>%
@@
 
-# Summary of Comparison Between Implementations
-
-Although the spline-smoothing improved the correlations with wastewater concentrations and the correlations with Covidestim estimates, it did not improve the proportion containing the Covidestim median.
-
-The implementation centering $\Pr(S_1|\text{untested})$ at the survey value along with the implementation that does not vary the priors by state or date are the most correlated with Covidestim estimates, wastewater concentrations, and  have the highest proportion of county-biweeks containing the Covidestim median.
+# Faceted by Version
+
+```{r, fig.width=15, fig.height=16}
+
+county %>% 
+  mutate(vlabel=fct_reorder(vlabel,as.numeric(vlabel),.desc=TRUE)) %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub),
+             alpha = .8,
+             show.legend=FALSE,
+             fill="#596281") +
+  facet_grid(name~vlabel, 
+             labeller=labeller(vlabel =as_labeller(
+               TeX, default=label_parsed)),
+             scales="free") +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(
+          legend.position = "top",
+          ) +
+  scale_fill_manual(values = pal)  +
+  labs(x="",
+       y = "Estimated Infections",
+       title = "Simulation Intervals Over Time",
+       subtitle = "Counties in Massachusetts") +
+  guides(color = guide_legend(override.aes = list(linewidth =2)))
+
+```
+
+```{r,fig.width=15, fig.height=16, eval=save_figures}
+
+ggsave(here('figures/simulation-intervals-faceted-version-county.pdf'))
+
+```
+
+```{r, fig.height=8, fig.width=14}
+
+county$vlabel %>% unique()
+
+pal
+
+county %>% 
+    mutate(vlabel=fct_reorder(vlabel,as.numeric(vlabel),.desc=TRUE)) %>%
+  filter(biweek>=6) %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub,
+             fill=vlabel),
+             alpha = .5) +
+  facet_wrap(~name,
+             scales="free_y") +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(
+          legend.position = "top",
+          legend.text=element_text(hjust=0)
+          ) +
+  scale_fill_manual(values = pal,
+                    labels=TeX(unique(as.character(county$vlabel))))  +
+  labs(x="",
+       fill="",
+       y = "Estimated Infections",
+       title = "Simulation Intervals Over Time",
+       subtitle = "Counties in Massachusetts") +
+  guides(fill=guide_legend(nrow=2))
+
+```
+
+```{r,fig.height=8, fig.width=14, eval=save_figures}
+
+ggsave(here('figures/simulation-intervals-color-version-county.pdf'))
+
+```

### 2023-07-03 reports/priors.Rmd
update figures

@@
   viridis::scale_fill_viridis(discrete=TRUE,
-                             option = "rocket", begin=.3,end=.8) +
+                             option = "rocket", 
+                             begin=.3,
+                             end=.8) +
   labs(x = "Value",

### 2023-07-08 reports/county_figures.Rmd
add heatmap

@@
       legend.position = "right",
-      strip.text.x = element_text(size = 11, face = "bold", color="white"),
-      strip.text.y = element_text(size = 11, face = "bold", color="white",angle=270),
+      strip.text.x = element_text(size = 11, color="white"),
+      strip.text.y = element_text(size = 11,  color="white",angle=270),
       strip.background = element_rect(fill = "#3E3D3D")
@@
     mutate(vlabel=case_when(
-      version == "v7" ~  "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
+      version == "v7" ~  "$\\overset{Priors\\,Do\\,Not\\,Vary}{\\,by\\,State\\,and\\,Date}$",
       version == "v5" ~ "$\\beta$ Centered at Survey Value",
-      version =="v6" ~  "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
-      TRUE ~ vlabel
+      version =="v6" ~  "$\\overset{P(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}",
+      version == "v3" ~ "$P(S_1|untested)$ Centered at Survey Value",
     )) %>%
     mutate(vlabel =factor(vlabel,levels= c(
-        "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
+        "$\\overset{Priors\\,Do\\,Not\\,Vary}{\\,by\\,State\\,and\\,Date}$", 
          "$P(S_1|untested)$ Centered at Survey Value",
          "$\\beta$ Centered at Survey Value",
-        "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values"))) %>%
+        "$\\overset{P(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}"))) %>%
     left_join(county_names) %>%
@@
     ungroup()
+
+  county$vlabel %>% unique()
+
+  wjoined <- wjoined %>%
+    mutate(vlabel=case_when(
+      version == "v7" ~  "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
+      version == "v5" ~ "$\\beta$ Centered at Survey Value",
+      version =="v6" ~  "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+      version == "v3" ~ "$P(S_1|untested)$ Centered at Survey Value"
+    ))
 
@@
             vlabel =unique(vlabel)) %>%
-  mutate(x=2.8,
+  mutate(x=3.5,
          y=.1)
@@
   theme_c() +
+  theme(strip.text.x = element_text(size=12)) +
   labs(x = "Probabilistic Bias Infections\n(Standardized by County)",
@@
                               use = "complete.obs")) %>%
-  mutate(x=2.8,
+  mutate(x=3.5,
          y=.1)
-
-w_data %>%
-  ggplot(aes(x=exp_cases_median_std, y = infections_std)) +
-  geom_point(alpha=.5) +
-  geom_linerange(aes(xmin = exp_cases_lb_std,
-                     xmax=exp_cases_ub_std, 
-                     y = infections_std),
-                 alpha = .3) +
-  geom_text(
-           aes(x=x,y=y,  label=paste0("Correlation: ", 
-                                      round(Correlation,3))), 
-                data= covidestim_correlations ) +
-  facet_wrap(~vlabel,
-              ncol=2,
-             labeller= as_labeller(TeX, 
-                                   default = label_parsed)) +
-  theme_c() +
-  labs(x = "Probabilistic Bias Infections\n(Standardized by County)",
-       y = "Covidestim Estimates\n(Standardized by County)",
-       title = "Correlations Between Probabilistic Bias Infections\nand Covidestim Estimates")
 
@@
 
+plt_covidestim_pb
+
 ```
@@
 
-```{r,eval=save_figures}
+```{r, fig.height=10, fig.width=14, eval=save_figures}
 
@@
   mutate(vlabel = gsub(
-    "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
+     "$\\overset{Priors\\,Do\\,Not\\,Vary}{\\,by\\,State\\,and\\,Date}$", 
     "Priors Do Not Vary by State and Date",
@@
                 '$P(S_1|untested)$ Centered at Survey Value',
-                '$P(S_1|untested)$ and $\\beta$ Centered at Survey Values',
+                "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
                "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
@@
 county %>% 
+  mutate(name=gsub("Nantucket and Dukes", "Nantucket\nand Dukes", name)) %>%
   mutate(vlabel=fct_reorder(vlabel,as.numeric(vlabel),.desc=TRUE)) %>%
@@
           ) +
+  theme_c(strip.text.x=element_text(size=13,color='white')) +
   scale_fill_manual(values = pal)  +
@@
 
+## Fill by Version
+
 ```{r, fig.height=8, fig.width=14}
 
-county$vlabel %>% unique()
-
 pal
 
-county %>% 
+county_relab <- county %>%
+  mutate(vlabel=case_when(
+      version == "v7" ~  "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
+      version == "v5" ~ "$\\beta$ Centered at Survey Value",
+      version =="v6" ~  "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+      version == "v3" ~ "$P(S_1|untested)$ Centered at Survey Value",
+    )) %>%
+    mutate(vlabel =factor(vlabel,levels= c(
+        "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
+         "$P(S_1|untested)$ Centered at Survey Value",
+         "$\\beta$ Centered at Survey Value",
+       "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values"))) %>%
     mutate(vlabel=fct_reorder(vlabel,as.numeric(vlabel),.desc=TRUE)) %>%
-  filter(biweek>=6) %>%
+  filter(biweek>=6)
+
+county_relab$vlabel %>% unique()
+
+county_relab %>%
    ggplot() +
@@
   scale_fill_manual(values = pal,
-                    labels=TeX(unique(as.character(county$vlabel))))  +
+                    labels=TeX(unique(as.character(county_relab$vlabel))))  +
   labs(x="",
@@
 ```
+
+# Heatmap
+
+```{r}
+
+######################################################
+# HEATMAP OF RATIO OF ESTIMATED TO OBSERVED
+######################################################
+
+heatmap_dat <- county %>%
+  filter(version == "v7") %>%
+  group_by(biweek, name, exp_cases_median, positive) %>%
+  summarize(date = min(date)) %>%
+  ungroup()
+
+wide <- heatmap_dat %>%
+    mutate(ratio= exp_cases_median/positive) %>%
+  select(-c(exp_cases_median, positive)) %>%
+  pivot_wider(names_from=name,values_from =ratio)
+
+heatmap_dat %>%
+  # mutate(county_name = gsub("Nantucket", "Nantucket and", name)) %>%
+  mutate(ratio=exp_cases_median/positive) %>%
+  mutate(date_lab = format(as.POSIXct(date),format="%b %d %Y")) %>%
+  group_by(name) %>%
+  mutate(m = median(ratio)) %>%
+  ungroup() %>%
+  ggplot(aes(x=fct_reorder(date_lab, date),
+             y = fct_reorder(name,m),
+             fill =exp_cases_median/positive)) +
+  geom_tile() +
+ # scale_x_discrete(breaks=breaks_date, labels = breaks_date)  +
+  viridis::scale_fill_viridis(option="rocket", direction=-1) +
+  labs(x="",
+       y= "",
+       fill = "Ratio of Estimated Infections\nto Observed Infections",
+       title = "Ratio of Median Estimated Infections to Observed Infections\nfor Counties in Massachusetts") +
+  theme_c() +
+  theme(axis.text.x = element_text(size=11, angle=90),
+        axis.text.y = element_text(size = 17),
+        axis.title = element_text(size=16),
+        legend.title = element_text(face="bold", size=18),
+        plot.title =element_text(size=22),
+        plot.subtitle = element_text(size=20))
+
+# which biweeks are highest?
+heatmap_dat %>%
+  mutate(ratio=exp_cases_median/positive) %>%
+  group_by(biweek) %>%
+  summarize(m=median(ratio),
+            date=min(date)) %>%
+  arrange(desc(m))
+
+ggsave(here('figures/ma_county_heatmap.pdf'),
+       height=10,
+       width =13,
+       dpi=400)
+
+```

### 2023-07-08 reports/county_figures.Rmd
update county figures

@@
   summarize(date = min(date)) %>%
-  ungroup()
-
-wide <- heatmap_dat %>%
-    mutate(ratio= exp_cases_median/positive) %>%
-  select(-c(exp_cases_median, positive)) %>%
-  pivot_wider(names_from=name,values_from =ratio)
-
-heatmap_dat %>%
-  # mutate(county_name = gsub("Nantucket", "Nantucket and", name)) %>%
+  ungroup() %>%
   mutate(ratio=exp_cases_median/positive) %>%
@@
   mutate(m = median(ratio)) %>%
-  ungroup() %>%
+  ungroup()
+
+date_labels_all <-  unique(heatmap_dat$date_lab)
+date_labels <- date_labels_all[seq(2, length(date_labels),by=2)]
+
+heatmap_dat %>%
   ggplot(aes(x=fct_reorder(date_lab, date),
@@
   theme_c() +
-  theme(axis.text.x = element_text(size=11, angle=90),
+  theme(axis.text.x = element_text(size=14, angle=90),
         axis.text.y = element_text(size = 17),
@@
         plot.title =element_text(size=22),
-        plot.subtitle = element_text(size=20))
+        plot.subtitle = element_text(size=20)) +
+  scale_x_discrete(breaks = date_labels_all, labels=date_labels)
 
@@
 ```
+
+# Testing Rate versus Ratio of Estimated to Observed
+
+```{r testrate-v-ratio, fig.cap  = 'The ratio of the median estimated infections to observed infections plotted against the testing rate, where the testing rate is calculated as the total number tested in a two-week interval over the population size.
+When the priors are the same for all time intervals and states, there is minimal variability relationship between the testing rate and the ratio of estimated to observed infections, since the correction for incomplete testing and diagnostic test inaccuracy is identical for each time-interval and location.
+However, when we allow $\\beta$ or $\\Pr(S_1|\\text{untested})$ to vary by state and time interval, there is much variability in the relationship.
+A horizontal line in red at 1 is included to reference; a ratio of exactly one would indicate no infections went unobserved.'}
+
+lb <- county %>% mutate(testrate = total/population ) %>% pull(testrate) %>% min()
+
+###########################################################################
+# plot relationship between ratio of estimated cases to observed 
+# against testing rate 
+###########################################################################
+county %>%
+  ggplot(aes(x=total/population, y = exp_cases_median/positive)) +
+  geom_point(alpha=.03, size=.9) +
+  facet_wrap(~vlabel,
+               labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
+  theme_c(axis.text.x = element_text(size =9),
+          axis.title = element_text(size=16) ) +
+  theme(strip.text.x = element_text(size=11, color="white")) +
+  scale_y_continuous(n.breaks=10) +
+  geom_hline(yintercept=1, linetype=2,color="darkred", alpha=.8, linewidth=.5) +
+  labs(y = "Ratio of Estimated Infections to Observed Infections",
+       x = "Testing Rate",
+       subtitle = "All Time Intervals for Counties in Massachusetts",
+       title = "Relationship Between Testing Rate and\nRatio of Estimated Infections to Observed Infections")+
+  scale_x_continuous(limits=c(lb,.3), n.breaks=4)
+
+ggsave(here('figures/testing_rate_ratio-county.pdf'), width=11,height=7, dpi=400)
+
+```
+
+```{r}
+
+###########################################################################
+# plot relationship between ratio of estimated cases to observed 
+# against testing rate 
+###########################################################################
+county %>%
+  ggplot(aes(x=total/population, y = exp_cases_median/positive, color=population)) +
+  geom_point(alpha=.03, size=.9) +
+  geom_linerange(aes(ymin = exp_cases_lb/positive,
+                     ymax=exp_cases_ub/positive ),
+                 alpha=.03) +
+  facet_wrap(~vlabel,
+               labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
+  theme_c(axis.text.x = element_text(size =9),
+          axis.title = element_text(size=16) ) +
+  theme(strip.text.x = element_text(size=11, color="white")) +
+  scale_y_continuous(n.breaks=10) +
+  geom_hline(yintercept=1, linetype=2,color="darkred", alpha=.8, linewidth=.5) +
+  labs(y = "Ratio of Estimated Infections to Observed Infections",
+       x = "Testing Rate",
+       subtitle = "All Time Intervals and All States",
+       title = "Relationship Between Testing Rate and\nRatio of Estimated Infections to Observed Infections")+
+  scale_x_continuous( n.breaks=4, limits=c(lb,.3))
+
+```

### 2023-07-09 reports/county_figures.Rmd
update figures

@@
 
+FILT_VERSIONS <- c("v3", "v5", "v6", "v7")
+
 if(filter_versions) {
   county <- county %>%
-    filter(version %in% c("v3", "v5", "v6", "v7")) 
+    filter(version %in% FILT_VERSIONS) 
 }
@@
 date_labels_all <-  unique(heatmap_dat$date_lab)
-date_labels <- date_labels_all[seq(2, length(date_labels),by=2)]
+date_labels <- date_labels_all
+date_labels[seq(1, length(date_labels_all),by=2)] <- ""
+
+#date_labels <- date_labels_all[seq(1, length(date_labels_all),by=2)]
 
@@
   theme_c() +
-  theme(axis.text.x = element_text(size=14, angle=90),
+  theme(axis.text.x = element_text(size=13, angle=90),
         axis.text.y = element_text(size = 17),
@@
        title = "Relationship Between Testing Rate and\nRatio of Estimated Infections to Observed Infections")+
-  scale_x_continuous(limits=c(lb,.3), n.breaks=4)
+  scale_x_continuous(limits=c(lb,.3), n.breaks=6)
 
@@
 ```
+
+# Table Showing Covidestim Coverage
+
+```{r}
+
+versions <- tibble(
+  version = c( "v3", "v5", "v6", "v7"),
+  vlabel = factor(
+    c(  "$P(S_1|$untested) Centered at Survey Value"  ,
+    "$\\beta$ Centered at Survey Value",
+    "$P(S_1|$untested) and $\\beta$ Centered at Survey Values",
+     "Priors Do Not Vary by State or Date"
+    ),
+    levels =   c(  "$P(S_1|$untested) Centered at Survey Value"  ,
+    "$\\beta$ Centered at Survey Value",
+    "$P(S_1|$untested) and $\\beta$ Centered at Survey Values",
+     "Priors Do Not Vary by State or Date"
+    ))
+)
+```
+
+```{r load-state}
+
+state <- tar_read(state_v1, 
+                   store =targets_dir) %>% 
+  bind_rows(
+    tar_read(state_v2,
+             store =targets_dir)
+  ) %>%
+   bind_rows(
+    tar_read(state_v3, 
+             store =targets_dir)
+  ) %>%
+   bind_rows(
+    tar_read(state_v4, 
+             store =targets_dir)
+   )%>%
+   bind_rows(
+    tar_read(state_v5,
+             store =targets_dir)
+  )%>%
+   bind_rows(
+    tar_read(state_v6,
+             store =targets_dir)
+  ) %>%
+   bind_rows(
+    tar_read(state_v7,
+             store =targets_dir)
+  )
+
+covidestim_state <- tar_read(covidestim_biweekly_state,
+                        store =targets_dir)
+
+state_covidestim <- state %>%
+  inner_join(covidestim_state, relationship="many-to-many",
+             by = c('fips'='state', 'biweek'='biweek')) %>%
+  distinct() %>%
+  filter(version %in% FILT_VERSIONS)
+
+```
+
+```{r}
+
+county_covidestim <- county %>%
+  inner_join(covidestim, relationship = "many-to-many") %>%
+  select(-date) %>%
+  distinct()
+
+get_coverage <- function(df, geo_scale) {
+  df %>%
+    mutate(in_interval = case_when(
+    exp_cases_lb <= infections & 
+      infections <= exp_cases_ub ~ "Contained in Interval",
+    exp_cases_lb  > infections  ~ "Below Interval",
+    exp_cases_ub < infections ~ "Above Interval"
+  ) ) %>%
+  group_by(version, in_interval) %>%
+  summarize(n=n()) %>%
+  group_by(version) %>%
+  mutate(total = sum(n)) %>%
+  ungroup() %>%
+  mutate(prop = n/total) %>%
+  select(version, prop, in_interval) %>%
+  pivot_wider(names_from = in_interval, values_from=prop, names_prefix="Percent ") %>%
+  mutate(across(contains("Percent"), ~.x*100 )) %>%
+ # mutate(across(contains("Percent"), ~paste0(round(.x,3),"%")))
+  left_join(versions) %>%
+    mutate(`Geographic Scale` = geo_scale) %>%
+  select( `Geographic Scale`,
+         Implementation = vlabel, 
+          `Percent Below Interval`,
+         `Percent Contained in Interval`,
+           `Percent Above Interval`)
+}
+
+county_coverage <- county_covidestim %>% get_coverage(geo_scale="County")
+saveRDS(county_coverage, here("figures/table_data/county_coverage.RDS"))
+
+state_coverage <- state_covidestim %>% get_coverage(geo_scale="State")
+saveRDS(state_coverage, here("figures/table_data/state_coverage.RDS"))
+
+```
+
+\@ref(table:coverage) is
+
+```{r}
+
+county_coverage %>%
+  bind_rows(state_coverage) %>%
+  group_by(`Geographic Scale`) %>%
+  arrange(desc(`Percent Contained in Interval`),
+          .by_group=TRUE) %>%
+  ungroup() %>%
+  select(-`Geographic Scale`) %>%
+  kbl(caption = "\\label{table:coverage}Coverage of Covidestim Medians",
+      booktabs=TRUE,  
+      escape=FALSE,
+      digits=3) %>%
+  kable_styling() %>%
+  group_rows(start_row=1, end_row =4, group_label= "County") %>%
+  group_rows(start_row=5, end_row =8, group_label= "State")
+
+```
+
+```{r,eval=FALSE}
+
+county_coverage <- county_covidestim %>%
+ mutate(in_interval = case_when(
+    exp_cases_lb <= infections & 
+      infections <= exp_cases_ub ~ "Contained in Interval",
+    exp_cases_lb  > infections  ~ "Below Interval",
+    exp_cases_ub < infections ~ "Above Interval"
+  ) ) %>%
+  group_by(version, in_interval) %>%
+  summarize(n=n()) %>%
+  group_by(version) %>%
+  mutate(total = sum(n)) %>%
+  ungroup() %>%
+  mutate(prop = n/total) %>%
+  select(version, prop, in_interval) %>%
+  pivot_wider(names_from = in_interval, values_from=prop, names_prefix="Percent ") %>%
+  mutate(across(contains("Percent"), ~.x*100 )) %>%
+ # mutate(across(contains("Percent"), ~paste0(round(.x,3),"%")))
+  left_join(versions) %>%
+  select( # `Geographic Scale`,
+         Implementation = vlabel, 
+         `Percent Above Interval`,
+         `Percent Contained in Interval`,
+          `Percent Below Interval`)
+
+```

### 2023-07-09 reports/state_figures.Rmd
update figures

@@
   pull(state)
+
 ```
@@
 
-### Original Priors versus CTIS Priors (that do not vary)
+### Wu *et al.*'s Priors versus CTIS Priors (that do not vary)
 
-```{r}
+```{r, fig.width=14,fig.height=9}
+
+colors_labs <- state_res %>%
+  filter(version %in% c("v1","v7")) %>%
+  pull(vlabel) %>% unique()
 
@@
              fill=vlabel),
-             alpha = .6,
-             show.legend=FALSE,
+             alpha = .6
              ) +
@@
           ) +
-  viridis::scale_fill_viridis(option="magma", begin=.1,end=.9, discrete=TRUE)  +
+  viridis::scale_fill_viridis(option="magma", begin=.1,end=.9, discrete=TRUE,
+                              breaks =(colors_labs),
+                              labels=TeX(as.character(colors_labs)))  +
   labs(x="",
@@
        title = "Simulation Intervals Over Time",
-       subtitle = "By State")
+       subtitle = "By State",
+       fill = "")
 
@@
 
-```{r,eval=save_figures, fig.height=10,fig.width=12}
+```{r,eval=save_figures, fig.height=10,fig.width=14}
 

### 2023-07-10 writing/manuscript.Rmd
update figures
update county figures
add asymptomatic rate prior writing

@@
  \usepackage{float}
-
+ \usepackage{amsmath}
+bibliography: "`r rbbt::bbt_write_bib('bib/manuscript.bib', overwrite = TRUE)`"
+csl: "csl/nature.csl"
 ---
@@
 Covidestim intervals summed over the same two-week time-scale are shown in red.
-The scale on the $y$-axis is distinct across states to highlight differences across time within each state."}
+The scale on the $y$-axis is distinct across states to highlight differences across time within each state.", fig.height=10,fig.width=14}
 
@@
 
+```{r ma-heatmap, fig.cap=" The ratio of estimated to observed infections across time for counties in Massachusetts.
+Counties are ordered by the median ratio across time intervals, from the highest ratio (Barnstable) to the lowest (Hampshire).
+Similar to what we see at the state level, the highest ratios were during the summer of 2021 during the Delta wave -- a period of decreased testing.
+The span of time with the highest ratio of estimated to observed infections was July 2, 2021 through July 30, 2021."}
+
+include_graphics(here('figures/ma_county_heatmap.pdf'))
+
+```
+
 ## County Level Estimates in Massachusetts
+
+```{r, fig.width=15, fig.height=16,  fig.cap = "Simulation intervals for counties in Massachusetts.
+Each interval corresponds to a 95\\% simulation interval for the total number of estimated infections for that county in that two-week time interval.
+The columns represent different implementations of the probabilistic bias analysis.
+The first column corresponds to the implementation where we specify priors that are the same for all dates.
+For the implementation in the second column, we center the distribution of $\\beta$ at the ratio of the screening test positivity to the overall test positivity from the survey.
+For the third column, we center the distribution of $\\Pr(S_1|\\text{untested})$ at the percentage of the population experiencing COVID-19-like illness from the survey.
+The fourth column centers both  $\\beta$ and $\\Pr(S_1|\\text{untested})$ at the aforementioned values."}
+
+include_graphics(here('figures/simulation-intervals-faceted-version-county.pdf'))
+
+```
+
+```{r, fig.width=15,  fig.cap = "Simulation intervals for counties in Massachusetts, colored by the implementation of probabilistic bias analysis.
+The implementation that centers both $\\Pr(S_1|\\text{untested})$ and $\\beta$ at their empirical values is consistently the highest among the implementations, and the version only centering $\\Pr(S_1|\\text{untested})$ at the survey value is highly concordant with the version where the priors do not vary by date."}
+
+include_graphics(here('figures/simulation-intervals-color-version-county.pdf'))
+
+```
+
+```{r testrate-v-ratio, fig.cap  = 'The ratio of the median estimated infections to observed infections plotted against the testing rate, where the testing rate is calculated as the total number tested in a two-week interval over the population size.
+When the priors are the same for all time intervals, there is minimal variability relationship between the testing rate and the ratio of estimated to observed infections, since the correction for incomplete testing and diagnostic test inaccuracy is identical for each time-interval.
+However, when we allow $\\beta$ or $\\Pr(S_1|\\text{untested})$ to vary over time, there is more variability in the relationship.
+A horizontal line in red at 1 is included to reference; a ratio of exactly 1 would indicate no infections went unobserved.'}
+
+include_graphics(here('figures/testing_rate_ratio-county.pdf'))
+
+```
 
@@
 
-NOTE add (a), (b) to this figure
+Table \@ref(table:coverage) is
+
+\renewcommand{\arraystretch}{2}
 
 ```{r}
+
+library(tidyverse)
+library(kableExtra)
+
+county_coverage <- readRDS(here("figures/table_data/county_coverage.RDS"))
+state_coverage <- readRDS(here("figures/table_data/state_coverage.RDS"))
+
+row_num <- nrow(county_coverage) + nrow(state_coverage)
+
+county_coverage %>%
+  bind_rows(state_coverage) %>%
+  group_by(`Geographic Scale`) %>%
+  arrange(desc(`Percent Contained in Interval`),
+          .by_group=TRUE) %>%
+  ungroup() %>%
+  select(-`Geographic Scale`) %>%
+  kbl(caption = "\\label{table:coverage}Coverage of Covidestim Medians at the County and State Levels",
+      booktabs=TRUE,  
+      escape=FALSE,
+      digits=3) %>%
+  kable_styling(latex_options=c("scale_down","HOLD_position"))  %>%
+  group_rows(start_row=1, end_row =4, group_label= "County", 
+             latex_gap_space = "0.7em",
+             hline_after = TRUE) %>%
+  group_rows(start_row=5, end_row =8, group_label= "State",
+             latex_gap_space = "0.7em",
+             hline_after = TRUE) %>%
+  row_spec(1:(row_num-1),  hline_after = TRUE, 
+           extra_css = "border-bottom: 1px solid;") %>%
+  row_spec(0, font_size=11, bold=TRUE) %>%
+  footnote(general = "The percent of simulation intervals where the Covidestim median falls below, within, or above the interval, when considering all simulation intervals for that implementation and geographic scale.", general_title ="", footnote_as_chunk=TRUE, threeparttable=TRUE)
+
+```
+
+```{r,eval=FALSE}
 
@@
 
+```{r,out.width='70%', fig.align='center'}
+
+include_graphics(here('figures/analysis_diagram.png'))
+
+```
+
 ### Probabilistic bias analysis
@@
 ### Specification of Priors
+
+#### $\Pr(S_0|\text{test}_+,\text{untested})$ \newline\newline
+
+There is substantial heterogeneity in estimates of the percent of infections that are asymptomatic.
+This is in part due to distinct study populations and selection criteria.
+In particular, estimates from screening studies may be better estimates of the asymptomatic rate among the untested population: estimates from studies where the population was not screened, and as such was comprised of individuals that sought out a PCR test, may include a higher proportion of symptomatic individuals, biasing estimates of the asymptomatic rate downwards.
+
+One meta-analysis included studies across the globe as of February 4, 2021, and estimated the pooled percent of asymptomatic infections among confirmed infections to be 40.50% (95% CI 33.50%-47.50%) [@ma2021].
+This analysis did not restrict to screening studies.
+Another meta-analysis, when restricting to screening studies, found the pooled asymptomatic percentage to be 47.3% (95% CI, 34.0 - 61.0%) [@sah2021].
+Both meta-analyses noted the substantial amount of heterogeneity in the percent of asymptomatic infections.
+
+In a large screening study where the sample was individuals arriving from overseas, the asymptomatic rate was 76.8% [@fang2023].
+A screening study among children admitted to a pediatric emergency department between May 2020 and January 2021 found  the asymptomatic rate to be 51.7% [@ford2022].
+
+Several studies were conducted among university students.
+Among students at the University of Arizona in the fall semester of 2020, including students who sought testing and who were required to test, the asymptomatic rate of infection was 79.2%.
+A study at the University of Notre Dame distinguished between presymptomatic infection and asymptomatic infection, and found found 32% to be asymptomatic throughout the entire course of infection, 27.0% to be presymptomatic, and 40.5% to be symptomatic.
+The asymptomatic rate among nonresidential students participating in the surveillance testing system at Clemson University was 69%.The generalizability of these studies is limited given that students are typically healthy and as such may be more likely to experience asymptomatic infection.
+
+Vaccine coverage also may influence the asymptomatic rate.
+In a study in Israel on the effectiveness of the Pfizer–BioNTech mRNA COVID-19 vaccine BNT162b2, 55.7% (49,138 out of 88203) of infections were asymptomatic in the unvaccinated group, and 68.2% of infections  (3632 out of 5324) of infections were asymptomatic in the vaccinated group.
+
+Numerous additional factors contribute to the heterogeneity we see among estimates of the percent of infections that are asymptomatic, including community prevalence, the study population, and the time period when the study population was tested.
+The use of different definitions also may contribute.
+This includes the definition of a symptomatic infection, since our understanding of the clinical presentation of COVID-19 has evolved over time [@sah2021], as well as the definition of asymptomatic, since some define asymptomatic to include presymptomatic cases, where people that had no symptoms upon testing positive but may have went on to develop symptoms at a later date, and truly asymptomatic cases, where an infected individual never goes on to develop symptoms.
+
+Because of the heterogeneity in estimates of the percent of infections that are asymptomatic, for this prior we specified a beta distribution with the majority of the density between 0.3 and 0.8, with a mean of 0.55 and standard deviation of 0.12.
+
+### Limitations
+
+# References

### 2023-07-10 reports/county_figures.Rmd
update figures

@@
 # plot relationship between ratio of estimated cases to observed 
-# against testing rate 
+# against testing rate -- colored by county 
 ###########################################################################
 county %>%
-  ggplot(aes(x=total/population, y = exp_cases_median/positive, color=population)) +
+  ggplot(aes(x=total/population, y = exp_cases_median/positive, color=name)) +
   geom_point(alpha=.03, size=.9) +
@@
   scale_x_continuous( n.breaks=4, limits=c(lb,.3))
+
+```
+
+```{r, fig.height=12}
+
+###########################################################################
+# plot relationship between ratio of estimated cases to observed 
+# against testing rate  -- faceted by county 
+###########################################################################
+county %>%
+  ggplot(aes(x=total/population, y = exp_cases_median/positive)) +
+  geom_point(alpha=.03, size=.9) +
+  geom_linerange(aes(ymin = exp_cases_lb/positive,
+                     ymax=exp_cases_ub/positive ),
+                 alpha=.03) +
+  facet_grid(name~vlabel,
+               labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
+  theme_c(axis.text.x = element_text(size =9),
+          axis.title = element_text(size=16) ) +
+  theme(strip.text.x = element_text(size=11, color="white")) +
+  scale_y_continuous(n.breaks=10) +
+  geom_hline(yintercept=1, linetype=2,color="darkred", alpha=.8, linewidth=.5) +
+  labs(y = "Ratio of Estimated Infections to Observed Infections",
+       x = "Testing Rate",
+       subtitle = "All Time Intervals and All States",
+       title = "Relationship Between Testing Rate and\nRatio of Estimated Infections to Observed Infections")+
+  scale_x_continuous( n.breaks=4, limits=c(lb-.02,.3))
 
@@
 
-county_coverage <- county_covidestim %>% get_coverage(geo_scale="County")
+county_coverage <- county_covidestim %>% get_coverage(geo_scale="County Level (in Massachusetts)")
 saveRDS(county_coverage, here("figures/table_data/county_coverage.RDS"))
 
-state_coverage <- state_covidestim %>% get_coverage(geo_scale="State")
+state_coverage <- state_covidestim %>% get_coverage(geo_scale="State Level")
 saveRDS(state_coverage, here("figures/table_data/state_coverage.RDS"))
@@
 ```
-
-\@ref(table:coverage) is
 

### 2023-07-10 reports/priors.Rmd
update figures

@@
 
+tibble(x = seq(0, 1, length = 10^(5)),
+       y = with(prior_params, beta_density(x,
+                                     mean = s_untested_mean,
+                                     sd = s_untested_sd,
+                                     bounds = s_untested_bounds))) %>%
+   ggplot(aes(x=x, y = y)) +
+  geom_line(alpha = .8) + 
+  geom_ribbon(aes(x=x,ymin=0,ymax=y),
+              fill="black",
+              alpha=.6) +
+  theme_c(legend.text=element_text(size = 10)) +
+  viridis::scale_fill_viridis(discrete=TRUE,
+                             option = "rocket", begin=.3,end=.8) +
+  labs(x = "Value",
+       y = "Probability Density", 
+       title = latex2exp::TeX("Definition of Prior for $P(S_1|untested)$"),
+       fill ='',
+       subtitle = paste0("Mean: ", prior_params$s_untested_mean,
+                         ", SD: ", prior_params$s_untested_sd))
+
 ```
@@
                          ", SD: ", prior_params$p_s0_pos_sd))
+
+tibble(x = seq(0, 1, length = 10^(5)),
+       y = with(prior_params, beta_density(x,
+                                     mean = .55,
+                                     sd = .12,
+                                     bounds = NA))) %>%
+   ggplot(aes(x=x, y = y)) +
+  geom_line(alpha = .8) + 
+  geom_ribbon(aes(x=x,ymin=0,ymax=y),
+              fill="black",
+              alpha=.6) +
+  theme_c(legend.text=element_text(size = 10)) +
+  viridis::scale_fill_viridis(discrete=TRUE,
+                             option = "rocket", begin=.3,end=.8) +
+  labs(x = "Value",
+       y = "Probability Density", 
+       title = latex2exp::TeX("Definition of Prior for $P(S_0|test_+,untested)$"),
+       fill ='',
+       subtitle = paste0("Mean: ", .55,
+                         ", SD: ", .12)) +
+  scale_x_continuous(n.breaks=11)
+
+tibble(samp =sample_beta_density(1e4, mean = .55,
+                                     sd = .12,
+                                     bounds = NA))
+
+qbeta(.025, shape1= get_beta_params( mu = .55,
+                                     sd = .12)$a,
+      shape2= get_beta_params( mu = .55,
+                                     sd = .12)$b
+      )
+
+qbeta(.975, shape1= get_beta_params( mu = .55,
+                                     sd = .12)$a,
+      shape2= get_beta_params( mu = .55,
+                                     sd = .12)$b
+      )
 
@@
 ```
+
+# Bayesian Melding
+
+```{r}
+
+library(patchwork)
+library(truncdist)
+
+source(here('R/04-melding.R'))
+
+params <- prior_params
+params$beta_shape1 <- beta_shape[1]
+params$beta_shape2 <- beta_shape[2]
+params$s_untested_shape1 <- s_untested_shape[1]
+params$s_untested_shape2 <- s_untested_shape[2]
+params$direct_params <- TRUE
+# 
+# prior_params <- list(
+#     alpha_mean = .95,
+#     alpha_sd = 0.08,
+#     alpha_bounds = NA,
+#    # alpha_bounds = c(.8,1),
+#     beta_mean = .15,
+#     beta_sd =.09,
+#     beta_bounds = NA,
+#   #  beta_bounds = c(0.002, 0.4),
+#     s_untested_mean = .03,
+#     s_untested_sd = .0225,
+#   #  s_untested_bounds = c(0.0018, Inf),
+#     s_untested_bounds = NA,
+#     p_s0_pos_mean = .4,
+#     p_s0_pos_sd = .1225,
+#  #  p_s0_pos_bounds = NA,
+#     p_s0_pos_bounds = c(.25, .7),
+#     pre_nsamp = 1e4,
+#     post_nsamp = 1e4)
+
+melded <- do.call(get_melded, prior_params)
+
+theta_wide <- with(prior_params,  
+ theta <- tibble(x= seq(0,1.4, length = 200),
+                 alpha = gamma_density(x,
+                                       mean = alpha_mean,
+                                       sd = alpha_sd,
+                                       bounds = alpha_bounds),
+                    # beta= beta_density(x,
+                    #                           mean = beta_mean,
+                    #                           sd = beta_sd,
+                    #                           bounds = beta_bounds),
+                    # P_S_untested = beta_density(x,
+                    #                             mean = s_untested_mean,
+                    #                             sd = s_untested_sd,
+                    #                             bounds = s_untested_bounds),
+                 P_S0_pos = beta_density(x,
+                                         mean = p_s0_pos_mean,
+                                         sd = p_s0_pos_sd,
+                                         bounds = p_s0_pos_bounds)
+                 )) %>%
+  mutate(beta = dbeta(x, shape1= beta_shape[1], shape2=beta_shape[2]),
+         P_S_untested = dbeta(x, shape1=s_untested_shape[1],
+                              shape2=s_untested_shape[2] ))
+
+theta_samp<- with(prior_params,  
+ theta <- tibble(alpha = sample_gamma_density(pre_nsamp,
+                                                mean = alpha_mean,
+                                                sd = alpha_sd,
+                                                bounds = alpha_bounds),
+                    beta= rbeta(pre_nsamp,
+                                shape1= beta_shape[1], 
+                                shape2=beta_shape[2]),
+                    P_S_untested = rbeta(pre_nsamp,
+                                         shape1=s_untested_shape[1],
+                                         shape2=s_untested_shape[2])) %>%
+        mutate(phi_induced = est_P_A_testpos(P_S_untested = P_S_untested,
+                                             alpha = alpha,
+                                             beta=beta))) %>%
+  select(phi_induced)
+
+theta <- theta_wide %>% 
+  pivot_longer(-c(x)) %>%
+    mutate(name = case_when(
+      name == "alpha" ~"$\\alpha$",
+      name == "beta" ~"$\\beta$",
+      name == "phi_induced" ~ "$M(\\theta) = Pr(S_0|test+,untested)$",
+      name == "P_S_untested" ~ "$Pr(S_1|untested)$",
+      name == "P_S0_pos" ~ "$Pr(S_0|test+,untested)$")
+    ) %>%
+    mutate(name = factor(name,
+                         levels = c(
+                           "$\\alpha$",
+                           "$\\beta$",
+                           "$Pr(S_1|untested)$",
+                           "$Pr(S_0|test+,untested)$",
+                           "$M(\\theta) = Pr(S_0|test+,untested)$")))
+
+```
+
+```{r, fig.width=9,fig.height=9}
+
+melded_long <- reformat_melded(melded$post_melding,
+                               melded$pre_melding,
+                               pre_nsamp = prior_params$pre_nsamp,
+                              p_s0_pos_mean = prior_params$p_s0_pos_mean,
+                              p_s0_pos_sd = prior_params$p_s0_pos_sd,
+                              p_s0_pos_bounds = c(.3,.7)) %>%
+  filter(type=="After Melding") %>%
+  mutate(name = gsub("$P(S_1|untested)$",
+                     "$Pr(S_1|untested)$", name,fixed=TRUE),
+         name = gsub("$P(S_0|test+,untested)$",
+                     "$Pr(S_0|test+,untested)$", name, fixed=TRUE)) %>%
+    mutate(name = factor(name,
+                         levels = c(
+                           "$\\alpha$",
+                           "$\\beta$",
+                           "$Pr(S_1|untested)$",
+                           "$Pr(S_0|test+,untested)$")))
+
+plt1 <- theta %>% 
+  filter(name != "$M(\\theta) = Pr(S_0|test+,untested)$" & name !=  "$Pr(S_0|test+,untested)$") %>%
+  ggplot() +
+  geom_ribbon(aes(x=x, ymin=0, ymax= value, fill='Before Melding'),alpha=.7, color=NA)+
+  geom_density(aes(x=value, fill = 'After Melding'), 
+               color=NA,
+               data = melded_long[melded_long$name !=
+                                    "$Pr(S_0|test+,untested)$",]) +
+  facet_wrap(~name,
+             labeller=  as_labeller(TeX, default = label_parsed),
+             ncol = 3, scales="free_y") +
+  theme_c()+
+  theme(legend.position="right",
+          axis.text.x=element_text(size=6),
+          axis.title.x=element_text(size=12),
+          axis.title.y=element_text(size=12)) +
+  theme(legend.position="none") +
+  scale_x_continuous(n.breaks=5) +
+  labs(x="Value")+
+scale_fill_manual(values = c("Before Melding"= "#5670BF", "Induced" = "#B28542",
+                                 "After Melding" = "#418F6A"))
+
+theta_samp <- theta_samp %>% mutate(name ="$Pr(S_0|test+,untested)$" )
+
+plt2 <- theta %>% 
+  filter(name==  "$Pr(S_0|test+,untested)$") %>%
+  ggplot() +
+  geom_ribbon(aes(x=x, ymin=0, ymax= value, fill='Before Melding'),alpha=.7)+
+  geom_density(aes(x=value, fill = 'After Melding'),
+               alpha=.7,color=NA,
+               data = melded_long[melded_long$name ==
+                                    "$Pr(S_0|test+,untested)$",]) +
+    geom_density(aes(x= phi_induced, fill="Induced"), 
+                 color = NA,
+                 data = theta_samp,
+                 alpha =.7) +
+  facet_wrap(~name,
+             labeller=  as_labeller(TeX, default = label_parsed),
+             ncol = 3, scales="free_y") +
+  theme_c() +
+  theme(legend.position="right",
+          axis.text.x=element_text(size=6),
+          axis.title.x=element_text(size=12),
+          axis.title.y=element_text(size=12))+
+  scale_x_continuous(n.breaks=5, limits =c(0,1)) +
+  labs(x="Value",
+       fill = "")+
+    scale_fill_manual(values = c("Before Melding"= "#5670BF", "Induced" = "#B28542",
+                                 "After Melding" = "#418F6A"))
+
+plt1 / plt2
+
+ggsave(here('figures/melding-priors.pdf'))
+
+```

### 2023-07-10 reports/state_figures.Rmd
update figures

@@
 ```
+
+```{r,eval=FALSE, include=FALSE}
+
+dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
+
+v7_1 <- tar_read(ma_v1, store="_targets_7_08") %>%
+  mutate(targets_version="1")
+
+v7_2 <- tar_read(ma_v1, store="_targets") %>%
+  mutate(targets_version="2")
+
+v7_1 <- tar_read(state_v1, store="_targets_7_08") %>%
+  mutate(targets_version="1")
+
+v7_2 <- tar_read(state_v1, store="_targets") %>%
+  mutate(targets_version="2")
+
+v7_1 %>%
+  bind_rows(v7_2) %>%
+  filter(fips %in% sample(unique(v7_1$fips),5)) %>%
+  left_join(dates,relationship = "many-to-many" ) %>%
+  ggplot(aes(x=date, 
+             ymin = exp_cases_lb,
+             ymax=exp_cases_ub,
+             fill = targets_version)) +
+  geom_ribbon(alpha=.6) +
+  facet_wrap(~fips,scales="free")
+
+v7_1 %>%
+  bind_rows(v7_2) %>%
+  select(fips, exp_cases_median, biweek, targets_version) %>%
+  pivot_wider(names_from=targets_version,
+              values_from = exp_cases_median,
+              names_prefix="exp_cases_targets_v") %>%
+  mutate(diff =exp_cases_targets_v1 - exp_cases_targets_v2) %>% 
+  summarize(diff =mean(diff))
+
+```

### 2023-07-10 R/02-priors.R

@@ -1,6 +1,26 @@
 
 get_priors <- function() { 
   
+  # list(
+  #   alpha_mean = .95,
+  #   alpha_sd = 0.08,
+  #   alpha_bounds = NA,
+  #   # alpha_bounds = c(.8,1),
+  #   beta_mean = .15,
+  #   beta_sd =.09,
+  #   beta_bounds = NA,
+  #   #  beta_bounds = c(0.002, 0.4),
+  #   s_untested_mean = .03,
+  #   s_untested_sd = .0225,
+  #   #  s_untested_bounds = c(0.0018, Inf),
+  #   s_untested_bounds = NA,
+  #   p_s0_pos_mean = .4,
+  #   p_s0_pos_sd = .1225,
+  #   p_s0_pos_bounds = NA,
+  #   #  p_s0_pos_bounds = c(.25, .7),
+  #   pre_nsamp = 1e6,
+  #   post_nsamp = 1e5)
+  
   list(
     alpha_mean = .95,
     alpha_sd = 0.08,
@@ -14,12 +34,13 @@ get_priors <- function() {
     s_untested_sd = .0225,
     #  s_untested_bounds = c(0.0018, Inf),
     s_untested_bounds = NA,
-    p_s0_pos_mean = .4,
-    p_s0_pos_sd = .1225,
+    p_s0_pos_mean = .55,
+    p_s0_pos_sd = .12,
     p_s0_pos_bounds = NA,
     #  p_s0_pos_bounds = c(.25, .7),
     pre_nsamp = 1e6,
     post_nsamp = 1e5)
+  
   
 }
 

### 2023-07-10 R/03-base-functions.R

@@ -66,12 +66,8 @@ sample_beta_density <- function(n, mean, sd, bounds = NA) {
     mu = mean,
     sd = sd)
   
-  rbeta(n,
-        shape1 = shape_params$a,
-        shape2 = shape_params$b)
-  
   if(!length(bounds) == 1){
-    # message("here")
+    message(paste("Truncating with bounds: ", paste(bounds,collapse=", ")))
     rtrunc(n,
            spec = "beta",
            a = bounds[1],

### 2023-07-10 R/04-melding.R

@@ -21,7 +21,7 @@ get_melded <- function(alpha_mean = 0.9,
                        beta_shape1=NULL,
                        beta_shape2=NULL,
                        s_untested_shape1=NULL,
-                       s_untested_shape2) {
+                       s_untested_shape2=NULL) {
   
   
   given_args <- as.list(environment())

### 2023-07-10 _targets.R
update targets

@@ -5,6 +5,7 @@
 
 # Load packages required to define the pipeline:
 library(targets)
+library(crew)
 # library(tarchetypes) # Load other packages as needed. # nolint
 
 # Set target options:
@@ -12,12 +13,13 @@ tar_option_set(
   # packages needed for targets
   packages = c("tidyverse", "lubridate", 
                "here", "httr", "imputeTS"), 
+  controller=crew_controller_local(workers=4),
   format = "rds" # default storage format
   # Set other options as needed.
 )
 
 # tar_make_clustermq() configuration (okay to leave alone):
-options(clustermq.scheduler = "multicore")
+# options(clustermq.scheduler = "multicore")
 
 # Run the R scripts in the R/ folder with your custom functions:
 #tar_source("R/priors.R")

### 2023-07-10 reports/county_figures.Rmd
update county figures

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,cache=FALSE, fig.width=10, fig.height=7)
+knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,cache=TRUE, fig.width=10, fig.height=7)
 
@@
 
-ggsave(here('figures/testing_rate_ratio-county.pdf'), width=11,height=7, dpi=400)
+ggsave(here('figures/testing-rate-ratio-county.pdf'), width=11,height=7, dpi=400)
 

### 2023-07-10 writing/manuscript.Rmd
update county figures

@@
 
-For simplicity, at the state-level we focus on the implementation that does not vary by state or date.
+For simplicity, at the state-level we focus on the implementation where the priors do not vary by state or date.
 This also allows us to consider the entirety of 2021, since survey data is only available for dates after March 20, 2021.
@@
 
-include_graphics(here('figures/testing_rate_ratio-county.pdf'))
+include_graphics(here('figures/testing-rate-ratio-county.pdf'))
 
@@
   row_spec(0, font_size=11, bold=TRUE) %>%
-  footnote(general = "The percent of simulation intervals where the Covidestim median falls below, within, or above the interval, when considering all simulation intervals for that implementation and geographic scale.", general_title ="", footnote_as_chunk=TRUE, threeparttable=TRUE)
+  footnote(general = paste0("The percent of simulation intervals where the Covidestim median falls below, within, or above the interval,",
+  "when considering all simulation intervals for that implementation and geographic scale.",
+  "Implementations are ordered from those with the highest proportion of Covidestim medians contained in the interval, to the lowest."),
+  general_title ="", footnote_as_chunk=TRUE, threeparttable=TRUE)
 
@@
 
+```{r}
+
+```
+
+#### $\alpha$
+
+#### $\beta$
+
+#### $\Pr(S_1| \text{untested})$
+
 #### $\Pr(S_0|\text{test}_+,\text{untested})$ \newline\newline
+
+\vspace{4 mm}
 

### 2023-07-10 R/01-data_cleaning.R
add variants target

@@ -607,6 +607,63 @@ get_biweekly_wastewater <- function() {
 }
 
 
+#------------- Variant Data -----------------
+
+get_variant_proportions <- function() {
+  
+  variant <- "https://data.cdc.gov/resource/jr58-6ysp.json?$limit=50000&$where=week_ending between '2021-02-18T00:00:00.000' and '2022-03-01T00:00:00.000'&usa_or_hhsregion=USA"
+  variant <- httr::GET(URLencode(variant))
+  
+  
+  # only go to the 6th because these are already weekly counts
+  # get the last couple weeks of 2021
+  variant <-jsonlite::fromJSON(
+    httr::content(variant,
+                  as = "text",
+                  encoding = "UTF-8"),
+    simplifyVector = TRUE,
+    flatten = TRUE)  %>%
+    as_tibble() 
+  
+  saveRDS(variant, 'data/data_raw/variant_prop.RDS')
+  
+  
+  # https://www.cdc.gov/coronavirus/2019-ncov/variants/variant-classifications.html
+  variant <- variant %>% 
+    mutate(week_end_date = substr(week_ending, 1,10),
+           week_end_date=ymd(week_end_date)) %>%
+    mutate(variant_category = case_when(
+      grepl("B[.]1[.]1[.]529|BA[.]1|BA[.]2|BA[.]4|BA[.]5", variant) ~ "Omicron",
+      variant %in% c("B.1.621", "B.1.621.1") ~ "Mu",
+      variant == "B.1.1.7" ~ "Alpha",
+      grepl("B[.]1[.]351", variant) ~ "Beta",
+      grepl("P.1", variant) ~"Gamma",
+      grepl("B[.]1[.]617[.]2", variant) ~ "Delta",
+      variant %in% c("B.1.427" ,"B.1.429") ~ "Epsilon",
+      variant == "B.1.525" ~ "Eta",
+      variant == "Other" ~ "Other"
+    )) %>%
+    mutate(share = as.numeric(share),
+           creation_date = ymd(substr(creation_date,1,10))) %>%
+    filter(modeltype == "weighted" & time_interval=="weekly") %>%
+    group_by(variant, week_end_date, variant_category, time_interval) %>% 
+    slice_max(n=1, order_by=creation_date) %>%
+    group_by(variant_category, week_end_date, time_interval) %>%
+    summarize(share =sum(share, na.rm=TRUE))
+  
+  
+  variant <- variant %>%
[…]

### 2023-07-10 _targets.R

@@ -57,6 +57,9 @@ list(
   tar_target(
     name = ma_biweekly_county,
     command = get_mass_county()),
+  tar_target(
+    name = variant,
+    command = get_variant_proportions()),
   
   #---------------- state-level analysis -------------------
   # set priors

### 2023-07-10 reports/county_figures.Rmd
update manuscript

@@
                                    default = label_parsed)) +
-  theme_c() +
+  theme_c(axis.title = element_text(size=15),
+          strip.text=element_text(size=15, color='white', margin=margin(2,0,2,0))) +
   labs(x = "Probabilistic Bias Infections\n(Standardized by County)",
@@
                                    default = label_parsed)) +
-  theme_c() +
-  theme(strip.text.x = element_text(size=12)) +
+  theme_c(axis.title = element_text(size=15),
+          strip.text=element_text(size=15, color='white',margin=margin(2,0,2,0))) +
+ # theme(strip.text.x = element_text(size=12)) +
   labs(x = "Probabilistic Bias Infections\n(Standardized by County)",
@@
                                    default = label_parsed)) +
-  theme_c() +
-  theme(strip.text.x = element_text(size=12)) +
+ theme_c(axis.title = element_text(size=15),
+          strip.text=element_text(size=15, color='white',margin=margin(2,0,2,0))) +
+ # theme(strip.text.x = element_text(size=12)) +
   labs(x = "Probabilistic Bias Infections\n(Standardized by County)",
@@
 
+cat("Maximum ratio, on average, occured: ")
+county %>% 
+  filter(version=="v7") %>% 
+  group_by(biweek) %>% 
+  summarize(m = mean(exp_cases_median/positive),dates=paste(range(date),collapse=",")) %>%
+  slice_max(n=1, m)
+
 ######################################################
@@
   geom_hline(yintercept=1, linetype=2,color="darkred", alpha=.8, linewidth=.5) +
-  labs(y = "Ratio of Estimated Infections to Observed Infections",
+  labs(y = "Ratio of Estimated Infections\nto Observed Infections",
        x = "Testing Rate",

### 2023-07-10 reports/priors.Rmd
update manuscript

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)
+knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, cache=TRUE)
 
@@
 
+```{r beta-spline-smoothed,fig.cap="The ratio of the screening test positivity over the overall test positivity from the COVID-19 Trends and Impact Survey is taken to be the estimate of $\\beta$.
+Because of the noise present inthe data, we use the spline smoothed values to inform the priors.
+More specifically, we use cubic spline smoothing with 2 knots (July 15th, 2021 and December 1st, 2021).", fig.width=12,fig.height=12}
+
+num_states <- ctis_smoothed  %>%
+  filter(keep) %>%
+  pull(state) %>%
+  unique() %>%
+  length()
+
+cat("There are", num_states, "states with sufficient data available to estimate $\\beta$ across the time period considered.")
+
+ctis_smoothed  %>%
+  filter(keep) %>%
+  mutate(state=toupper(state)) %>%
+  select(date,
+         state, 
+         imputed_beta,
+        # beta_estimate_smoothed, 
+         beta_estimate_spline_smoothed) %>%
+  pivot_longer(contains("beta")) %>%
+  mutate(name = case_when(
+    name == "beta_estimate_spline_smoothed" ~ "Spline smoothed",
+    name == "imputed_beta" ~ "Survey Value"
+  )) %>%
+  ggplot(aes(x=date, y=value, color = name, 
+             alpha = name, linewidth=name)) +
+  geom_line() +
+  facet_wrap(~state, ncol=4) +
+  scale_alpha_manual(values=c(
+                            "Spline smoothed" = .9, 
+                              "Survey Value"=.3),
+                     name='') +
+  scale_linewidth_manual(values = c(
+                            "Spline smoothed" = 1.05, 
+                              "Survey Value"=.5)) +
+  scale_color_manual(values=c("#3381FF", "#26900F"),
+                     name ='') +
+  guides(alpha="none",
+         linewidth="none",
+         color = guide_legend(override.aes = list(linewidth = 3,
+                                                  alpha = c(
+                            "Spline smoothed" = .9, 
+                              "Survey Value"=.3)),
+                            nrow=3)) +
+  theme_c(legend.position="top") +
+  ylim(0,1) +
+  scale_x_date(date_breaks="3 months",
+               date_labels = "%b %Y") + 
+  labs(y = TeX("Survey Estimate of $\\beta$"),
+       x= "",
+       title = TeX("Smoothing Survey Estimates of $\\beta$"))
+
+ggsave(here('figures/smoothing-beta.pdf'))
+
+```
+
 ## $\Pr(S_1 | \text{untested})$
@@
                               "Survey Value"=.9)) +
-  scale_color_manual(values=c("Survey Value" ="black", "LOESS smoothed"="darkred"),
+  scale_color_manual(values=c("Survey Value" = "#26900F",
+                              "LOESS smoothed"="#3381FF"),
                      name ='') +
@@
        x= "",
-       title = TeX("Comparing Approaches for Smoothing Survey Estimates of $Pr(S_1|untested)$"))
+       title = TeX("Smoothing Survey Estimates of $Pr(S_1|untested)$"))
+
+ggsave(here('figures/smoothing-s-untested.pdf'))
 
@@
 
-```{r, fig.width=9,fig.height=9}
+```{r, fig.width=9,fig.height=6}
 
@@
                                     "$Pr(S_0|test+,untested)$",]) +
+    geom_point(aes(x=1,y=0),size=0, color=NA) +
   facet_wrap(~name,
              labeller=  as_labeller(TeX, default = label_parsed),
-             ncol = 3, scales="free_y") +
+             ncol = 3, scales="free") +
   theme_c()+
@@
              labeller=  as_labeller(TeX, default = label_parsed),
-             ncol = 3, scales="free_y") +
+             ncol = 3, scales="free") +
   theme_c() +
@@
        fill = "")+
-    scale_fill_manual(values = c("Before Melding"= "#5670BF", "Induced" = "#B28542",
+    scale_fill_manual(values = c("Before Melding"= "#5670BF", 
+                                 "Induced" = "#B28542",
                                  "After Melding" = "#418F6A"))

### 2023-07-10 reports/state_figures.Rmd
update county figures
update manuscript

@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE,cache=FALSE)
+knitr::opts_chunk$set(echo = TRUE, message=FALSE,warning=FALSE,cache=TRUE)
 
@@
 
-filter_versions <- FALSE
-save_figures <- FALSE
+filter_versions <- TRUE
+save_figures <- TRUE
 
@@
 
-```{r, fig.height=14, fig.width=11}
+```{r, fig.height=14, fig.width=12}
+
+comp_state_pal <- c("#247C90", "red")
+
+names(comp_state_pal) <- c('Probabilistic Bias Analysis', "Covidestim")
 
 subset %>%
+   # mutate(vlabel = gsub("$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$",
+   #                     "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", as.character(vlabel))) %>% 
    ggplot() +
@@
              ymax = exp_cases_ub,
-             fill = vlabel),
+             fill = 'Probabilistic Bias Analysis'),
              alpha = .8,
@@
                date_labels = "%b %Y") +
-  theme_c(axis.text.x=element_text(size=5),
+  theme_c(axis.text.x=element_text(size=7),
           axis.text.y = element_text(size=4.5),
@@
           legend.text=element_text(size=12),
-          strip.text.x=element_text(size=8,
+          strip.text.x=element_text(size=11,
                                     face='plain',
                                     color='white')) +
-  scale_fill_manual(values =pal,
-                    labels = TeX(names(pal)),
-                    name=''
-                    #,
-                  # , breaks="Covidestim"
-                    )  +
+  scale_fill_manual(values =comp_state_pal, name='')  +
   labs(y = "Estimated Infections",
@@
 
-```{r,eval=save_figures, fig.height=14, fig.width=9}
+```{r,eval=save_figures, fig.height=14, fig.width=12}
 
@@
 
+max_val <- subset %>% 
+  # left_join(codes) %>%
+  filter(biweek %in% c(7, 13,27)) %>%
+  mutate(ratio=exp_cases_ub/positive) %>%
+  pull(ratio) %>%
+  max()
+
 plt1 <- subset %>% 
@@
        subtitle = "March 26, 2021 through April 8, 2021") +
-  xlim(0,65)
+  xlim(0,max_val+2)
 
@@
        subtitle = "June 18, 2021 through July 1, 2021") +
-  xlim(0,65)
+  xlim(0,max_val+2)
 
@@
        subtitle = "December 31, 2021 through January 14, 2022") +
-  xlim(0,65)
+  xlim(0,max_val+2)
 
@@
 ggsave(here('figures/ratio-estimated-to-observed-multiple-waves.pdf'))
+
+```
+
+```{r}
+
+cat("Time interval with the highest ratio of estimated cases to observed: ")
+
+subset %>%
+  group_by(biweek) %>%
+  mutate(ratio=exp_cases_median/positive) %>%
+  summarize(mean_ratio=mean(ratio),
+         date = paste(range(date),collapse=" to ")) %>%
+  slice_max(n=1, order_by=mean_ratio)
 
@@
 
+# Summer Testing Rate
+
+```{r}
+
+subset %>% 
+  mutate(testrate = total/population) %>%
+  group_by(biweek) %>% 
+  mutate(m = median(testrate),
+            date=min(date)) %>%
+  ggplot(aes(x=date,y=testrate, group=state)) +
+  geom_line(alpha=.2) +
+  geom_line(aes(y=m, color='Median'), size=2) +
+  labs(y = "Total Number of Tests\nOver Population Size",
+       title ="Testing Rate Over Time",
+       x="") +
+  theme_c() +
+  theme(axis.title.x = element_text(size=16),
+          axis.title.y  = element_text(size=16),
+        legend.position='top') +
+  geom_vline(xintercept = ymd("2021-07-16"), linetype = 2, linewidth=.5) +
+  geom_vline(xintercept = ymd("2021-06-18"), linetype = 2, linewidth=.5) +
+  scale_x_date(date_breaks="2 months",
+               date_labels = "%b %Y") +
+  scale_color_manual(values=c(Median="darkred"),name='')
+
+ggsave(here('figures/testrate-low-summer.pdf'), height=6,width=12, dpi=400)
+
+```
+
 ```{r,eval=FALSE, include=FALSE}
@@
 ```
+
+# Variant Data
+
+```{r, fig.heihgt=7}
+
+m <- state_res %>%
+  filter(state == "MI" & version=="v7") %>%
+  pull(exp_cases_ub) %>%
+  max()
+
+variant <- tar_read(variant, store =targets_dir)
+
+varpal <- c("#7BD8DA","#709f0f", "#e71761", "#9245c8", 
+            "#97481c", "#5054b1", "#DA7BA8", "#26B392")
+
+state_res %>%
+  filter(state == "MI" & version %in% c("v7")) %>%
+ # filter(state %in% sample(corrected$state,5)) %>%
+  filter(biweek >=6) %>%
+  group_by(biweek) %>%
+  mutate(xmin = min(date),
+         xmax = max(date)) %>%
+  ungroup() %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub),
+             fill = "#3E3D3D",
+             alpha = .5,
+             show.legend=FALSE) +
+  # geom_ribbon(aes(x = date,
+  #                 fill = 'Covidestim',
+  #                 ymin = infections.lo,
+  #                 ymax = infections.hi),
+  #             alpha = .5) +
+  # facet_grid(~vlabel, scales = "free_y",
+  #             labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
+  scale_x_date(date_breaks = "2 months", 
+               date_labels = "%b %Y") +
+  theme_c(legend.position="top") +
+  # theme_c(axis.text.x = element_text(size = 16),
+  #         axis.title.y = element_text(size=13),
+  #         plot.title=element_text(face = "bold",
+  #                                 hjust = .5, 
+  #                                 size = 24,
+  #                                 margin=margin(5,5,15,5,'pt')),
+  #         strip.background = element_rect(fill = "#393939"),
+  #         strip.text = element_text(color =  "white", size = 16),
+  #         strip.text.y = element_text(margin = margin(5,5,5,5,'pt'),
+  #                                     size = 22,
+  #                                     face="bold"),
+  #         strip.text.x =  element_text(margin = margin(5,3,5,3,'pt'),
+  #                                      face = "bold",
+  #                                      size = 22),
+  #         plot.subtitle = ggtext::element_markdown(size = 25,margin = margin(3,5,10,5,'pt'),
+  #                                      face = "italic"),
+  #         legend.position = "top",
+  #         legend.text =element_text(size = 17),
+  #         legend.title = element_text(size=18, face="bold")) +
+  geom_line(aes(x=week, y=share*m,
+                color =variant_category,
+                group=variant_category),
+            data=variant,
+            linewidth=1.3) +
+  # scale_fill_manual(values =pal,
+  #                   labels = TeX(names(pal)),
+  #                    breaks='Covidestim',
+  #                   name='')  +
+  scale_y_continuous(sec.axis = sec_axis( trans=~./m, name="Variant Proportion"),
+                     labels = comma) +
+  labs(y = "Estimated Infections",
+       title = paste0("Estimated Infections by Version in Michigan"),
+       color = "SARS-CoV-2 Variant",
+       x="Date") +
+  guides(color = guide_legend(override.aes = list(linewidth =3),
+                              ncol=2,title.position="top")) +
+ # scale_color_brewer(palette="Accent") 
+  scale_color_manual(values=varpal)
+
+ggsave(here('figures/michigan_variant.pdf'), dpi=400)
+
+```

### 2023-07-10 writing/manuscript.Rmd
update manuscript

@@
 
-For simplicity, at the state-level we focus on the implementation where the priors do not vary by state or date.
+At both the state and county levels, we considered multiple implementations of the probabilistic bias analysis to compare estimates produced using different specification of the priors using survey data from the COVID-19 Trends and Impact Survey [@salomon2021a].
+
+However, for simplicity, at the state-level we focus on the implementation where the priors do not vary by state or date.
 This also allows us to consider the entirety of 2021, since survey data is only available for dates after March 20, 2021.
@@
 
-Although prevalence of COVID-19 was highest during the time interval during the Omicron wave, we see that the ratio of estimated infections to observed infections is higher during the time intervals in the alpha and delta waves.
-This distinction is explained by the differences in testing rates during these period: the testing rate during this two-week interval during the omicron wave was 2.4 times that of the alpha wave and 4.9 times that of the delta wave.
+Although incidence of COVID-19 was highest during the time interval during the Omicron wave, as is clear in Figure  \@ref(fig:simulation-intervals-state),  we see that the ratio of estimated infections to observed infections is higher during the time intervals in the alpha and delta waves.
+This distinction is explained by the differences in testing rates during these period: on average, the testing rate during this two-week interval during the omicron wave was 2.4 times that of the alpha wave and 4.9 times that of the delta wave.
+This particular time interval during the delta wave; June 16, 2021 through July 1st, 2021; had the maximum ratio of estimated to observed infections, and corresponded to the minimum testing rate (Supplementary Figure testrate-low-summer).
 
@@
 
-In Figure \@ref(fig:simulation-intervals-state), we see the simulation intervals for all two-week intervals and all states.
+### Comparison to Covidestim
+
+Because there is no established ground truth for the true number of infections for any time-interval, a useful source of comparison is other models that also estimate the true number of incident infections.
+Covidestim is a notable Bayesian nowcasting model that has been maintained throughout the course of the pandemic, sharing publicly available estimates at the state and county levels [@chitwood2022].
+Consequently, we compare our estimates to Covidestim, with recognition that the Covidestim model is not a ground truth, and this comparison serves primarily to gain insight into the times and locations where estimates are concordant or discordant rather than to validate the approach.
+Given the true number of infections is unobserved, we are to some extent operating in the unknown, and must rely on assumptions based on available data.
+
+In Figure \@ref(fig:simulation-intervals-state), we see the simulation intervals for all two-week intervals and all states as well as the 95% Covidestim credible intervals summed to be on the same time scale.
+We see agreement is much higher before the time period spanning December 2021 through January 2022, where Covidestim intervals tend to be higher than the probabilistic bias intervals.
+This period corresponds to the Omicron wave of the pandemic (Supplementary Figure michigan-variant).
+This was a major shift in the pandemic: the Omicron variant is highly transmissible compared to previous variants, and it has immune invasion capacity, which means vaccines provides less protection against infection [@andrews2022].
+Omicron also is associated with a lower infection fatality ratio [@liu2022], as well as milder infection overall, which may influence testing behavior.
+Omicron's rise was rapid too, presumably as a result of this enhanced transmissibility and immune escape;  it became the dominant variant over the course of a single month.
+
+As a result of these dramatic changes, Chitwood *et al.*@chitwood2022 made substantial changes to Covidestim, as noted in the model change log.
+Because the variant causes much milder infections, the infection fatality for Omicron infections is lower than previous variants, and death counts were much lower relative to the number of infections.
+To handle this change, rather than fitting model with deaths, they switched to using hospitalizations.
+They also allowed for the possibility of reinfections, since although reinfections were more rare with previous variants, Omicron is associated with higher reinfection rates [@pulliam2022].
+The changes in the model may contribute to the differences we see between the probabilistic bias intervals and Covidestim intervals during the Omicron wave.
+
+Another difference we see is in the months of the summer of 2021, where the increase in Covidestim estimates appears lagged in comparison to the probabilistic bias simulation intervals.
+The difference we see between the Covidestim estimates and probabilistic bias intervals is likely a result of the way these approaches treat incomplete testing.
+The focus of the probabilistic bias analysis is to correct for incomplete testing, and as such the method is sensitive to changes in the total number tested and the positivity rate.
+By contrast, while Covidestim models probabilities of diagnosis by symptom state to allow for variation in case ascertainment, the total number of tests is not an input into the model, so model estimates are not affected by changes in testing rate.
+This is particularly relevant during the summer of 2021, when an increase in test positivity preceded the increase in observed tests, leading to an increase in the probabilistic bias simulation intervals that precedes the increase in Covidestim estimates.
 
@@
 
+## County Level Estimates in Massachusetts
+
+At the county level, this approach to approximating the true number of infections is only possible in a subset of the counties in the United States due to the need for both positive tests and total tests.
+While positive tests are frequently reported at the county level, total tests are not.
+
+We focus on Massachussetts because this state is the only state that reports both positive and total tests and has the most counties where wastewater data was reported throughout the entire time period of interest.
+Wastewater data here is useful in the sense that it is inherently not subject to the same biases as voluntary testing data, since it captures anyone in the wastewater catchment area, and as such is not influenced by more symptomatic people being more likely to be tested or differences in access to testing.
+
+In Figure (CHOOSE BETWEEN VERSION BY COLOR AND FACETED), we present simulation intervals for each implementation of probabilistic bias analysis.
+
+In the first three implementations, the prior distributions specified for the bias parameters vary by state and date.
+In particular, we use survey data from the COVID-19 Trends and Impact Survey [@salomon2021a] to inform the prior distributions for two of the bias parameters: the probability of having symptoms among the untested population ($\Pr(S_1|\text{untested})$), and the parameter $\beta$, which represents the ratio of the test positivity among members of the untested population who have no symptoms to the overall test positivity.
+More detail on these implementations is provided in the [Methods](#Methods) section.
+In the last implementation, the prior distributions are informed by the distributions from the survey data across all dates and states, but the same prior distributions are used for all states and time intervals considered.
+While there is substantial overlap among all of the versions, we see the implementation allowing $\beta$ and $\Pr(S_1|\text{untested})$ to vary by state and date tends to be the highest, followed by that only allowing $\beta$ to vary by state and date,  followed by that only allowing $\Pr(S_1|\text{untested})$ to vary by state and date, and lastly the implementation that does not vary the priors by state or date tends to produce the lowest estimates.
+
+```{r sim-intervals-faceted-county, fig.width=15, fig.height=16,  fig.cap = "Simulation intervals for counties in Massachusetts.
+Each interval corresponds to a 95\\% simulation interval for the total number of estimated infections for that county in that two-week time interval.
+The columns represent different implementations of the probabilistic bias analysis: in the first through the third columns, the priors vary by state and date, while for the fourth, they are the same for all states and time intervals considered.
+For the implementation in the first column, we center the distribution of $\\beta$ at the ratio of the screening test positivity to the overall test positivity from the survey, and we center the distribution of $\\Pr(S_1|\\text{untested})$ at the percentage of the population experiencing COVID-19-like illness from the survey for each two-week interval.
+The second column centers only $\\beta$ at the aformentioned value, and the third column only centers $\\Pr(S_1|\\text{untested})$ at the aforementioned value.
+The fourth column corresponds to the implementation where we specify priors that are the same for all dates.
+The implementation that centers both $\\Pr(S_1|\\text{untested})$ and $\\beta$ at the survey values is consistently the highest among the implementations, followed by the implementation that centers only $\\beta$ at the survey value, followed then by the implementation only centering $\\Pr(S_1|\\text{untested})$ at the survey value, and then the lowest among the implementations is that where the priors do not vary by date."}
+
+include_graphics(here('figures/simulation-intervals-faceted-version-county.pdf'))
+
+```
+
+```{r sim-intervals-color-county, fig.width=15,  fig.cap = "Simulation intervals for counties in Massachusetts, colored by the implementation of probabilistic bias analysis.
+In the first three implementations, the priors vary by state and date, while for the fourth, they are the same for all states and time intervals considered.
+For the first implementation, we center the distribution of $\\beta$ at the ratio of the screening test positivity to the overall test positivity from the survey, and we center the distribution of $\\Pr(S_1|\\text{untested})$ at the percentage of the population experiencing COVID-19-like illness from the survey for each two-week interval.
+The second implementation, we center only $\\beta$ at the aformentioned value, and the for the next we only center $\\Pr(S_1|\\text{untested})$ at the aforementioned value.
+The last implementation corresponds to the implementation where we specify priors that are the same for all dates.
+The implementation that centers both $\\Pr(S_1|\\text{untested})$ and $\\beta$ at the survey values is consistently the highest among the implementations, followed by the implementation that centers only $\\beta$ at the survey value, followed then by the implementation only centering $\\Pr(S_1|\\text{untested})$ at the survey value, and then the lowest among the implementations is that where the priors do not vary by date."}
+
+include_graphics(here('figures/simulation-intervals-color-version-county.pdf'))
+
+```
+
+Also of interest beyond the total numbers of infections is the ratio of the infections that are estimated to those that were observed, as this gives us a sense for the case ascertainment at different time periods in the pandemic.
+We see in Figure \@ref(fig:ma-heatmap) that the two-week interval estimated to have the highest ratio of estimated to observed infections was July 2, 2021 through July 30, 2021, during the delta wave.
+This corresponds to a similar time frame that we saw at the state-level in Figure \@ref(fig:ratio-estimated-to-observed-by-wave).
+Hamsphire County and Suffolk County were consistently among the lowest with regard to the ratio of estimated to observed infections, which may be explained by the screening testing occurring at universities in these counties.
+
 ```{r ma-heatmap, fig.cap=" The ratio of estimated to observed infections across time for counties in Massachusetts.
@@
 
-## County Level Estimates in Massachusetts
+As we see in Figure \@ref(fig:testrate-v-ratio), the nature of the relationship between the testing rate and the ratio of estimated infections to observed infections depends on whether we allow $\beta$ and $\Pr(S_1|\text{untested})$ to vary by location and date.
+In particular, when we sample from the same priors for every correction (the first panel of Figure \@ref(fig:testrate-v-ratio), we see there is little variability in the relationship between the testing rate and median estimated infections, because the form of the correction is identical for each two-week interval and state considered.
+Allowing $\beta$ and/or $\Pr(S_1|\text{untested})$ to vary by time and location introduces additional variability in the relationship between the ratio of estimated infections to observed and testing rate.
 
-```{r, fig.width=15, fig.height=16,  fig.cap = "Simulation intervals for counties in Massachusetts.
-Each interval corresponds to a 95\\% simulation interval for the total number of estimated infections for that county in that two-week time interval.
-The columns represent different implementations of the probabilistic bias analysis.
-The first column corresponds to the implementation where we specify priors that are the same for all dates.
-For the implementation in the second column, we center the distribution of $\\beta$ at the ratio of the screening test positivity to the overall test positivity from the survey.
-For the third column, we center the distribution of $\\Pr(S_1|\\text{untested})$ at the percentage of the population experiencing COVID-19-like illness from the survey.
-The fourth column centers both  $\\beta$ and $\\Pr(S_1|\\text{untested})$ at the aforementioned values."}
+The nonlinearity of the relationship between the testing rate and the ratio of estimated infections to observed infections is more clear when we consider how we calculate the positives among the untested population.
+To do this, we split up the population $N$ into the number tested, $N_{untested}$, and untested, $N_{tested}$, for that two-week interval.
 
-include_graphics(here('figures/simulation-intervals-faceted-version-county.pdf'))
+Denoting $N^*$ to be the number who would test positive for COVID-19 if they were tested, on the $y$-axis, the ratio of estimated infections to observed infections is approximately^[This isn't exactly the estimated infections, because for simplicity of notation we are not writing out the correction for test inaccuracy.]
+$\frac{N^*_{\text{tested}} + N^*_{\text{untested}}}{N^*_{\text{tested}}},$ where we calculate $N^*_{\text{untested}}$ using the specified priors, and $N^*_{\text{untested}}$ is the observed positive tests.
 
-```
-
-```{r, fig.width=15,  fig.cap = "Simulation intervals for counties in Massachusetts, colored by the implementation of probabilistic bias analysis.
-The implementation that centers both $\\Pr(S_1|\\text{untested})$ and $\\beta$ at their empirical values is consistently the highest among the implementations, and the version only centering $\\Pr(S_1|\\text{untested})$ at the survey value is highly concordant with the version where the priors do not vary by date."}
-
-include_graphics(here('figures/simulation-intervals-color-version-county.pdf'))
-
-```
+On the $x$-axis, we have the number tested over the population size, $\dfrac{N_{\text{tested}}}{N}$.
+Thus, we see the trend in each panel where for small changes in testing rate when the testing rate is very low, the ratio of unobserved to unobserved is very high since $N^*_{\text{untested}}$ will be large relative  $N^*_{\text{tested}}$, that is, a large proportion of infections are going undetected.
+However, with higher testing rates,  $N^*_{\text{untested}}$ will be smaller relative  $N^*_{\text{tested}}$, and the ratio of estimated to observed infections nears one.
 
@@
 
-Because there is no established ground truth to compare to regarding the true number of infections for any time-interval, at the county level we compare our results to two distincts sources of information: wastewater data aggregated at the county-level, and results from a previously published Bayesian evidence synthesis model, Covidestim.
-This also allows us to compare how different implementations of probabilistic bias analysis compare to Covidestim estimates and wastewater concentrations.
+As with the state level results, we also compare estimates at the county-level to Covidestim estimates.
+However, at the county-level, we proceeded with Massachusetts specifically with the aim of providing another source of comparison: wastewater concentrations.
 
-Our first comparison considers the correlations between wastewater concentrations and the probabilistic bias estimates as well as between the Covidestim estimates and the probabilistic bias estimates (Figure \@ref(fig:correlations)).
+We see in Figure \@ref(fig:correlations)a) that for all implementations, the probabilistic bias estimates are highly correlated with wastewater concentrations, and in Figure \@ref(fig:correlations)b) we see they also highly correlated with the Covidestim estimates.
+The correlations in both cases are very similar between implementations, differing by less than 0.01.
 
-```{r correlations, fig.cap ="Considering the correlations between probablistic bias estimates and wastewater concentrations (a) and between probablistic bias estimates and Covidestim estimates (b).
+```{r correlations, fig.cap ="Considering the correlations between the probabilistic bias estimates and wastewater concentrations (a) and between probablistic bias estimates and Covidestim estimates (b).
 We see that all implementations considered are highly correlated with both wastewater concentrations and Covidestim estimates.
-The implementation that does not allow priors to vary by state or date has the highest correlations."}
+In both panels, each point is a county-biweek, where the value on the $x$-axis is the median of the simulation interval for that county in that two-week interval.
+For (a), the value on the $y$-axis is the mean wastewater concentration for that county and two-week interval.
+For (b), the value on the $y$-axis is the median of the Covidestim credible interval for that county, summed to be on the same two-week-interval time scale.
+The correlations are highly similar between implementations, differing by less than 0.01."}
 
@@
 
-```{r}
+```{r melding-priors, fig.width=9,fig.height=9}
+
+include_graphics(here('figures/melding-priors.pdf'))
 
@@
 
+- Validation:
+  - Wastewater encatchments do not fall exactly on county lines, so aggregation here contains limited information
+  - Covidestim -- sum of daily medians != median for entire two-week time interval
+- Repeat testing may bias estimates in places where this form of testing comprises a substantial proportion of total testing, but data on people-viral-total and people-viral-positive is not widely available
+
 # References

### 2023-07-10 writing/supplement.Rmd
update manuscript

@@
 ---
-title: "Supplement"
+title: "Supplementary Materials"
 author: "Quinn White"
@@
 
+# Testing Rate (2021-06-18 - 2021-07-01)
+
+```{r testrate-low-summer, fig.cap="Testing rate, calculated as the etotal number of tests for a 2-week interval in a state over the census population in that state, over time.
+The median across states for each two-week interval is shown in red.
+In relation to Figure 1 in the main text, where we see that the period with the highest ratio of estimated to observed infections is June 18, 2021 through July 1, 2021, we see here that testing rates across states are at a minimum during this time period."}
+
+include_graphics(here('figures/testrate-low-summer.pdf'))
+
+```
+
 ## COVID-19 Trends and Impact Survey
 
+```{r, fig.cap="The ratio of the screening test positivity over the overall test positivity from the COVID-19 Trends and Impact Survey is taken to be the estimate of $\\beta$.
+Because of the noise present inthe data, we use the spline smoothed values to inform the priors.
+More specifically, we use cubic spline smoothing with 2 knots (July 15th, 2021 and December 1st, 2021).
+While there are estimates of $\\Pr(S_1|\\text{untested})$ for all states (seen in the following figure), there are 27 states with estimates of $\\beta$.", fig.width=14,fig.height=12}
+include_graphics(here('figures/smoothing-beta.pdf'))
+```
+
+```{r,fig.cap="The percentage of the population experiencing COVID-19-like illness is taken to be the estimate of $\\Pr(S_1|\\text{untested})$.
+The LOESS smoothed estimate with a span of 0.2 is shown in red.
+Survey data on this variable is available for all states.", fig.width=14,fig.height=12}
+
+include_graphics(here('figures/smoothing-s-untested.pdf'))
+
+```
+
 # County Results by Implementation
+
+# Variant Data
+
+```{r michigan-variant, fig.cap = "Simulation intervals for Michigan along with the variant proportions for variants designated as *Variant of Interest* or *Variant of Concern* by the World Health Organization.
+These variant proportions are not specific to Michigan; the estimates are for the United States.
+The rise in Omicron corresponds to the dramatic increase in both the probabilistic bias intervals and Covidestim estimates in December of 2021 through January of 2022.
+We also see the increase in the Delta variant that precedes the Delta wave in the summer of 2021."}
+
+include_graphics(here('figures/michigan_variant.pdf'))
+
+```

### 2023-07-17 reports/county_figures.Rmd
consistent coloring

@@
     filter(version %in% FILT_VERSIONS) 
+}
+
+```
+
+```{r}
+
+convert_labels <- function(labels, to_newlines=TRUE) {
+
+  if(to_newlines) {
+    gsub("$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
+         "$\\overset{Priors\\,Do\\,Not\\,Vary}{\\,by\\,State\\,and\\,Date}$", 
+         as.character(labels),
+         fixed=TRUE) %>%
+    gsub("$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+         "$\\overset{P(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}", 
+         .,
+         fixed=TRUE)
+  }
+  else {
+      gsub(
+         "$\\overset{Priors\\,Do\\,Not\\,Vary}{\\,by\\,State\\,and\\,Date}$", 
+         "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
+         as.character(labels),
+         fixed=TRUE) %>%
+    gsub("$\\overset{P(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}",
+         "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+         .,
+         fixed=TRUE)
+  }
+
 }
@@
 
+names(pal) <- convert_labels(labels(pal), to_newlines=TRUE)
+
 county %>% 
@@
              ymin = exp_cases_lb,
-             ymax = exp_cases_ub),
-             alpha = .8,
-             show.legend=FALSE,
-             fill="#596281") +
+             ymax = exp_cases_ub,
+             fill = vlabel),
+             alpha = .7,
+             show.legend=FALSE) +
   facet_grid(name~vlabel, 
@@
           legend.position = "top",
-          ) +
-  theme_c(strip.text.x=element_text(size=13,color='white')) +
+          strip.text.x=element_text(size=13,color='white'),
+              axis.title = element_text(size=17),
+          plot.title = element_text(size=24, hjust = .5, face="bold", 
+                                    margin=margin(2,0,2,0)),
+          plot.subtitle = element_text(size=20, face="italic",
+                                       margin=margin(1,0,5,0))) +
   scale_fill_manual(values = pal)  +
@@
        title = "Simulation Intervals Over Time",
-       subtitle = "Counties in Massachusetts") +
-  guides(color = guide_legend(override.aes = list(linewidth =2)))
+       subtitle = "Counties in Massachusetts")
 
@@
 
-pal
+# switch labels to version without line breaks for clearer legend
+names(pal) <- convert_labels(labels(pal), to_newlines=FALSE)
 
@@
              fill=vlabel),
-             alpha = .5) +
+             alpha = .7) +
   facet_wrap(~name,
@@
 
-cat("Maximum ratio, on average, occured: ")
-county %>% 
+max_ratio_date <- county %>% 
   filter(version=="v7") %>% 
   group_by(biweek) %>% 
-  summarize(m = mean(exp_cases_median/positive),dates=paste(range(date),collapse=",")) %>%
+  summarize(m = mean(exp_cases_median/positive),
+            dates=paste(range(date),collapse=" to ")) %>%
   slice_max(n=1, m)
+
+cat("Maximum ratio, on average, occured on interval: ",
+    max_ratio_date$dates, "on biweek", 
+    max_ratio_date$biweek)
 

### 2023-07-17 reports/state_figures.Rmd
consistent coloring

@@
 save_figures <- TRUE
+
+pal <- c("red", "#10BAC5", "#1B10C5", "#EFB719", "#900C3F")
+
+names(pal) <- c("Covidestim",
+                '$P(S_1|untested)$ Centered at Survey Value',
+                '$P(S_1|untested)$ and $\\beta$ Centered at Survey Values',
+               "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
+               "$\\beta$ Centered at Survey Value"
+               )
 
@@
   state_res <- state_res %>%
-  filter(version %in% c("v3", "v5", "v6", "v7"))
+  filter(version %in% c("v3", "v5", "v6", "v7"))  %>%
+    mutate(vlabel=case_when(
+      version == "v7" ~  "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
+      version == "v5" ~ "$\\beta$ Centered at Survey Value",
+      version =="v6" ~  "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+      version == "v3" ~ "$P(S_1|untested)$ Centered at Survey Value"
+    ))
 
@@
 subset <- state_res %>%
-  filter(vlabel ==  "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$")
+  filter(vlabel ==  "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$")
 
-pal <- c("#247C90", "red")
-
-names(pal) <- c(as.character(unique(subset$vlabel)), "Covidestim")
+# pal <- c("#247C90", "red")
+# 
+# names(pal) <- c(as.character(unique(subset$vlabel)), "Covidestim")
 
@@
 
-comp_state_pal <- c("#247C90", "red")
+# comp_state_pal <- c("#247C90", "red")
+# 
+# names(comp_state_pal) <- c('Probabilistic Bias Analysis', "Covidestim")
+# 
+# 
+# vlab_for_pal <- unique(subset$vlabel) %>% as.character()
 
+color_for_bias <- pal[[which(names(pal)=="$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$")]]
+
+comp_state_pal<- c(color_for_bias,"red")
 names(comp_state_pal) <- c('Probabilistic Bias Analysis', "Covidestim")
@@
              ymin = exp_cases_lb,
-             ymax = exp_cases_ub),
-             fill = "#3E3D3D",
+             ymax = exp_cases_ub,
+             fill =vlabel),
+           #  fill = "#3E3D3D",
              alpha = .5,
@@
  # scale_color_brewer(palette="Accent") 
-  scale_color_manual(values=varpal)
+  scale_color_manual(values=varpal) +
+  scale_fill_manual(values=pal)
 

### 2023-07-17 writing/supplement.Rmd
consistent coloring

@@
 
+```{r sim-intervals-faceted-county, fig.width=15, fig.height=16,  fig.cap = "Simulation intervals for counties in Massachusetts.
+Each interval corresponds to a 95\\% simulation interval for the total number of estimated infections for that county in that two-week time interval.
+The columns represent different implementations of the probabilistic bias analysis: in the first through the third columns, the priors vary by state and date, while for the fourth, they are the same for all states and time intervals considered.
+For the implementation in the first column, we center the distribution of $\\beta$ at the ratio of the screening test positivity to the overall test positivity from the survey, and we center the distribution of $\\Pr(S_1|\\text{untested})$ at the percentage of the population experiencing COVID-19-like illness from the survey for each two-week interval.
+The second column centers only $\\beta$ at the aformentioned value, and the third column only centers $\\Pr(S_1|\\text{untested})$ at the aforementioned value.
+The fourth column corresponds to the implementation where we specify priors that are the same for all dates.
+The implementation that centers both $\\Pr(S_1|\\text{untested})$ and $\\beta$ at the survey values is consistently the highest among the implementations, followed by the implementation that centers only $\\beta$ at the survey value, followed then by the implementation only centering $\\Pr(S_1|\\text{untested})$ at the survey value, and then the lowest among the implementations is that where the priors do not vary by date."}
+
+include_graphics(here('figures/simulation-intervals-faceted-version-county.pdf'))
+
+```
+
 # Variant Data

### 2023-07-17 reports/compare_models.Rmd
comparing pre and post omicron covidestim models

---
title: "Comparing Covidestim Models"
author: "Quinn White"
date: "`r Sys.Date()`"
output:
  rmdformats::readthedown:
    df_print: paged
    code_folding: show
    css: !expr here::here('css/template.css')
  html_document:
    df_print: paged
    toc-depth: 4
knit: (function(input, ...) {
    rmarkdown::render(
      input,
      output_dir = "../output_html")})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Relevant sections of `covidestim/inst/stan/stan_program_default.stan` from before the adjustments for Omicron (the file [here](https://github.com/covidestim/covidestim/blob/6151c275fdfa085f750fb20d2e63a5e784c76b15/inst/stan/stan_program_default.stan)) and the most recent model version (the file [here](https://github.com/covidestim/covidestim/blob/master/inst/stan/stan_program_default.stan)).

# Handling reinfections differently

## Previous Version

Subtracting infections from population size.

```{stan output.var="", eval=FALSE}

 pop_uninf = pop_size;
  for(i in 1:N_days_tot) {
    if(i==1){
      logRt[i] = logRt0[i];
    } else{
      logRt[i] = logRt0[i] + log(pop_uninf/pop_size);
    }
    deriv1_log_new_inf[i] = logRt[i]/serial_i;
    log_new_inf[i] = sum(deriv1_log_new_inf[1:i]) + log_new_inf_0;
    new_inf[i] = (1-exp(-exp(log_new_inf[i])/pop_uninf)) * pop_uninf;
    pop_uninf -= new_inf[i];
    if (pop_uninf < 1) {
      // print("WARNING pop_uninf preliminary value was ", pop_uninf);
      pop_uninf = 1;
    }
  }

```

## New Version

Incorporating reinfections

```{stan output.var="", eval=FALSE}

  // NEW INCIDENT CASES

  // modeled with a spline
  //     spl_par_rt[2:N_spl_par_rt] = spl_par_rt0; 
  // spl_par_rt[1] = (spl_basis_rt[2+N_days_before,2:N_spl_par_rt]-
  //                  spl_basis_rt[1+N_days_before,2:N_spl_par_rt]) * spl_par_rt0 / 
  //                 (spl_basis_rt[1+N_days_before,1]-spl_basis_rt[2+N_days_before,1]);
  logRt0 = spl_basis_rt * spl_par_rt;
  ever_inf = pop_size * cum_p_inf_init; // initial ever infected population
  population_protection_init[1] = pop_size * start_p_imm; // initial (waning) protection from vax/inf/boost
  if(N_weeks_before > 0){
    // if there are before weeks modeled, the initial effective protection
    // equals the population that will never be infected by the start date of the data
    // i.e., the date for which the cum_p_inf_init and start_p_imm are reported.
    effective_protection_prvl[1] = pop_size * (1 - cum_p_inf_init);
  } else {
    // if there are no before weeks modeled, the initial effecive protection is 
    // the complete immune population
    effective_protection_prvl[1] = pop_size * start_p_imm;
  }
  susceptible_prvl[1] = pop_size - effective_protection_prvl[1];
  // effective_protection_inf_prvl[1] = pop_size * cum_p_inf_init;  //
  // effective protection from infections; for the first timepoint include everyone
  // with a historic infection;

  // for(i in 1:N_days_tot) {
  for(i in 1:N_weeks_tot) {
   if(i > 1){
       susceptible_prvl[i] = pop_size - effective_protection_prvl[i-1];
   }

    if (susceptible_prvl[i] < 1) {
      // print("WARNING num_uninf preliminary value was ", num_uninf);
      susceptible_prvl[i] = 1;
    }
    logRt[i] = logRt0[i] + log(susceptible_prvl[i]/pop_size);

    deriv1_log_infections[i] = logRt[i]/serial_i;

    log_infections[i] = sum(deriv1_log_infections[1:i]) + log_infections_0;

    infections[i] = exp(log_infections[i]);

    // prot_boost[i] = sum(obs_boost[1:i] * 0.8);
    // if(i > N_days_before){
    // wane_boost[i] = sum(obs_boost[1:i-N_days_before] .* (.35 * exp(-.008 * idx3[N_days_tot-(i-N_days_before)+1:N_days_tot]) + .45) );
    if(i > N_weeks_before){
      // for the observed data weeks, population protection is calculated for the 
      // waning of the initial immunity since the observed data started;
      // waning of booster immunity since the observed data started;
      // waning of the modeled infections since the observed data started 
      // (note that any infections prior to this date are implicitly included in the start_p_imm)
    // wane_boost[i] = sum(obs_boost[1:i-N_weeks_before] .* (.35 * exp(-.008 * idx3[N_weeks_tot-(i-N_weeks_before)+1:N_weeks_tot]*7) + .45) );
    population_protection_init[i] = pop_size * start_p_imm * exp(-.008 * idx3[N_weeks_tot-i+N_weeks_before+1]*7) ; // increasse from 75% susceptible to 90% susceptible (10% remains rpotected)
    population_protection_boost[i] = sum(obs_boost[1:i-N_weeks_before]*.8 .* (exp(-.008 * idx3[N_weeks_tot-i+N_weeks_before+1:N_weeks_tot]*7)) );
    population_protection_inf[i] = sum(infections[N_weeks_before +1:i] .* ( exp(-.008 * idx3[N_weeks_tot-i+N_weeks_before+1:N_weeks_tot] * 7)));
    } else {
      // for the before period, population protected is only the sum of the initial population protected (never infected prior to start date)
      // and those infected before the start of the data.
population_protection_init[i] = pop_size * (1 - cum_p_inf_init);
population_protection_boost[i] = 0;
population_protection_inf[i] = sum(infections[1:i] .* ( exp(-.008 * idx3[N_weeks_before-i+1:N_weeks_before] * 7)));
    }
```

# Fitting with Hospitalizations Instead of Deaths

## Previous Version (Deaths)

Deaths in the Likelihood:

```{stan  output.var="", eval= FALSE}

///// LIKELIHOOD
// Before data
  if(pre_period_zero==1){
    if(N_days_before>0){
      tmp_sum_cas_pre = sum(occur_cas[1:N_days_before]);
      tmp_sum_die_pre = sum(occur_die[1:N_days_before]);

      if (tmp_sum_cas_pre <= 0) // Account for floating point error
        tmp_sum_cas_pre = 0.0000000001;
      if (tmp_sum_die_pre <= 0) // Account for floating point error
        tmp_sum_die_pre = 0.0000000001;

      target += neg_binomial_2_lpmf( 0 | tmp_sum_cas_pre, phi_cas);
      target += neg_binomial_2_lpmf( 0 | tmp_sum_die_pre, phi_die);
    }
  }

// Observed data
  if (cas_yes == 1) {
    tmp_obs_cas = obs_cas[1];
    tmp_occur_cas = occur_cas[1 + N_days_before];

    for(i in 1:N_days) {

      if (tmp_occur_cas <= 0) // Account for floating point error
        tmp_occur_cas = 0.000000001;
      //
      //don't add to target if no observations are present
      if(i > lastCaseDate) {
        break;
      }

      // Don't add to `target` unless we have `N_days_av` of data accumulated
      // in `tmp_occur_cas`
      if (i >= N_days_av)
        target += neg_binomial_2_lpmf(tmp_obs_cas | tmp_occur_cas, phi_cas)/
          N_days_av;

      if(i>nda0){
        tmp_obs_cas   -= obs_cas[i - nda0];
        tmp_occur_cas -= occur_cas[i + N_days_before - nda0];
      }
      if(i<N_days){
        tmp_obs_cas   += obs_cas[i + 1];
        tmp_occur_cas += occur_cas[i + N_days_before + 1];
      }
    }
  }

  if (die_yes == 1){

    tmp_obs_die = obs_die[1];
    tmp_occur_die = occur_die[1 + N_days_before];

    for(i in 1:N_days) {

      if (tmp_occur_die <= 0) // Account for floating point error
        tmp_occur_die = 0.000000001;

      //don't add to target if no observations are present
      if(i > lastDeathDate) {
        break;
      }

      // Don't add to `target` unless we have `N_days_av` of data accumulated
      // in `tmp_occur_die`
      if (i >= N_days_av)
        target += neg_binomial_2_lpmf(tmp_obs_die | tmp_occur_die, phi_die)/
          N_days_av;

      if(i>nda0){  
        tmp_obs_die   -= obs_die[i - nda0];
        tmp_occur_die -= occur_die[i + N_days_before - nda0];
      }
      if(i<N_days){
        tmp_obs_die   += obs_die[i + 1];
        tmp_occur_die += occur_die[i + N_days_before + 1];
      }
    }
  }
    }

```

## New Version (Hospitalizations)

Fitted hospitalizations are taken to be the number of severe diagnoses -- relevant lines:

```{stan  output.var="", eval=FALSE}
// DIAGNOSIS // 
  // rate ratio of diagnosis at asymptomatic vs symptomatic, symptomatic vs severe
  rr_diag_sym_vs_sev = inv_logit(spl_basis_dx * logit(spl_par_sym_dx));
  p_diag_if_sym = p_diag_if_sev * rr_diag_sym_vs_sev;
  diagnoses_severe = p_diag_if_sev * conv1d(severe - dx_sym_sev, sev_diag_delay_rv);
fitted_hospitalizations = diagnoses_severe;

```

Hospitalizations in the Likelihood:

```{stan  output.var="", eval=FALSE}

// LIKELIHOOD
  // Before data
  if(pre_period_zero==1){
    // if(N_days_before>0){
    if(N_weeks_before>0){

      // if (sum(fitted_cases[1:N_days_before]) < 0)
      //   reject("`sum(fitted_cases[1:N_days_before])` had a negative value");
      // 
      // if (sum(fitted_deaths[1:N_days_before]) < 0)
      //   reject("`sum(fitted_deaths[1:N_days_before])` had a negative value");
      if (sum(fitted_cases[1:N_weeks_before]) < 0)
        reject("`sum(fitted_cases[1:N_weeks_before])` had a negative value");

      if (sum(fitted_hospitalizations[1:N_weeks_before]) < 0)
        reject("`sum(fitted_hospitalizations[1:N_weeks_before])` had a negative value");

      // target += neg_binomial_2_lpmf( 0 | sum(fitted_cases[1:N_days_before]), phi_cas);
      // target += neg_binomial_2_lpmf( 0 | sum(fitted_deaths[1:N_days_before]), phi_die);
      target += neg_binomial_2_lpmf( 0 | sum(fitted_cases[1:N_weeks_before]), phi_cas);
      target += neg_binomial_2_lpmf( 0 | sum(fitted_hospitalizations[1:N_weeks_before+4]), phi_hosp);
    }
  }

  if (min(fitted_cases) < 0)
    reject("`fitted_cases` had a negative value");

  if (min(fitted_hospitalizations) < 0)
    reject("`fitted_hospitalizations` had a negative value");

  // LIKELIHOOD
  // During data
  // target += neg_binomial_2_lpmf(
  //   // `obs_cas` from the first observed day to the last death date
  //   obs_cas_mvs[N_days_av:lastCaseDate] |
  //     // `fitted_cases` from the first observed day (`N_days_before+1`) to the
  //     // last death date
  //     fitted_cases_mvs[N_days_before+N_days_av : N_days_before+lastCaseDate],
  //   phi_cas
  // ) ;// Optional, but likely unncessesary: / N_days_av;
  // 
  // target += neg_binomial_2_lpmf(
  //   // `obs_die` from the first observed day to the last death date
  //   obs_die[N_days_av:lastDeathDate] |
  //     // `fitted_deaths` from the first observed day (`N_days_before+1`) to the
  //     // last death date
  //     fitted_deaths[N_days_before+N_days_av : N_days_before+lastDeathDate],
  //   phi_die
  // ); // optional, but likelie unnecessary: / N_days_av;
  target += neg_binomial_2_lpmf(
    // `obs_cas` from the first observed day to the last death date
    obs_cas_mvs[1:lastCaseWeek] |
      // `fitted_cases` from the first observed day (`N_days_before+1`) to the
      // last death date
      fitted_cases_mvs[N_weeks_before+1 : N_weeks_before+lastCaseWeek],
    phi_cas
  ) ;// Optional, but likely unncessesary: / N_days_av;

// for hospitalizations: use lastCaseWeek as last observation 
// this is still to adjust in the future and add a lastHospiWeek variable
  target += neg_binomial_2_lpmf(
    // `obs_die` from the first observed day to the last death date
    obs_hosp_mvs[5:lastHospWeek] |
      // `fitted_deaths` from the first observed day (`N_days_before+1`) to the
      // last death date
      fitted_hospitalizations_mvs[N_weeks_before+5 : N_weeks_before+lastHospWeek],
    phi_hosp
  ); // optional, but likelie unnecessary: / N_days_av;
}
```

# How the IFR is Used

## Previous Version

In the earlier implementation, those who die is calculated as those who die among the symptomatic and severe: `new_die_dx = dx_sym_die + dx_sev_die;`, where `dx_sym_die` and `dx_sev_die` are informed by `p_die_if_sympt` and `p_die_if_sevt` respectively:

```{stan  output.var="", eval=FALSE}

for(i in 1:N_days_tot){
  p_die_if_sevt[i] = p_die_if_sevt[i] * pow(ifr_vac_adj[i], prob_vac[1]);
  p_sev_if_symt[i] = p_sev_if_sym * pow(ifr_vac_adj[i], prob_vac[2]);
  p_sym_if_inft[i] = p_sym_if_inf * pow(ifr_vac_adj[i], prob_vac[3]);
  }

```

and `dx_sym_die` is calculated as

```{stan  output.var="", eval=FALSE}
  for(i in 1:N_days_tot) {
    dx_sym_die[i] = dot_product(dx_sym_sev[idx1[i]:i],
      sev_prg_delay_rv[idx2[i]:Max_delay]) * p_die_if_sevt[i];
  }
```

It is then `new_die_dx` that is used, after convolution with the reporting delay (to produce `occur_die`), that is used in the likelihood:

```{stan  output.var="", eval=FALSE}

 target += neg_binomial_2_lpmf(tmp_obs_die | tmp_occur_die, phi_die)/
          N_days_av;

```

where  `neg_binomial_2_lpmf(ints n | reals mu, reals phi)` is the negative binomial probability mass of $n$ given location $\mu$ and precision $\phi$.

## New Version

Here, the likelihood instead considers hospitalizations rather than deaths:

```{stan  output.var="", eval=FALSE}
 target += neg_binomial_2_lpmf(
    // `obs_die` from the first observed day to the last death date
    obs_hosp_mvs[5:lastHospWeek] |
      // `fitted_deaths` from the first observed day (`N_days_before+1`) to the
      // last death date
      fitted_hospitalizations_mvs[N_weeks_before+5 : N_weeks_before+lastHospWeek],
    phi_hosp
  )
```

where `fitted_hospitalizations` is the number of diagnoses among the severe (`_mv` indicates a moving sum), and relies on `p_diag_if_sym`:

```{stan  output.var="", eval=FALSE}
// DIAGNOSIS // 
  // rate ratio of diagnosis at asymptomatic vs symptomatic, symptomatic vs severe
  rr_diag_sym_vs_sev = inv_logit(spl_basis_dx * logit(spl_par_sym_dx));
  p_diag_if_sym = p_diag_if_sev * rr_diag_sym_vs_sev;
  diagnoses_severe = p_diag_if_sev * conv1d(severe - dx_sym_sev, sev_diag_delay_rv);
fitted_hospitalizations = diagnoses_severe;

```

The probability `p_sev_if_symt[i] = p_sev_if_sym  * pow(ifr_vac_adj[i], prob_vac[2]);` is used to calculate the number of severe cases; so the fitted hospitalizations in this sense is related to the IFR since `severe = p_sev_if_symt .*conv1d(symptomatic, sym_prg_delay_rv);` (and diagnoses among the severe is then calculated from this value). The prior for `prob_vac` is the same for both implementations.

# Summary

* Reinfections are handled differently in that the earlier implementation did not allow reinfections and as such infections were subtracted from the population at each time step, while the newer implementation does allow for reinfections by modeling the protection from immunity to the virus.
* The earlier model used a likelihood on cases and on deaths, while the newer implementation includes a likelihood on cases and hospitalizations.
* Both use the IFR in the estimation of fitted hospitalizations or fitted deaths, since both use the IFR in the calculation of the probabilities of transitioning between states:

```{stan  output.var="", eval=FALSE}
 for( i in 1:N_weeks_tot){
  p_die_if_sevt[i]     = p_die_if_sevt[i]   .* pow(ifr_vac_adj[i], prob_vac[1]);
  p_sev_if_symt[i]     = p_sev_if_sym        * pow(ifr_vac_adj[i], prob_vac[2]);
  p_sym_if_inft[i]     = p_sym_if_inf        * pow(ifr_vac_adj[i], prob_vac[3]);
  }
```

* Newer model is at the weekly time step instead of daily.

### 2023-07-17 reports/state_figures.Rmd
include variant data from across the entirety of 2021

@@
  # filter(state %in% sample(corrected$state,5)) %>%
-  filter(biweek >=6) %>%
+ # filter(biweek >=6) %>%
   group_by(biweek) %>%

### 2023-07-17 R/01-data_cleaning.R

@@ -611,7 +611,7 @@ get_biweekly_wastewater <- function() {
 
 get_variant_proportions <- function() {
   
-  variant <- "https://data.cdc.gov/resource/jr58-6ysp.json?$limit=50000&$where=week_ending between '2021-02-18T00:00:00.000' and '2022-03-01T00:00:00.000'&usa_or_hhsregion=USA"
+  variant <- "https://data.cdc.gov/resource/jr58-6ysp.json?$limit=50000&$where=week_ending between '2021-01-01T00:00:00.000' and '2022-03-01T00:00:00.000'&usa_or_hhsregion=USA"
   variant <- httr::GET(URLencode(variant))
   
   
@@ -641,15 +641,21 @@ get_variant_proportions <- function() {
       grepl("B[.]1[.]617[.]2", variant) ~ "Delta",
       variant %in% c("B.1.427" ,"B.1.429") ~ "Epsilon",
       variant == "B.1.525" ~ "Eta",
-      variant == "Other" ~ "Other"
+      TRUE ~ "Other"
     )) %>%
     mutate(share = as.numeric(share),
            creation_date = ymd(substr(creation_date,1,10))) %>%
-    filter(modeltype == "weighted" & time_interval=="weekly") %>%
+    filter(modeltype == "weighted") %>% 
+    mutate(week_val = ifelse(time_interval=="weekly", 1, 2)) %>%
     group_by(variant, week_end_date, variant_category, time_interval) %>% 
+    # take most recently created
     slice_max(n=1, order_by=creation_date) %>%
+    group_by(variant, week_end_date) %>%
+    slice_min(n=1, week_val ) %>%
+    # take weekly if available, otherwise biweekly 
     group_by(variant_category, week_end_date, time_interval) %>%
     summarize(share =sum(share, na.rm=TRUE))
+  
   
   
   variant <- variant %>%
@@ -658,8 +664,6 @@ get_variant_proportions <- function() {
     # week beginning rather than week end
     mutate(week = week_end_date - days(7)) %>%
     filter(!is.na(variant_category)) 
-  
-
   
   return(variant)
   

### 2023-07-18 reports/county_figures.Rmd
update manuscript & supplement

@@
 names(pal) <- c("Covidestim",
-                '$P(S_1|untested)$ Centered at Survey Value',
-                '$P(S_1|untested)$ and $\\beta$ Centered at Survey Values',
+                '$Pr(S_1|untested)$ Centered at Survey Value',
+                '$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values',
                "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
@@
          fixed=TRUE) %>%
-    gsub("$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
-         "$\\overset{P(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}", 
+    gsub("$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+         "$\\overset{Pr(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}$", 
          .,
@@
          fixed=TRUE) %>%
-    gsub("$\\overset{P(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}",
-         "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+    gsub("$\\overset{Pr(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}$",
+         "$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values",
          .,
@@
     "$\\beta$ Centered at Survey Value",
-    "$P(S_1|untested)$ Centered at Survey Value",
-    "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+    "$Pr(S_1|untested)$ Centered at Survey Value",
+    "$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values",
         "$\\beta$ Centered at Survey Value (Spline Smoothed)",
-         "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values ($\\beta$ Spline Smoothed)",
+         "$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values ($\\beta$ Spline Smoothed)",
         "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$"
@@
         "$\\beta$ Centered at Survey Value",
-         "$P(S_1|untested)$ Centered at Survey Value",
-        "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+         "$Pr(S_1|untested)$ Centered at Survey Value",
+        "$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values",
         "$\\beta$ Centered at Survey Value (Spline Smoothed)",
-         "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values ($\\beta$ Spline Smoothed)",
+         "$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values ($\\beta$ Spline Smoothed)",
         "$\\,$CTIS Priors that Do Not Vary by State or Date$\\,$"
@@
       version == "v5" ~ "$\\beta$ Centered at Survey Value",
-      version =="v6" ~  "$\\overset{P(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}",
-      version == "v3" ~ "$P(S_1|untested)$ Centered at Survey Value",
+      version =="v6" ~  "$\\overset{Pr(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}$",
+      version == "v3" ~ "$Pr(S_1|untested)$ Centered at Survey Value",
     )) %>%
@@
         "$\\overset{Priors\\,Do\\,Not\\,Vary}{\\,by\\,State\\,and\\,Date}$", 
-         "$P(S_1|untested)$ Centered at Survey Value",
+         "$Pr(S_1|untested)$ Centered at Survey Value",
          "$\\beta$ Centered at Survey Value",
-        "$\\overset{P(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}"))) %>%
+        "$\\overset{Pr(S_1|untested)\\,and\\,beta}{Centered\\,at\\,Survey\\,Values}$"))) %>%
     left_join(county_names) %>%
@@
       version == "v5" ~ "$\\beta$ Centered at Survey Value",
-      version =="v6" ~  "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
-      version == "v3" ~ "$P(S_1|untested)$ Centered at Survey Value"
+      version =="v6" ~  "$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+      version == "v3" ~ "$Pr(S_1|untested)$ Centered at Survey Value"
     ))
@@
 names(pal) <- c(#"$observed$", 
-                '$P(S_1|untested)$ Centered at Survey Value',
-                "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+                '$Pr(S_1|untested)$ Centered at Survey Value',
+                "$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values",
                "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
@@
 names(pal) <- c(#"$observed$", 
-                '$P(S_1|untested)$ Centered at Survey Value',
-                '$P(S_1|untested)$ and $\\beta$ Centered at Survey Values',
+                '$Pr(S_1|untested)$ Centered at Survey Value',
+                '$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values',
                "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
@@
 ma_concordance_by_county %>%
-  filter(vlabel=="$P(S_1|untested)$ Centered at Survey Value") %>%
+  filter(vlabel=="$Pr(S_1|untested)$ Centered at Survey Value") %>%
   group_by(name) %>%
@@
        title = "Massachusetts: Proportion Where Covidestim Median\nWas Within, Above, or Below the Median",
-       subtitle = TeX("For Implementation with $P(S_1|untested)$ Centered at Survey Value")) +
+       subtitle = TeX("For Implementation with $Pr(S_1|untested)$ Centered at Survey Value")) +
   theme_c() +
@@
 county %>% 
+  # only include dates where there were estimates for all implementations 
+  filter(biweek >=6) %>%
   mutate(name=gsub("Nantucket and Dukes", "Nantucket\nand Dukes", name)) %>%
@@
 
+```{r,results='asis'}
+
+version_with_max <- function(variable) {
+   county %>%
+    mutate(vlabel = convert_labels(vlabel, to_newlines=FALSE)) %>%
+    select(-date) %>%
+    distinct() %>% 
+    # only include biweeks where all implementations have estimates
+    filter(biweek >=6) %>%
+    group_by(biweek,fips) %>%
+    slice_max(n=1,order_by=!!sym(variable), with_ties=FALSE) %>%
+    group_by(vlabel) %>%
+    summarize(n=n()) %>%
+    ungroup() %>%
+    mutate(total=sum(n),
+           prop=n/total)
+}
+
+version_with_min <- function(variable) {
+   county %>%
+    mutate(vlabel = convert_labels(vlabel, to_newlines=FALSE)) %>%
+    select(-date) %>%
+    distinct() %>% 
+    # only include biweeks where all implementations have estimates
+    filter(biweek >=6) %>%
+    group_by(biweek,  fips) %>%
+    slice_min(n=1,order_by=!!sym(variable), with_ties=FALSE) %>%
+    group_by(vlabel) %>%
+    summarize(n=n()) %>%
+    ungroup() %>%
+    mutate(total=sum(n),
+           prop=n/total) 
+}
+
+# 
+# county %>%
+#     select(-date) %>%
+#     distinct() %>% 
+#     # only include biweeks where all implementations have estimates
+#     filter(biweek >=6) %>%
+#     group_by(biweek,fips) %>%
+#    # slice_min(n=1,order_by=!!sym(variable), with_ties=FALSE)  %>%
+#   select(exp_cases_lb, biweek, fips,version) %>%
+#   pivot_wider(values_from = exp_cases_lb, names_from=version ) %>%
+#   mutate(diff=v3-v7) %>%
+#   pull(diff) %>%
+#   mean()
+
+version_with_max_median <- version_with_max("exp_cases_median")
+version_with_min_median <- version_with_min("exp_cases_median")
+
+version_with_max_lb <- version_with_max("exp_cases_lb")
+version_with_min_lb <- version_with_min("exp_cases_lb")
+
+version_with_max_ub <- version_with_max("exp_cases_ub")
+version_with_min_ub <- version_with_min("exp_cases_ub")
+
+#############
+# MEDIAN
+#############
+cat("* The implementatation that most frequently had the maximum median cases across the two-week intervals was",
+    as.character(version_with_max_median$vlabel[which.max(version_with_max_median$prop)]), 
+    "where the median of", 
+    version_with_max_median$prop[which.max(version_with_max_median$prop)] *100 %>% round(3),
+    "percent of simulation intervals was highest.")
+
+cat("* The implementatation that most frequently had the minimum median cases across the two-week intervals was",
+    as.character(version_with_min_median$vlabel[which.max(version_with_min_median$prop)]), 
+    "where the median of", 
+    version_with_min_median$prop[which.max(version_with_min_median$prop)] *100 %>% round(3),
+    "percent of simulation intervals was lowest.")
+
+#################
+# LOWER QUANTILE
+#################
+
+cat("* The implementatation that most frequently had the maximum 2.7% percentile in infections across the two-week intervals was",
+    as.character(version_with_max_lb$vlabel[which.max(version_with_max_lb$prop)]), 
+    "where the median of", 
+    version_with_max_lb$prop[which.max(version_with_max_lb$prop)] *100 %>% round(3),
+    "percent of simulation intervals was highest.")
+
+cat("* The implementatation that most frequently had the minimum 97.5% percentile in infections across the two-week intervals was",
+    as.character(version_with_min_lb$vlabel[which.max(version_with_min_lb$prop)]), 
+    "where the median of", 
+    version_with_min_lb$prop[which.max(version_with_min_lb$prop)] *100 %>% round(3),
+    "percent of simulation intervals was lowest.")
+
+```
+
 ```{r, fig.height=8, fig.width=14}
@@
       version == "v5" ~ "$\\beta$ Centered at Survey Value",
-      version =="v6" ~  "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values",
-      version == "v3" ~ "$P(S_1|untested)$ Centered at Survey Value",
+      version =="v6" ~  "$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+      version == "v3" ~ "$Pr(S_1|untested)$ Centered at Survey Value",
     )) %>%
@@
         "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
-         "$P(S_1|untested)$ Centered at Survey Value",
+         "$Pr(S_1|untested)$ Centered at Survey Value",
          "$\\beta$ Centered at Survey Value",
-       "$P(S_1|untested)$ and $\\beta$ Centered at Survey Values"))) %>%
+       "$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values"))) %>%
     mutate(vlabel=fct_reorder(vlabel,as.numeric(vlabel),.desc=TRUE)) %>%
@@
 
+```{r}
+
+# only two implementations included
+
+versions_to_include <- c(version_with_max_median$vlabel[which.max(version_with_max_median$prop)],
+                         version_with_min_median$vlabel[which.max(version_with_min_median$prop)]) %>%
+  unlist() %>% as.character()
+
+versions_to_include
+
+# switch labels to version without line breaks for clearer legend
+# names(pal) <- convert_labels(labels(pal), to_newlines=FALSE)
+names(pal)
+
+county_relab <- county %>%
+  mutate(vlabel=convert_labels(vlabel, to_newlines=FALSE)) %>%
+  filter(vlabel %in% versions_to_include) %>%
+  mutate(vlabel=case_when(
+      version == "v7" ~  "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
+      version == "v5" ~ "$\\beta$ Centered at Survey Value",
+      version =="v6" ~  "$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values",
+      version == "v3" ~ "$Pr(S_1|untested)$ Centered at Survey Value",
+    )) %>%
+    mutate(vlabel =factor(vlabel,levels= c(
+        "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$", 
+         "$Pr(S_1|untested)$ Centered at Survey Value",
+         "$\\beta$ Centered at Survey Value",
+       "$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values"))) %>%
+    mutate(vlabel=fct_reorder(vlabel,as.numeric(vlabel),.desc=TRUE)) %>%
+  filter(biweek>=6)
+
+county_relab$vlabel %>% unique()
+
+county_relab %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub,
+             fill=vlabel),
+             alpha = .7) +
+  facet_wrap(~name,
+             scales="free_y") +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(
+          legend.position = "top",
+          legend.text=element_text(hjust=0)
+          ) +
+  scale_fill_manual(values = pal,
+                    labels=TeX(unique(as.character(county_relab$vlabel))))  +
+  labs(x="",
+       fill="",
+       y = "Estimated Infections",
+       title = "Simulation Intervals Over Time",
+       subtitle = "Counties in Massachusetts") +
+  guides(fill=guide_legend(nrow=2))
+
+```
+
+```{r,fig.height=8, fig.width=14, eval=save_figures}
+
+ggsave(here('figures/simulation-intervals-color-version-only-2-county.pdf'))
+
+```
+
 # Heatmap
@@
   vlabel = factor(
-    c(  "$P(S_1|$untested) Centered at Survey Value"  ,
+    c(  "$Pr(S_1|$untested) Centered at Survey Value"  ,
     "$\\beta$ Centered at Survey Value",
-    "$P(S_1|$untested) and $\\beta$ Centered at Survey Values",
+    "$Pr(S_1|$untested) and $\\beta$ Centered at Survey Values",
      "Priors Do Not Vary by State or Date"
     ),
-    levels =   c(  "$P(S_1|$untested) Centered at Survey Value"  ,
+    levels =   c(  "$Pr(S_1|$untested) Centered at Survey Value"  ,
     "$\\beta$ Centered at Survey Value",
-    "$P(S_1|$untested) and $\\beta$ Centered at Survey Values",
+    "$Pr(S_1|$untested) and $\\beta$ Centered at Survey Values",
      "Priors Do Not Vary by State or Date"

### 2023-07-18 reports/state_figures.Rmd
update manuscript & supplement

@@
 
+dates %>% 
+  group_by(biweek) %>%
+  summarize(mindate=min(date),maxdate=max(date))
+
 state_res <- state_res %>%

### 2023-07-18 writing/manuscript.Rmd
update manuscript & supplement

@@
 
-As a result of these dramatic changes, Chitwood *et al.*@chitwood2022 made substantial changes to Covidestim, as noted in the model change log.
+As a result of these dramatic changes, Chitwood *et al.* @chitwood2022 made substantial changes to Covidestim, as noted in the model change log.
 Because the variant causes much milder infections, the infection fatality for Omicron infections is lower than previous variants, and death counts were much lower relative to the number of infections.
@@
 
-In Figure (CHOOSE BETWEEN VERSION BY COLOR AND FACETED), we present simulation intervals for each implementation of probabilistic bias analysis.
+In Figure \@ref(fig:sim-intervals-color-county), we present simulation intervals for two of the four implementations of probabilistic bias analysis.
+For most time intervals, the other two implementations fall between those shown in the figure.
 
 In the first three implementations, the prior distributions specified for the bias parameters vary by state and date.
-In particular, we use survey data from the COVID-19 Trends and Impact Survey [@salomon2021a] to inform the prior distributions for two of the bias parameters: the probability of having symptoms among the untested population ($\Pr(S_1|\text{untested})$), and the parameter $\beta$, which represents the ratio of the test positivity among members of the untested population who have no symptoms to the overall test positivity.
+In particular, we use survey data from the COVID-19 Trends and Impact Survey [@salomon2021a] to inform the prior distributions for one to two of the bias parameters: the probability of having symptoms among the untested population ($\Pr(S_1|\text{untested})$), and the parameter $\beta$, which represents the ratio of the test positivity among members of the untested population who have no symptoms to the overall test positivity.
 More detail on these implementations is provided in the [Methods](#Methods) section.
-In the last implementation, the prior distributions are informed by the distributions from the survey data across all dates and states, but the same prior distributions are used for all states and time intervals considered.
-While there is substantial overlap among all of the versions, we see the implementation allowing $\beta$ and $\Pr(S_1|\text{untested})$ to vary by state and date tends to be the highest, followed by that only allowing $\beta$ to vary by state and date,  followed by that only allowing $\Pr(S_1|\text{untested})$ to vary by state and date, and lastly the implementation that does not vary the priors by state or date tends to produce the lowest estimates.
+In another, more simple, implementation, which is the only implementation presented at the state-level, the prior distributions are informed by the distributions from the survey data across all dates and states, but the same prior distributions are used for all states and time intervals considered.
+
+While there is substantial overlap among all of the versions, the implementation allowing $\beta$ and $\Pr(S_1|\text{untested})$ to vary by state and date tends to be the highest, followed by that only allowing $\beta$ to vary by state and date,  followed by that only allowing $\Pr(S_1|\text{untested})$ to vary by state and date, and lastly the implementation that does not vary the priors by state or date tends to produce the lowest estimates (Supplementary Figure sim-intervals-faceted-county).
 
@@
 The fourth column corresponds to the implementation where we specify priors that are the same for all dates.
-The implementation that centers both $\\Pr(S_1|\\text{untested})$ and $\\beta$ at the survey values is consistently the highest among the implementations, followed by the implementation that centers only $\\beta$ at the survey value, followed then by the implementation only centering $\\Pr(S_1|\\text{untested})$ at the survey value, and then the lowest among the implementations is that where the priors do not vary by date."}
+The implementation that centers both $\\Pr(S_1|\\text{untested})$ and $\\beta$ at the survey values is consistently the highest among the implementations, followed by the implementation that centers only $\\beta$ at the survey value, followed then by the implementation only centering $\\Pr(S_1|\\text{untested})$ at the survey value, and then the lowest among the implementations is that where the priors do not vary by date.", include=FALSE}
 
@@
 ```{r sim-intervals-color-county, fig.width=15,  fig.cap = "Simulation intervals for counties in Massachusetts, colored by the implementation of probabilistic bias analysis.
-In the first three implementations, the priors vary by state and date, while for the fourth, they are the same for all states and time intervals considered.
+Only the two implementations that were consistently the highest and lowest among the implementations are included for clarity.
+In the   In the first three implementations, the priors vary by state and date, while for the fourth, they are the same for all states and time intervals considered.
 For the first implementation, we center the distribution of $\\beta$ at the ratio of the screening test positivity to the overall test positivity from the survey, and we center the distribution of $\\Pr(S_1|\\text{untested})$ at the percentage of the population experiencing COVID-19-like illness from the survey for each two-week interval.
@@
 
-include_graphics(here('figures/simulation-intervals-color-version-county.pdf'))
+include_graphics(here('figures/simulation-intervals-color-version-only-2-county.pdf'))
 
@@
 This corresponds to a similar time frame that we saw at the state-level in Figure \@ref(fig:ratio-estimated-to-observed-by-wave).
-Hamsphire County and Suffolk County were consistently among the lowest with regard to the ratio of estimated to observed infections, which may be explained by the screening testing occurring at universities in these counties.
+Hampsphire County and Suffolk County were consistently among the lowest with regard to the ratio of estimated to observed infections, which may be explained by the screening testing occurring at universities in these counties.
 

### 2023-07-18 reports/state_figures.Rmd
add dotted line at model change (state level)

@@
 
-dates %>% 
+# last biweek for first model is biweek 24
+last_biweek_old_model <- dates %>% 
   group_by(biweek) %>%
-  summarize(mindate=min(date),maxdate=max(date))
+  summarize(mindate=min(date),maxdate=max(date))  %>%
+  filter("2021-11-30" >= mindate & "2021-11-30" <= maxdate)
+
+cat("Last two week interval with previous model was biweek", last_biweek_old_model$biweek,
+    "\ndates", paste0("(", last_biweek_old_model$mindate, ", ",
+                    last_biweek_old_model$maxdate, ")"))
+
+end_old_model <- last_biweek_old_model$maxdate
 
@@
               alpha = .6) +
+  geom_vline(xintercept=end_old_model,
+             linetype=2,
+             color="darkred",
+             linewidth=.5) +
   facet_wrap(~name, ncol=4, scales = "free",

### 2023-07-23 reports/state_figures.Rmd
look into the ratio of cases to deaths pre and post omicron

@@
              show.legend=FALSE) +
-  # geom_ribbon(aes(x = date,
-  #                 fill = 'Covidestim',
-  #                 ymin = infections.lo,
-  #                 ymax = infections.hi),
-  #             alpha = .5) +
-  # facet_grid(~vlabel, scales = "free_y",
-  #             labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
   scale_x_date(date_breaks = "2 months", 
@@
   theme_c(legend.position="top") +
-  # theme_c(axis.text.x = element_text(size = 16),
-  #         axis.title.y = element_text(size=13),
-  #         plot.title=element_text(face = "bold",
-  #                                 hjust = .5, 
-  #                                 size = 24,
-  #                                 margin=margin(5,5,15,5,'pt')),
-  #         strip.background = element_rect(fill = "#393939"),
-  #         strip.text = element_text(color =  "white", size = 16),
-  #         strip.text.y = element_text(margin = margin(5,5,5,5,'pt'),
-  #                                     size = 22,
-  #                                     face="bold"),
-  #         strip.text.x =  element_text(margin = margin(5,3,5,3,'pt'),
-  #                                      face = "bold",
-  #                                      size = 22),
-  #         plot.subtitle = ggtext::element_markdown(size = 25,margin = margin(3,5,10,5,'pt'),
-  #                                      face = "italic"),
-  #         legend.position = "top",
-  #         legend.text =element_text(size = 17),
-  #         legend.title = element_text(size=18, face="bold")) +
   geom_line(aes(x=week, y=share*m,
@@
 ```
+
+# Comparing Overlap
+
+Weaknesses in this approach:
+
+- it doesn't distinguish between a very small interval being covered (i.e. when there are very few cases) and a larger interval being covered (i.e. when there are many cases)
+- it doesn't consider the width of the PB interval
+
+```{r, fig.height=7.5}
+
+state_overlaps <- subset %>%
+  select(-date) %>%
+  distinct() %>%
+  mutate(lower_overlap = map2_dbl(infections.lo, exp_cases_lb, ~max(.x,.y)),
+         upper_overlap = map2_dbl(infections.hi, exp_cases_ub, ~min(.x,.y)),
+         overlap = case_when(
+           upper_overlap-lower_overlap > 0 ~   upper_overlap-lower_overlap,
+           TRUE ~ 0
+         )) %>%
+  select(contains('overlap'), name, biweek, 
+         population, positive, total, exp_cases_lb,
+         exp_cases_ub, contains('infections')) %>%
+  mutate(overlap_not_norm = overlap,
+         overlap = overlap/population,
+         fraction_overlap = overlap_not_norm/(infections.hi-infections.lo))
+
+#############
+# BY STATE
+#############
+state_overlaps %>%
+  group_by(name) %>%
+  summarize(mean_overlap=mean(fraction_overlap)) %>%
+  ggplot(aes(y= fct_reorder(name, mean_overlap),
+             x =  mean_overlap)) +
+  geom_bar(stat="identity") +
+  theme_c() +
+  labs(y="",
+       x="Mean Overlap",
+       title = "Mean Overlap with Covidestim by State",
+       subtitle = "Overlap Normalized by Width of Covidestim Interval")
+
+##########################
+# BY TWO-WEEK INTERVAL
+##########################
+state_overlaps %>%
+  group_by(biweek) %>%
+  summarize(mean_overlap=mean(fraction_overlap)) %>%
+  left_join(dates) %>%
+  group_by(biweek) %>%
+  summarize(date = median(date),
+            mean_overlap=unique(mean_overlap)) %>%
+  ggplot(aes(x=date,
+             y =  mean_overlap)) +
+  geom_bar(stat="identity", alpha =.7) +
+  geom_vline(xintercept=end_old_model,linetype=2,
+             color='darkred',
+             linewidth=1.1) +
+  theme_c() +
+  labs(y="",
+       x="Mean Overlap",
+       title = "Mean Overlap with Covidestim by Time Interval",
+       subtitle = "Overlap Normalized by Width of Covidestim Interval") +
+  scale_x_date(date_breaks="2 months", date_labels="%b %Y")
+
+##########################
+# BY TWO-WEEK INTERVAL
+##########################
+# state_overlaps %>%
+#   mutate(posrate=positive/total) %>%
+#   filter(name %in% sample(unique(state_overlaps$name), 5)) %>%
+#   ggplot(aes(x=posrate, y = fraction_overlap)) +
+#   geom_point(alpha=.8) +
+#   theme_c() +
+#   # facet_wrap(~name, scales="free_y") +
+#   labs(x = "Positivity Rate", y = "Overlap") +
+#   scale_y_continuous(labels=comma)
+```
+
+```{r, fig.height=12}
+
+set.seed(123)
+state_sample <- sample(unique(state_overlaps$name), 16)
+
+state_overlaps %>%
+  mutate(posrate=positive/total) %>%
+ # filter(name %in% state_sample) %>%
+  ggplot(aes(x=posrate, y = fraction_overlap)) +
+  geom_point(alpha=.8) +
+  theme_c() +
+  facet_wrap(~name, scales="free_y") +
+  labs(x = "Positivity Rate", y = "Overlap",
+       title = "Comparing Fraction of Interval Covered by Positivity Rate",
+        subtitle = "Each Point Corresponds to a State-Biweek")
+
+state_overlaps %>%
+  mutate(posrate=positive/total) %>%
+#  filter(name %in% state_sample) %>%
+  ggplot(aes(x=posrate, y = fraction_overlap)) +
+  geom_point(alpha=.8) +
+  theme_c() +
+ # facet_wrap(~name, scales="free_y") +
+  labs(x = "Positivity Rate", y = "Overlap",
+       title = "Comparing Fraction of Interval Covered by Positivity Rate",
+        subtitle = "Each Point Corresponds to a State-Biweek")
+
+```
+
+```{r, fig.height=7}
+
+state_overlaps %>%
+  mutate(testrate=total/population) %>%
+  filter(name %in% state_sample) %>%
+  ggplot(aes(x=testrate, y = fraction_overlap)) +
+  geom_point(alpha=.8) +
+  theme_c() +
+  facet_wrap(~name, scales="free") +
+  labs( x= "Total Tests / Population Size", 
+        y = "Fraction of Covidestim Interval\nOverlapping with the PB interval",
+        title = "Comparing Fraction of Interval Covered by Testing Rate",
+        subtitle = "Each Point Corresponds to a State-Biweek")
+
+```
+
+```{r}
+
+state_overlaps %>%
+  mutate(testrate=total/population) %>%
+  filter(name %in% state_sample) %>%
+  ggplot(aes(x=testrate, y = fraction_overlap)) +
+  geom_point(alpha=.8) +
+  theme_c() +
+  labs( x= "Total Tests / Population Size", 
+        y = "Fraction of Covidestim Interval\nOverlapping with the PB interval",
+        title = "Comparing Fraction of Interval Covered by Testing Rate",
+        subtitle = "Each Point Corresponds to a State-Biweek")
+
+```
+
+## Relationship Between Agreement and Cumulative Incidence (Pre-Omicron)
+
+```{r, fig.height=11}
+
+state_population_link <- "https://www2.census.gov/programs-surveys/popest/datasets/2010-2019/state/detail/SCPRC-EST2019-18+POP-RES.csv"
+state_pop <- read_csv(state_population_link)
+statecodes <- read_csv(here("data/demographic/statecodes.csv"))
+
+state_pop <- state_pop %>%
+    left_join(statecodes, by = c("NAME" = "state")) %>%
+    select(population = POPESTIMATE2019,
+           state = code,
+           name = NAME) %>%
+    filter(!is.na(state))
+
+state_deaths <- tar_read(state_deaths, store=targets_dir)
+
+set.seed(123)
+state_sample <- sample(unique(state_deaths$state), 5)
+
+state_deaths %>%
+#  filter(state %in% state_sample) %>%
+  ggplot(aes(x=date,y=deaths/cases)) +
+  geom_line() +
+  facet_wrap(~state) +
+  theme_c() +
+  geom_vline(xintercept=end_old_model, color='darkred', linetype=2)
+
+comp_deaths <- state_deaths %>%
+  mutate(omicron=ifelse(date <= mdy("12-02-2021"),
+                        "pre_omicron", "omicron")) %>%
+  group_by(state, omicron) %>%
+  summarize(cumulative_deaths = sum(deaths),
+            cumulative_cases= sum(cases)) %>%
+  pivot_wider(names_from = omicron, 
+              values_from = c(cumulative_deaths, cumulative_cases)) %>%
+  mutate(pre_deaths_over_cases = cumulative_deaths_pre_omicron/cumulative_cases_pre_omicron,
+         omicron_deaths_over_cases = cumulative_deaths_omicron/cumulative_cases_omicron)
+
+comp_deaths %>%
+  select(state,pre_deaths_over_cases,omicron_deaths_over_cases) %>%
+  pivot_longer(contains("over_cases")) %>%
+  group_by(state) %>%
+  mutate(pre = value[which(name=="pre_deaths_over_cases")]) %>%
+  mutate(name = case_when(
+    grepl("omicron", name) ~ "Omicron",
+    grepl("pre", name) ~ "Before Omicron"
+  )) %>%
+  ggplot(aes(y=fct_reorder(state,pre),  x= value, fill=name)) +
+  geom_bar(stat='identity', position='dodge') +
+  theme_c(legend.title = element_text(face="bold", size=14)) +
+  labs(x = TeX("$\\frac{Total\\,Deaths}{Total\\,Cases}$"),
+       y="",
+       fill = "Time Period",
+       title = "Total Deaths Over Total Cases in the Time Period Before Omicron\n Compared to During Omicron") +
+  scale_x_continuous(n.breaks=7)
+
+comp_deaths %>%
+  select(state,pre_deaths_over_cases,omicron_deaths_over_cases) %>%
+  mutate(pct_change =( omicron_deaths_over_cases -pre_deaths_over_cases ) / pre_deaths_over_cases ) %>% 
+  ggplot(aes(y=fct_reorder(state,pct_change),  x= pct_change)) +
+  geom_bar(stat='identity', position='dodge') +
+  theme_c(legend.title = element_text(face="bold", size=11),
+          axis.text.x=element_text(size=14),
+          plot.subtitle = element_text(size=16,
+                                       face="plain", 
+                                       margin=margin(0,0,4,0))) +
+  labs(x = TeX("$\\frac{Total\\,Deaths}{Total\\,Cases}$"),
+       y="",
+       fill = "Time Period",
+       title = TeX("Percent Change in $\\frac{Total\\,Deaths}{Total\\,Cases}$ from the Time Period Before Omicron"),
+       subtitle="Compared to During Omicron") +
+  scale_x_continuous(labels=scales::percent)
+
+```
+
+```{r}
+
+comp_deaths %>%
+  select(state,pre_deaths_over_cases,omicron_deaths_over_cases) %>%
+  mutate(pct_change =( omicron_deaths_over_cases - pre_deaths_over_cases ) / pre_deaths_over_cases ) %>%
+  left_join(state_overlaps, by = c('state' = 'name')) %>%
+  group_by(state) %>%
+  summarize(mean_overlap=mean(fraction_overlap),
+            pct_change=unique(pct_change)) %>%
+  ggplot(aes(x=pct_change,y=mean_overlap)) +
+  geom_point(size=2) +
+  theme_c(  plot.subtitle = element_text(size=16,
+                                       face="plain", 
+                                       margin=margin(0,0,4,0))) +
+  scale_x_continuous(labels=scales::percent) +
+  labs(x="Percent Change", y = "Mean Overlap\n(By State)", 
+        title = TeX("Relationship Between the Percent Change in $\\frac{Total\\,Deaths}{Total\\,Cases}$ from the Time Period Before Omicron"),
+       subtitle="Compared to During Omicron versus the Mean Overlap")
+
+```
+
+```{r}
+
+comp_deaths %>%
+  rename(name=state) %>%
+  ungroup() %>%
+  left_join(state_pop[,c("name", "population")]) %>%
+  mutate(frac_infected = cumulative_cases_pre_omicron/population) %>%
+  left_join(state_overlaps[,c("name","fraction_overlap", "biweek")]) %>%
+  filter(biweek >= 25) %>%
+  group_by(name) %>%
+  summarize(mean_overlap=mean(fraction_overlap),
+            frac_infected=unique(frac_infected)) %>%
+  ggplot(aes(x=frac_infected, y =mean_overlap)) +
+  geom_point() +
+  theme_c()
+
+```
+
+```{r}
+
+covidestim_link <- "https://covidestim.s3.us-east-2.amazonaws.com/latest/state/estimates.csv"
+
+covidestim <- read_csv(covidestim_link)
+
+# join data from each source to include dates before and after 2021-12-02
+covidestim <- covidestim %>%
+    select(date, contains("infections"), state) %>%
+    filter(year(date) > 2020) %>%
+    mutate(week = week(date), year = year(date))
+
+covidestim %>%
+  group_by(state) %>%
+  summarize(infections=sum(infections))%>%
+  rename(name=state) %>%
+  left_join(state_pop[,c("name", "population")]) %>%
+  mutate(frac_infected = infections/population) %>%
+  left_join(state_overlaps[,c("name","fraction_overlap", "biweek")]) %>%
+  filter(biweek >= 25) %>%
+  group_by(name) %>%
+  summarize(mean_overlap=mean(fraction_overlap),
+            frac_infected=unique(frac_infected)) %>%
+  ggplot(aes(x=frac_infected, y =mean_overlap)) +
+  geom_point() +
+  theme_c()
+
+```

### 2023-07-23 R/01-data_cleaning.R

@@ -669,5 +669,49 @@ get_variant_proportions <- function() {
   
 }
 
+#------------- State Deaths -------------
+
+get_state_deaths <- function() {
+  
+  
+  if(!file.exists(here("data/data_raw/state-deaths-jhu.RDS"))) {
+    dates <- seq.Date(mdy("12-31-2020"), mdy("02-26-2022"), by="1 day") %>%
+      format("%m-%d-%Y") 
+    
+    links <- paste0("https://raw.githubusercontent.com/CSSEGISandData/COVID-19/master/csse_covid_19_data/csse_covid_19_daily_reports_us/",
+                    dates, ".csv")
+    
+    state_deaths <- map2_df(links, dates, ~read_csv(.x) %>% mutate(date = .y))
+    saveRDS(state_deaths, here("data/data_raw/state-deaths-jhu.RDS"))
+    
+  }
+ 
+  else {
+    state_deaths <- readRDS(here("data/data_raw/state-deaths-jhu.RDS"))
+  }
+
+  state_deaths <- state_deaths %>%
+    filter(FIPS <= 56) %>%
+    # exclude outside of continental U.S.
+    select(date=Date, cases=Confirmed, deaths = Deaths, state=Province_State) %>%
+    group_by(state) %>%
+    mutate(cases=cases-lag(cases, n=1, order_by=date),
+           deaths=deaths-lag(deaths, n=1, order_by=date)) %>%
+    ungroup() %>%
+    filter(!is.na(cases) & !is.na(deaths)) %>%
+    mutate(across(c(cases,deaths), ~ifelse(.x <0, 0, .x)))
+  
+  
+  dates <- readRDS(here("data/data_raw/date_to_biweek.RDS"))
+  
+  state_deaths <- state_deaths %>%
+    left_join(dates) %>%
+    group_by(state,biweek) %>%
+    summarize(across(c(deaths, cases), sum),
+              date=median(date))
+  
+  return(state_deaths)
+ 
+}
 
 

### 2023-07-23 _targets.R

@@ -60,6 +60,9 @@ list(
   tar_target(
     name = variant,
     command = get_variant_proportions()),
+  tar_target(
+    name = state_deaths,
+    command = get_state_deaths()),
   
   #---------------- state-level analysis -------------------
   # set priors

### 2023-07-24 reports/county_figures.Rmd
cross correlation figure

@@
 ```
+
+# Cross Correlations
+
+```{r plot ccf functions}
+
+objs <- ls()
+remove <- objs[which(!objs %in% c("theme_c", "wjoined", "pal"))]
+rm(remove)
+
+pal <- c("red", "#10BAC5", "#1B10C5", "#EFB719", "#900C3F", "#FF7400")
+
+names(pal) <- c( "Covidestim",
+                '$Pr(S_1|untested)$ Centered at Survey Value',
+                '$Pr(S_1|untested)$ and $\\beta$ Centered at Survey Values',
+               "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$",
+               "$\\beta$ Centered at Survey Value",
+               "observed"
+               )
+
+plot_all_ccf <- function(input_df_for_fips, 
+                         varnames= c("differenced_mean_conc",
+                                  "differenced_exp_median")) {
+
+  versions <- input_df_for_fips %>% 
+    group_split(version)
+
+  message(length(versions))
+
+  plt1 <- plot_ccf(versions[[1]], varnames)
+  plt2 <- plot_ccf(versions[[2]], varnames)
+  plt3 <- plot_ccf(versions[[3]], varnames)
+  plt4 <- plot_ccf(versions[[4]], varnames)
+
+  cowplot::plot_grid(plt1, plt2, plt3, plt4, nrow = 1)
+
+}
+
+plot_ccf <- function(input_df_for_version,
+                     varnames = c("differenced_mean_conc",
+                                  "differenced_exp_median")) {
+
+  var1 <- varnames[1]
+  var2 <-varnames[2]
+  lab <- unique(input_df_for_version$vlabel)
+
+  cross_correlations <- ccf(
+    pull(input_df_for_version, var1),
+     pull(input_df_for_version, var2), 
+                            plot = FALSE)
+
+  ccf_res <- tibble(lag = cross_correlations$lag[,,1], 
+                  correlation=cross_correlations$acf[,,1])
+
+  county_name <- gsub("County, MA", "", unique(input_df_for_version$name))
+
+  ccf_res %>%  
+    mutate(max = ifelse(correlation == max(correlation), 
+                        lag, NA),
+           maxcorr = round(max(correlation),3)) %>%
+    ggplot(aes(x=lag, y = correlation)) +
+    geom_point() +
+    geom_linerange(aes(ymin = 0, ymax=correlation)) +
+    theme_c() +
+    theme(plot.title = element_text(size = 12, hjust = .5,
+                                    margin=margin(1,1,8,1)),
+          plot.margin = margin(10,1,1,1),
+            plot.subtitle = element_text(size = 12, hjust = .5, 
+                                         margin = margin(4,1,1,1))) +
+    labs(title =TeX(paste0(
+      "Cross Correlation between $(Y_t)$ and $(Z_t)$, County: ", 
+      county_name)),
+         subtitle = TeX(paste0("Version: ", lab))) +
+    geom_label(aes(x = max, y = correlation,
+                   label = paste0("Max Correlation: ",
+                                  maxcorr, "\n Lag: ",
+                                  max)),
+               hjust = -.1) +
+    ylim(-1,1) +
+    scale_x_continuous(n.breaks = 10)
+
+}
+
+get_max_lag_by_version <- function(input_df_for_version,
+                                   varnames = c("differenced_mean_conc",
+                                  "differenced_exp_median"),
+                                  comparison = "bias corrected") {
+
+  var1 <- varnames[1]
+  var2 <-varnames[2]
+  lab <- unique(input_df_for_version$vlabel)
+
+  cross_correlations <- ccf(
+    pull(input_df_for_version, var1),
+     pull(input_df_for_version, var2), 
+                            plot = FALSE)
+
+  # cross_correlations <- ccf(input_df_for_version$differenced_mean_conc,
+  #                           input_df_for_version$differenced_exp_median, 
+  #                           plot = FALSE)
+
+  ccf_res <- tibble(lag = cross_correlations$lag[,,1], 
+                  correlation=cross_correlations$acf[,,1])
+
+  m <- ccf_res %>% filter(correlation == max(correlation))
+
+  tibble(vlabel = unique(input_df_for_version$vlabel),
+         fips = unique(input_df_for_version$fips),
+         max_lag = m$lag,
+         max_corr = m$correlation,
+         comparison = comparison,
+         county_name = unique(input_df_for_version$name))
+
+}
+
+```
+
+```{r generate differenced df}
+
+differenced_all <- wjoined %>%
+  filter(!is.na(mean_conc)) %>%
+  select(-date) %>%
+  distinct() %>%
+  group_by(vlabel, fips) %>%
+  arrange(biweek) %>%
+  mutate(differenced_exp_median = exp_cases_median -
+           lag(exp_cases_median, n =1),
+         differenced_mean_conc = mean_conc - lag(mean_conc, n = 1),
+         differenced_positive = positive - lag(positive, n = 1),
+         differenced_covidestim = infections - lag(infections, n = 1)) %>%
+  ungroup() %>%
+  filter(!is.na(differenced_exp_median) & !is.na(differenced_mean_conc)) %>%
+  ungroup()
+
+```
+
+### Cross Correlation (Wastewater and Covidestim Estimates)
+
+```{r wastewater covidestim, fig.width = 16}
+
+# glimpse(differenced_all)
+
+differenced_all %>% 
+  filter(version == "v7" & !is.na(differenced_covidestim)) %>%
+  group_by(fips) %>%
+  group_split() %>% 
+  walk(~ { 
+    plt <- .x %>% plot_ccf(varnames= c("differenced_mean_conc",
+                                  "differenced_covidestim"))
+    print(plt)})
+
+```
+
+### Cross Correlation (Wastewater and Observed Cases)
+
+```{r, error = FALSE, out.width = "80%"}
+
+# cross correlation plots for each county
+differenced_all %>% 
+  filter(version == "v7") %>%
+  group_by(fips) %>%
+  group_split() %>% 
+  walk(~ { 
+    plt <- .x %>% plot_ccf(varnames= c("differenced_mean_conc",
+                                  "differenced_positive"))
+    print(plt)})
+
+```
+
+### Cross Correlation (Wastewater and Bias-Corrected Cases)
+
+```{r, fig.width = 23}
+
+# cross correlation plots for each county
+differenced_all %>% 
+  group_by(fips) %>%
+  group_split() %>%
+  walk(~ { 
+    plt <- .x %>% plot_all_ccf()
+    print(plt)})
+
+```
+
+# Compare Lags with Maximum Correlation Between Covidestim, Wastewater, and Bias Corrected Counts
+
+## Compare Lags Among versions of Bias Corrected Counts
+
+For each county, lag with the maximum correlation:
+
+```{r, fig.width = 12.7, fig.height =7}
+
+# lags for bias-corrected counts
+lags_all_pb <- differenced_all %>%
+  group_by(fips, vlabel) %>%
+  group_split() %>%
+  map_df(get_max_lag_by_version)
+
+lags_all_pb %>%
+  group_by(fips) %>%
+  summarize(`Lags with Maximum Correlation in Some Version` =
+              paste0(unique(max_lag), collapse =", "))
+
+# lags for observed counts
+lags_all_observed <- differenced_all %>% 
+  filter(version == "v7") %>%
+  group_by(fips) %>%
+  group_split()%>%
+  map_df(~get_max_lag_by_version(
+    input_df_for_version = .x,
+    varnames= c("differenced_mean_conc",
+                "differenced_positive"),
+   comparison = "observed" )) %>%
+  select(-vlabel)
+
+# compare pb max correlations to those for observed counts 
+lags_all_pb %>%
+  bind_rows(lags_all_observed) %>%
+  mutate(vlabel=as.character(vlabel)) %>%
+   mutate(source = ifelse(!is.na(vlabel), vlabel, comparison)) %>%
+  group_by(fips) %>%
+  mutate(m = median(max_corr)) %>%
+  ungroup() %>%
+  mutate(max_lag=factor(max_lag)) %>%
+  mutate(county_name = gsub("County, MA", "", county_name)) %>%
+   ggplot(aes(x=fct_reorder(county_name,m, 
+                           .desc = TRUE),
+             y = max_corr,
+             color = source, 
+             shape = max_lag)) +
+  geom_jitter(aes(fill=source),color="black",
+              height =0, width =.15, size = 3,
+              stroke=.1,
+              alpha=.9) +
+  scale_shape_manual(values=c(21,22,24)) +
+  scale_fill_manual(values=pal,
+                    name = 'Time Series Compared with\nWastewater:',
+                    labels=TeX(names(pal))) +
+  theme_bw() +
+  theme_c() +
+  theme(plot.title = element_text(size=16,hjust=.2, face="bold"),
+    # axis.text.x = element_text(angle  = 60, vjust = .4),
+    axis.text.x = element_text(angle  = 50, vjust = .95, hjust=1, size =14),
+    axis.title.x = element_text(margin = margin(3,3,3,3)),
+        legend.text = element_text(size = 14, hjust=0),
+        axis.text.y = element_text(size = 11),
+        legend.title = element_text(face="bold", size = 16)) +
+  labs(y = "Maximum Correlation") +
+  scale_size_discrete(breaks = c(1.9,2),
+                      range = c(1.9,2),
+                      labels = c("bias corrected", "observed")) +
+  labs(x = "",
+       y = "Maximum Correlation",
+       title = "Comparing Maximum Correlation of Probabilistic Bias Counts with Wastewater\nto that Between Observed Cases and Wastewater",
+       shape = "Lag Where Maximum\nCorrelation Was Obtained:",
+       color = "Time Series Compared with Wastewater:") +
+  guides(shape = guide_legend(override.aes = list(size = 4, stroke=.4)),
+         fill= guide_legend(override.aes = list(size = 3.5, 
+                                                shape=24, 
+                                                stroke=.8)))
+
+ggsave(here("figures/cross_correlation_observed_pb.pdf"), 
+       height = 8, width=12)
+
+# max series by rank; use for following comparison
+max_series_by_rank <- lags_all_pb %>%
+  group_by(fips) %>%
+  mutate(r = rank(max_corr),
+            series = vlabel[which.max(r)]) %>%
+  group_by(series) %>%
+  summarize(n=n()) %>%
+  filter(n==max(n)) %>%
+  pull(series)
+
+```

### 2023-07-24 writing/manuscript.Rmd
cross correlation figure

@@
 
-In the first three implementations, the prior distributions specified for the bias parameters vary by state and date.
+In the three out of the four implementations, the prior distributions specified for the bias parameters vary by state and date.
 In particular, we use survey data from the COVID-19 Trends and Impact Survey [@salomon2021a] to inform the prior distributions for one to two of the bias parameters: the probability of having symptoms among the untested population ($\Pr(S_1|\text{untested})$), and the parameter $\beta$, which represents the ratio of the test positivity among members of the untested population who have no symptoms to the overall test positivity.
 More detail on these implementations is provided in the [Methods](#Methods) section.
-In another, more simple, implementation, which is the only implementation presented at the state-level, the prior distributions are informed by the distributions from the survey data across all dates and states, but the same prior distributions are used for all states and time intervals considered.
+In another, more simple, implementation, which is the only implementation presented at the state-level (Figure \@ref(fig:ratio-estimated-to-observed-by-wave) and \@ref(fig:simulation-intervals-state)), the prior distributions are informed by the distributions from the survey data across all dates and states, but the same prior distributions are used for all states and time intervals considered.
 
@@
 Only the two implementations that were consistently the highest and lowest among the implementations are included for clarity.
-In the   In the first three implementations, the priors vary by state and date, while for the fourth, they are the same for all states and time intervals considered.
+In the   In the three out of the four implementations, the priors vary by state and date, while for the fourth, they are the same for all states and time intervals considered.
 For the first implementation, we center the distribution of $\\beta$ at the ratio of the screening test positivity to the overall test positivity from the survey, and we center the distribution of $\\Pr(S_1|\\text{untested})$ at the percentage of the population experiencing COVID-19-like illness from the survey for each two-week interval.
@@
 This corresponds to a similar time frame that we saw at the state-level in Figure \@ref(fig:ratio-estimated-to-observed-by-wave).
-Hampsphire County and Suffolk County were consistently among the lowest with regard to the ratio of estimated to observed infections, which may be explained by the screening testing occurring at universities in these counties.
+Hampshire County and Suffolk County were consistently among the lowest with regard to the ratio of estimated to observed infections, which may be explained by the screening testing occurring at universities in these counties.
 
@@
 
-```{r,eval=FALSE}
+Also of interest is whether there is a lag between the wastewater concentrations and bias corrected estimates.
+Although wastewater concentrations can be a leading indicator since people may shed viral material before developing symptoms and subsequently getting tested, the lead time is a result of multiple epidemiological factors, including viral shedding dynamics that may differ with evolution of the virus, access to testing, and testing behavior [@olesen2021a].
+Indeed, the lead or lag time between wastewater concentrations and hospitalizations or cases has varies substantially between waves, where the lead time is higher in earlier waves and near zero at later waves, a difference that may be attributable to changes in diagnostic test availability over time  [@xiao2022a; @hopkins2023a].
 
-include_graphics(here('figures/simulation-intervals-county.pdf'))
+In Figure \@ref(fig:cross-corr-wastewater), we see that among all counties except one in Massachsuetts, the maximum correlation between mean wastewater concentration and biweekly infection estimates was obtained at a lag of zero, that is, there  typically there was no lag among the biweekly infection estimates and mean wastewater concentrations.
+This is to be expected given the lead time for wastewater concentrations, when observed, is less than 2 weeks [@olesen2021a].
+In two counties, Hamsphire and Barnstable, the lag between the probabilistic bias counts and wastewater concentration was less than that between the observed positive tests and wastewater concentrations.
+
+```{r cross-corr-wastewater, out.width="70%"}
+
+include_graphics(here("figures/cross_correlation_observed_pb.jpeg"))
 

### 2023-07-24 reports/state_figures.Rmd
add interpretation of table + cross correlations figure

@@
 # Summarizing Concordance with Covidestim
+
+Below, we see that although the implementation where $\Pr(S_1|untested)$ was at the survey as well as the implementation did not vary by state or date, the mean simulation interval width was actually lower, indicating the greater concordance was not merely a result of wider bounds.
+
+```{r}
+
+state_res %>%
+  select(-c(date)) %>%
+  distinct() %>%
+  mutate(interval_length =exp_cases_ub - exp_cases_lb) %>%
+  group_by(vlabel) %>%
+  summarize(mean_interval_length = mean(interval_length))
+
+```
 

### 2023-07-24 writing/manuscript.Rmd
add interpretation of table + cross correlations figure

@@
 Omicron also is associated with a lower infection fatality ratio [@liu2022], as well as milder infection overall, which may influence testing behavior.
-Omicron's rise was rapid too, presumably as a result of this enhanced transmissibility and immune escape;  it became the dominant variant over the course of a single month.
+Its rise to dominance was rapid, presumably as a result of this enhanced transmissibility and immune escape;  it became the dominant variant over the course of a single month.
 
@@
 The changes in the model may contribute to the differences we see between the probabilistic bias intervals and Covidestim intervals during the Omicron wave.
+
+ADD TEST POSITIVITY FIGURE
+FIGURE OUT WHICH STATES HAVE LAGS
 
@@
 
-We focus on Massachussetts because this state is the only state that reports both positive and total tests and has the most counties where wastewater data was reported throughout the entire time period of interest.
+We focus on Massachussetts because this state is the only state that reports both positive and total tests at the county-level and has wastewater data consistently reported throughout the entire time period of interest.
 Wastewater data here is useful in the sense that it is inherently not subject to the same biases as voluntary testing data, since it captures anyone in the wastewater catchment area, and as such is not influenced by more symptomatic people being more likely to be tested or differences in access to testing.
@@
 
-```{r testrate-v-ratio, fig.cap  = 'The ratio of the median estimated infections to observed infections plotted against the testing rate, where the testing rate is calculated as the total number tested in a two-week interval over the population size.
+```{r testrate-v-ratio, fig.cap ="The ratio of the median estimated infections to observed infections plotted against the testing rate, where the testing rate is calculated as the total number tested in a two-week interval over the population size.
 When the priors are the same for all time intervals, there is minimal variability relationship between the testing rate and the ratio of estimated to observed infections, since the correction for incomplete testing and diagnostic test inaccuracy is identical for each time-interval.
 However, when we allow $\\beta$ or $\\Pr(S_1|\\text{untested})$ to vary over time, there is more variability in the relationship.
-A horizontal line in red at 1 is included to reference; a ratio of exactly 1 would indicate no infections went unobserved.'}
+A horizontal line in red at 1 is included to reference; a ratio of exactly 1 would indicate no infections went unobserved."}
 
@@
 
-We see in Figure \@ref(fig:correlations)a) that for all implementations, the probabilistic bias estimates are highly correlated with wastewater concentrations, and in Figure \@ref(fig:correlations)b) we see they also highly correlated with the Covidestim estimates.
+We see in Figure \@ref(fig:correlations) (a) that for all implementations, the probabilistic bias estimates are highly correlated with wastewater concentrations, and in Figure \@ref(fig:correlations) (b) we see they also highly correlated with the Covidestim estimates.
 The correlations in both cases are very similar between implementations, differing by less than 0.01.
@@
 
-Table \@ref(table:coverage) is
+In Table \@ref(table:coverage), we summarize the concordance with Covidestim at the county and state levels by comparing the percent of simulation intervals where the Covidestim median falls above, below, or within the interval.
+At both geographic scales, we see that the probabilistic bias implementation that is most concordant with Covidestim is the that where the prior for $\Pr(S_1|\text{untested})$ is centered at the survey value of the percent of the population experiencing COVID-19-like illness, followed by the implementation that does not vary by state or date.
+This agreement was not a result of wider bounds, as the simulation intervals for the other two implementations were larger on average.
+For both of these implementations, when the Covidestim median was not contained, this was typically because it fell above the simulation interval.
+As is clear from Figure \@ref(fig:simulation-intervals-state) at the state-level, this often occurred during the Omicron wave, when the Covidestim estimates are consistently above the probabilistic bias counts.
 
@@
           .by_group=TRUE) %>%
+  mutate(across(contains("percent"),~ paste0(round(.,2), "\\%"))) %>%
   ungroup() %>%
@@
   footnote(general = paste0("The percent of simulation intervals where the Covidestim median falls below, within, or above the interval,",
-  "when considering all simulation intervals for that implementation and geographic scale.",
+  "when considering all simulation intervals for that implementation and geographic scale. ",
   "Implementations are ordered from those with the highest proportion of Covidestim medians contained in the interval, to the lowest."),
@@
 This is to be expected given the lead time for wastewater concentrations, when observed, is less than 2 weeks [@olesen2021a].
-In two counties, Hamsphire and Barnstable, the lag between the probabilistic bias counts and wastewater concentration was less than that between the observed positive tests and wastewater concentrations.
+In two counties, Hampsphire and Barnstable, the lag between the probabilistic bias counts and wastewater concentration was less than that between the observed positive tests and wastewater concentrations.
 
-```{r cross-corr-wastewater, out.width="70%"}
+```{r cross-corr-wastewater, fig.width=8, fig.cap="Results of cross correlation analysis between probabilistic bias estimates and wastewater concentrations aggregated to the county level for counties in Massachusetts.
+For each county, the maximum correlation between the probablistic bias estimates and wastewater concentration is shown, colored by implementation; also included is the correlation between the observed positive tests and wastewater concentration.
+The shape indicates the lag at which the maximum correlation occurred, with units in biweeks.
+For example, a lag of -1 indicates the maximum correlation is obtained when the biweekly concentrations are lagged by 1 biweek, or equivalently, that the wastewater concentrations lead the observed positive tests by one biweek.
+For all counties except Barnstable, the maximum correlations between wastewater concentrations and probablistic bias estimates occur at no lag.
+In Barnstable and Hampshire, observed tests lagged wastewater concentrations, but the probablistic bias estimates for these counties did not."}
 
-include_graphics(here("figures/cross_correlation_observed_pb.jpeg"))
+include_graphics(here("figures/cross_correlation_observed_pb.pdf"))
 
@@
 
-# Methods
+# Methods {#Methods}
 

### 2023-08-26 reports/state_figures.Rmd
update figures

@@
   group_by(vlabel) %>%
-  summarize(mean_interval_length = mean(interval_length))
+  summarize(mean_interval_length = mean(interval_length)) %>%
+  arrange(desc(mean_interval_length))
+
+```
+
+```{r}
+
+state_res_differenced <- state_res %>%
+  select(-date) %>% distinct() %>%
+  group_by(state, vlabel) %>%
+  mutate(exp_cases_median_differenced = exp_cases_median - lag(exp_cases_median, n =1, order_by = biweek),
+         infections_differenced = infections - lag(infections, n = 1, order_by=biweek)) %>%
+  select(state, name, vlabel, contains('differenced'), biweek) %>%
+  filter(!is.na(exp_cases_median_differenced)) %>%
+  filter(biweek <=25)
+
+# if lag is NEGATIVE, covidestim LEADS wastewater
+# if lag is POSITIVE, covidestim LAGS wastewater
+# since by the documentation, 'the lag k value returned by ccf(x, y) estimates
+# the correlation between x[t+k] and y[t].'
+
+get_max_lag_by_version <- function(input_df_for_version,
+                                   varnames = c("infections_differenced",
+                                  "exp_cases_median_differenced"),
+                                  comparison = "bias corrected") {
+
+  var1 <- varnames[1]
+  var2 <-varnames[2]
+  lab <- unique(input_df_for_version$vlabel)
+
+  cross_correlations <- ccf(
+    pull(input_df_for_version, var1),
+     pull(input_df_for_version, var2), 
+                            plot = FALSE,
+    lag.max=3)
+
+  ccf_res <- tibble(lag = cross_correlations$lag[,,1], 
+                  correlation=cross_correlations$acf[,,1])
+
+  m <- ccf_res %>% filter(correlation == max(correlation))
+
+  tibble(vlabel = unique(input_df_for_version$vlabel),
+         fips = unique(input_df_for_version$state),
+         max_lag = m$lag,
+         max_corr = m$correlation,
+         comparison = comparison,
+         name = unique(input_df_for_version$name))
+
+}
+
+cross_corr <- state_res_differenced %>%
+  group_by(state, vlabel) %>%
+  group_split() %>%
+  map_df(get_max_lag_by_version)
+
+states_with_lag <- cross_corr %>%
+  filter(max_lag >  0 & 
+           vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$" & 
+           max_corr > .55) %>% 
+  pull(fips) %>%
+  unique()
+
+```
+
+```{r, include=FALSE}
+
+state_res_differenced %>%
+  filter(state == "AL" & vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
+  ggplot(aes(x=exp_cases_median_differenced, y = infections_differenced)) +
+  geom_point()
+
+state_res %>%
+  filter(state == "AL" & vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
+  ggplot(aes(x=exp_cases_median, y = infections)) +
+  geom_point()
 
@@
 ggsave(here('figures/intervals-and-covidestim-all-states.pdf'))
+
+```
+
+## Compare States that Have a Lag
+
+```{r,  fig.height=11, fig.width=12}
+
+# states_with_lag
+
+subset %>%
+    filter(state %in% states_with_lag & biweek <= 25) %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub,
+             fill = 'Probabilistic Bias Analysis'),
+             alpha = .8,
+             show.legend=FALSE) +
+  geom_ribbon(aes(x = date,
+                  fill = 'Covidestim',
+                  ymin = infections.lo,
+                  ymax = infections.hi),
+              alpha = .6) +
+  # geom_vline(xintercept=end_old_model,
+  #            linetype=2,
+  #            color="darkred",
+  #            linewidth=.5) +
+  facet_wrap(~name, ncol=3, scales = "free",
+              labeller= labeller(vlabel =as_labeller(TeX, default=label_parsed))) +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3.5 months", 
+               date_labels = "%b %Y") +
+  theme_c(axis.text.x=element_text(size=7),
+          axis.text.y = element_text(size=4.5),
+          legend.position = "top",
+          legend.text=element_text(size=12),
+          strip.text.x=element_text(size=11,
+                                    face='plain',
+                                    color='white')) +
+  scale_fill_manual(values =comp_state_pal, name='')  +
+  labs(y = "Estimated Infections",
+       title = paste0("Estimated Infections by State\nin States with a Lag")) +
+  guides(color = guide_legend(override.aes = list(linewidth =2),
+                              nrow=2,
+                              byrow=TRUE))
+
+```
+
+```{r, fig.height=11, fig.width=12, eval=save_figures}
+
+ggsave(here('figures/intervals-and-covidestim-states-lag.pdf'))
+
+```
+
+## Test Positivity and Positive Tests in States with a Lag
+
+```{r testpos-lag, fig.height=14, fig.width=12}
+
+############################################
+# POSITIVE TESTS AND TEST POSITIVITY
+############################################
+pltlist <- subset %>%
+  filter(biweek<=25) %>%
+  filter(state %in% states_with_lag) %>%
+    mutate(testpos = positive/total) %>%
+  group_by(across(-date)) %>%
+  summarize(date=min(date) + days(7)) %>%
+  group_by(state) %>%
+  mutate(adj = (1/max(testpos)) * max(positive)) %>%
+  group_split() %>%
+  map(~ {
+    adj <-unique(.x$adj)
+    .x %>%
+      ggplot(aes(x=date))+
+    geom_line(aes(y=positive, color = 'Positive Tests')) +
+      geom_point(aes(y=positive, color = 'Positive Tests'),
+              alpha=.5,
+              size=.8) +
+    geom_line(aes(y=testpos*adj, color='Test Positivity'),
+              linetype=2) +
+      geom_point(aes(y=testpos*adj, color='Test Positivity'),
+              alpha=.5,
+              size=.8) +
+    #scale_linetype_discrete(breaks = c(1,1,2)) +
+    scale_color_manual(values=c(
+                                "Positive Tests"= "#279143",
+                                "Test Positivity"= '#E38E4F')) +
+    guides(color = guide_legend(override.aes = list(linetype = c(1,2),
+                                                    linewidth=c(2,2)))) +
+    theme_c() +
+    theme(axis.title = element_text(size=10),
+          axis.text =element_text(size=9),
+          legend.position="none") +
+    scale_y_continuous(labels=comma, 
+                       sec.axis = sec_axis(
+                         ~./adj,
+                         name = 'Test Positivity')) +
+    labs(color ='',
+         y="Positive Tests",
+         x= "Date") +
+      facet_wrap(~name)
+
+  })
+
+legend <- get_legend(
+  pltlist[[1]] +
+    theme(legend.position = "top",
+          legend.text=element_text(size=16))
+)
+
+title_gg <- ggplot() + 
+  labs(title = "Comparing Trends in Positive Tests and Test Positivity") + 
+  theme(plot.title=element_text(face="bold",
+                                hjust = .5, 
+                                size = 18,
+                                margin =margin(5,0,2,0)))
+
+plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)
+
+plot_grid(title_gg, legend, plt, rel_heights=c(.03, .03,.94), ncol=1)
+
+```
+
+```{r, eval=save_figures, fig.height=14, fig.width=12}
+
+ggsave(here('figures/testpos-lag.pdf'))
 

### 2023-08-26 writing/supplement.Rmd
update manuscript & supplement
update figures

@@
 
-## COVID-19 Trends and Impact Survey
+# COVID-19 Trends and Impact Survey
 
@@
 While there are estimates of $\\Pr(S_1|\\text{untested})$ for all states (seen in the following figure), there are 27 states with estimates of $\\beta$.", fig.width=14,fig.height=12}
+
 include_graphics(here('figures/smoothing-beta.pdf'))
+
 ```
@@
 
+# Comparing Lag Between Probabilistic Bias Estimates and Covidestim Infections
+
+```{r}
+
+include_graphics(here('figures/intervals-and-covidestim-states-lag.pdf'))
+
+```
+
+```{r}
+
+include_graphics(here('figures/testpos-lag.pdf'))
+
+```
+
 # County Results by Implementation
 
-```{r sim-intervals-faceted-county, fig.width=15, fig.height=16,  fig.cap = "Simulation intervals for counties in Massachusetts.
+```{r sim-intervals-faceted-county, fig.width=15.5, fig.height=16,  fig.cap = "Simulation intervals for counties in Massachusetts.
 Each interval corresponds to a 95\\% simulation interval for the total number of estimated infections for that county in that two-week time interval.

### 2023-08-28 reports/county_figures.Rmd
update lag figures

@@
       input,
-      output_dir = "../output_html")})
+      output_dir = "../output")})
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE,cache=TRUE, fig.width=10, fig.height=7)
+knitr::opts_chunk$set(echo = TRUE, 
+                      message=FALSE, 
+                      warning=FALSE,
+                      cache=TRUE, 
+                      fig.width=10, 
+                      fig.height=7)
 

### 2023-08-28 reports/state_figures.Rmd
update lag figures

--- 2023-07-24 reports/state_figures.Rmd
@@
     css: !expr here::here('css/template.css')
+    toc_depth: 4
   html_document:
@@
   group_by(vlabel) %>%
-  summarize(mean_interval_length = mean(interval_length))
+  summarize(mean_interval_length = mean(interval_length)) %>%
+  arrange(desc(mean_interval_length)) %>%
+  kableExtra::kbl()
+
+```
+
+```{r}
+
+state_res_differenced <- state_res %>%
+  select(-date) %>% distinct() %>%
+  mutate(testpos=positive/total) %>%
+  group_by(state, vlabel) %>%
+  mutate(exp_cases_median_differenced =
+           exp_cases_median - lag(exp_cases_median,
+                                  n =1,
+                                  order_by = biweek),
+         infections_differenced = infections - lag(infections,
+                                                   n = 1,
+                                                   order_by=biweek),
+         testpos_differenced = lag(testpos, n=1, order_by=biweek),
+         positive_differenced =lag(positive,n=1,order_by=biweek)) %>%
+  select(state, name, vlabel, contains('differenced'), biweek) %>%
+  filter(!is.na(exp_cases_median_differenced)) %>%
+  filter( biweek <=25)
+
+# if lag is NEGATIVE, covidestim LEADS wastewater
+# if lag is POSITIVE, covidestim LAGS wastewater
+# since by the documentation, 'the lag k value returned by ccf(x, y) estimates
+# the correlation between x[t+k] and y[t].'
+
+get_max_lag_by_version <- function(input_df_for_version,
+                                   varnames = c("infections_differenced",
+                                  "exp_cases_median_differenced"),
+                                  comparison = "bias corrected") {
+
+  var1 <- varnames[1]
+  var2 <-varnames[2]
+  lab <- unique(input_df_for_version$vlabel)
+
+  cross_correlations <- ccf(
+    pull(input_df_for_version, var1),
+     pull(input_df_for_version, var2), 
+                            plot = FALSE,
+    lag.max=3)
+
+  ccf_res <- tibble(lag = cross_correlations$lag[,,1], 
+                  correlation=cross_correlations$acf[,,1])
+
+  m <- ccf_res %>% filter(correlation == max(correlation))
+
+  tibble(vlabel = unique(input_df_for_version$vlabel),
+         fips = unique(input_df_for_version$state),
+         max_lag = m$lag,
+         max_corr = m$correlation,
+         comparison = comparison,
+         name = unique(input_df_for_version$name))
+
+}
+
+cross_corr <- state_res_differenced %>%
+  group_by(state, vlabel) %>%
+  group_split() %>%
+  map_df(get_max_lag_by_version)
+
+states_with_lag <- cross_corr %>%
+  filter(max_lag >  0 & vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
+  # remove those where lag appears to be due to low correlation rather 
+  # than true lag
+  filter(max_corr > .6) %>% 
+  pull(fips) %>%
+  unique()
+
+cross_corr_testpos <- state_res_differenced %>%
+  filter(biweek >=6 & biweek <= 25) %>%
+  # positive tests and positivity are not dependent on vlabel; use this
+  # one because it goes to earlier dates
+  filter(vlabel=="$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
+  group_by(state, vlabel) %>%
+  group_split() %>%
+  map_df(~get_max_lag_by_version(.x,
+    varnames = c("positive_differenced", 
+                 "testpos_differenced"))) %>%
+  filter(max_lag > 0) %>%
+  pull(fips) %>%
+  unique()
+
+# length(states_with_lag)
+# setdiff(states_with_lag, cross_corr_testpos) %>% length()
+
+```
+
+```{r, include=FALSE}
+
+state_res_differenced %>%
+  filter(state == "AL" & vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
+  ggplot(aes(x=exp_cases_median_differenced, y = infections_differenced)) +
+  geom_point()
+
+state_res %>%
+  filter(state == "AL" & vlabel == "$Priors\\,Do\\,Not\\,Vary\\,by\\,State\\,and\\,Date$") %>%
+  ggplot(aes(x=exp_cases_median, y = infections)) +
+  geom_point()
 
@@
 ggsave(here('figures/intervals-and-covidestim-all-states.pdf'))
+
+```
+
+## Compare States that Have a Lag
+
+States where the Covidestim estimates are lagged relative to the probabilistic bias counts are `r subset %>% filter(state %in% states_with_lag) %>% pull(name) %>% unique()`.
+
+```{r,  fig.height=11, fig.width=12}
+
+# states_with_lag
+
+lagged_fig_a <- subset %>%
+    filter(state %in% states_with_lag & biweek <= 25 & biweek >=6) %>%
+   ggplot() +
+  geom_ribbon(aes(x = date, 
+             ymin = exp_cases_lb,
+             ymax = exp_cases_ub,
+             fill = 'Probabilistic Bias Analysis'),
+             alpha = .8,
+             show.legend=FALSE) +
+  geom_ribbon(aes(x = date,
+                  fill = 'Covidestim',
+                  ymin = infections.lo,
+                  ymax = infections.hi),
+              alpha = .6) +
+  # geom_vline(xintercept=end_old_model,
+  #            linetype=2,
+  #            color="darkred",
+  #            linewidth=.5) +
+  facet_wrap(~name, ncol=3, scales = "free",
+              labeller= labeller(
+                vlabel =as_labeller(TeX, default=label_parsed))) +
+  scale_y_continuous(labels = comma) +
+  scale_x_date(date_breaks = "3 months", 
+               date_labels = "%b %Y") +
+  theme_c(axis.text.x=element_text(size=7),
+          axis.text.y = element_text(size=4.5),
+          legend.position = "top",
+          legend.text=element_text(size=12),
+          strip.text.x=element_text(size=11,
+                                    face='plain',
+                                    color='white')) +
+  theme(plot.title = element_text(face="bold",
+                                 hjust = 0, 
+                                 size = 13,
+                                 margin=margin(0,0,0,0))) +
+  scale_fill_manual(values =comp_state_pal, name='')  +
+  labs(y = "Estimated Infections",
+       title = paste0(
+         "(a) Estimated Infections by State in States with a Lag")) +
+  guides(color = guide_legend(override.aes = list(linewidth =2),
+                              nrow=2,
+                              byrow=TRUE))
+lagged_fig_a
+
+```
+
+```{r, fig.height=11, fig.width=12, eval=save_figures}
+
+ggsave(here('figures/intervals-and-covidestim-states-lag.pdf'))
+
+```
+
+## Test Positivity and Positive Tests in States with a Lag
+
+```{r testpos-lag, fig.height=14, fig.width=12}
+
+############################################
+# POSITIVE TESTS AND TEST POSITIVITY
+############################################
+
+ subset %>%
+  filter(biweek<=25) %>%
+  filter(state %in% states_with_lag) %>%
+    mutate(testpos = positive/total)
+
+pltlist <- subset %>%
+  filter(biweek >= 6 & biweek<=25) %>%
+  filter(state %in% states_with_lag) %>%
+    mutate(testpos = positive/total) %>%
+  group_by(across(-date)) %>%
+  summarize(date=min(date) + days(7)) %>%
+  group_by(state) %>%
+  mutate(adj = (1/max(testpos)) * max(positive)) %>%
+  group_split() %>%
+  map(~ {
+    adj <-unique(.x$adj)
+    .x %>%
+      ggplot(aes(x=date))+
+    geom_line(aes(y=positive, color = 'Positive Tests'),
+              alpha=.8) +
+      geom_point(aes(y=positive, color = 'Positive Tests'),
+              alpha=.5,
+              size=.9) +
+    geom_line(aes(y=testpos*adj, color='Test Positivity'),
+              alpha=.8
+              #,linetype=2
+              ) +
+      geom_point(aes(y=testpos*adj, color='Test Positivity'),
+              alpha=.5,
+              size=.9) +
+    #scale_linetype_discrete(breaks = c(1,1,2)) +
+    scale_color_manual(values=c(
+                                "Positive Tests"= "#279143",
+                                "Test Positivity"= '#E38E4F')) +
+    guides(color = guide_legend(override.aes = list(linetype = c(1,2),
+                                                    linewidth=c(2,2)))) +
+    theme_c() +
+    theme(axis.title = element_text(size=10),
+          axis.text =element_text(size=9),
+          legend.position="none") +
+    scale_y_continuous(labels=comma, 
+                       sec.axis = sec_axis(
+                         ~./adj,
+                         name = 'Test Positivity')) +
+    labs(color ='',
+         y="Positive Tests",
+         x= "Date") +
+      facet_wrap(~name) +
+      scale_x_date(date_labels="%b %Y",
+                   date_breaks = "3 months")
+
+  })
+
+legend <- get_legend(
+  pltlist[[1]] +
+    theme(legend.position = "top",
+          legend.text=element_text(size=13, margin=margin(6,3,0,0)))
+)
+
+title_gg <- ggplot() +
+  labs(
+    #title = "(b) Comparing Trends in Positive Tests and Test Positivity"
+    title = "") +
+  theme(plot.title=element_text(face="bold",
+                                hjust = 0,
+                                size = 13))
+
+plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)
+
+lagged_fig_b <- plot_grid(title_gg, legend, plt, rel_heights=c(.03, .03,.94), ncol=1)
+
+# lagged_fig_b <- plot_grid( legend, plt, rel_heights=c(.05,.95), ncol=1)
+
+lagged_fig_b
+
+```
+
+### States Where Covidestim Estimates Appear to be Lagged Relative to Probabilistic Bias Counts
+```{r, eval=save_figures, fig.height=14.8, fig.width=12}
+
+lag_title <- ggplot() +
+  labs(title = "States Where Covidestim Estimates Appear to be Lagged\n Relative to Probabilistic Bias Counts") +
+  theme(plot.title=element_text(face="bold", hjust = .5, size=18))
+
+plot_grid(# lag_title, 
+          lagged_fig_a + labs(title=""), 
+          lagged_fig_b,
+          labels=c(
+           # "",
+            "(a) Estimated Infections by State in States with a Lag",
+            "(b) Comparing Trends in Positive Tests and Test Positivity in States with a Lag"),
+        #  rel_heights = c(.04, .47,.49),
+          label_x=0,
+         # vjust=c(1.5,1.5,.1),
+        vjust=c(1.5,.1),
+          hjust=-.05,
+          ncol=1)
+
+ggsave(here('figures/testpos-lag.pdf'))
+
+```
+
+```{r}
+
+########################################################
+# PERCENT CHANGE IN POSITIVE TESTS AND TEST POSITIVITY
+########################################################
+
+ subset %>%
+  filter(biweek<=25) %>%
+  filter(state %in% states_with_lag) %>%
+    mutate(testpos = positive/total)
+
+pltlist <- subset %>%
+  filter(biweek >= 6 & biweek<=25) %>%
+  filter(state %in% states_with_lag) %>%
+    mutate(testpos = positive/total) %>%
+  group_by(across(-date)) %>%
+  # put date in midpoint
+  summarize(date=min(date) + days(7)) %>%
+  group_by(state) %>%
+  mutate(pct_change_testpos = (testpos-lag(testpos, n=1, order_by=biweek)) /
+                                 lag(testpos, n=1,
+                                     order_by=biweek),
+         pct_change_positive = (positive - lag(positive,
+                                               n=1,
+                                               order_by=biweek)) / 
+                                 lag(positive,
+                                               n=1,
+                                               order_by=biweek)) %>%
+ #  mutate(adj = (1/max(testpos)) * max(positive)) %>%
+  group_split() %>%
+  map(~ {
+   # adj <-unique(.x$adj)
+    .x %>%
+      ggplot(aes(x=date))+
+    geom_line(aes(y=pct_change_positive, color = 'Positive Tests'),
+              alpha=.8) +
+      geom_point(aes(y=pct_change_positive, color = 'Positive Tests'),
+              alpha=.5,
+              size=.9) +
+    geom_line(aes(y=pct_change_testpos, color='Test Positivity'),
+              alpha=.8
+              #,linetype=2
+              ) +
+      geom_point(aes(y=pct_change_testpos, color='Test Positivity'),
+              alpha=.5,
+              size=.9) +
+    #scale_linetype_discrete(breaks = c(1,1,2)) +
+    scale_color_manual(values=c(
+                                "Positive Tests"= "#279143",
+                                "Test Positivity"= '#E38E4F')) +
+    guides(color = guide_legend(override.aes = list(linetype = c(1,2),
+                                                    linewidth=c(2,2)))) +
+    theme_c() +
+    theme(axis.title = element_text(size=10),
+          axis.text =element_text(size=9),
+          legend.position="none") +
+    # scale_y_continuous(labels=comma, 
+    #                    sec.axis = sec_axis(
+    #                      ~./adj,
+    #                      name = 'Test Positivity')) +
+    labs(color ='',
+         y="Positive Tests",
+         x= "Date") +
+      facet_wrap(~name) +
+      scale_x_date(date_labels="%b %Y")
+
+  })
+
+legend <- get_legend(
+  pltlist[[1]] +
+    theme(legend.position = "top",
+          legend.text=element_text(size=16))
+)
+
+title_gg <- ggplot() + 
+  labs(title = "Comparing Trends in Positive Tests and Test Positivity") + 
+  theme(plot.title=element_text(face="bold",
+                                hjust = .5, 
+                                size = 18,
+                                margin =margin(5,0,2,0)))
+
+plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)
+
+plot_grid(title_gg, legend, plt, rel_heights=c(.03, .03,.94), ncol=1)
+
+```
+
+```{r}
+
+########################################################
+# PERCENT CHANGE IN POSITIVE TESTS AND TOTAL TESTS
+########################################################
+
+pltlist <- subset %>%
+  filter(biweek >= 6 & biweek<=25) %>%
+  filter(state %in% states_with_lag) %>%
+    mutate(testpos = positive/total) %>%
+  group_by(across(-date)) %>%
+  # put date in midpoint
+  summarize(date=min(date) + days(7)) %>%
+  group_by(state) %>%
+  mutate(pct_change_testpos = (total-lag(total, n=1, order_by=biweek)) /
+                                 lag(total, n=1,
+                                     order_by=biweek),
+         pct_change_positive = (positive - lag(positive,
+                                               n=1,
+                                               order_by=biweek)) / 
+                                 lag(positive,
+                                               n=1,
+                                               order_by=biweek)) %>%
+ #  mutate(adj = (1/max(testpos)) * max(positive)) %>%
+  group_split() %>%
+  map(~ {
+   # adj <-unique(.x$adj)
+    .x %>%
+      ggplot(aes(x=date))+
+    geom_line(aes(y=pct_change_positive, color = 'Positive Tests'),
+              alpha=.8) +
+      geom_point(aes(y=pct_change_positive, color = 'Positive Tests'),
+              alpha=.5,
+              size=.9) +
+    geom_line(aes(y=pct_change_testpos, color='Total Tests'),
+              alpha=.8
+              #,linetype=2
+              ) +
+      geom_point(aes(y=pct_change_testpos, color='Total Tests'),
+              alpha=.5,
+              size=.9) +
+    #scale_linetype_discrete(breaks = c(1,1,2)) +
+    scale_color_manual(values=c(
+                                "Positive Tests"= "#279143",
+                                "Total Tests"= '#E38E4F')) +
+    guides(color = guide_legend(override.aes = list(linetype = c(1,2),
+                                                    linewidth=c(2,2)))) +
+    theme_c() +
+    theme(axis.title = element_text(size=10),
+          axis.text =element_text(size=9),
+          legend.position="none") +
+    # scale_y_continuous(labels=comma, 
+    #                    sec.axis = sec_axis(
+    #                      ~./adj,
+    #                      name = 'Test Positivity')) +
+    labs(color ='',
+         y="Total Tests",
+         x= "Date") +
+      facet_wrap(~name) +
+      scale_x_date(date_labels="%b %Y")
+
+  })
+
+legend <- get_legend(
+  pltlist[[1]] +
+    theme(legend.position = "top",
+          legend.text=element_text(size=16))
+)
+
+title_gg <- ggplot() + 
+  labs(title = "Comparing Trends in Positive Tests and Total Tests") + 
+  theme(plot.title=element_text(face="bold",
+                                hjust = .5, 
+                                size = 18,
+                                margin =margin(5,0,2,0)))
+
+plt <- cowplot::plot_grid(plotlist=pltlist, ncol=3)
+
+plot_grid(title_gg, legend, plt, rel_heights=c(.03, .03,.94), ncol=1)
 

### 2023-08-28 writing/supplement.Rmd
update lag figures

@@
     number_sections: false
-
+header-includes:
+- \usepackage{float}
 ---
@@
 ```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE)
+knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.pos="H",out.extra = "")
 library(knitr)
@@
 Here, we consider the 2.5\\% and 97.5\\% percentiles and mean for each state.
-Most have a mean near $20%$."}
+Most have a mean near 0.20."}
 
@@
 
-```{r testrate-low-summer, fig.cap="Testing rate, calculated as the etotal number of tests for a 2-week interval in a state over the census population in that state, over time.
+```{r testrate-low-summer, fig.cap="Testing rate, calculated as the total number of tests for a 2-week interval in a state over the census population in that state, over time.
 The median across states for each two-week interval is shown in red.
@@
 
-include_graphics(here('figures/intervals-and-covidestim-states-lag.pdf'))
+# include_graphics(here('figures/intervals-and-covidestim-states-lag.pdf'))
 
@@
 
-```{r}
+```{r lagged-figs, fig.cap="States where the Covidestim infection estimates are lagged relative to the probabilistic bias estimates, that is, they are most highly correlated when the Covidestim infections are lagged by one or more biweeks.
+(a) Simulation intervals demonstrating the lag seen around the delta peak in summer 2021.
+(b) Test positivity and positive tests for each state.
+Increases in test positivity tend to precede increases in positive tests, and these increases in test positivity result in increases in the estimates probduced by the probabilistic bias analysis seen in part (a)."}
 
@@
 
-```{r michigan-variant, fig.cap = "Simulation intervals for Michigan along with the variant proportions for variants designated as *Variant of Interest* or *Variant of Concern* by the World Health Organization.
+```{r michigan-variant, fig.cap = "Simulation intervals for Michigan along with the variant proportions for variants designated as \\emph{Variant of Interest} or \\emph{Variant of Concern} by the World Health Organization.
 These variant proportions are not specific to Michigan; the estimates are for the United States.

### 2023-09-03 reports/state_figures.Rmd
update figs

@@
   scale_y_continuous(labels = comma) +
-  scale_x_date(date_breaks = "3 months", 
+  scale_x_date(
+    #date_breaks = "3 months", 
+    breaks=ymd("2021-03-12", "2021-05-21", "2021-07-30", "2021-10-08"),
                date_labels = "%b %Y") +

### 2023-09-03 writing/manuscript.Rmd
update figs

@@
 
-ADD TEST POSITIVITY FIGURE
-FIGURE OUT WHICH STATES HAVE LAGS
-
 Another difference we see is in the months of the summer of 2021, where the increase in Covidestim estimates appears lagged in comparison to the probabilistic bias simulation intervals.
 The difference we see between the Covidestim estimates and probabilistic bias intervals is likely a result of the way these approaches treat incomplete testing.
-The focus of the probabilistic bias analysis is to correct for incomplete testing, and as such the method is sensitive to changes in the total number tested and the positivity rate.
+The focus of the probabilistic bias analysis is to correct for incomplete testing, and as such the method is sensitive to changes in the total number tested and the positivity rate (see Figure DIAGRAM in the methods section).
 By contrast, while Covidestim models probabilities of diagnosis by symptom state to allow for variation in case ascertainment, the total number of tests is not an input into the model, so model estimates are not affected by changes in testing rate.
-This is particularly relevant during the summer of 2021, when an increase in test positivity preceded the increase in observed tests, leading to an increase in the probabilistic bias simulation intervals that precedes the increase in Covidestim estimates.
+This is particularly relevant during the summer of 2021, when an increase in test positivity preceded the increase in observed tests, leading to an increase in the probabilistic bias simulation intervals that precedes the increase in Covidestim estimates in many states (Supplementary Figure lagged-figs).
+Differences attributed to how these methods consider test positivity are likely to be heightened during times where increases in infections outpace testing capacity or test utilization.
 
@@
 
+# Discussion
+
+### Limitations
+
+- Validation:
+  - Wastewater encatchments do not fall exactly on county lines, so aggregation here contains limited information
+  - Covidestim -- sum of daily medians != median for entire two-week time interval
+- Repeat testing may bias estimates in places where this form of testing comprises a substantial proportion of total testing, but data on people-viral-total and people-viral-positive is not widely available
+
 # Methods {#Methods}
@@
 
-### Massachusetts County Level
+We acquired state-level PCR testing data from public-use data on HealthData.gov, which includes both positive and negative test counts [@unitedstatesdepartmentofhealthandhumanservices].
+At the county level, we obtained publicly available data from Massachusetts published on the state government website [@departmentofpublichealth].
+
+Covidestim estimates for dates after December 2021 were obtained from the Covidestim API, and those  for dates after December 2021 were obtained from the legacy page.
 
 ### State Level
+
+### Massachusetts County Level
 
@@
 
-### Limitations
+# Data Availability
 
-- Validation:
-  - Wastewater encatchments do not fall exactly on county lines, so aggregation here contains limited information
-  - Covidestim -- sum of daily medians != median for entire two-week time interval
-- Repeat testing may bias estimates in places where this form of testing comprises a substantial proportion of total testing, but data on people-viral-total and people-viral-positive is not widely available
+# Code Availability
 

### 2023-09-03 writing/supplement.Rmd
update figs

@@
 
-```{r lagged-figs, fig.cap="States where the Covidestim infection estimates are lagged relative to the probabilistic bias estimates, that is, they are most highly correlated when the Covidestim infections are lagged by one or more biweeks.
+```{r lagged-figs, fig.cap="States where the Covidestim infection estimates appear to be lagged relative to the probabilistic bias estimates, that is, they are most highly correlated when the Covidestim infections are lagged by one or more biweeks.
 (a) Simulation intervals demonstrating the lag seen around the delta peak in summer 2021.
